### 2018-11-21 ITiCSE2019.tex
Initial commit

%%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
\documentclass[sigconf]{acmart}
\usepackage{paralist}
\usepackage{url}
\usepackage[hyphenbreaks]{breakurl}

\def\UrlBreaks{\do\/\do-}

%%%% As of March 2017, [siggraph] is no longer used. Please use sigconf (above) for SIGGRAPH conferences.

%%%% Proceedings format for SIGPLAN conferences 
% \documentclass[sigplan, anonymous, review]{acmart}

%%%% Proceedings format for SIGCHI conferences
% \documentclass[sigchi, review]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\copyrightyear{2019}
\acmYear{2019}
\setcopyright{acmlicensed}
\acmConference[ITiCSE '19]{The 24th Annual Conference on Innovation
  and Technology in Computer Science Education}{Jul. 15-17,
  2019}{Aberdeen, UK}
%\acmBooktitle{}
%\acmPrice{15.00}
%\acmDOI{10.1145/3159450.3159547}
%\acmISBN{978-1-4503-5103-4/18/02}
% This slight change to the code may also save 1 or 2 lines of space.

% removes the headers from each page per the preparation instructions, as these are not needed and will be updated with the chairs' actual session names during the pagination/indexing process:
\fancyhead{}

\begin{document}
\title{What Do We Mean By Cybersecurity?}
%\titlenote{}
%\subtitle{Extended Abstract}
%\subtitlenote{}
\author{James H. Davenport}
\orcid{0000-0002-3982-7545}
\affiliation{%
  \institution{University of Bath}
  \streetaddress{}
  \city{Bath} 
  \country{United Kingdom}
}
\email{j.h.davenport@bath.ac.uk}

\author{Tom Crick}
\orcid{0000-0001-5196-9389}
\affiliation{%
  \institution{Swansea University}
  \streetaddress{}
  \city{Swansea} 
  \country{United Kingdom}
}
\email{thomas.crick@swansea.ac.uk}

% The default list of authors is too long for headers}
%\renewcommand{\shortauthors}{Moller and Crick}

\begin{abstract}
Abstract here...
\end{abstract}

\keywords{TBC}

\maketitle

\section{introduction}

And here we go...

%\section*{Acknowledgements}

\bibliographystyle{ACM-Reference-Format}
\bibliography{ITiCSE2019}

\end{document}

### 2018-11-24 ITiCSE2019.tex
JHD had a start

@@
 \begin{document}
-\title{What Do We Mean By Cybersecurity?}
+\title{How Well do we Teach Cybersecurity?}
+%\title{What Do We Mean By Cybersecurity?}  Tom's original was this
 %\titlenote{}
@@
 
-\section{introduction}
+\section{Introduction}
 
-And here we go...
+Cybersecurity has been in the news for several years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
+
+There are calls for education to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}.
+
+Nevertheless, it is one thing to write curricula and requirements, and another thing to deliver appropriate education, and one could reasonably ask how well this is done in practice.
+
+\section{Challenges}
+
+\subsection{Educational Resources}
+JHD to write about Database books.
+
+\subsection{Staff}
+JHD possibly to write about his recruitment experiences.  ? more IoC input?
+
+\subsection{Accreditation}
+?AI/TP to write something.
 

### 2018-11-25 ITiCSE2019.tex
More narrative

@@
 
-Nevertheless, it is one thing to write curricula and requirements, and another thing to deliver appropriate education, and one could reasonably ask how well this is done in practice.
+Nevertheless, it is one thing to write curricula and requirements, and
+another thing to deliver appropriate education, and one could
+reasonably ask how well this is done in practice.
+
+\section{Policy Context}
+
+Substantial CS/digital curriculum reform across the
+UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
+-- but what about focus on cybersecurity?
+
+UK national economic skills priority
+e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
+(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
+and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
+\url{https://www.cybersecuritychallenge.org.uk/}). Large-scale media
+attention on the ``cybersecurity skills gap''
+e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
+and
+\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
+etc -- link to UK Digital Strategy and UK Industrial Strategy.
 
 \section{Challenges}
+
+% should we frame this as one or more case studies?
+% do we have any national survey data?
 

### 2018-11-26 ITiCSE2019.tex
Format changes
Added database reference
Added PCI DSS
More SQL stuff

@@
 and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
-\url{https://www.cybersecuritychallenge.org.uk/}). Large-scale media
+\url{https://www.cybersecuritychallenge.org.uk/}).
+
+There is large-scale media
 attention on the ``cybersecurity skills gap''
@@
 
+\subsection{PCI DSS}
+\cite{PCI2018b}.
+
 \subsection{Educational Resources}
-JHD to write about Database books.
+It is 15 years since \cite{Guimaraesetal2004} 
+wrote ``All the topics listed above should be presented in the first
+Database Course'', and the first such topic was SQL injection \cite{Anonymous2018b}.
+SQL injection has been around fro twenty years \cite{HornerHyslip2017a}.
+Nevertheless SQL injection is still a major weakness: number one in the OWASP Top 10 \cite{OWASP2017a}.
+
+JHD to write about Database books. \cite{Drop2019}.
 

### 2018-12-13 ITiCSE2019.tex
More placeholders
Added co-authors
More updates
Add conference dates

@@
 \begin{document}
-\title{How Well do we Teach Cybersecurity?}
+\title{How Well do we Teach Cybersecurity?\\(Abstract 20 January, Ppaer 27th}
 %\title{What Do We Mean By Cybersecurity?}  Tom's original was this
@@
 \email{thomas.crick@swansea.ac.uk}
+
+\author{Alastair Irons}
+\orcid{}
+\affiliation{%
+  \institution{}
+  \streetaddress{}
+  \city{} 
+  \country{United Kingdom}
+}
+\email{@lastair.irons@sunderland.ac.uk}
+
+\author{Tom Prickett}
+\orcid{}
+\affiliation{%
+  \institution{}
+  \streetaddress{}
+  \city{} 
+  \country{United Kingdom}
+}
+\email{tom.prickett@northumbria.ac.uk}
 
@@
 
+Alastair to write about their Credit Fraud experience (MSc CyberSecurity).
+HE Academy grant (to be referenced).
+
+What should be in the general curriculum.
+
 \subsection{Educational Resources}
@@
 
+Alastair's experiences.
+
+? Poll CPHC, and/or survey job adverts.
 \subsection{Accreditation}

### 2018-12-20 ITiCSE2019.tex
First draft of Accreditation section
To Do
expand on Roadshow
evaluate success?

@@
 \affiliation{%
-  \institution{}
+  \institution{Sunderland University}
   \streetaddress{}
-  \city{} 
+  \city{Sunderland} 
   \country{United Kingdom}
 }
-\email{@lastair.irons@sunderland.ac.uk}
+\email{@alastair.irons@sunderland.ac.uk}
 
@@
 \affiliation{%
-  \institution{}
+  \institution{Northumbria University}
   \streetaddress{}
-  \city{} 
+  \city{Newcastle upon Tyne} 
   \country{United Kingdom}
@@
 
+For a few years now there has been a recognized need to build knowledge, skills and capacity in the area of Cybersecurity.
+This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
+\cite{UKCabinetOffice} or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.
+
+The teaching of Cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of Cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
+There has  been a debate as to whether Cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}.
+The consensus now is that Cybersecurity is both a discipline in its own right and that Cybersecurity should be taught within Computer Science and related degrees.
+There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite{ACM2013a} and specialised Cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+
+In the United Kingdom, Higher Education provision addresses both approaches.
+A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and Cybersecurity ( and closely related fields Computer Security, Digital / Computer Forensics, etc).
+In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to Cybersecurity for entry in September 2019.
+An even larger number of HEI provide study opportunities related to more general Computer Science.
+UCAS lists 246 provides for undergraduate programmes related to Computer Science.
+
+Accreditation has evolved to directly addresses the Cybersecurity challenges in both general Computer Science programmes and specialist Cybersecurity programmes.
+In the UK, accreditation in the broad computing area is being performed by a few different agencies.
+These include:
+
+\textbf{1. Not for Profit Organisations}
+Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements.
+One of the curricula elements is related to Cybersecurity.
+Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.
+Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and 5 Software Engineering for Business Programmes.
+As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education and none upon specialist Cybersecurity education.
+
+The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK.
+This could potentially include Cybersecurity related skills.
+Like the Tech Partnership the focus is upon job market requirements.
+Given the size of this initiative this potentially has a significant role to play however at time of writing it is not clear what that shall be.
+
+\textbf{2. National Cyber Security Centre (NCSC)}
+The NCSC is a UK Government organisation tasked with enhancing the Cybersecurity of the UK.
+The NCSC publishes and accredits to a number of Cybersecurity standards \cite{NCCS2018a}.
+These standard are linked to the ACM recommendations for curricula.
+To date the major focus of NSSC acreditation has been upon Masters degrees specializing in Cybersecurity.
+More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in Cybersecurity and Computer Science degrees with a significant Cybersecurity focus \cite{NCCS2018b}.
+
+Currently accredited are 15 Cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters Cybersecurity programmes and 1 Cybersecurity Degree Programme with a further 2 provisionally accredited.
+Hence the extent of accreditation is currently reasonably limited in terms of reach to Computer Science programmes.
+This appears to be a positive initiative that will hopefully further develop over time.
+
+\textbf{3. Professional Bodies}
+The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of Cybersecurity discipline areas.
+The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord and Seoul Accord (do we need to reference these).
+These memorandums support the internationalising of the curriculum and promote consistency and parity in Computer Science Education globally.
+These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
+Both the IET and The BCS have a long history of expecting coverage environmental factors within the programmes they accredit.
+The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}.
+Clearly Cybersecurity has been and continues to be part of these expectations.
+
+This ambition of this approach is to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area.)  The approach taken is not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
+In this context this is realized as an expectation Cybersecurity is an inclusion in all degrees accredited by the BCS.
+e.g.
+the expectation for coverage is true for Computer Science as well as Cybersecurity programmes.
+Two criteria are expected to be covered by all programmes seeking accreditation.
+These are \cite{BCS2018a}:
+
+2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices
+
+2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system
+
+Additionally programmes seeking Chartered Information Technology Professional Accreditation also have to cover:
+
+3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security
+
+In the context of BCS accreditation, these requirements imply an exit standard that all students on a programme must be able to demonstrate irrespective of the option choices they have made.
+This means a HEI applying for accreditation is expected to provide evidence that the criteria are taught and assessed in a non-trivial manner, to and by all students upon the programme seeking accreditation .
+A HEI is expected to provide evidence in the form of programme and module specification documentation and example assessment specifications (coursework and examinations).
+These criteria and the expectation that they are taught and assessed has been present for a number of years.
+
+Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis.
+Building on the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
+This development was a partnership between the BCS,  The Council of Heads and Professors of Computing, (ISC)2 (the largest not-for-profit membership body of certified information and software security professionals worldwide, with over 130,000 members) and the Office for Cyber Security and Information Assurance in the Cabinet Office.
+This project lead to the agreement of a set of Cybersecurity Principles and Learning Outcomes for Computer Science and related IT Degrees in the UK \cite{CPHCISC2}.
+These guidelines are closely tied to  the ACM recommendations for curricula \cite{ACM2013a}.
+
+The agreed Cybersecurity Principles and Learning Outcomes were included in the BCS Guidelines for accreditation in 2016.
+This resulted in an extended expectation of the coverage of Cybersecurity in programmes accredited by the Institute.
+The expectation is now that all students upon all accredited programmes will graduate with skills, knowledge and experience related to:
+
+1. Information and risk
+2. Threats and attacks
+3. Cybersecurity architecture and operations
+4. Secure systems and products
+5. Cybersecurity management
+
+This was discussed with the wider education community by a road-show lead by CPHC. (Alastair could you add something here?)
+
+This change to the accreditation guidelines is now in an implementation period.
+The accreditation process adopted by the BCS is cyclic in nature.
+Formally the cycle is 5 years in duration.
+The new expectations have been implemented as follows.
+To ensure continuous accreditation,  accreditation visits are normally scheduled every 4 years.
+At the time of the next visit in this accreditation cycle accreditation is conditional upon a HEI having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
+It is anticipate from 2020(?) the expectation will be all accredited programmes have the new expectations fully embedded.
+
+TO DO - Has this worked? Should we ask the BCS accreditation team a favour to see if they will let us know: 
+1. Total number of accredited programmes
+2. impact of changes to  BCS Accreditation Cybersecurity requirements - maybe number of HEIs with a 90 day related to it or possibly number with "At Threshold"
+3. Would this give us enough evidence of some impact? Clearly it is a work in progress but I think some progress has been made?
+4. Any things else?
+
 %\section*{Acknowledgements}

### 2018-12-20 ITiCSE2019.tex
JHD additions/bibliography edits

@@
 \begin{document}
-\title{How Well do we Teach Cybersecurity?\\(Abstract 20 January, Ppaer 27th}
+\title{How Well do we Teach Cybersecurity?\\(Abstract 20 January, Paper 27th)}
 %\title{What Do We Mean By Cybersecurity?}  Tom's original was this
@@
 reasonably ask how well this is done in practice.
+\subsection{CyberSecurity: for all or for specialists}
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} CyberSecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
 
+Specialist curricula abound: a recent one is \cite{ACM2017b}.
 \section{Policy Context}

### 2018-12-20 ITiCSE2019.tex
JHD updates during talk

--- 2018-12-20 ITiCSE2019.tex
@@
 reasonably ask how well this is done in practice.
+\subsection{CyberSecurity: for all or for specialists?}
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} CyberSecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
 
+AI: GDPR essentially expects CyberSecurity to be in all coding. 
+JHD: PCI DSS insists on a Web Application Firewall, which is only needed if you don't trust your applications.
+
+Specialist curricula abound: a recent one is \cite{ACM2017b}.
+However, given the shortage of specialists, in practice a good generalist with some CYberSecurity expertise can often get a ``specialist''.
 \section{Policy Context}
@@
 The expectation is now that all students upon all accredited programmes will graduate with skills, knowledge and experience related to:
-
-1. Information and risk
-2. Threats and attacks
-3. Cybersecurity architecture and operations
-4. Secure systems and products
-5. Cybersecurity management
-
+\begin{enumerate}
+\item Information and risk
+\item Threats and attacks
+\item Cybersecurity architecture and operations
+\item Secure systems and products
+\item Cybersecurity management
+\end{enumerate}
 This was discussed with the wider education community by a road-show lead by CPHC. (Alastair could you add something here?)
@@
 The new expectations have been implemented as follows.
-To ensure continuous accreditation,  accreditation visits are normally scheduled every 4 years.
+To ensure continuous accreditation,  accreditation visits are normally scheduled every 5 years.
 At the time of the next visit in this accreditation cycle accreditation is conditional upon a HEI having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.

### 2018-12-20 ITiCSE2019.tex
More fixes

@@
 }
-\email{@alastair.irons@sunderland.ac.uk}
+\email{alastair.irons@sunderland.ac.uk}
 
@@
 
-Specialist curricula abound: a recent one is \cite{ACM2017b}.
-However, given the shortage of specialists, in practice a good generalist with some CYberSecurity expertise can often get a ``specialist''.
+Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}.
+However, given the shortage of specialists, in practice a good generalist with some CyberSecurity expertise can often get a ``specialist''.
+\subsection{Research Questions}
+\begin{enumerate}
+\item The person in teh street --- probably not.
+\item The general CS graduate.
+\item The general CS masters graduate.
+\item The specialist CS graduate.
+\item The specialist CS masters graduate.
+\end{enumerate}
 \section{Policy Context}

### 2018-12-20 ITiCSE2019.tex
Added JCNSS2018a and reference

@@
 \subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} CyberSecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} CyberSecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
 
@@
 Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}.
-However, given the shortage of specialists, in practice a good generalist with some CyberSecurity expertise can often get a ``specialist''.
+However, given the shortage of specialists, in practice a good generalist with some CyberSecurity expertise can often get a ``specialist'' job.
 \subsection{Research Questions}

### 2018-12-20 ITiCSE2019.tex
Further JHD

@@
 e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
-(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
+%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
+(especially \cite{JCNSS2018a}
 and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and

### 2018-12-25 ITiCSE2019.tex
More SQL

@@
 wrote ``All the topics listed above should be presented in the first
-Database Course'', and the first such topic was SQL injection \cite{Anonymous2018b}.
-SQL injection has been around fro twenty years \cite{HornerHyslip2017a}.
-Nevertheless SQL injection is still a major weakness: number one in the OWASP Top 10 \cite{OWASP2017a}.
+Database Course'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}.
+SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon (\url{https://xkcd.com/327/}, dating back to 2007 according to the Internet Archive) and website (\url{http://bobby-tables.com/}).
+Nevertheless SQL injection is still a major weakness: number one in the OWASP Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.
 
-JHD to write about Database books. \cite{Drop2019}.
+Clearly such a major weakness should be well-taught.
+One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cyber-security course.
+It's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.
+
+JHD to write about Database books. \cite{Drop2019}. using \cite{StangerMartin2015a}
 

### 2018-12-26 ITiCSE2019.tex
More NCSC stuff

@@
 
-\subsection{Educational Resources}
+\subsection{Educational Resources: SQL Injection}
 It is 15 years since \cite{Guimaraesetal2004} 
@@
 One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cyber-security course.
-It's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.
+In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.
 
-JHD to write about Database books. \cite{Drop2019}. using \cite{StangerMartin2015a}
+Hence \cite{Drop2019} analyzed the database textbooks used by  44 of the top 50 Computer Science
+departments in the United States (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed).
+There were seven such boks, but three accounted for the 36 of the 44.
+Five of the seven (30 of the 44) had no mention of SQL injection.
+On the other two, the more popular one has a seriously flawed discussion\footnote{``However, they imply that using parameters is equivalent to using a function to add escape characters
+around user input. This is incorrect, as using parameters allows
+SQL statements to be pre-compiled, and prevents any user input
+from being interpreted as code, while escaping user input is not
+recommended as a sole defense since imperfect escape functions
+can easily be subverted.'' \cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``However, the fact that the first
+example should not be used is not discussed until two pages after
+the example in the text, and is not mentioned at all in the caption or
+on the page where the figure appears. This means a student who is
+skimming the text looking for an example to modify for their own
+code could simply copy the code that first appears in the example,
+without being aware that this is in fact an example of what they
+shouldn't do.'' \cite{Drop2019}}.
+
+\subsection{Informal Resources}
+The web abounds with informal resources: tutorials and code snippets.
+How good are these, and how good are people at using these?
+This has been looked at by \cite{Unruhetal2017a}, who took the top five results from Google for six queries.
+Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in WASP's Top Ten \cite{OWASP2017a}.} weaknesses.
+Searching for these fragments in PHP projects GitHub found 82 instances of these fragments, of which 117 were verified by manual to be vulnerable --- 80\% of which were vulnerable to SQL injection.
 
@@
 The NCSC is a UK Government organisation tasked with enhancing the Cybersecurity of the UK.
-The NCSC publishes and accredits to a number of Cybersecurity standards \cite{NCCS2018a}.
+The NCSC publishes and accredits to a number of Cybersecurity standards \cite{NCSC2018a}.
 These standard are linked to the ACM recommendations for curricula.
 To date the major focus of NSSC acreditation has been upon Masters degrees specializing in Cybersecurity.
-More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in Cybersecurity and Computer Science degrees with a significant Cybersecurity focus \cite{NCCS2018b}.
+More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in Cybersecurity and Computer Science degrees with a significant Cybersecurity focus \cite{NCSC2018b}.
 

### 2019-01-02 ITiCSE2019.tex
Added Michael Page quote on skil shortage
Minor comments from paper chat

@@
 \begin{enumerate}
-\item The person in teh street --- probably not.
+\item The person in the street --- probably not.
+({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
+Minor point but maybe worth mentioning e.g.
+being safe online, accessing public services, etc}})
 \item The general CS graduate.
@@
 \subsection{Staff}
+It is well known that CyberSecurity skills are in short supply, for example \cite{Page2018a}.
+\begin{quote}
+
+\end{quote}
 JHD possibly to write about his recruitment experiences.  ? more IoC input?

### 2019-01-02 ITiCSE2019.tex
Changes from telco

@@
 \begin{document}
-\title{How Well do we Teach Cybersecurity?\\(Abstract 20 January, Paper 27th)}
+\title{Are We Teaching Cybersecurity (and how well)?\\(Abstract 20 January, Paper 27th)}
 %\title{What Do We Mean By Cybersecurity?}  Tom's original was this
@@
 \item The person in the street --- probably not.
-({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
-Minor point but maybe worth mentioning e.g.
-being safe online, accessing public services, etc}})
 \item The general CS graduate.
@@
 
+What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
+{\bf JHD to research ACM general curriculum; also what is status of CYBOOK etc.}
+
+Is teaching CyberSecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways)
+
+AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.
+{\bf AI to write up}.
+  Should we be teaching it because it's underpinning? because there's a skill shortage?  JHD would argue both.
+
 % should we frame this as one or more case studies?
 % do we have any national survey data?
+
+\subsection{What does BCS tell Universities}
+and what happens.  {\bf Action TP}.
 
@@
 HE Academy grant (to be referenced).
-
-What should be in the general curriculum.
 
@@
 
-Alastair's experiences.
+{\bf Alastair's experiences.}
 
-? Poll CPHC, and/or survey job adverts.
+{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
 \subsection{Accreditation}

### 2019-01-02 ITiCSE2019.tex

@@
 \item The person in the street --- probably not.
+({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
+Minor point but maybe worth mentioning e.g.
+being safe online, accessing public services, etc}})
+\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design.
+BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
 \item The general CS graduate.

### 2019-01-03 ITiCSE2019.tex
Add files via upload
added "what does the BCS tell Universities" and "Accreditation - what progress has been made"

@@
 
-\subsection{What does BCS tell Universities}
-and what happens.  {\bf Action TP}.
-
 \subsection{PCI DSS}
@@
 \subsection{Accreditation}
-?AI/TP to write something.
+{\bf{Alastair could you check this and adjust if appropriate?}}
 
@@
 
-The agreed Cybersecurity Principles and Learning Outcomes were included in the BCS Guidelines for accreditation in 2016.
+The agreed Cybersecurity Principles and Learning Outcomes were included in the BCS Guidelines for accreditation in 2015.
 This resulted in an extended expectation of the coverage of Cybersecurity in programmes accredited by the Institute.
@@
 \end{enumerate}
-This was discussed with the wider education community by a road-show lead by CPHC. (Alastair could you add something here?)
+
+\subsection{What does BCS tell Universities}
+The agreed Cybersecurity Principles and Learning Outcome were discussed with the wider education community by a road-show lead by CPHC. {\bf{(Alastair could you add something here?)}}
+
+The BCS Guidelines on Course Accreditation are published upon the BCS website \cite{BCS2018a}.
+The BCS also publishes the changes that have been made \cite{BCS2018b}.
+When changes are made, the BCS communicates the changes by email and in writing to all the BCS Educational Affiliates, that is all the HEIs that seek accreditation from the BCS.
+The expectations for Cybersecurity were extended in the June 2015 version of the guidelines for consideration at Accreditation Visits that took place from September 2015 or later.
 
@@
 The accreditation process adopted by the BCS is cyclic in nature.
-Formally the cycle is 5 years in duration.
+Formally, the cycle is 5 years in duration.
 The new expectations have been implemented as follows.
 To ensure continuous accreditation, accreditation visits are normally scheduled every 5 years.
-At the time of the next visit in this accreditation cycle accreditation is conditional upon a HEI having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
-It is anticipate from 2020(?) the expectation will be all accredited programmes have the new expectations fully embedded.
+At the time of the next visit in this accreditation cycle, accreditation is conditional upon a HEI having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
+It is anticipated that from 2020(?) the expectation will be all accredited programmes have the new expectations fully embedded.
 
-TO DO - Has this worked? Should we ask the BCS accreditation team a favour to see if they will let us know: 
-1. Total number of accredited programmes
-2. impact of changes to  BCS Accreditation Cybersecurity requirements - maybe number of HEIs with a 90 day related to it or possibly number with "At Threshold"
-3. Would this give us enough evidence of some impact? Clearly it is a work in progress but I think some progress has been made?
-4. Any things else?
+In the year prior to an Accreditation Visit, HEIs are invited to an Accreditation Briefing from the BCS.
+The intention of the briefing is to help ensure Accreditation Visits go smoothly from the perspective of both the BCS and visited HEI.
+The briefings take place virtually.
+The briefing includes a summary of the process, discussion of recent changes, guidance regarding the application and a summary of common issues that are being seen in other HEIs.
+Significant opportunity for seeking clarification is provided.
+One of the issues that is highlighted is not all institutions have yet evolved their programmes to fully address the increased expectation for Cybersecurity.
+This is resulting in accreditation being contingent upon a HEI taking action to address this short fall or in some cases the withdrawal of accreditation.
+A number of HEIs are in the process of adjusting their curricula to meet the new expectations.
+In this case, the BCS notes the changes to programme design, the outputs from which will be scrutinized at the next accreditation visit.
+
+\subsection{Accreditation - what progress has been made}
+
+This initiative is a collective attempt to formally include Cybersecurity in all BCS Accredited programmes.
+Some of these programmes will be specialist Cybersecurity programmes, however the majority will take a different emphasis.
+This is a work in progress.
+A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines.
+What is being observed is the majority of visited HEIs have now either adjusted their curricula to extend the coverage of Cybersecurity or have a plan in place to do so.
+However, a minority are requiring encouragement to do so.
+
+{\bf {
+TO DO
+
+TP made information request on 03/01/19. Education team are on holiday to 7/01/19. They also are currently short staffed.
+
+What I have requested is:
+
+1.	The HEIs / number of programmes we have refused accreditation on the basis of Cybersecurity
+2.	The number of HEIs / programmes which we have requested a 90 day or an action plan on the basis of Cybersecurity ( that was then successfully addressed)
+3.	The number of HEIs / programme which are marked “At Threshold” for Cyber Security
+
+This clearly may or may not be forthcoming.
+
+}}
 

### 2019-01-05 ITiCSE2019.tex
JHD added quote from Michael Page, but pointed to ESG

@@
 \begin{quote}
+% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
+Research into the state of IT conducted annually by ESG\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
+In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
+This had climbed to 51\% at the beginning of this year.
+Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
+\end{quote}
+\begin{quote}
 

### 2019-01-10 ITiCSE2019.tex
Added Parr2014a
First hack at abstract

@@
 \begin{abstract}
-Abstract here...
+There are many facets to CyberSecurity education: we address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
+We consider particularly the UK perspective, where the British Computer Society accredits computing degrees, and insists on a certain CyberSecurity element in all such.
 \end{abstract}
@@
 In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} CyberSecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
+
+This is not a brand-new concern: see \cite{Parr2014a}
 

### 2019-01-11 ITiCSE2019.tex
Tom's IET source

@@
 Cybersecurity has been in the news for several years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
+It needs attention across the spectrum:
+\begin{quote}
+change the culture in your organisation around cyber security; to t ry to do for cyber what has been done so successfully for health and safety, for example, over the last ten years - to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
+\cite{Hannigan2019a}
+\end{quote}
 

### 2019-01-15 ITiCSE2019.tex
JAVA comments
added accreditation outcomes

@@
 
+\cite{Mengetal2018a,Chenetal2019a,Zorz2019a}
+
 \subsection{Staff}
@@
 
-{\bf {
-TO DO
+From the start of the Autumn 2015 term, up to and including the Autumn 2018 term, the BCS have carried out 70 accreditation visits (including 4 international visits).
+The BCS identified action was required to address concerns related to Cybersecurity at 16 HEIs.
+So 54 HEIs were already delivering Cybersecurity in keeping with the BCS expectations.
 
-TP made information request on 03/01/19. Education team are on holiday to 7/01/19. They also are currently short staffed.
+Long term actions were expected from 12 HEIs (6 in 2015/16, 3 in 2016/17 and 3 in the Autumn of 2018) who were awarded 'At Threshold' judgments.
+10 of these judgments were across all programmes.
+1 was specifically against a generalist masters programme only.
+This indicates that the BCS will expect adjustments to have taken place before the next accreditation visit.
+As indicated earlier, this was commonly the case that adjustments had been made to approved programmes of study, however the adjusted modules had not yet been delivered so the evidence base was incomplete in terms of how Cybersecurity was assessed.
 
-What I have requested is:
+Short terms 90 Day Responses where required from 4 HEIs. The outcomes of these actions were as follows:
 
-1.	The HEIs / number of programmes we have refused accreditation on the basis of Cybersecurity
-2.	The number of HEIs / programmes which we have requested a 90 day or an action plan on the basis of Cybersecurity ( that was then successfully addressed)
-3.	The number of HEIs / programme which are marked “At Threshold” for Cyber Security
+1)	Of the 11 UG programmes involved all were approved 'At Threshold'
 
-This clearly may or may not be forthcoming.
+2)	Of the 9 UG programmes involved, 8 were approved and 1 refused
 
-}}
+3)	Of the 5 UG programmes involved, all approved 'At Threshold'
 
-%\section*{Acknowledgements}
+4)	Of the 3 UG programmes involved, all 3 were refused.
+
+Good practice was identified at one university by the commendation:
+
+"The second year project provides an opportunity for exploring security aspects in depth with an industrial use case"
+
+In summation, this shows that many UK HEIs have now embedded Cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
+Clearly not all HEIs in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of Cybersecurity to an agreed standard.
+
+\section*{Acknowledgements}
+Thanks to Sally Pearce, Academic Accreditation Manager, BCS, The Chartered Institute for IT for supplying the summary information related to Accreditation.
 

### 2019-01-16 ITiCSE2019.tex
Added Parrishetal2018a

@@
 
+\def\IntroYear{2014}  % when BCS started asking for CyberSecurity
 \begin{abstract}
 There are many facets to CyberSecurity education: we address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
-We consider particularly the UK perspective, where the British Computer Society accredits computing degrees, and insists on a certain CyberSecurity element in all such.
+We consider particularly the UK perspective, where the British Computer Society accredits computing degrees, and (since \IntroYear) insists on a certain CyberSecurity element in all such.
 \end{abstract}
@@
 There are calls for education to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}.
+There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.
 

### 2019-01-16 ITiCSE2019.tex
Minor tweaks
Abstract

@@
 \begin{document}
-\title{Are We Teaching Cybersecurity (and how well)?\\(Abstract 20 January, Paper 27th)}
+\title{Cybersecurity is too important to be left to the specialists
+\\(Abstract 20 January, Paper 27th)}
 %\title{What Do We Mean By Cybersecurity?}  Tom's original was this
@@
 There are many facets to CyberSecurity education: we address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
+(Cyber)Security (if thought about at all) has hisorically been left to specialists, sometimes called ``Information Assurance'', which worked, even if not well, in a waterfall model.
+The Agile and DevOps trends have changed devleopment, but not themselves the view of security.
+Over the last few years, there has been a recogition of the need for change, exemplified by ``DevSecOps''.
+But if security is not to be left to the experts, then the generalist must know about it.
+
 We consider particularly the UK perspective, where the British Computer Society accredits computing degrees, and (since \IntroYear) insists on a certain CyberSecurity element in all such.
@@
 \end{enumerate}
+\pagebreak
 \section{Policy Context}

### 2019-01-16 ITiCSE2019.tex
Explain Hannigan

@@
 
+\author{Tom Prickett}
+\orcid{}
+\affiliation{%
+  \institution{Northumbria University}
+  \streetaddress{}
+  \city{Newcastle upon Tyne} 
+  \country{United Kingdom}
+}
+\email{tom.prickett@northumbria.ac.uk}
+
 \author{Alastair Irons}
@@
 
-\author{Tom Prickett}
-\orcid{}
-\affiliation{%
-  \institution{Northumbria University}
-  \streetaddress{}
-  \city{Newcastle upon Tyne} 
-  \country{United Kingdom}
-}
-\email{tom.prickett@northumbria.ac.uk}
-
 % The default list of authors is too long for headers}
-%\renewcommand{\shortauthors}{Moller and Crick}
+%\renewcommand{\shortauthors}{Davenport et al.}
 
@@
 \begin{quote}
-change the culture in your organisation around cyber security; to t ry to do for cyber what has been done so successfully for health and safety, for example, over the last ten years - to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
-\cite{Hannigan2019a}
+change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years - to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
+\cite{Hannigan2019a}\footnote{Former Director of GCHQ: Brittain equivalent of NSA.}
 \end{quote}

### 2019-01-16 ITiCSE2019.tex
Changes to abstract

@@
 \begin{document}
-\title{Cybersecurity is too important to be left to the specialists
-\\(Abstract 20 January, Paper 27th)}
-%\title{What Do We Mean By Cybersecurity?}  Tom's original was this
-%\titlenote{}
-%\subtitle{Extended Abstract}
-%\subtitlenote{}
+\title{Cybersecurity Is Too Important To Be Left To The Specialists}
+%\title{What Do We Mean By Cybersecurity Education?}  Tom's original
 \author{James H. Davenport}
@@
 \begin{abstract}
-There are many facets to CyberSecurity education: we address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
-(Cyber)Security (if thought about at all) has hisorically been left to specialists, sometimes called ``Information Assurance'', which worked, even if not well, in a waterfall model.
-The Agile and DevOps trends have changed devleopment, but not themselves the view of security.
-Over the last few years, there has been a recogition of the need for change, exemplified by ``DevSecOps''.
-But if security is not to be left to the experts, then the generalist must know about it.
+There are numerous facets to cybersecurity education, from theory to practice, hardware and software, social and technical, to various other dimensions.
+Whilst various model curricula and associated recommendations exist, in this paper we address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters level discipline or left to professional practice, sometimes called ``information assurance'', which worked, even if not well, in a waterfall model.
+We have seen Agile and DevOps changing software and systems devleopment, but not necessarily the view of security; however, over recent years, there has been a recogition of the need for change, exemplified by the emergence of ``DevSecOps''.
+If security is not to be left to the experts, then the generalist must know about it.
+In the context of significant international computer science curriculum reform -- both in compulsory, as well as post-compulsory education -- what does this trend more generally for computer science and informatics education, and how we teach it?
 
-We consider particularly the UK perspective, where the British Computer Society accredits computing degrees, and (since \IntroYear) insists on a certain CyberSecurity element in all such.
+In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
+Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, with potential adoption and adaption in other jurisdictions.
+
+% in which a professional body\/learned society accredits university degree programmes, and (since \IntroYear) insists on a certain CyberSecurity element in all such.
 \end{abstract}

### 2019-01-16 ITiCSE2019.tex
Text around Java; conclusions

--- 2019-01-16 ITiCSE2019.tex
@@
 
-\subsection{Informal Resources}
+\subsection{Informal Resources}\label{sec:informal}
 The web abounds with informal resources: tutorials and code snippets.
@@
 This has been looked at by \cite{Unruhetal2017a}, who took the top five results from Google for six queries.
-Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in WASP's Top Ten \cite{OWASP2017a}.} weaknesses.
-Searching for these fragments in PHP projects GitHub found 82 instances of these fragments, of which 117 were verified by manual to be vulnerable --- 80\% of which were vulnerable to SQL injection.
+Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
+Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified by manual to be vulnerable --- 80\% of which were vulnerable to SQL injection.
 
-\cite{Mengetal2018a,Chenetal2019a,Zorz2019a}
+\subsection{The case of Java}\label{sec:Java}
+To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for Java textbooks.
+Indeed, many such books go nowhere near security applications.
+But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resuorces.
+\cite{Mengetal2018a} analysed 503 CyberSecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.
+53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%).
+The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.
+By default, Spring implcitly enables protection against this.
+But all the accepted answers to CSRF-related failures simply suggested disabling they check.
+There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
+solved the issue!!! I have no idea why it was enabled by default''.
+
+\cite{Chenetal2019a,Zorz2019a}
 
@@
 
+\section{Conclusions}
+
+We have the following recommendations.
+\begin{enumerate}
+\item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
+\item All CS courses houdl emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
+\end{enumerate}
+
 \section*{Acknowledgements}

### 2019-01-16 ITiCSE2019.tex
Removed pagebreak
Added recomendations to abstract
Further tweaks to abstract

@@
 \def\IntroYear{2014}  % when BCS started asking for CyberSecurity
+
 \begin{abstract}
-There are numerous facets to cybersecurity education, from theory to practice, hardware and software, social and technical, to various other dimensions.
-Whilst various model curricula and associated recommendations exist, in this paper we address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters level discipline or left to professional practice, sometimes called ``information assurance'', which worked, even if not well, in a waterfall model.
-We have seen Agile and DevOps changing software and systems devleopment, but not necessarily the view of security; however, over recent years, there has been a recogition of the need for change, exemplified by the emergence of ``DevSecOps''.
+There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
+Whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
+
+In the world of software and systems development, we have seen the disruption of Agile and DevOps, but not necessarily through the lens of security; however, more recently there has been a recogition of the need for change, exemplified by the emergence of ``DevSecOps''.
 If security is not to be left to the experts, then the generalist must know about it.
-In the context of significant international computer science curriculum reform -- both in compulsory, as well as post-compulsory education -- what does this trend more generally for computer science and informatics education, and how we teach it?
+Thus, in the context of widespread international computer science curriculum reform -- both in compulsory, as well as post-compulsory education -- what does this trend mean more generally for institutions and educators, and how do we teach it?
 
 In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
-Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, with potential adoption and adaption in other jurisdictions.
+Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
+
+We also make some
 
@@
 \end{enumerate}
-\pagebreak
+%\pagebreak - to fix URL issues
 \section{Policy Context}

### 2019-01-16 ITiCSE2019.tex
Typo in abstract
Updated authors in alpha format, added 250 word abstract

@@
 %\title{What Do We Mean By Cybersecurity Education?}  Tom's original
-\author{James H. Davenport}
-\orcid{0000-0002-3982-7545}
-\affiliation{%
-  \institution{University of Bath}
-  \streetaddress{}
-  \city{Bath} 
-  \country{United Kingdom}
-}
-\email{j.h.davenport@bath.ac.uk}
 
@@
 
-\author{Tom Prickett}
-\orcid{}
+\author{James H. Davenport}
+\orcid{0000-0002-3982-7545}
 \affiliation{%
-  \institution{Northumbria University}
+  \institution{University of Bath}
   \streetaddress{}
-  \city{Newcastle upon Tyne} 
+  \city{Bath} 
   \country{United Kingdom}
 }
-\email{tom.prickett@northumbria.ac.uk}
+\email{j.h.davenport@bath.ac.uk}
 
@@
 
+\author{Tom Prickett}
+\orcid{}
+\affiliation{%
+  \institution{Northumbria University}
+  \streetaddress{}
+  \city{Newcastle upon Tyne} 
+  \country{United Kingdom}
+}
+\email{tom.prickett@northumbria.ac.uk}
+
 % The default list of authors is too long for headers}
-%\renewcommand{\shortauthors}{Davenport et al.}
+%\renewcommand{\shortauthors}{Crick et al.}
 
@@
 Whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
 
-In the world of software and systems development, we have seen the disruption of Agile and DevOps, but not necessarily through the lens of security; however, more recently there has been a recogition of the need for change, exemplified by the emergence of ``DevSecOps''.
+In the world of software and systems development, we have seen the disruption of Agile and DevOps, but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
 If security is not to be left to the experts, then the generalist must know about it.
@@
 Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
+%292 words
 
-We also make some
+% submitted abstract (hard 250 word limit):
+% There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical; whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
+But if security is not to be left to the experts, then the generalist must know about it.
+Thus, in the context of widespread international computer science curriculum reform, what does this trend mean more generally for institutions and educators, and can do we teach it?
 
-% in which a professional body\/learned society accredits university degree programmes, and (since \IntroYear) insists on a certain CyberSecurity element in all such.
+% In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
+Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
 \end{abstract}

### 2019-01-17 ITiCSE2019.tex
More Java thinking
Added dates to Agile and DevOps

@@
 
-In the world of software and systems development, we have seen the disruption of Agile and DevOps, but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
+	In the world of software and systems development, we have seen the disruption of Agile (invented 2001) and DevOps (invented 2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
 If security is not to be left to the experts, then the generalist must know about it.
@@
 There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
-solved the issue!!! I have no idea why it was enabled by default''.
+solved the issue!!! I have no idea why it was enabled by default''. As of writing (16 January 2018) there were no negative comments about this disabling of a vital security feature.
 
-\cite{Chenetal2019a,Zorz2019a}
+\cite{Chenetal2019a,Zorz2019a,Fischeretal2017a}
 

### 2019-01-21 ITiCSE2019.tex
Add files via upload

@@
 
-{\bf Alastair's experiences.}
+In the Authors experience, it is proving very difficult to recruit academic staff with specialisms in cybersecurity.
+The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
+As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.
+A recent example had a professor of Cybersecurity, two senior academic positions and 2 junior academic positions in one advert.
+There are examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year.
+There are also examples of  cybersecurity subject groups moving en masse from one university to another.
+
+Not sure if we want to raise something about funding of posts ??
+
+Poaching of staff ??
+
+Is this too contentious
 
@@
 \subsection{Accreditation}
-{\bf{Alastair could you check this and adjust if appropriate?}}
 
@@
 
-This ambition of this approach is to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area.)  The approach taken is not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
+Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis.
+In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
+In 2013, an initiative was set up by (ISC)2, CPHC and the Cabinet Office to examine embedding Cybersecurity into undergraduate degrees in the UK.
+Three workshops in 2013 and 2014 attempted to define the principles of Cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula.
+Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate Computing Science degrees.
+The BCS agreed to adopt the outputs into their accreditation criteria.
+This was the first time that Cybersecurity has been explicitly referenced within accreditation criteria for computing and IT-related degrees.
+The fact that cybersecurity is a is included as a  component of the BCS accreditation criteria, reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards Chartered status.
+
+The final workshop resulted in the production of reference guidelines ("Cybersecurity Principles and Learning Outcomes") \cite{CPHCISC2} that established a baseline of common knowledge, example learning outcome domains for Cybersecurity within the Computing Science courses and guidance on embedding the concepts.
+The document provides specific guidance for embedding and enhancing relevant Cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.
+The document suggested 5 areas of coverage
+
+\begin{itemize}
+    \item Information and risk 
+    \item Threats and attacks 
+    \item Cybersecurity architecture and operations
+    \item Secure systems and products; and 
+    \item Cybersecurity management
+\end{itemize}
+
+The ambition of this approach is to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area.)  The approach taken is not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
 In this context this is realized as an expectation Cybersecurity is an inclusion in all degrees accredited by the BCS.
@@
 
-Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis.
-Building on the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
-This development was a partnership between the BCS,  The Council of Heads and Professors of Computing, (ISC)2 (the largest not-for-profit membership body of certified information and software security professionals worldwide, with over 130,000 members) and the Office for Cyber Security and Information Assurance in the Cabinet Office.
-This project lead to the agreement of a set of Cybersecurity Principles and Learning Outcomes for Computer Science and related IT Degrees in the UK \cite{CPHCISC2}.
-These guidelines are closely tied to  the ACM recommendations for curricula \cite{ACM2013a}.
-
-The agreed Cybersecurity Principles and Learning Outcomes were included in the BCS Guidelines for accreditation in 2015.
-This resulted in an extended expectation of the coverage of Cybersecurity in programmes accredited by the Institute.
-The expectation is now that all students upon all accredited programmes will graduate with skills, knowledge and experience related to:
-\begin{enumerate}
-\item Information and risk
-\item Threats and attacks
-\item Cybersecurity architecture and operations
-\item Secure systems and products
-\item Cybersecurity management
-\end{enumerate}
+The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages HEIs to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
+In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals - either as a stand alone module or as a significant component(s) of other modules.
+This approach differs from the ACM approach where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
 
 \subsection{What does BCS tell Universities}
-The agreed Cybersecurity Principles and Learning Outcome were discussed with the wider education community by a road-show lead by CPHC. {\bf{(Alastair could you add something here?)}}
+The agreed Cybersecurity Principles and Learning Outcome \cite{CPHCISC2} were discussed with the wider education community by a road-show lead by CPHC.
+A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of Computing Science degrees.
+The workshops included case studies from universities who had embedded cybersecurity Computer Science curriculum illustrating different approaches to implementation.
+The workshops had 102 attendees from the academic Computing Science community representing 60 UK Universities.
 
@@
 At the time of the next visit in this accreditation cycle, accreditation is conditional upon a HEI having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
-It is anticipated that from 2020(?) the expectation will be all accredited programmes have the new expectations fully embedded.
+It is anticipated that from 2020 the expectation will be all accredited programmes have the new expectations fully embedded.
 

### 2019-01-21 ITiCSE2019.tex
Comments on Parrishetal

@@
 There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.
+We note that part of its third aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these (see section \ref{sec:SQL}).
 
@@
 
-\subsection{Educational Resources: SQL Injection}
+\subsection{Educational Resources: SQL Injection}\label{sec:SQL}
 It is 15 years since \cite{Guimaraesetal2004} 

### 2019-01-26 ITiCSE2019.tex
Added references to CyBOK

@@
 What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
-{\bf JHD to research ACM general curriculum; also what is status of CYBOOK etc.}
+{\bf JHD to research ACM general curriculum;}
+
+The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way frm being a textbook (which is not their aim).
 
@@
 Nevertheless SQL injection is still a major weakness: number one in the OWASP Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.
+\cite[the UK's definitive reference]{Bristol2018a} states ``a wide range of attack techniques for exploiting SQL
+injection or script injection are known and documented.''.
 

### 2019-01-27 ITiCSE2019.tex
JHD edits

@@
 %%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
-\documentclass[sigconf]{acmart}
+%\documentclass[sigconf]{acmart}
+\documentclass[sigconf,anonymous]{acmart}
 \usepackage{paralist}
@@
 change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years - to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
-\cite{Hannigan2019a}\footnote{Former Director of GCHQ: Brittain equivalent of NSA.}
+\cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}
 \end{quote}
@@
 \subsection{Research Questions}
+There are various levels of specialism at which CyberSecurity education can be addressed.
 \begin{enumerate}
-\item The person in the street --- probably not.
-({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
+\item The person in the street --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
+% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
 Minor point but maybe worth mentioning e.g.
 being safe online, accessing public services, etc}})
-\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design.
+%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design.
 BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
@@
 \end{enumerate}
+The focus of this paper is on 2--3: the general CS graduate.
 %\pagebreak - to fix URL issues

### 2019-01-27 ITiCSE2019.tex
Various JHD

@@
 \usepackage{paralist}
-\usepackage{url}
-\usepackage[hyphenbreaks]{breakurl}
+%\usepackage{url}
+%\usepackage{enumitem}
+%\usepackage[hyphenbreaks]{breakurl}
 
@@
 
-This is not a brand-new concern: see \cite{Parr2014a}
+This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
+On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
+It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implcitly by causing boards to invest more in cybersecurity.
+On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
 
@@
 The focus of this paper is on 2--3: the general CS graduate.
+\begin{description}
+\item[RQ1]What should be taught to the generalist?
+\item[RQ2]Should this be taught stand-alone or integrated?
+\item[RQ3]How might accreditation regard CyberSecurity
+\end{description}
 %\pagebreak - to fix URL issues
@@
 
-Substantial CS/digital curriculum reform across the
+There has been substantial CS/digital curriculum reform across the
 UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
--- but what about focus on cybersecurity?
+--- but what has happened to the focus on cybersecurity?
+The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention CyberSecurity, but merely as a sub-item in one work package (1.2a).
 
@@
 Good practice was identified at one university by the commendation:
-
-"The second year project provides an opportunity for exploring security aspects in depth with an industrial use case"
-
-In summation, this shows that many UK HEIs have now embedded Cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
+ \begin{quote}
+``The second year project provides an opportunity for exploring security aspects in depth with an industrial use case''.
+\end{quote}
+In sum, this shows that many UK HEIs have now embedded Cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
 Clearly not all HEIs in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of Cybersecurity to an agreed standard.
@@
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item All CS courses houdl emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
+\item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
 \end{enumerate}

### 2019-01-27 ITiCSE2019.tex
Tried to fix underscores in bib
Edits, move Accreditation into a separate section

@@
 \begin{abstract}
-There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
+There are numerous facets to Cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
 Whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
@@
 
-In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
-Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
+In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on Cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
+Through this analysis of policy and practice in the UK, we make a number of recommendations for future Cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
 %292 words
@@
 \subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} CyberSecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
+In one sense, this title is a false dichotomy: there is a serious need for Cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} Cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
 
 This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
-On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
-It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implcitly by causing boards to invest more in cybersecurity.
+On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of Cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
+It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in Cybersecurity.
 On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
@@
 Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}.
-However, given the shortage of specialists, in practice a good generalist with some CyberSecurity expertise can often get a ``specialist'' job.
+However, given the shortage of specialists, in practice a good generalist with some Cybersecurity expertise can often get a ``specialist'' job.
 \subsection{Research Questions}
-There are various levels of specialism at which CyberSecurity education can be addressed.
+There are various levels of specialism at which Cybersecurity education can be addressed.
 \begin{enumerate}
@@
 \begin{description}
-\item[RQ1]What should be taught to the generalist?
+\item[RQ1]What should be taught to the generalist, and how?
 \item[RQ2]Should this be taught stand-alone or integrated? 
-\item[RQ3]How might accreditation regard CyberSecurity
+\item[RQ3]How might accreditation regard Cybersecurity Education, or help with it?
 \end{description}
+\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
+\begin{quote}
+The Information Assurance and Security KA is unique among the set of KAs presented here
+given the manner in which the topics are pervasive throughout other Knowledge Areas.
+\end{quote}
+It proposes 9 ``core'' hours and 63.5 distributed across the other Knowledge Areas.
 %\pagebreak - to fix URL issues
@@
 UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
---- but what has happened to the focus on cybersecurity?
-The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention CyberSecurity, but merely as a sub-item in one work package (1.2a).
+--- but what has happened to the focus on Cybersecurity?
+The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention Cybersecurity, but merely as a sub-item in one work package (1.2a).
 
@@
 There is large-scale media
-attention on the ``cybersecurity skills gap''
+attention on the ``Cybersecurity skills gap''
 e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
@@
 
-Is teaching CyberSecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways)
+Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways)
 
-AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.
-{\bf AI to write up}.
+%AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
   Should we be teaching it because it's underpinning? because there's a skill shortage?  JHD would argue both.
@@
 
-Alastair to write about their Credit Fraud experience (MSc CyberSecurity).
+Alastair to write about their Credit Fraud experience (MSc Cybersecurity).
 HE Academy grant (to be referenced).
@@
 Clearly such a major weakness should be well-taught.
-One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cyber-security course.
+One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist Cybersecurity course.
 In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.
@@
 departments in the United States (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed).
-There were seven such boks, but three accounted for the 36 of the 44.
+There were seven such books, but three accounted for the 36 of the 44.
 Five of the seven (30 of the 44) had no mention of SQL injection.
@@
 But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resuorces.
-\cite{Mengetal2018a} analysed 503 CyberSecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.
+\cite{Mengetal2018a} analysed 503 Cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.
 53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%).
@@
 \subsection{Staff}
-It is well known that CyberSecurity skills are in short supply, for example \cite{Page2018a}.
+It is well known that Cybersecurity skills are in short supply, for example \cite{Page2018a}.
 \begin{quote}
@@
 
-In the Authors experience, it is proving very difficult to recruit academic staff with specialisms in cybersecurity.
-The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
-As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.
+In the authors' experience, it is proving very difficult to recruit academic staff with specialisms in Cybersecurity.
+The demand for Cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
+As universities expand their Cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.
 A recent example had a professor of Cybersecurity, two senior academic positions and 2 junior academic positions in one advert.
-There are examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year.
-There are also examples of  cybersecurity subject groups moving en masse from one university to another.
+There are examples in the UK of Cybersecurity lecturing jobs remaining unfilled for longer than a year.
+There are also examples of  Cybersecurity subject groups moving en masse from one university to another.
 
-Not sure if we want to raise something about funding of posts ??
+%Not sure if we want to raise something about funding of posts ??
+%Poaching of staff ??
+%Is this too contentious
 
-Poaching of staff ??
+%{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
+\section{Accreditation: RQ3}
 
-Is this too contentious
-
-{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
-\subsection{Accreditation}
-
-For a few years now there has been a recognized need to build knowledge, skills and capacity in the area of Cybersecurity.
+For a few years now there has been a recognized need in the UK to build knowledge, skills and capacity in the area of Cybersecurity.
 This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
@@
 The consensus now is that Cybersecurity is both a discipline in its own right and that Cybersecurity should be taught within Computer Science and related degrees.
-There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite{ACM2013a} and specialised Cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised Cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
 
 In the United Kingdom, Higher Education provision addresses both approaches.
-A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and Cybersecurity ( and closely related fields Computer Security, Digital / Computer Forensics, etc).
+A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and Cybersecurity (and closely related fields Computer Security, Digital / Computer Forensics, etc).
 In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to Cybersecurity for entry in September 2019.
@@
 The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of Cybersecurity discipline areas.
-The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord and Seoul Accord (do we need to reference these).
-These memorandums support the internationalising of the curriculum and promote consistency and parity in Computer Science Education globally.
+The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}.
+These memoranda support the internationalising of the curriculum and promote consistency and parity in Computer Science Education globally.
 These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
@@
 In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
-In 2013, an initiative was set up by (ISC)2, CPHC and the Cabinet Office to examine embedding Cybersecurity into undergraduate degrees in the UK.
+In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of Computer Science Departments) and the Cabinet Office to examine embedding Cybersecurity into undergraduate degrees in the UK.
 Three workshops in 2013 and 2014 attempted to define the principles of Cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula.
@@
 This was the first time that Cybersecurity has been explicitly referenced within accreditation criteria for computing and IT-related degrees.
-The fact that cybersecurity is a is included as a  component of the BCS accreditation criteria, reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards Chartered status.
+The fact that Cybersecurity is included as a  component of the BCS accreditation criteria, reflects the importance placed on Cybersecurity and the expectation that all computing graduates should have knowledge and skills in Cybersecurity as they move towards Chartered status.
 
@@
 
-The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages HEIs to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
-In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals - either as a stand alone module or as a significant component(s) of other modules.
+The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages HEIs to embed Cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
+In addition there an expectation that there is significant coverage of Cybersecurity principles and fundamentals - either as a stand alone module or as a significant component(s) of other modules.
 This approach differs from the ACM approach where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
@@
 \subsection{What does BCS tell Universities}
-The agreed Cybersecurity Principles and Learning Outcome \cite{CPHCISC2} were discussed with the wider education community by a road-show lead by CPHC.
-A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of Computing Science degrees.
-The workshops included case studies from universities who had embedded cybersecurity Computer Science curriculum illustrating different approaches to implementation.
+The agreed Cybersecurity Principles and Learning Outcome \cite{CPHCISC2} were discussed with the wider education community by a road-show led by CPHC.
+A series of workshops took place in 2015 which presented the rationale for embedding Cybersecurity in the curriculum of Computing Science degrees.
+The workshops included case studies from universities who had embedded Cybersecurity Computer Science curriculum illustrating different approaches to implementation.
 The workshops had 102 attendees from the academic Computing Science community representing 60 UK Universities.
@@
 
-1)	Of the 11 UG programmes involved all were approved 'At Threshold'
+A)	Of the 11 UG programmes involved all were approved 'At Threshold'
 
-2)	Of the 9 UG programmes involved, 8 were approved and 1 refused
+B)	Of the 9 UG programmes involved, 8 were approved and 1 refused
 
-3)	Of the 5 UG programmes involved, all approved 'At Threshold'
+C)	Of the 5 UG programmes involved, all approved 'At Threshold'
 
-4)	Of the 3 UG programmes involved, all 3 were refused.
+D)	Of the 3 UG programmes involved, all 3 were refused.
 
@@
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
+\item Accreditation is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.
 \end{enumerate}

### 2019-01-27 ITiCSE2019.tex
More restructuring

@@
 \usepackage{paralist}
-%\usepackage{url}
+\usepackage{url}
 %\usepackage{enumitem}
-%\usepackage[hyphenbreaks]{breakurl}
+\usepackage[hyphenbreaks]{breakurl}
+\usepackage[T1]{fontenc}
+\usepackage{hyperref}
 
@@
 Clearly Cybersecurity has been and continues to be part of these expectations.
-
+\subsection{Developing Expectations}
 Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis.

### 2019-01-27 ITiCSE2019.tex
Txt around vulnerabilities

@@
 Indeed, many such books go nowhere near security applications.
-But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resuorces.
+But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources.
 \cite{Mengetal2018a} analysed 503 Cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.
@@
 solved the issue!!! I have no idea why it was enabled by default''. As of writing (16 January 2018) there were no negative comments about this disabling of a vital security feature.
-
-\cite{Chenetal2019a,Zorz2019a,Fischeretal2017a}
+\par
+\cite{Chenetal2019a,Zorz2019a} 
+\subsection{Android}
+\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
+\begin{quote}
+We found that 15.4\% of all 1.3 million Android applications
+contained security-related code snippets from
+Stack Overflow. Out of these 97.9\% contain at least one
+insecure code snippet.
+\end{quote}
+We should note two caveats (in opposite directions).
+Their labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure.
+On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
 
@@
 \end{quote}
-\begin{quote}
-
-\end{quote}
-JHD possibly to write about his recruitment experiences.  ? more IoC input?
+The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
+%JHD possibly to write about his recruitment experiences.  ? more IoC input?
 

### 2019-01-28 ITiCSE2019.tex
Improved Java, and conclusion 2

@@
 \par
-\cite{Chenetal2019a,Zorz2019a} 
+This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was
+\begin{quote}
+644 out of the 1,429 inspected answer posts
+(45\%) are insecure, meaning that insecure suggestions
+popularly exist on SO. Insecure answers dominate, in
+particular, the SSL/TLS category [355 insecure versus 150 secure]].
+\end{quote}
+
 \subsection{Android}
@@
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
+One should probably use the data frm \cite{Chenetal2019a}: ``If ou pikc up a SSL/TLS andwer from Stack Overflow, there's a 70\% chance it's insecure''.
 \item Accreditation is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.

### 2019-01-28 ITiCSE2019.tex
JHD changes on RQs and PCI DSS

@@
 
-\subsection{PCI DSS}
-\cite{PCI2018b}.
+\subsection{PCI DSS}\label{sec:PCIDSS}
+The Payments Card Industry Data Security Standards \cite{PCI2018b} underpin all processing of credit/debit cards.
+Nevertheless, they are very rarely mentioned in generalist computer scientist courses --- one payments industry person commented ``I've given up even asking if recruits have heard of PCI DSS: it's so rare''.
 
-Alastair to write about their Credit Fraud experience (MSc Cybersecurity).
-HE Academy grant (to be referenced).
+This would not matter so much if everyone handling payments data wer sent by their employers on an effective\footnote{``Effective'' is important: the Processing Director at one major acquirer commented ``I just caught two developers in the elevator discussing how to meet a customer requirement, and I had to tell them this was contrary to PCI DSS.
+Their response was that the customer wanted it!''.} PCI DSS course.
+However, the payments business is now so spread across websites, often run by SMEs, or non-specialists, that some provision should be made in courses.
+Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.
+
+%Alastair to write about their Credit Fraud experience (MSc Cybersecurity).  HE Academy grant (to be referenced).
 
@@
 \section{Conclusions}
+As regards our research questions, we can make the following comments.
+\begin{description}
+\item[RQ1]
+\item[RQ2]The recommendation in \cite[p.
+98]{ACM2013a} that Cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea.
+Indeed, a recnt discussion on Cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.
+However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.
+It is mor eimportant that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than tthat they not be taught at all.
+\item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.
+\end{description}
 
@@
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
-One should probably use the data frm \cite{Chenetal2019a}: ``If ou pikc up a SSL/TLS andwer from Stack Overflow, there's a 70\% chance it's insecure''.
-\item Accreditation is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.
+One should probably use the data frm \cite{Chenetal2019a}: ``If you pick up a SSL/TLS answer from Stack Overflow, there's a 70\% chance it's insecure''.
 \end{enumerate}

### 2019-01-28 ITiCSE2019.tex
Conclusions

@@
 
-\subsection{Android}
+\subsection{Android}\label{sec:Android}
 \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
@@
 \begin{description}
-\item[RQ1]
+\item[RQ1]The guidelines from both ACM and BCS are good for general education.
+However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}--) resources.
 \item[RQ2]The recommendation in \cite[p.
@@
 However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.
-It is mor eimportant that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than tthat they not be taught at all.
+It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than tthat they not be taught at all.
+\par
+We comend the BCS-identified good practice of exploring Cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
 \item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.

### 2019-01-28 ITiCSE2019.tex
Bug fixes
Added FIE abstract

--- 2019-01-28 ITiCSE2019.tex
@@
 \begin{abstract}
-There are numerous facets to Cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
-Whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
+% submitted abstract (100-500 word limit):
+There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
+Whilst various national and international model curricula and recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline; or left for development in professional practice, sometimes called “information assurance” in the UK and Europe, which worked, even if not well, in a sequential process model.
 
-	In the world of software and systems development, we have seen the disruption of Agile (invented 2001) and DevOps (invented 2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
+In the world of software and systems development, we have seen the disruption of Agile (invented c.2001) and DevOps (invented c.2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
 If security is not to be left to the experts, then the generalist must know about it.
-Thus, in the context of widespread international computer science curriculum reform -- both in compulsory, as well as post-compulsory education -- what does this trend mean more generally for institutions and educators, and how do we teach it?
+Thus, in the context of widespread international computer science/engineering curriculum reform – both in compulsory, as well as post-compulsory education – what does this trend mean more generally for institutions and educators, and how do we teach it?
 
-In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on Cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
-Through this analysis of policy and practice in the UK, we make a number of recommendations for future Cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
-%292 words
-
-% submitted abstract (hard 250 word limit):
-% There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical; whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
-But if security is not to be left to the experts, then the generalist must know about it.
-Thus, in the context of widespread international computer science curriculum reform, what does this trend mean more generally for institutions and educators, and can do we teach it?
-
-% In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
+In this paper, we frame a key social, cultural and economic global challenge in computing and engineering education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
 Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.

### 2019-01-30 ITiCSE2019.tex

@@
 However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.
-It is mor eimportant that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than tthat they not be taught at all.
+It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than tthat they not be taught at all.
+\par
+We commend the BCS-identified good practice of exploring Cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
 \item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.

### 2019-02-02 ITiCSE2019.tex
Added policy link and updated submitted FIE abstract

@@
 There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
-Whilst various national and international model curricula and recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline; or left for development in professional practice, sometimes called “information assurance” in the UK and Europe, which worked, even if not well, in a sequential process model.
+A multitude of national and international model curricula and recommendations - from national academies, learned societies and even governments -- have been presented and discussed in recent years, with varying levels of impact on policy and practice.
+In this paper we attempt to address the key questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline; or left for development in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a sequential process model.
 
-In the world of software and systems development, we have seen the disruption of Agile (invented c.2001) and DevOps (invented c.2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
+For example, in the world of software and systems development, we have seen the disruption of Agile (invented c.2001) and DevOps (invented c.2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
 If security is not to be left to the experts, then the generalist must know about it.
-Thus, in the context of widespread international computer science/engineering curriculum reform – both in compulsory, as well as post-compulsory education – what does this trend mean more generally for institutions and educators, and how do we teach it?
+Thus, in the context of widespread international computer science/engineering curriculum reform - both in compulsory, as well as post-compulsory education - what does this trend mean more generally for institutions and educators, and how do we teach it?
 
-In this paper, we frame a key social, cultural and economic global challenge in computing and engineering education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
-Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
+In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by analysing the UK's national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities (and into industry and professional practice).
+Through this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
 \end{abstract}
@@
 %\pagebreak - to fix URL issues
+
 \section{Policy Context}
+
+% TC to add policy+skills context
+% inc. new EU Cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
 

### 2019-02-05 ITiCSE2019.tex
Added Ackerman2019a and quote

--- 2019-01-30 ITiCSE2019.tex
@@
 \begin{abstract}
-% submitted abstract (100-500 word limit):
-There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
-Whilst various national and international model curricula and recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline; or left for development in professional practice, sometimes called “information assurance” in the UK and Europe, which worked, even if not well, in a sequential process model.
+There are numerous facets to Cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
+Whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
 
-In the world of software and systems development, we have seen the disruption of Agile (invented c.2001) and DevOps (invented c.2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
+	In the world of software and systems development, we have seen the disruption of Agile (invented 2001) and DevOps (invented 2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
 If security is not to be left to the experts, then the generalist must know about it.
-Thus, in the context of widespread international computer science/engineering curriculum reform – both in compulsory, as well as post-compulsory education – what does this trend mean more generally for institutions and educators, and how do we teach it?
+Thus, in the context of widespread international computer science curriculum reform -- both in compulsory, as well as post-compulsory education -- what does this trend mean more generally for institutions and educators, and how do we teach it?
 
-In this paper, we frame a key social, cultural and economic global challenge in computing and engineering education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
+In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on Cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
+Through this analysis of policy and practice in the UK, we make a number of recommendations for future Cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
+%292 words
+
+% submitted abstract (hard 250 word limit):
+% There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical; whilst various model curricula and associated recommendations have been presented and discussed in recent years, in this paper we attempt to address the questions ``what should the generalist computer scientist know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline or left to develop in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a waterfall model.
+But if security is not to be left to the experts, then the generalist must know about it.
+Thus, in the context of widespread international computer science curriculum reform, what does this trend mean more generally for institutions and educators, and can do we teach it?
+
+% In this paper, we frame a key social and global challenge in computing education by analysing the UK perspective on cybersecurity education, with a substantial national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities.
 Through this analysis of policy and practice in the UK, we make a number of recommendations for future cybersecurity educational initiatives, including practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions.
@@
 %\pagebreak - to fix URL issues
+\par
+Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, writes as follows
+\begin{quote}
+Universities suffer shortcomings, as well.
+Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity.
+There is a big catch, however.
+Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.
+\end{quote}
 \section{Policy Context}

### 2019-02-05 ITiCSE2019.tex

--- 2019-02-02 ITiCSE2019.tex
@@
 %\pagebreak - to fix URL issues
-
+\par
+Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, writes as follows
+\begin{quote}
+Universities suffer shortcomings, as well.
+Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity.
+There is a big catch, however.
+Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.
+\end{quote}
 \section{Policy Context}

### 2019-02-13 ITiCSE2019.tex
Added Votipkaetal2019a and use of it.
Text for Ruiz2019a

@@
 \end{quote}
+
+But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: he quotes that 61\% of UK courses offer mandatory cybersecurity content, and his research was based on web scraping, it represents a lower bound.
 \section{Policy Context}
@@
 On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
-
+\par
+The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}.
+Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''.
 \subsection{Staff}

### 2019-02-20 ITiCSE2019.tex
Added Timeline

@@
 Clearly Cybersecurity has been and continues to be part of these expectations.
+
+In recent years the BCS has evolved its Accreditation Practices to promote and mandate the inclusion of Cybersecurity within the programmes the body accredits.
+The timeline which this process has followed in Table \ref{table:1} .
+\begin{center}
+  \begin{table}[h!]
+  \begin{tabular}{ | p{6cm} |p{1.5cm} |}
+    \hline
+    Developing Expectations &   \\ \hline
+    UK Government Cybersecurity Strategy \cite{UKCabinetOffice} & November 2011 \\ \hline
+    Three workshops of a consortium of industry, academia and government bodies - led by CPHC and (ISC)2 - leading to the development of Cybersecurity learning guidelines to be embedded into BCS accredited UK Computer Science and IT-related degree \cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
+    UK Government report Cybersecurity Skills, Business Perspectives and Government's Next Steps Report Released \cite{UKCabinetOffice2014} & March 2014  \\ \hline
+    Council of Professors and Heads of Computing (CPHC) Identifies Cybersecurity as one the top 3 concerns in Computing & April 2014 \\ \hline
+    Joint Development of White Paper from CPHC and The International Information Systems Security Certification Consortium(ISC)2 \cite{CPHCISC2014} & April -November 2014 \\ \hline
+    Extended Cybersecurity Criteria included in BCS Accreditation Guidelines \cite{BCS2018a}& June 2015 \\
+    \hline
+    What does the BCS tell Universities? & \\ \hline
+    Cybersecurity Principles Roadshow & March-April 2016 \\ \hline
+    Accreditation - what progress has been made? &  \\ \hline
+    All visited HEIs expected to be fully compliant & September 2020 (BCS follow a 5-year Accreditation Cycle)  \\ \hline
+  \end{tabular}
+  \caption{Timeline of the development of Cybersecurity expectations}
+  \label{table:1}
+  \end{table}
+\end{center}
 \subsection{Developing Expectations}
 Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis.
+Like many Governments, the UK Goverment has been actively been seeking ways to address this \cite{UKCabinetOffice}, \cite{UKCabinetOffice2014}.
 In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
 In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of Computer Science Departments) and the Cabinet Office to examine embedding Cybersecurity into undergraduate degrees in the UK.
-Three workshops in 2013 and 2014 attempted to define the principles of Cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula.
+Three workshops in 2013 and 2015 attempted to define the principles of Cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula.
 Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate Computing Science degrees.
+This work intially lead to a white paper related to a proposal in the form of a whitepaper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}.
 The BCS agreed to adopt the outputs into their accreditation criteria.
-This was the first time that Cybersecurity has been explicitly referenced within accreditation criteria for computing and IT-related degrees.
+This was the first time that Cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees.
 The fact that Cybersecurity is included as a  component of the BCS accreditation criteria, reflects the importance placed on Cybersecurity and the expectation that all computing graduates should have knowledge and skills in Cybersecurity as they move towards Chartered status.
 
-The final workshop resulted in the production of reference guidelines ("Cybersecurity Principles and Learning Outcomes") \cite{CPHCISC2} that established a baseline of common knowledge, example learning outcome domains for Cybersecurity within the Computing Science courses and guidance on embedding the concepts.
+The produced reference guidelines ("Cybersecurity Principles and Learning Outcomes") \cite{CPHCISC2} established a baseline of common knowledge, example learning outcome domains for Cybersecurity within the Computing Science courses and guidance on embedding the concepts.
 The document provides specific guidance for embedding and enhancing relevant Cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.

### 2019-02-25 ITiCSE2019.tex
Added SQL/Wikipedia
JHD on Wikiedia (bug fix)
Typo etc in Recommendations

@@
 shouldn't do.'' \cite{Drop2019}}.
-
+\par
+This blindness is not limited to textbooks: although Wikipedia has an excellent article on SQL Injection, it was not linked from the SQL page itself.\footnote{The authors intend to fix this after the paper is processed, so as not to break anonymity.}
 \subsection{Informal Resources}\label{sec:informal}
@@
 
-We have the following recommendations.
+We have the following specific recommendations.
 \begin{enumerate}
@@
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
-One should probably use the data frm \cite{Chenetal2019a}: ``If you pick up a SSL/TLS answer from Stack Overflow, there's a 70\% chance it's insecure''.
+One should probably use the data from \cite{Chenetal2019a}: ``If you pick up a SSL/TLS answer from Stack Overflow, there's a 70\% chance it's insecure''.
 \end{enumerate}

### 2019-02-25 FIE2019.tex
Merged to IEEE style

--- 2019-02-25 ITiCSE2019.tex
@@
-%%%% Proceedings format for most of ACM conferences (with the exceptions listed below) and all ICPS volumes.
-%\documentclass[sigconf]{acmart}
-\documentclass[sigconf,anonymous]{acmart}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
 \usepackage{paralist}
 \usepackage{url}
-%\usepackage{enumitem}
+\usepackage{enumitem}
 \usepackage[hyphenbreaks]{breakurl}
 \usepackage[T1]{fontenc}
+\usepackage{booktabs} % For formal tables
+
 \usepackage{hyperref}
@@
 
-%%%% As of March 2017, [siggraph] is no longer used. Please use sigconf (above) for SIGGRAPH conferences.
+\begin{document}
 
-%%%% Proceedings format for SIGPLAN conferences 
-% \documentclass[sigplan, anonymous, review]{acmart}
-
-%%%% Proceedings format for SIGCHI conferences
-% \documentclass[sigchi, review]{acmart}
-
-\usepackage{booktabs} % For formal tables
-
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmlicensed}
-\acmConference[ITiCSE '19]{The 24th Annual Conference on Innovation
-  and Technology in Computer Science Education}{Jul. 15-17,
-  2019}{Aberdeen, UK}
-%\acmBooktitle{}
-%\acmPrice{15.00}
-%\acmDOI{10.1145/3159450.3159547}
-%\acmISBN{978-1-4503-5103-4/18/02}
-% This slight change to the code may also save 1 or 2 lines of space.
-
-% removes the headers from each page per the preparation instructions, as these are not needed and will be updated with the chairs' actual session names during the pagination/indexing process:
-\fancyhead{}
-
-\begin{document}
-\title{Cybersecurity Is Too Important To Be Left To The Specialists}
+\title{Cybersecurity Is Too Important\\To Be Left To The Specialists}
 %\title{What Do We Mean By Cybersecurity Education?}  Tom's original
 
-\author{Tom Crick}
-\orcid{0000-0001-5196-9389}
-\affiliation{%
-  \institution{Swansea University}
-  \streetaddress{}
-  \city{Swansea} 
-  \country{United Kingdom}
-}
-\email{thomas.crick@swansea.ac.uk}
+% \author{\IEEEauthorblockN{Tom Crick}
+% \IEEEauthorblockA{\textit{School of Education} \\
+% \textit{Swansea University}\\
+% Swansea, UK \\
+% thomas.crick@swansea.ac.uk}
+% \and
+% \IEEEauthorblockN{James H. Davenport}
+% \IEEEauthorblockA{\textit{Department of Computer Science} \\
+% \textit{University of Bath}\\
+% Bath, UK \\
+% j.h.davenport@bath.ac.uk}
+% \and
+% \IEEEauthorblockN{Alastair Irons}
+% \IEEEauthorblockA{\textit{Faculty of Technology} \\
+% \textit{name of organization (of Aff.)}\\
+% Sunderland University, Sunderland, UK\\
+% alastair.irons@sunderland.ac.uk}
+% \and
+% \IEEEauthorblockN{Tom Prickett}
+% \IEEEauthorblockA{\textit{School of Computing} \\
+% \textit{name of organization (of Aff.)}\\
+% Northumbria University, Newcastle upon Tyne, UK \\
+% tom.prickett@northumbria.ac.uk}
+% }
 
-\author{James H. Davenport}
-\orcid{0000-0002-3982-7545}
-\affiliation{%
-  \institution{University of Bath}
-  \streetaddress{}
-  \city{Bath} 
-  \country{United Kingdom}
-}
-\email{j.h.davenport@bath.ac.uk}
+% compact IEEE author format for more than three authors...
+\author{\IEEEauthorblockN{Tom Crick\IEEEauthorrefmark{1}, James
+    H. Davenport\IEEEauthorrefmark{2}, Alastair
+    Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
+\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Education, 
+Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, 
+University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
+\IEEEauthorblockA{\IEEEauthorrefmark{3}Faculty of Technology, 
+Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
+\IEEEauthorblockA{\IEEEauthorrefmark{4}School of Computing, 
+Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
 
-\author{Alastair Irons}
-\orcid{}
-\affiliation{%
-  \institution{Sunderland University}
-  \streetaddress{}
-  \city{Sunderland} 
-  \country{United Kingdom}
-}
-\email{alastair.irons@sunderland.ac.uk}
-
-\author{Tom Prickett}
-\orcid{}
-\affiliation{%
-  \institution{Northumbria University}
-  \streetaddress{}
-  \city{Newcastle upon Tyne} 
-  \country{United Kingdom}
-}
-\email{tom.prickett@northumbria.ac.uk}
-
-% The default list of authors is too long for headers}
-%\renewcommand{\shortauthors}{Crick et al.}
+\maketitle
 
@@
 % submitted abstract (100-500 word limit):
-There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
-A multitude of national and international model curricula and recommendations - from national academies, learned societies and even governments -- have been presented and discussed in recent years, with varying levels of impact on policy and practice.
-In this paper we attempt to address the key questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline; or left for development in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a sequential process model.
+There are numerous facets to cybersecurity education, from theory to
+practice, hardware and software, to social and technical (as well as
+other important dimensions). A multitude of national and international
+model curricula and recommendations - from national academies, learned
+societies and even governments -- have been presented and discussed in
+recent years, with varying levels of impact on policy and practice. In
+this paper we attempt to address the key questions ``what should the
+generalist computer scientist/engineer know'' and ``how well is that
+being taught''. (Cyber)security -- if thought about at all -- has
+historically been left to specialists, frequently viewed as a
+masters-level discipline; or left for development in professional
+practice, sometimes called ``information assurance'' in the UK and
+Europe, which worked, even if not well, in a sequential process model.
 
-For example, in the world of software and systems development, we have seen the disruption of Agile (invented c.2001) and DevOps (invented c.2009), but not necessarily through the lens of security; however, more recently there has been a recognition of the need for change, exemplified by the emergence of ``DevSecOps''.
-If security is not to be left to the experts, then the generalist must know about it.
-Thus, in the context of widespread international computer science/engineering curriculum reform - both in compulsory, as well as post-compulsory education - what does this trend mean more generally for institutions and educators, and how do we teach it?
+For example, in the world of software and systems development, we have
+seen the disruption of Agile (invented c.2001) and DevOps (invented
+c.2009), but not necessarily through the lens of security; however,
+more recently there has been a recognition of the need for change,
+exemplified by the emergence of ``DevSecOps''. If security is not to
+be left to the experts, then the generalist must know about it. Thus,
+in the context of widespread international computer
+science/engineering curriculum reform - both in compulsory, as well as
+post-compulsory education - what does this trend mean more generally
+for institutions and educators, and how do we teach it?
 
-In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by analysing the UK's national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities (and into industry and professional practice).
-Through this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
+In this paper, we frame a key social, cultural and economic global
+challenge of cybersecurity in computing and engineering education by
+analysing the UK's national security, economic, and skills policy
+context, shaping the practical considerations for education
+professionals from schools through to universities (and into industry
+and professional practice). Through this analysis of UK educational
+policy and practice, we make a number of recommendations for future
+cybersecurity educational initiatives and interventions, including a
+number of innovative practical suggestions for educators and
+curriculum designers, with a view for potential adoption and adaption
+in other jurisdictions internationally.
 \end{abstract}
 
-\keywords{TBC}
-
-\maketitle
+\begin{IEEEkeywords}
+TBC
+\end{IEEEkeywords}
 
@@
 \section*{Acknowledgements}
-Thanks to Sally Pearce, Academic Accreditation Manager, BCS, The Chartered Institute for IT for supplying the summary information related to Accreditation.
+Thanks to Sally Pearce, Academic Accreditation Manager, BCS--The
+Chartered Institute for IT for supplying the summary information
+related to Accreditation.
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{ITiCSE2019}
+\bibliographystyle{IEEEtran}
+\bibliography{FIE2019}
 

### 2019-02-26 FIE2019.tex
Added Sections to table

@@
 On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
+% JHD what are these?  Also the next paragraph needs thinking about
 
@@
     \hline
-    Developing Expectations &   \\ \hline
+    IV. ACCREDITATION:RQ3 - A. Developing Expectations &   \\ \hline
     UK Government Cybersecurity Strategy \cite{UKCabinetOffice} & November 2011 \\ \hline
@@
     \hline
-    What does the BCS tell Universities? & \\ \hline
+    IV. ACCREDITATION:RQ3 - B. What does the BCS tell Universities? & \\ \hline
     Cybersecurity Principles Roadshow & March-April 2016 \\ \hline
-    Accreditation - what progress has been made? &  \\ \hline
+    IV. ACCREDITATION:RQ3 - C. Accreditation - what progress has been made? &  \\ \hline
     All visited HEIs expected to be fully compliant & September 2020 (BCS follow a 5-year Accreditation Cycle)  \\ \hline

### 2019-03-10 FIE2019.tex
DROP2019 got best paper award
JHD added acceditation stuff

@@
 AI: GDPR essentially expects CyberSecurity to be in all coding.
-JHD: PCI DSS insists on a Web Application Firewall, which is only needed if you don't trust your applications.
+
+The Payments Industy \cite[6.6]{PCI2018b} insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secore coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
 
@@
 
-What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
-{\bf JHD to research ACM general curriculum;}
+%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
+%{\bf JHD to research ACM general curriculum;}
+The ACM general curriculum \cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
 
@@
 
-This would not matter so much if everyone handling payments data wer sent by their employers on an effective\footnote{``Effective'' is important: the Processing Director at one major acquirer commented ``I just caught two developers in the elevator discussing how to meet a customer requirement, and I had to tell them this was contrary to PCI DSS.
+This would not matter so much if everyone handling payments data were sent by their employers on an effective\footnote{``Effective'' is important: the Processing Director at one major acquirer commented ``I just caught two developers in the elevator discussing how to meet a customer requirement, and I had to tell them this was contrary to PCI DSS.
 Their response was that the customer wanted it!''.} PCI DSS course.
@@
 
+% Drop2019 was awarded the (unique) best paper at SIGCSE2019 - not sure what to make of this.
 Hence \cite{Drop2019} analyzed the database textbooks used by  44 of the top 50 Computer Science

### 2019-03-14 FIE2019.tex
JHD added his ToDo
Added a GDPR paragraph plus couple of typos

@@
+% JHD to
+% a) Add the Bonn work
+% b) Add ABET stuff and ask Tom P to cross-check.
 \documentclass[conference]{IEEEtran}
@@
 
-The Payments Industy \cite[6.6]{PCI2018b} insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secore coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
+In particular the expectation of ''Privacy by Design'' and and by "Privacy by Default" has significant implications for all software development with a European context that processes personal data.
+Whilst both ''Privacy by Design'' and ''Privacy by Default' have been expected good practices for a number of years they expectations now have a legal dimension.
+A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}.
+This further emphasizes the need for mainstream software developers and by extension Computer Science graduates to be versed in relevant security practices.
+
+The Payments Industry \cite[6.6]{PCI2018b} insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
 

### 2019-03-16 FIE2019.tex
JHD edits, including Naiakshima

@@
 % a) Add the Bonn work
-% b) Add ABET stuff and ask Tom P to cross-check.
+% b) Add ABET stuff (done) and ask Tom P to cross-check.
 \documentclass[conference]{IEEEtran}
@@
 The Payments Industry \cite[6.6]{PCI2018b} insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
+This would seem to be particularly the case with ``agile'' processes \cite{Bartsch2011a}.
 
@@
 But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: he quotes that 61\% of UK courses offer mandatory cybersecurity content, and his research was based on web scraping, it represents a lower bound.
+
+It is at least plausible to attribute this difference to differences in the accreditation regimes.
+\begin{description}
+\item[UK]BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accedited universities should be compliant by 2020 (because of the five-year cycle).
+\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
+ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 Juy 2019) \cite[chapter 7]{Davenportminor2019a}.
+% need a better reference
+\end{description}
 \section{Policy Context}
@@
 
-Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways)
+Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a})
 
@@
 popularly exist on SO. Insecure answers dominate, in
-particular, the SSL/TLS category [355 insecure versus 150 secure]].
+particular, the SSL/TLS category [355 insecure versus 150 secure, i.e. $>70$\%].
 \end{quote}
@@
 Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''.
+\subsection{Agile}
+To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for ``Agile'' textbooks.
+It is the authors' experience that the vast majority of students' individual software projects are developed on Agile lines.
+Many authorshave found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
+\par
+Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
+were part of a team working on creating a social networking site for
+our university and to implement a password storage mechanism for this.
+10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not.
+None of the unprimed ones implemented any security.
 \subsection{Staff}
@@
 In recent years the BCS has evolved its Accreditation Practices to promote and mandate the inclusion of Cybersecurity within the programmes the body accredits.
-The timeline which this process has followed in Table \ref{table:1} .
+The timeline which this process has followed in Table \ref{table:1}.
 \begin{center}

### 2019-03-16 FIE2019.tex
Typo

@@
 
-The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way frm being a textbook (which is not their aim).
+The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
 

### 2019-03-20 FIE2019.tex
Added comments on what to fix
small tidy

@@
 % \IEEEauthorblockN{Tom Prickett}
-% \IEEEauthorblockA{\textit{School of Computing} \\
+% \IEEEauthorblockA{\textit{Department of Computing and Information Sciences} \\
 % \textit{name of organization (of Aff.)}\\
@@
 Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}School of Computing, 
+\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Computer and Information Sciences, 
 Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
@@
 On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
-% JHD what are these?  Also the next paragraph needs thinking about
+% JHD what are these?
+Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.
 
@@
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
-ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 Juy 2019) \cite[chapter 7]{Davenportminor2019a}.
+ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019) \cite[chapter 7]{Davenportminor2019a}.
 % need a better reference
@@
 
-Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a})
+Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.
 
@@
 It is the authors' experience that the vast majority of students' individual software projects are developed on Agile lines.
-Many authorshave found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
+Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
 \par
@@
 
-The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK.
-This could potentially include Cybersecurity related skills.
+The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK \cite{Davenportetal2019a}.
+This is likely to include Cybersecurity related skills.
 Like the Tech Partnership the focus is upon job market requirements.
-Given the size of this initiative this potentially has a significant role to play however at time of writing it is not clear what that shall be.
+Additionally the Institute is looking at potentially endorsing the demonstrable capabilities of graduates as shown by their university studies, work experience and work placements.
+Given the size of this initiative this potentially has a significant role to play however at time of writing it is clearly a work in progress.
 
@@
 Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate Computing Science degrees.
-This work intially lead to a white paper related to a proposal in the form of a whitepaper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}.
+This work initially lead to a white paper related to a proposal in the form of a whitepaper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}.
 The BCS agreed to adopt the outputs into their accreditation criteria.
@@
 98]{ACM2013a} that Cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea.
-Indeed, a recnt discussion on Cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.
+Indeed, a recent discussion on Cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.
 However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.
-It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than tthat they not be taught at all.
+It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
 \par

### 2019-03-23 FIE2019.tex
Added references to Oudshoornetal2018a (and deleted JHD's notes)

@@
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
-ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019) \cite[chapter 7]{Davenportminor2019a}.
-% need a better reference
+ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots.
+%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
 \end{description}

### 2019-04-03 FIE2019.tex
Edited UK/US Accreditation

@@
 In particular the expectation of ''Privacy by Design'' and and by "Privacy by Default" has significant implications for all software development with a European context that processes personal data.
-Whilst both ''Privacy by Design'' and ''Privacy by Default' have been expected good practices for a number of years they expectations now have a legal dimension.
+Whilst both ''Privacy by Design'' and ''Privacy by Default' have been expected good practices for a number of years, these expectations now have a legal dimension.
 A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}.
@@
 \begin{description}
-\item[UK]BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accedited universities should be compliant by 2020 (because of the five-year cycle).
+\item[UK]BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
 ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots.
+This should mean that  all accredited universities should be compliant by 2020 (because of the five-year cycle).
 %\cite[chapter 7]{Davenportminor2019a}.  % need a better reference

### 2019-04-03 FIE2019.tex
Shortened Tom P's affiliation

@@
 Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, 
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Dept.~Computer Science, 
 University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
@@
 Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Computer and Information Sciences, 
+\IEEEauthorblockA{\IEEEauthorrefmark{4}Dept.~Computer\&Information Sciences, 
 Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}

### 2019-04-14 FIE2019.tex
TomP's acceditation changes; a couple of bug fixes

@@
 \begin{quote}
-change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years - to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
+change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
 \cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}
@@
 
-AI: GDPR essentially expects CyberSecurity to be in all coding.
+%AI: GDPR essentially expects CyberSecurity to be in all coding.
 
 In particular the expectation of ''Privacy by Design'' and and by "Privacy by Default" has significant implications for all software development with a European context that processes personal data.
-Whilst both ''Privacy by Design'' and ''Privacy by Default' have been expected good practices for a number of years, these expectations now have a legal dimension.
+Whilst both ''Privacy by Design'' and ''Privacy by Default' have been expected good practices for a number of years, these expectations now have a legal dimension, at least in Europe.
 A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}.
@@
 \begin{description}
-\item[UK]BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
+\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed Cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
+More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced Cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
+%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
 ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots.
-This should mean that  all accredited universities should be compliant by 2020 (because of the five-year cycle).
+This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
 %\cite[chapter 7]{Davenportminor2019a}.  % need a better reference

### 2019-04-16 FIE2019.tex
Made anonymous
E-mailed version, also thanks to accreditors
A few bug fixes and ``Security by Implementation'' added on page 1.

@@
-% JHD to
-% a) Add the Bonn work
-% b) Add ABET stuff (done) and ask Tom P to cross-check.
+\newif\ifanon
+\anontrue
 \documentclass[conference]{IEEEtran}
@@
 % }
-
+\ifanon
+\author{Anonymous}
+\else
 % compact IEEE author format for more than three authors...
@@
 Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
-
+\fi
 \maketitle
@@
 
-In particular the expectation of ''Privacy by Design'' and and by "Privacy by Default" has significant implications for all software development with a European context that processes personal data.
-Whilst both ''Privacy by Design'' and ''Privacy by Default' have been expected good practices for a number of years, these expectations now have a legal dimension, at least in Europe.
+In particular the expectation of ``Privacy by Design'' and and by ``Privacy by Default'' has significant implications for all software development with a European context that processes personal data.
+Whilst both ``Privacy by Design'' and ``Privacy by Default'' have been expected good practices for a number of years, these expectations now have a legal dimension, at least in Europe.
+
+However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
 A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}.
@@
 Chartered Institute for IT for supplying the summary information
-related to Accreditation.
+related to Accreditation. Many people, accreditors and accredited, contributed to  improving the standard of Cybersecurity teaching, and spreading good practice.
 

### 2019-04-16 FIE2019.tex
Trivia, but must re-run BiBTeX

@@
 Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
-Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified by manual to be vulnerable --- 80\% of which were vulnerable to SQL injection.
+Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
 

### 2019-04-16 FIE2019.tex
More conclusions

@@
 It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
+Nevertheless, we believe it is wrong for a complete curriculum to ignore CyberSecurity issues and:
+\begin{itemize}
+	\item teach SQL without teaching SQL Injection (\cite{Drop2019} and the second author's experience);
+	\item teach ``web programming'' without teaching Cross-Site Scripting \cite[(XSS)]{OWASP2017a};
+	\item teach \verb+strcpy+ etc. without teaching the necessity of guards.
+\end{itemize}
 \par
@@
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
+\item Web Programming courses shoud probably do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.

### 2019-06-09 Response.tex
JHD starts the process

\font\manual=manfnt
\def\dbend{{\manual\char127}}
\def\eqq{{\buildrel?\over=}}
\def\C{\mathbf{ C}}
\def\N{\mathbf{ N}}
\def\Q{\mathbf{ Q}}
\def\R{\mathbf{ R}}
\def\Z{\mathbf{Z}}
\def\F{\mathbf{F}}
\def\softO{\tilde{\cal O}}
\def\O{{\cal O}}
%\def\N{{\bf N}}
%\def\Q{{\bf Q}}
%\def\R{{\bf R}}
%\def\Z{{\bf Z}}
\def\action#1{\hfil\rlap{\bf#1}\break}
\def\pcite#1{[#1]}
\def\Res{\mathop{\rm Res}\nolimits}
\def\r#1#2{``#1''$\rightarrow$``#2''}
\documentclass{article}
\bibliographystyle{alphaurl}
\usepackage[hyphens]{url}
%\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage[show]{ed}
\usepackage{graphicx}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{notation}{Notation}
\newtheorem{example}{Example}
\newtheorem{problem}{Problem}
\def\decision#1{\par{\bf #1}}
\pagestyle{empty}
\begin{document}
\author{Crick/Davenport/Irons/Prickett}
\title{Response to Referees}
\date{June 2019}
\maketitle

\section{Referee 1}
No comments that require addressing in detail.
\section{Referee 2}
No comments that require addressing in detail.
\section{Referee 3}
\begin{enumerate}
\item ``The research appears to be poorly structured and the analysis/argument is hard to interpret''.
\item[A]We have done what we can, but note that the other referees disagree.
\item ``The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.''
\item[A]
\item 
\item[A]
\item 
\item[A]
\item 
\item[A]
\item 
\item[A]
\end{enumerate}
\section{Programme Committee}
No comments that require addressing in detail.

\end{document}

\end{document}
\bibliography{../../../../jhd}
\section{}
\begin{description}
\item
\end{description}
\begin{itemize}
\item
\end{itemize}
\begin{example}
\end{example}
\begin{definition}
\end{definition}
\begin{theorem}
\end{theorem}
\begin{enumerate}
\end{enumerate}
\begin{description}
\item[Theme 1]
\item[Theme 2]
\item[Theme 3]
\item[Theme 4]
\item[Theme 5]
\end{description}

### 2019-06-09 FIE2019-final.tex
JHD starts the process

--- 2019-04-16 FIE2019.tex
@@
 \begin{IEEEkeywords}
-TBC
+Cybersecurity, accreditation, curricula, computer science education, UK
 \end{IEEEkeywords}
@@
 
-Cybersecurity has been in the news for several years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
+Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
 It needs attention across the spectrum:
@@
 
-There are calls for education to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}.
+With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,brown-et-al:toce2014,wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,murphy-et-al:programming2017,simon-et-al:sigcse2018,tryfonas+crick:petra2018}.
 There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.
@@
 
-Nevertheless, it is one thing to write curricula and requirements, and
-another thing to deliver appropriate education, and one could
+Nevertheless, it is one thing to write national curricula, specifications and requirements, and
+another thing to deliver appropriate and relevant education and skills; furthermore, one could
 reasonably ask how well this is done in practice.
+
 \subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for Cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} Cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
 
 This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
-On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of Cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
-It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in Cybersecurity.
+On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
+It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity.
 On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
@@
 Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}.
-However, given the shortage of specialists, in practice a good generalist with some Cybersecurity expertise can often get a ``specialist'' job.
+However, given the shortage of specialists, in practice a good generalist with some cybersecurity expertise can often get a ``specialist'' job.
 \subsection{Research Questions}
-There are various levels of specialism at which Cybersecurity education can be addressed.
+There are various levels of specialism at which cybersecurity education can be addressed.
 \begin{enumerate}
@@
 \item[RQ2]Should this be taught stand-alone or integrated? 
-\item[RQ3]How might accreditation regard Cybersecurity Education, or help with it?
+\item[RQ3]How might accreditation regard cybersecurity Education, or help with it?
 \end{description}
@@
 \begin{description}
-\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed Cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
-More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced Cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
+\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
+More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
 %BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
@@
 % TC to add policy+skills context
-% inc. new EU Cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
+% inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
 
@@
 UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
---- but what has happened to the focus on Cybersecurity?
-The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention Cybersecurity, but merely as a sub-item in one work package (1.2a).
+--- but what has happened to the focus on cybersecurity?
+The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
 
@@
 There is large-scale media
-attention on the ``Cybersecurity skills gap''
+attention on the ``cybersecurity skills gap''
 e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
@@
 
-Is teaching Cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.
+Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.
 
@@
 
-%Alastair to write about their Credit Fraud experience (MSc Cybersecurity).  HE Academy grant (to be referenced).
+%Alastair to write about their Credit Fraud experience (MSc cybersecurity).  HE Academy grant (to be referenced).
 
@@
 Clearly such a major weakness should be well-taught.
-One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist Cybersecurity course.
+One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cybersecurity course.
 In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.
@@
 But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources.
-\cite{Mengetal2018a} analysed 503 Cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.
+\cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.
 53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%).
 The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.
-By default, Spring implcitly enables protection against this.
+By default, Spring implicitly enables protection against this.
 But all the accepted answers to CSRF-related failures simply suggested disabling they check.
@@
 \subsection{Staff}
-It is well known that Cybersecurity skills are in short supply, for example \cite{Page2018a}.
+It is well known that cybersecurity skills are in short supply, for example \cite{Page2018a}.
 \begin{quote}
@@
 
-In the authors' experience, it is proving very difficult to recruit academic staff with specialisms in Cybersecurity.
-The demand for Cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
-As universities expand their Cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.
-A recent example had a professor of Cybersecurity, two senior academic positions and 2 junior academic positions in one advert.
-There are examples in the UK of Cybersecurity lecturing jobs remaining unfilled for longer than a year.
-There are also examples of  Cybersecurity subject groups moving en masse from one university to another.
+In the authors' experience, it is proving very difficult to recruit academic staff with specialisms in cybersecurity.
+The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
+As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.
+A recent example had a professor of cybersecurity, two senior academic positions and 2 junior academic positions in one advert.
+There are examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year.
+There are also examples of  cybersecurity subject groups moving en masse from one university to another.
 
@@
 
-For a few years now there has been a recognized need in the UK to build knowledge, skills and capacity in the area of Cybersecurity.
+For a few years now there has been a recognized need in the UK to build knowledge, skills and capacity in the area of cybersecurity.
 This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
-\cite{UKCabinetOffice} or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.
+\cite{UKCabinetOffice} or National Initiative for cybersecurity Education (NICE) in the USA \cite{NICE}.
 
-The teaching of Cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of Cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
-There has  been a debate as to whether Cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}.
-The consensus now is that Cybersecurity is both a discipline in its own right and that Cybersecurity should be taught within Computer Science and related degrees.
-There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised Cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
+There has  been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}.
+The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
+There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
 
 In the United Kingdom, Higher Education provision addresses both approaches.
-A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and Cybersecurity (and closely related fields Computer Security, Digital / Computer Forensics, etc).
-In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to Cybersecurity for entry in September 2019.
+A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and cybersecurity (and closely related fields Computer Security, Digital / Computer Forensics, etc).
+In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to cybersecurity for entry in September 2019.
 An even larger number of HEI provide study opportunities related to more general Computer Science.
@@
 
-Accreditation has evolved to directly addresses the Cybersecurity challenges in both general Computer Science programmes and specialist Cybersecurity programmes.
+Accreditation has evolved to directly addresses the cybersecurity challenges in both general Computer Science programmes and specialist cybersecurity programmes.
 In the UK, accreditation in the broad computing area is being performed by a few different agencies.
@@
 Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements.
-One of the curricula elements is related to Cybersecurity.
+One of the curricula elements is related to cybersecurity.
 Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.
 Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and 5 Software Engineering for Business Programmes.
-As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education and none upon specialist Cybersecurity education.
+As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education and none upon specialist cybersecurity education.
 
 The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK \cite{Davenportetal2019a}.
-This is likely to include Cybersecurity related skills.
+This is likely to include cybersecurity related skills.
 Like the Tech Partnership the focus is upon job market requirements.
@@
 \textbf{2. National Cyber Security Centre (NCSC)}
-The NCSC is a UK Government organisation tasked with enhancing the Cybersecurity of the UK.
-The NCSC publishes and accredits to a number of Cybersecurity standards \cite{NCSC2018a}.
+The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
+The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}.
 These standard are linked to the ACM recommendations for curricula.
-To date the major focus of NSSC acreditation has been upon Masters degrees specializing in Cybersecurity.
-More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in Cybersecurity and Computer Science degrees with a significant Cybersecurity focus \cite{NCSC2018b}.
+To date the major focus of NSSC accreditation has been upon Masters degrees specializing in cybersecurity.
+More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and Computer Science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
 
-Currently accredited are 15 Cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters Cybersecurity programmes and 1 Cybersecurity Degree Programme with a further 2 provisionally accredited.
+Currently accredited are 15 cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters cybersecurity programmes and 1 cybersecurity Degree Programme with a further 2 provisionally accredited.
 Hence the extent of accreditation is currently reasonably limited in terms of reach to Computer Science programmes.
@@
 \textbf{3. Professional Bodies}
-The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of Cybersecurity discipline areas.
+The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of cybersecurity discipline areas.
 The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}.
@@
 The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}.
-Clearly Cybersecurity has been and continues to be part of these expectations.
+Clearly cybersecurity has been and continues to be part of these expectations.
 
-In recent years the BCS has evolved its Accreditation Practices to promote and mandate the inclusion of Cybersecurity within the programmes the body accredits.
+In recent years the BCS has evolved its Accreditation Practices to promote and mandate the inclusion of cybersecurity within the programmes the body accredits.
 The timeline which this process has followed in Table \ref{table:1}.
@@
 \end{center}
+
 \subsection{Developing Expectations}
-Internationally the expectations regarding both the breadth and depth of the expected Cybersecurity coverage has been the subject of much discussion, debate and analysis.
-Like many Governments, the UK Goverment has been actively been seeking ways to address this \cite{UKCabinetOffice}, \cite{UKCabinetOffice2014}.
+Internationally the expectations regarding both the breadth and depth of the expected cybersecurity coverage has been the subject of much discussion, debate and analysis.
+Like many Governments, the UK Government has been actively been seeking ways to address this \cite{UKCabinetOffice}, \cite{UKCabinetOffice2014}.
 In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
-In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of Computer Science Departments) and the Cabinet Office to examine embedding Cybersecurity into undergraduate degrees in the UK.
-Three workshops in 2013 and 2015 attempted to define the principles of Cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula.
+In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of Computer Science Departments) and the Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
+Three workshops in 2013 and 2015 attempted to define the principles of cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula.
 Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate Computing Science degrees.
-This work initially lead to a white paper related to a proposal in the form of a whitepaper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}.
+This work initially lead to a white paper related to a proposal in the form of a white paper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}.
 The BCS agreed to adopt the outputs into their accreditation criteria.
-This was the first time that Cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees.
-The fact that Cybersecurity is included as a  component of the BCS accreditation criteria, reflects the importance placed on Cybersecurity and the expectation that all computing graduates should have knowledge and skills in Cybersecurity as they move towards Chartered status.
+This was the first time that cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees.
+The fact that cybersecurity is included as a  component of the BCS accreditation criteria, reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards Chartered status.
 
-The produced reference guidelines ("Cybersecurity Principles and Learning Outcomes") \cite{CPHCISC2} established a baseline of common knowledge, example learning outcome domains for Cybersecurity within the Computing Science courses and guidance on embedding the concepts.
-The document provides specific guidance for embedding and enhancing relevant Cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.
+The produced reference guidelines ("Cybersecurity Principles and Learning Outcomes") \cite{CPHCISC2} established a baseline of common knowledge, example learning outcome domains for cybersecurity within the Computing Science courses and guidance on embedding the concepts.
+The document provides specific guidance for embedding and enhancing relevant cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.
 The document suggested 5 areas of coverage
@@
 The ambition of this approach is to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area.)  The approach taken is not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
-In this context this is realized as an expectation Cybersecurity is an inclusion in all degrees accredited by the BCS.
+In this context this is realized as an expectation cybersecurity is an inclusion in all degrees accredited by the BCS.
 e.g.
-the expectation for coverage is true for Computer Science as well as Cybersecurity programmes.
+the expectation for coverage is true for Computer Science as well as cybersecurity programmes.
 Two criteria are expected to be covered by all programmes seeking accreditation.
@@
 
-The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages HEIs to embed Cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
-In addition there an expectation that there is significant coverage of Cybersecurity principles and fundamentals - either as a stand alone module or as a significant component(s) of other modules.
+The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages HEIs to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
+In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals - either as a stand alone module or as a significant component(s) of other modules.
 This approach differs from the ACM approach where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
@@
 The agreed Cybersecurity Principles and Learning Outcome \cite{CPHCISC2} were discussed with the wider education community by a road-show led by CPHC.
-A series of workshops took place in 2015 which presented the rationale for embedding Cybersecurity in the curriculum of Computing Science degrees.
-The workshops included case studies from universities who had embedded Cybersecurity Computer Science curriculum illustrating different approaches to implementation.
+A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of Computing Science degrees.
+The workshops included case studies from universities who had embedded cybersecurity into their Computer Science curricula illustrating different approaches to implementation.
 The workshops had 102 attendees from the academic Computing Science community representing 60 UK Universities.
@@
 Significant opportunity for seeking clarification is provided.
-One of the issues that is highlighted is not all institutions have yet evolved their programmes to fully address the increased expectation for Cybersecurity.
+One of the issues that is highlighted is not all institutions have yet evolved their programmes to fully address the increased expectation for cybersecurity.
 This is resulting in accreditation being contingent upon a HEI taking action to address this short fall or in some cases the withdrawal of accreditation.
@@
 
-This initiative is a collective attempt to formally include Cybersecurity in all BCS Accredited programmes.
-Some of these programmes will be specialist Cybersecurity programmes, however the majority will take a different emphasis.
+This initiative is a collective attempt to formally include cybersecurity in all BCS Accredited programmes.
+Some of these programmes will be specialist cybersecurity programmes, however the majority will take a different emphasis.
 This is a work in progress.
 A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines.
-What is being observed is the majority of visited HEIs have now either adjusted their curricula to extend the coverage of Cybersecurity or have a plan in place to do so.
+What is being observed is the majority of visited HEIs have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
 However, a minority are requiring encouragement to do so.
@@
 From the start of the Autumn 2015 term, up to and including the Autumn 2018 term, the BCS have carried out 70 accreditation visits (including 4 international visits).
-The BCS identified action was required to address concerns related to Cybersecurity at 16 HEIs.
-So 54 HEIs were already delivering Cybersecurity in keeping with the BCS expectations.
+The BCS identified action was required to address concerns related to cybersecurity at 16 HEIs.
+So 54 HEIs were already delivering cybersecurity in keeping with the BCS expectations.
 
@@
 This indicates that the BCS will expect adjustments to have taken place before the next accreditation visit.
-As indicated earlier, this was commonly the case that adjustments had been made to approved programmes of study, however the adjusted modules had not yet been delivered so the evidence base was incomplete in terms of how Cybersecurity was assessed.
+As indicated earlier, this was commonly the case that adjustments had been made to approved programmes of study, however the adjusted modules had not yet been delivered so the evidence base was incomplete in terms of how cybersecurity was assessed.
 
@@
 \end{quote}
-In sum, this shows that many UK HEIs have now embedded Cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
-Clearly not all HEIs in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of Cybersecurity to an agreed standard.
+In sum, this shows that many UK HEIs have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
+Clearly not all HEIs in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
 
@@
 \item[RQ2]The recommendation in \cite[p.
-98]{ACM2013a} that Cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea.
-Indeed, a recent discussion on Cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.
+98]{ACM2013a} that cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea.
+Indeed, a recent discussion on cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.
 However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.
@@
 \par
-We commend the BCS-identified good practice of exploring Cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
-\item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of Cybersecurity teaching, and spreading good practice, and should continue this.
+We commend the BCS-identified good practice of exploring cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
+\item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of cybersecurity teaching, and spreading good practice, and should continue this.
 \end{description}
@@
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item Web Programming courses shoud probably do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
+\item Web Programming courses should probably do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
@@
 Chartered Institute for IT for supplying the summary information
-related to Accreditation. Many people, accreditors and accredited, contributed to  improving the standard of Cybersecurity teaching, and spreading good practice.
+related to Accreditation. Many people, accreditors and accredited, contributed to  improving the standard of cybersecurity teaching, and spreading good practice.
+
+\IEEEtriggeratref{51}
 

### 2019-06-09 Response.tex
Response should list all the actionable comments

@@
 \item[A]
+\item What do the authors mean by ``There has been a recent international working group but this has yet to report.'' 
+\item[A]
+\item It would add more clarity to introduce what these terms DevOps vs DevSecOps represent in the paper.
+Also, please use the acronym GDPR after introducing it during the first usage such as General Data Protection Regulation (GDPR).
+\item[A]
+\item There are repeated ``and'' in the sentence ``In particular the expectation of Privacy by design...'' Also, the ``by'' after ``and'' is not required.
+
+Please consider rephrasing to ``While both Privacy by design and privacy by defualt have been expected to be good practices''.
+
+\item[A]
+\item Please rephrase the following statement ``How might accreditation regard cybersecurity education, or help with it?'' as the sentence structure doesn't make much sense. 
+\item[A]
+\item It is not clear what is the message the authors are conveying in the 2nd paragraph?
+It lists three different web links and a reference.
+Also, the web links in 2nd and 3rd paragraphs could be moved to the reference and they could be referenced with a number in this paragraph.
+Again, please use the acronym KA after introducing it as Knowledge Areas (KA) at the first instance of use.
+
+\item[A]
+\item  What is JHD? Please expand?
+\item[A]
+\item Please expand OWASP. 
+\item[A]
+\item 
+The references to first author in the 2nd paragraph could be avoided and the sentence rephrased to highlight only the facts and not any personal experience. 
+\item[A]
+\item 
+ What does ``three accounted for the 36 of the 44'' mean? Please clarify. I am assuming the 36 and 44 refer to number of universities. Please add that clarification, if it is so.
+\item[A]
+\item 
+``If there are only 82 instances of these fragments, how can 117 of them be verified?'' Please add clarification to this discussion.
+\item[A]
+\end{enumerate}
+\section{Referee 4}
+\begin{enumerate}
+\item ``The problem is it is not a research paper in even the broadest interpretation of that sense.
+The RQs are invented based on topic areas the authors wish to riff on, only RQ3 comes (somewhat) close to a RQ with an associated presentation based on the literature and some analysis.
+The 1st 2 RQs, as well as the Introduction sections, are selectively argued, stream-of-consciousness almost writing.
+Entertaining, informative, arguable, debatable - but not research.
+No methodology is applied, no systematic review, etc.
+Instead a smattering of anecdotal information and quotes (and even more quotes in the footnotes) intended to convince the reader of a viewpoint that is never quite articulated clearly - until the end when it comes down to a small set of specific complaints about SQL Injection, XSS, and StackOverflow (which is overall a bit of a letdown as I felt as if this bloglike writing was building to a more impactful crescendo, some kind of kick-in-the-pants call to action for the community).''
+\item[A]
+\end{enumerate}
+\section{Programme Committee}
+No comments that require addressing in detail.
+
+\end{document}
 \item 
@@
 \item[A]
-\item 
-\item[A]
-\end{enumerate}
-\section{Programme Committee}
-No comments that require addressing in detail.
-
-\end{document}
 

### 2019-06-14 Response-TP.tex
Add files via upload

--- 2019-06-09 Response.tex
@@
 \item ``The research appears to be poorly structured and the analysis/argument is hard to interpret''.
-\item[A]We have done what we can, but note that the other referees disagree.
+\item[A1]We have done what we can, but note that the other referees disagree. We have attempted to make the structure clearer.
 \item ``The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.''
-\item[A]
+\item[A]We have adjusted the second paragraph of the abstract to provide an indication of the research design adopted and the rationale for doing so.
+Cybersecurity is now a commonly included curricula element within the delivery of general computer science.
+Evaluating how cybersecurity is taught in all aspects of computer science hence is considerably large task.
+We feel that is beyond the scope of our current research and the scope of this paper.
+We have instead adopted a case study based research approach to evaluate the evidence of the teaching of cybersecurity within general computer science.
+We have adjusted the abstract to indicate this and attempted to more clearly indicate the case study based approach adopted in the paper.
 \item What do the authors mean by ``There has been a recent international working group but this has yet to report.'' 
-\item[A]
+\item[A] In parallel to our research, their is a working group as part of the Association for Computing Machinery (ACM) Conference on Innovation and Technology in Computer Science Education intending to develop a taxonomy of cybersecurity education.
+We have clarified this in text.
 \item It would add more clarity to introduce what these terms DevOps vs DevSecOps represent in the paper.
 Also, please use the acronym GDPR after introducing it during the first usage such as General Data Protection Regulation (GDPR).
-\item[A]
+\item[A] Discussion related to DevOps vs DevSecOps has been reduced. We apologize for the manner in which we have employed abbreviations and have adopted the style recommended above.
 \item There are repeated ``and'' in the sentence ``In particular the expectation of Privacy by design...'' Also, the ``by'' after ``and'' is not required.
@@
 
-\item[A]
+\item[A] These have been rephrased, apologies for our weak use of the English language in this section. One of the authors is dyslexic and we should have better proof read some of the sections.
 \item Please rephrase the following statement ``How might accreditation regard cybersecurity education, or help with it?'' as the sentence structure doesn't make much sense. 
-\item[A]
+\item[A] We have rephrased to "Can accreditation enhance the provision of cybersecurity within an accrediting bodies jurisdiction?"
 \item It is not clear what is the message the authors are conveying in the 2nd paragraph?
@@
 \item  What is JHD? Please expand?
-\item[A]
+\item[A] Sorry we should have removed this. 
 \item Please expand OWASP. 
-\item[A]
+\item[A]We have done so. Also SQL.
 \item 
 The references to first author in the 2nd paragraph could be avoided and the sentence rephrased to highlight only the facts and not any personal experience. 
-\item[A]
+\item[A] We have removed the references to the authors throughout
 \item 
  What does ``three accounted for the 36 of the 44'' mean? Please clarify. I am assuming the 36 and 44 refer to number of universities. Please add that clarification, if it is so.
-\item[A]
+\item[A] We have clarified, we can confirm it did mean universities.
 \item 

### 2019-06-14 FIE2019-final-TP.tex
Add files via upload

--- 2019-06-09 FIE2019-final.tex
@@
 % submitted abstract (100-500 word limit):
-There are numerous facets to cybersecurity education, from theory to
-practice, hardware and software, to social and technical (as well as
-other important dimensions). A multitude of national and international
-model curricula and recommendations - from national academies, learned
-societies and even governments -- have been presented and discussed in
-recent years, with varying levels of impact on policy and practice. In
-this paper we attempt to address the key questions ``what should the
-generalist computer scientist/engineer know'' and ``how well is that
-being taught''. (Cyber)security -- if thought about at all -- has
-historically been left to specialists, frequently viewed as a
-masters-level discipline; or left for development in professional
-practice, sometimes called ``information assurance'' in the UK and
-Europe, which worked, even if not well, in a sequential process model.
+There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
+A multitude of national and international model curricula and recommendations - from national academies, learned societies and even governments -- have been presented and discussed in recent years, with varying levels of impact on policy and practice.
+In this paper we attempt to address the key questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
+(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline; or left for development in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a sequential process model.
 
-For example, in the world of software and systems development, we have
-seen the disruption of Agile (invented c.2001) and DevOps (invented
-c.2009), but not necessarily through the lens of security; however,
-more recently there has been a recognition of the need for change,
-exemplified by the emergence of ``DevSecOps''. If security is not to
-be left to the experts, then the generalist must know about it. Thus,
-in the context of widespread international computer
-science/engineering curriculum reform - both in compulsory, as well as
-post-compulsory education - what does this trend mean more generally
-for institutions and educators, and how do we teach it?
+Evaluating how cybersecurity is taught in all aspects of computer science is a  task of considerable size.
+We feel that to do this completely is beyond the scope of our current research and the scope of this paper.
+We have instead adopted a case study based research approach to evaluate the evidence of the teaching of cybersecurity within general computer science to undergraduate students.
+Thus, in the context of widespread international computer science/engineering curriculum reform - both in compulsory, as well as post-compulsory education - what does this need to embed cybersecurity mean more generally for institutions and educators, and how do we teach this curricula area?
 
-In this paper, we frame a key social, cultural and economic global
-challenge of cybersecurity in computing and engineering education by
-analysing the UK's national security, economic, and skills policy
-context, shaping the practical considerations for education
-professionals from schools through to universities (and into industry
-and professional practice). Through this analysis of UK educational
-policy and practice, we make a number of recommendations for future
-cybersecurity educational initiatives and interventions, including a
-number of innovative practical suggestions for educators and
-curriculum designers, with a view for potential adoption and adaption
-in other jurisdictions internationally.
+In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by analysing the UK's national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities (and into industry and professional practice).
+Through this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
 \end{abstract}
@@
 With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,brown-et-al:toce2014,wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,murphy-et-al:programming2017,simon-et-al:sigcse2018,tryfonas+crick:petra2018}.
-There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.
+There has been a recent Association for Computing Machinery (ACM)  working group as part of the Conference on Innovation and Technology in Computer Science Education \cite{Parrishetal2018a}, but this has yet to report.
+The intention of this group is to develop a taxonomy for approaches to cybersecurity education.
 We note that part of its third aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these (see section \ref{sec:SQL}).
@@
 \subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity.
+In this section we illustrate this by briefly exploring four cases, General Data Protection Regulation (GDPR), Agile development, one industry sector (with rigorous demands for secure software) and impact of these cases upon one another.
 
@@
 It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity.
-On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
+On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for expertise in this area.
 % JHD what are these?
@@
 %AI: GDPR essentially expects CyberSecurity to be in all coding. 
+In particular, for the development of software that will operate within Europe and that processes personal data, the expectation of ``Privacy by Design'' and ``Privacy by Default'' have significant legal implications.
+However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
 
-In particular the expectation of ``Privacy by Design'' and and by ``Privacy by Default'' has significant implications for all software development with a European context that processes personal data.
-Whilst both ``Privacy by Design'' and ``Privacy by Default'' have been expected good practices for a number of years, these expectations now have a legal dimension, at least in Europe.
-
-However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
 A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}.
@@
 
-The Payments Industry \cite[6.6]{PCI2018b} insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
+This can be further illustrated by the demands of particular sectors for secure software.
+Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
 This would seem to be particularly the case with ``agile'' processes \cite{Bartsch2011a}.
 
-Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}.
-However, given the shortage of specialists, in practice a good generalist with some cybersecurity expertise can often get a ``specialist'' job.
 \subsection{Research Questions}
@@
 BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
-\item The general CS graduate.
-\item The general CS masters graduate.
-\item The specialist CS graduate.
-\item The specialist CS masters graduate.
+\item The general Computer Science graduate.
+\item The general Computer Science masters graduate.
+\item The specialist Computer Science graduate.
+\item The specialist Computer Science masters graduate.
 \end{enumerate}
-The focus of this paper is on 2--3: the general CS graduate.
+The focus of this paper is on 2--3: the general Computer Science graduate.
 \begin{description}
@@
 \item[RQ2]Should this be taught stand-alone or integrated? 
-\item[RQ3]How might accreditation regard cybersecurity Education, or help with it?
+\item[RQ3]Can accreditation enhance the provision of cybersecurity within an accrediting bodies jurisdiction?
 \end{description}
-\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
+The ACM/Institute of Electrical and Electronics Engineers (IEEE)-CS Joint Task Force on Computing Curricula,\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
 \begin{quote}
@@
 \begin{description}
-\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
+\item[UK]The BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
 More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
@@
 
-There has been substantial CS/digital curriculum reform across the
+There has been substantial Computer Science /digital curriculum reform across the
 UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
@@
 
-\section{Challenges}
+\section{Challenges: RQ1 and RQ2}
 
@@
 %AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
-  Should we be teaching it because it's underpinning? because there's a skill shortage?  JHD would argue both.
+  This raises pedagogic questions which are beyond the scope of this research to resolve. Should we be teaching content because it's underpinning?; or because there's a skill shortage?
 
@@
 
-\subsection{PCI DSS}\label{sec:PCIDSS}
+In this section we will explore a number of case studies to evaluate the evidence of the challenges pertaining to cybersecurity.
+We explore an industry problem with evident cybersecurity implications.
+We could have selected many alternative industry areas which would have provided a similar analysis.
+We then explore the current state of educational resources with respect to cybersecurity.
+Again the cases are illustrative rather than complete in terms of the available case studies.
+Finally we consider challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
+
+\subsection{An industry example - PCI DSS}\label{sec:PCIDSS}
 The Payments Card Industry Data Security Standards \cite{PCI2018b} underpin all processing of credit/debit cards.
@@
 
-\subsection{Educational Resources: SQL Injection}\label{sec:SQL}
+\subsection{Educational Resources}\label{sec:EDResource}
+\subsubsection{Structured Query Language (SQL) Injection}\label{sec:SQL}
 It is 15 years since \cite{Guimaraesetal2004} 
@@
 SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon (\url{https://xkcd.com/327/}, dating back to 2007 according to the Internet Archive) and website (\url{http://bobby-tables.com/}).
-Nevertheless SQL injection is still a major weakness: number one in the OWASP Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.
+Nevertheless SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.
 \cite[the UK's definitive reference]{Bristol2018a} states ``a wide range of attack techniques for exploiting SQL
@@
 
-Clearly such a major weakness should be well-taught.
-One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cybersecurity course.
-In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.
+Clearly such a major weakness should be well-taught.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recommended textbooks.
 
@@
 departments in the United States (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed).
-There were seven such books, but three accounted for the 36 of the 44.
+There were seven such books, but three books accounted for the 36 of the 44 universities.
 Five of the seven (30 of the 44) had no mention of SQL injection.
@@
 \par
-This blindness is not limited to textbooks: although Wikipedia has an excellent article on SQL Injection, it was not linked from the SQL page itself.\footnote{The authors intend to fix this after the paper is processed, so as not to break anonymity.}
-\subsection{Informal Resources}\label{sec:informal}
-The web abounds with informal resources: tutorials and code snippets.
-How good are these, and how good are people at using these?
-This has been looked at by \cite{Unruhetal2017a}, who took the top five results from Google for six queries.
-Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
-Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
-
-\subsection{The case of Java}\label{sec:Java}
-To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for Java textbooks.
+This blindness is not limited to textbooks: although Wikipedia has an excellent article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
+\subsubsection{The case of Java}\label{sec:Java}
+A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for  Java textbooks.
 Indeed, many such books go nowhere near security applications.
@@
 
-\subsection{Android}\label{sec:Android}
-\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
+\subsubsection{Android}\label{sec:Android}
+A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks.
+\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
+Their key finding was this.
 \begin{quote}
@@
 Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''.
-\subsection{Agile}
-To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for ``Agile'' textbooks.
-It is the authors' experience that the vast majority of students' individual software projects are developed on Agile lines.
+\subsubsection{Agile}
+A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for ``Agile'' textbooks.
 Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
@@
 None of the unprimed ones implemented any security.
+\subsubsection{Informal Resources}\label{sec:informal}
+The web abounds with informal resources: tutorials and code snippets.
+How good are these, and how good are people at using these?
+This has been looked at by \cite{Unruhetal2017a}, who took the top five results from Google for six queries.
+Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
+Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
+Some students clearly make use of these resources, a recommendation of future work is to estimate students use of such sources.
+
 \subsection{Staff}
@@
 % JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
-Research into the state of IT conducted annually by ESG\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
+Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
 In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
@@
 
-In the authors' experience, it is proving very difficult to recruit academic staff with specialisms in cybersecurity.
+In the United Kingdom there is a prominence of job adverts to recruit academic staff with specialisms in cybersecurity.
 The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
@@
 \begin{itemize}
-	\item teach SQL without teaching SQL Injection (\cite{Drop2019} and the second author's experience);
+	\item teach SQL without teaching SQL Injection (\cite{Drop2019} );
 	\item teach ``web programming'' without teaching Cross-Site Scripting \cite[(XSS)]{OWASP2017a};

### 2019-06-15 FIE2019-final-TPJHD.tex
JHD's version

--- 2019-06-14 FIE2019-final-TP.tex
@@
 We have instead adopted a case study based research approach to evaluate the evidence of the teaching of cybersecurity within general computer science to undergraduate students.
-Thus, in the context of widespread international computer science/engineering curriculum reform - both in compulsory, as well as post-compulsory education - what does this need to embed cybersecurity mean more generally for institutions and educators, and how do we teach this curricula area?
+Thus, in the context of widespread international computer science/engineering curriculum reform, %- both in compulsory, as well as post-compulsory education -
+what does this need to embed cybersecurity mean more generally for institutions and educators, and how do we teach this subject?
 

### 2019-06-15 FIE2019-final-TPJHD.tex
tweaks to challenges intro and PCI section

@@
 
-Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.
+Is teaching cybersecurity different?
+Lecturing is probably not the best way.
+Should we use real-life case studies (e.g.
+British Airways \cite{Barth2018a}).
+\cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.
 
 %AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
-  This raises pedagogic questions which are beyond the scope of this research to resolve. Should we be teaching content because it's underpinning?; or because there's a skill shortage?
+  % Should we remove this?
+This raises pedagogic questions which are beyond the scope of this research to resolve.
+Should we be teaching content because it's underpinning?; or because there's a skill shortage?
 
@@
 
-\subsection{An industry example - PCI DSS}\label{sec:PCIDSS}
-The Payments Card Industry Data Security Standards \cite{PCI2018b} underpin all processing of credit/debit cards.
-Nevertheless, they are very rarely mentioned in generalist computer scientist courses --- one payments industry person commented ``I've given up even asking if recruits have heard of PCI DSS: it's so rare''.
-
-This would not matter so much if everyone handling payments data were sent by their employers on an effective\footnote{``Effective'' is important: the Processing Director at one major acquirer commented ``I just caught two developers in the elevator discussing how to meet a customer requirement, and I had to tell them this was contrary to PCI DSS.
-Their response was that the customer wanted it!''.} PCI DSS course.
-However, the payments business is now so spread across websites, often run by SMEs, or non-specialists, that some provision should be made in courses.
+\subsection{Industry standards - Payments Card Industry Data Security Standards (PCI DSS)}\label{sec:PCIDSS}
+The PCI DSS \cite{PCI2018b} underpin all processing of credit/debit cards.
+Nevertheless, they are very rarely mentioned in generalist computer scientist courses.
+This would not matter so much if everyone handling payments data were sent by their employers on an effective PCI DSS course.
+However, the payments business is now so spread across websites, often run by small and medium enterprises (SME), or non-specialists.
 Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.
-
-%Alastair to write about their Credit Fraud experience (MSc cybersecurity).  HE Academy grant (to be referenced).
+It is an interesting question as to whether standards such as PCI DSS should be addressed within degree courses (clearly degree courses can never cover all standards) or whether they should be addressed in professional training courses.
+However it appears the current situation is not ideal.
 

### 2019-06-15 FIE2019-final-TPJHD.tex
Attempt to fix the KA question

@@
 \begin{quote}
-The Information Assurance and Security KA is unique among the set of KAs presented here
-given the manner in which the topics are pervasive throughout other Knowledge Areas.
+The Information Assurance and Security Knowledge Area (KA) is unique among the set of KAs presented here
+given the manner in which the topics are pervasive throughout other KAs.
 \end{quote}
@@
 
-But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: he quotes that 61\% of UK courses offer mandatory cybersecurity content, and his research was based on web scraping, it represents a lower bound.
+But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: he quotes that 61\% of UK courses offer mandatory cybersecurity content, and, as his research was based on web scraping, this represents a lower bound.
 

### 2019-06-15 FIE2019-final-TPJHD.tex
Consistent (and shorter) bibliography

@@
 
-In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by analysing the UK's national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities (and into industry and professional practice).
-Through this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
+In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by dThrough this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
 \end{abstract}
@@
 \item[UK]The BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
-More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
+More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} %since 2010 redundant
+	with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
 %BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
-ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots.
-This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
+ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots.'' This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
 %\cite[chapter 7]{Davenportminor2019a}.  % need a better reference

### 2019-06-15 FIE2019-final-TPJHD.tex
JHD tidying

@@
 There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
-A multitude of national and international model curricula and recommendations - from national academies, learned societies and even governments -- have been presented and discussed in recent years, with varying levels of impact on policy and practice.
+A multitude of national and international model curricula and recommendations
+%-- from national academies, learned societies and even governments -- 
+have been presented and discussed in recent years, with varying levels of impact on policy and practice.
 In this paper we attempt to address the key questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
@@
 \subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity.
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large%: \cite{JCNSS2018a} has only anecdotal evidence   %JHD - do we need this, as its only anecdotal?
+), but also all in IT need to know \emph{some} cybersecurity.
 In this section we illustrate this by briefly exploring four cases, General Data Protection Regulation (GDPR), Agile development, one industry sector (with rigorous demands for secure software) and impact of these cases upon one another.
@@
 The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
-
-UK national economic skills priority
+There is little connection with
+UK national economic skills priority,
 e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
@@
 In the context of BCS accreditation, these requirements imply an exit standard that all students on a programme must be able to demonstrate irrespective of the option choices they have made.
-This means a HEI applying for accreditation is expected to provide evidence that the criteria are taught and assessed in a non-trivial manner, to and by all students upon the programme seeking accreditation .
+This means a HEI applying for accreditation is expected to provide evidence that the criteria are taught and assessed in a non-trivial manner, to and by all students upon the programme seeking accreditation.
 A HEI is expected to provide evidence in the form of programme and module specification documentation and example assessment specifications (coursework and examinations).

### 2019-06-19 FIE2019-final-TPJHD.tex
Some suggested updates

--- 2019-06-15 FIE2019-final-TPJHD.tex
@@
 A multitude of national and international model curricula and recommendations - from national academies, learned societies and even governments -- have been presented and discussed in recent years, with varying levels of impact on policy and practice.
-In this paper we attempt to address the key questions ``what should the generalist computer scientist/engineer know'' and ``how well is that being taught''.
-(Cyber)security -- if thought about at all -- has historically been left to specialists, frequently viewed as a masters-level discipline; or left for development in professional practice, sometimes called ``information assurance'' in the UK and Europe, which worked, even if not well, in a sequential process model.
+This paper address the key questions ``What cybersecurity is taught and what cybersecurity should be taught to the generalist Computer Science undergraduate students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to generalist Computer Science students?'' and ''Can accreditation by Professional, Statutory and Regulatory Bodies(PSRBs) enhance the provision of cybersecurity within a bodies jurisdiction?''.
 
 Evaluating how cybersecurity is taught in all aspects of computer science is a  task of considerable size.
-We feel that to do this completely is beyond the scope of our current research and the scope of this paper.
-We have instead adopted a case study based research approach to evaluate the evidence of the teaching of cybersecurity within general computer science to undergraduate students.
+This is beyond the scope the scope of this paper.
+Instead a case study based research approach has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to undergraduate students.
 Thus, in the context of widespread international computer science/engineering curriculum reform, %- both in compulsory, as well as post-compulsory education -
@@
 
-In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by analysing the UK's national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities (and into industry and professional practice).
-Through this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
+The paper proposes a number of recommendations for future cybersecurity educational initiatives, interventions and research, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
 \end{abstract}
@@
 There has been a recent Association for Computing Machinery (ACM)  working group as part of the Conference on Innovation and Technology in Computer Science Education \cite{Parrishetal2018a}, but this has yet to report.
-The intention of this group is to develop a taxonomy for approaches to cybersecurity education.
-We note that part of its third aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these (see section \ref{sec:SQL}).
+The intention of this working group is to develop a taxonomy for approaches to cybersecurity education.
+The third aim of this working group is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
 
 Nevertheless, it is one thing to write national curricula, specifications and requirements, and
-another thing to deliver appropriate and relevant education and skills; furthermore, one could
-reasonably ask how well this is done in practice.
+another thing to deliver appropriate and relevant education and skills; furthermore, this paper asks how well this is done in practice.
 
@@
 In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity.
-In this section we illustrate this by briefly exploring four cases, General Data Protection Regulation (GDPR), Agile development, one industry sector (with rigorous demands for secure software) and impact of these cases upon one another.
+This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
+In this section four case studies are briefly explored: General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
 
-This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
 On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
-It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity.
+It is hard to envisage one professioanl possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity.
 On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for expertise in this area.
@@
 \begin{description}
-\item[RQ1]What should be taught to the generalist, and how?
-\item[RQ2]Should this be taught stand-alone or integrated? 
-\item[RQ3]Can accreditation enhance the provision of cybersecurity within an accrediting bodies jurisdiction?
+\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the generalist Computer Science undergraduate students?
+\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to generalist Computer Science undergraduate students?
+\item[RQ3]Can accreditation by PSRBs enhance the provision of cybersecurity within a bodies jurisdiction?
 \end{description}
-The ACM/Institute of Electrical and Electronics Engineers (IEEE)-CS Joint Task Force on Computing Curricula,\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
+
+\subsection{Research Approach}
+A case study based approach is adopted in this research.
+As is common in case study based research, many alternative case studies could have been chosen.
+The cases are illustrative rather than complete in terms of the available case studies.
+The cases where evaluated to articulate the progress made and highlight some opportunities for future work.
+
+The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs,  together with the published evidence regarding compliance with these recommendations.
+This is followed by the evaluation of the policy context.
+Together these provide a context to the ongoing enhancement initiatives in the area of cybersecurity education.
+
+In order to address RQ1 and RQ2 a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate Computer Science students are evaluated.
+Namely,  an industry problem with evident cybersecurity implications;  the current state of educational resources with respect to cybersecurity;  and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
+
+Finally in order to address RQ3 cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate Computer Science in one jurisdiction (the United Kingdom).
+As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.
+
+\subsection{Current situation}
+The ACM/Institute of Electrical and Electronics Engineers (IEEE)-CS Joint Task Force on Computing Curricula,\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas(KAs) related to RQ2:
 \begin{quote}
-The Information Assurance and Security Knowledge Area (KA) is unique among the set of KAs presented here
-given the manner in which the topics are pervasive throughout other KAs.
+The Information Assurance and Security KA is unique among the set of KAs presented here
+given the manner in which the topics are pervasive throughout other KAs
 \end{quote}
-It proposes 9 ``core'' hours and 63.5 distributed across the other Knowledge Areas.
+It proposes 9 ``core'' hours and 63.5 distributed across the other KAs.
 %\pagebreak - to fix URL issues
 \par
-Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, writes as follows
+Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, indicates
 \begin{quote}
@@
 
-But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: he quotes that 61\% of UK courses offer mandatory cybersecurity content, and, as his research was based on web scraping, this represents a lower bound.
+But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: the paper quotes that 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping.
+As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publically available web documents.
 
@@
 \section{Policy Context}
+
+For a few years now there has been a recognized need to build knowledge, skills and capacity in the area of cybersecurity.
+This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
+\cite{UKCabinetOffice} or National Initiative for cybersecurity Education (NICE) in the USA \cite{NICE}.
+
+The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
+There has  been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}.
+The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
+There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
 
@@
 
-The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
-
-Is teaching cybersecurity different?
-Lecturing is probably not the best way.
-Should we use real-life case studies (e.g.
-British Airways \cite{Barth2018a}).
-\cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.
+The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterization, but a long way from being a textbook (which is not their aim).
 
@@
 
-In this section we will explore a number of case studies to evaluate the evidence of the challenges pertaining to cybersecurity.
-We explore an industry problem with evident cybersecurity implications.
-We could have selected many alternative industry areas which would have provided a similar analysis.
-We then explore the current state of educational resources with respect to cybersecurity.
-Again the cases are illustrative rather than complete in terms of the available case studies.
-Finally we consider challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
+\subsection{Industry standards - Payments Card Industry Data Security Standards (PCI DSS)}\label{sec:PCIDSS}
+Is teaching cybersecurity different?
+Lecturing is probably not the best way.
+Should we use real-life case studies (e.g.
+British Airways \cite{Barth2018a}).
+\cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.
+There are a number of ways a more practical manner could be action, one would be by the inclusion of pertinent cybersecurity standards in the curricula.
 
-\subsection{Industry standards - Payments Card Industry Data Security Standards (PCI DSS)}\label{sec:PCIDSS}
-The PCI DSS \cite{PCI2018b} underpin all processing of credit/debit cards.
+The PCI DSS \cite{PCI2018b} is one such standard.
+PCI DSS underpins all processing of credit/debit cards.
 Nevertheless, they are very rarely mentioned in generalist computer scientist courses.
@@
 Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.
+
 It is an interesting question as to whether standards such as PCI DSS should be addressed within degree courses (clearly degree courses can never cover all standards) or whether they should be addressed in professional training courses.
-However it appears the current situation is not ideal.
+However it appears the current situation is not ideal from the perspective of industry (or users of systems) and the inclusion of pertinent standards could be seen as a useful enhancement activity to the coverage of cybersecurity within generalist undergraduate computer science courses.
 
@@
 Five of the seven (30 of the 44) had no mention of SQL injection.
-On the other two, the more popular one has a seriously flawed discussion\footnote{``However, they imply that using parameters is equivalent to using a function to add escape characters
+On the other two, the more popular one has a seriously flawed discussion\footnote{``However, the paper implies that using parameters is equivalent to using a function to add escape characters
 around user input. This is incorrect, as using parameters allows
@@
 \par
-This blindness is not limited to textbooks: although Wikipedia has an excellent article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
+This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
 \subsubsection{The case of Java}\label{sec:Java}
@@
 \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
-Their key finding was this.
+The key finding was this.
 \begin{quote}
@@
 \end{quote}
-We should note two caveats (in opposite directions).
-Their labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure.
+Two caveats (in opposite directions) should be noted.
+The labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure.
 On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
@@
 How good are these, and how good are people at using these?
-This has been looked at by \cite{Unruhetal2017a}, who took the top five results from Google for six queries.
+This has been looked at by \cite{Unruhetal2017a}, This research took the top five results from Google for six queries.
 Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
 Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
-Some students clearly make use of these resources, a recommendation of future work is to estimate students use of such sources.
+Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students use of such Informal Resources.
 
@@
 %{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
-\section{Accreditation: RQ3}
+\section{Accreditation by PSRB: RQ3}
 
-For a few years now there has been a recognized need in the UK to build knowledge, skills and capacity in the area of cybersecurity.
-This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
-\cite{UKCabinetOffice} or National Initiative for cybersecurity Education (NICE) in the USA \cite{NICE}.
-
-The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
-There has  been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}.
-The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
-There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
-
-In the United Kingdom, Higher Education provision addresses both approaches.
+In the United Kingdom as in most jurisdictions, Higher Education provision addresses general Computer Science  and specialist cybersecurity courses.
 A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and cybersecurity (and closely related fields Computer Security, Digital / Computer Forensics, etc).
-In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEI) providing Undergraduate qualifications related to cybersecurity for entry in September 2019.
-An even larger number of HEI provide study opportunities related to more general Computer Science.
+In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEIs) providing Undergraduate qualifications related to cybersecurity for entry in September 2019.
+An even larger number of HEIs provide study opportunities related to more general Computer Science.
 UCAS lists 246 provides for undergraduate programmes related to Computer Science.
@@
 \section{Conclusions}
-As regards our research questions, we can make the following comments.
+As regards the research questions;.
 \begin{description}
-\item[RQ1]The guidelines from both ACM and BCS are good for general education.
-However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}--) resources.
-\item[RQ2]The recommendation in \cite[p.
+\item[RQ1] - What cybersecurity is taught and what cybersecurity should be taught to the generalist Computer Science undergraduate students?
+
+The guidelines from both ACM and BCS are good for general education.
+However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
+\item[RQ2] - Should cybersecurity be taught stand-alone or in an integrated manner to generalist Computer Science undergraduate students?
+
+The recommendation in \cite[p.
 98]{ACM2013a} that cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea.
-Indeed, a recent discussion on cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.
-However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.
+However, in the current state of education resources (sections \S\ref{sec:EDResource})  may be a counsel caution in this approach.
 It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
-Nevertheless, we believe it is wrong for a complete curriculum to ignore CyberSecurity issues and:
+Nevertheless, it is wrong for a complete curriculum to ignore CyberSecurity issues for example
 \begin{itemize}
@@
 	\item teach ``web programming'' without teaching Cross-Site Scripting \cite[(XSS)]{OWASP2017a};
-	\item teach \verb+strcpy+ etc. without teaching the necessity of guards.
+	\item impact of GDPR upon systems development
+	\item the use of Agile Development to develop secure systems
 \end{itemize}
 \par
-We commend the BCS-identified good practice of exploring cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
-\item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of cybersecurity teaching, and spreading good practice, and should continue this.
+The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly in PCI DSS is commended.(\S\ref{sec:PCIDSS}).
+\item[RQ3] Can accreditation by PSRBs enhance the provision of cybersecurity within a bodies jurisdiction?
+
+Accreditation (as practised by BCS) is a valuable tool in improving the standard of cybersecurity teaching, and spreading good practice, and should continue this. 
 \end{description}
@@
 \item Web Programming courses should probably do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
+\item A study similar to \cite{Drop2019} is recommended for future work in Web Programming, Agile Development, Android Development and potentially other areas of the curriculum.
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.

### 2019-06-19 FIE2019-final-TPJHD.tex
A couple of further typos fixed

@@
 A multitude of national and international model curricula and recommendations - from national academies, learned societies and even governments -- have been presented and discussed in recent years, with varying levels of impact on policy and practice.
-This paper address the key questions ``What cybersecurity is taught and what cybersecurity should be taught to the generalist Computer Science undergraduate students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to generalist Computer Science students?'' and ''Can accreditation by Professional, Statutory and Regulatory Bodies(PSRBs) enhance the provision of cybersecurity within a bodies jurisdiction?''.
+This paper address the key questions ``What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?'' and ''Can accreditation by Professional, Statutory and Regulatory Bodies(PSRBs) enhance the provision of cybersecurity within a bodies jurisdiction?''.
 
@@
 This is beyond the scope the scope of this paper.
-Instead a case study based research approach has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to undergraduate students.
+Instead a case study based research approach has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to students.
 Thus, in the context of widespread international computer science/engineering curriculum reform, %- both in compulsory, as well as post-compulsory education -
@@
 \item The general Computer Science graduate.
-\item The general Computer Science masters graduate.
+\item The generalist Computer Science masters graduate.
 \item The specialist Computer Science graduate.
@@
 \end{enumerate}
-The focus of this paper is on 2--3: the general Computer Science graduate.
+The focus of this paper is on 2-3: the general Computer Science graduate.
 \begin{description}
-\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the generalist Computer Science undergraduate students?
-\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to generalist Computer Science undergraduate students?
+\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science  students?
+\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
 \item[RQ3]Can accreditation by PSRBs enhance the provision of cybersecurity within a bodies jurisdiction?
@@
 \begin{description}
-\item[RQ1] - What cybersecurity is taught and what cybersecurity should be taught to the generalist Computer Science undergraduate students?
+\item[RQ1] - What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?
 
@@
 However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
-\item[RQ2] - Should cybersecurity be taught stand-alone or in an integrated manner to generalist Computer Science undergraduate students?
+\item[RQ2] - Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
 
@@
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item Web Programming courses should probably do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
+\item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
 \item A study similar to \cite{Drop2019} is recommended for future work in Web Programming, Agile Development, Android Development and potentially other areas of the curriculum.

### 2019-06-19 FIE2019-final-TPJHDsaved.tex
Push on references

--- 2019-06-14 FIE2019-final-TP.tex
@@
 We have instead adopted a case study based research approach to evaluate the evidence of the teaching of cybersecurity within general computer science to undergraduate students.
-Thus, in the context of widespread international computer science/engineering curriculum reform - both in compulsory, as well as post-compulsory education - what does this need to embed cybersecurity mean more generally for institutions and educators, and how do we teach this curricula area?
+Thus, in the context of widespread international computer science/engineering curriculum reform, %- both in compulsory, as well as post-compulsory education -
+what does this need to embed cybersecurity mean more generally for institutions and educators, and how do we teach this subject?
 
-In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by analysing the UK's national security, economic, and skills policy context, shaping the practical considerations for education professionals from schools through to universities (and into industry and professional practice).
-Through this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
+In this paper, we frame a key social, cultural and economic global challenge of cybersecurity in computing and engineering education by dThrough this analysis of UK educational policy and practice, we make a number of recommendations for future cybersecurity educational initiatives and interventions, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
 \end{abstract}
@@
 
-The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
+The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some
+%JHD:  was 2/3, now 5/7
+ (5 final, 7 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
 
-Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.
+Is teaching cybersecurity different?
+Lecturing is probably not the best way.
+Should we use real-life case studies (e.g.
+British Airways \cite{Barth2018a}).
+\cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.
 
 %AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
-  This raises pedagogic questions which are beyond the scope of this research to resolve. Should we be teaching content because it's underpinning?; or because there's a skill shortage?
+  % Should we remove this?
+This raises pedagogic questions which are beyond the scope of this research to resolve.
+Should we be teaching content because it's underpinning?; or because there's a skill shortage?
 

### 2019-06-19 FIE2019-final-TPJHD.tex
Shortened school section and abstract

@@
 There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
-A multitude of national and international model curricula and recommendations - from national academies, learned societies and even governments -- have been presented and discussed in recent years, with varying levels of impact on policy and practice.
-This paper address the key questions ``What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?'' and ''Can accreditation by Professional, Statutory and Regulatory Bodies(PSRBs) enhance the provision of cybersecurity within a bodies jurisdiction?''.
+A multitude of national and international model curricula and recommendations
+%- from national academies, learned societies and even governments -- 
+have been presented and discussed in recent years, with varying levels of impact on policy and practice.
+This paper address the key questions ``What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?'' and ''Can accreditation by Professional, Statutory and Regulatory Bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?''.
 
@@
 
-The paper proposes a number of recommendations for future cybersecurity educational initiatives, interventions and research, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
+%The paper proposes a number of recommendations for future cybersecurity educational initiatives, interventions and research, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
 \end{abstract}
@@
 
-With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,brown-et-al:toce2014,wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,murphy-et-al:programming2017,simon-et-al:sigcse2018,tryfonas+crick:petra2018}.
+With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{%crick+sentance:2011,brown-et-al:sigcse2013,
+brown-et-al:toce2014,%wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,
+murphy-et-al:programming2017%,simon-et-al:sigcse2018,tryfonas+crick:petra2018
+}.
 There has been a recent Association for Computing Machinery (ACM)  working group as part of the Conference on Innovation and Technology in Computer Science Education \cite{Parrishetal2018a}, but this has yet to report.
@@
 
-\IEEEtriggeratref{51}
+\IEEEtriggeratref{44}
 

### 2019-06-19 FIE2019-final-TPJHD.tex
More shortening of policy

@@
 \subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity.
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large),%: \cite{JCNSS2018a} has only anecdotal evidence), 
+but also all in IT need to know \emph{some} cybersecurity.
 This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
@@
 This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
-\cite{UKCabinetOffice} or National Initiative for cybersecurity Education (NICE) in the USA \cite{NICE}.
+\cite{UKCabinetOffice} or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.
 
@@
 There has been substantial Computer Science /digital curriculum reform across the
-UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
+UK~\cite{%crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,
+	brown-et-al:toce2014}%,moller+crick:jce2018}
 --- but what has happened to the focus on cybersecurity?
@@
 
-UK national economic skills priority
-e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
+%This is basically note form: JHD suggests deleting
+%UK national economic skills priority
+%e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
 %(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
-(especially \cite{JCNSS2018a}
-and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
-\url{https://www.cybersecuritychallenge.org.uk/}).
+%(especially \cite{JCNSS2018a}
+%and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
+%\url{https://www.cybersecuritychallenge.org.uk/}).
 
-There is large-scale media
-attention on the ``cybersecurity skills gap''
-e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
-and
-\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
-etc -- link to UK Digital Strategy and UK Industrial Strategy.
+%There is large-scale media attention on the ``cybersecurity skills gap''
+%e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
+%and
+%\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
+%etc -- link to UK Digital Strategy and UK Industrial Strategy.
 
@@
 Hence \cite{Drop2019} analyzed the database textbooks used by  44 of the top 50 Computer Science
-departments in the United States (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed).
+departments in the United States. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
 There were seven such books, but three books accounted for the 36 of the 44 universities.

### 2019-06-20 FIE2019-final-TPJHD.tex
Improve table
Tweaked title and abstract

@@
 
-\title{Cybersecurity Is Too Important\\To Be Left To The Specialists}
+\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists}
 %\title{What Do We Mean By Cybersecurity Education?}  Tom's original
@@
     Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
-\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Education, 
+\IEEEauthorblockA{\IEEEauthorrefmark{1}%School of Education, 
 Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Dept.~Computer Science, 
+\IEEEauthorblockA{\IEEEauthorrefmark{2}%Dept.~Computer Science, 
 University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Faculty of Technology, 
+\IEEEauthorblockA{\IEEEauthorrefmark{3}%Faculty of Technology, 
 Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}Dept.~Computer\&Information Sciences, 
+\IEEEauthorblockA{\IEEEauthorrefmark{4}%Dept.~Computer\&Information Sciences, 
 Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
@@
 % submitted abstract (100-500 word limit):
-There are numerous facets to cybersecurity education, from theory to practice, hardware and software, to social and technical (as well as other important dimensions).
-A multitude of national and international model curricula and recommendations
-%- from national academies, learned societies and even governments -- 
-have been presented and discussed in recent years, with varying levels of impact on policy and practice.
-This paper address the key questions ``What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?'' and ''Can accreditation by Professional, Statutory and Regulatory Bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?''.
+There are numerous dimensions to cybersecurity education: from theory to practice, hardware and software, to social and politics/policy (as well as other important facets).
+A multitude of national and international model curricula and recommendations have been presented and discussed in recent years, with varying levels of impact on education and practice.
+This paper address the key questions: ``What cybersecurity is taught and what cybersecurity should be taught to general computer science students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?'' and ``Can accreditation by professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?''.
 
-Evaluating how cybersecurity is taught in all aspects of computer science is a  task of considerable size.
-This is beyond the scope the scope of this paper.
-Instead a case study based research approach has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to students.
-Thus, in the context of widespread international computer science/engineering curriculum reform, %- both in compulsory, as well as post-compulsory education -
-what does this need to embed cybersecurity mean more generally for institutions and educators, and how do we teach this subject?
+Evaluating how cybersecurity is taught in all aspects of computer science is clearly a task of considerable size, one that is beyond the scope of this paper.
+Instead a case study-based research approach has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students.
+Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity mean more generally for institutions and educators, and how can we teach this subject more effectively?
 
@@
 The timeline which this process has followed in Table \ref{table:1}.
-\begin{center}
+%\begin{center}
   \begin{table}[h!]
+  \caption{Timeline of the development of Cybersecurity expectations}
+  \label{table:1}
   \begin{tabular}{ | p{6cm} |p{1.5cm} |}
@@
     IV. ACCREDITATION:RQ3 - C. Accreditation - what progress has been made? &  \\ \hline
-    All visited HEIs expected to be fully compliant & September 2020 (BCS follow a 5-year Accreditation Cycle)  \\ \hline
+    All visited HEIs expected to be fully compliant 
+ (BCS follow a 5-year Accreditation Cycle)  & September 2020\\ \hline
   \end{tabular}
-  \caption{Timeline of the development of Cybersecurity expectations}
-  \label{table:1}
   \end{table}
-\end{center}
+%\end{center}
 
@@
 
-\IEEEtriggeratref{44}
+\IEEEtriggeratref{45}
 

### 2019-06-20 Response-TP.tex
updated to reflect all changes I have made

--- 2019-06-09 Response.tex
@@
 \item ``The research appears to be poorly structured and the analysis/argument is hard to interpret''.
-\item[A]We have done what we can, but note that the other referees disagree.
+\item[A1]The paper has been re-written to  make the structure clearer.
+In particular, the abstract has been updated to more succinctly provide the essence of the work; increased the amount of signposting to guide the reader; added a research approach section to more clearly explain the approach followed; made the research questions more precise to communicate the intentions more clearly; and made the sections more self contained to better support the reading of the paper in a non-linear manner.
 \item ``The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.''
-\item[A]
+\item[A] The abstract is now two paragraphs long and has been enhanced to better present the essence of the work.
 \item What do the authors mean by ``There has been a recent international working group but this has yet to report.'' 
-\item[A]
+\item[A] In parallel to the research explored in the paper, their is a working group as part of the Association for Computing Machinery (ACM) Conference on Innovation and Technology in Computer Science Education intending to develop a taxonomy of cybersecurity education.
+This has been clarified.
 \item It would add more clarity to introduce what these terms DevOps vs DevSecOps represent in the paper.
 Also, please use the acronym GDPR after introducing it during the first usage such as General Data Protection Regulation (GDPR).
-\item[A]
+\item[A] Discussion related to DevOps vs DevSecOps has been reduced. Abbreviations have adopted the style recommended. 
 \item There are repeated ``and'' in the sentence ``In particular the expectation of Privacy by design...'' Also, the ``by'' after ``and'' is not required.
@@
 
-\item[A]
+\item[A] These sentences have been rephrased and other typographical errors addressed.
+
 \item Please rephrase the following statement ``How might accreditation regard cybersecurity education, or help with it?'' as the sentence structure doesn't make much sense. 
-\item[A]
+\item[A] This has been rephrased to "Can accreditation by Professional, Statutory and Regulatory Bodies enhance the provision of cybersecurity within a bodies jurisdiction?" The other research questions have also been rephrased to improve the clarity by which the intentions of the research are communicated.
 \item It is not clear what is the message the authors are conveying in the 2nd paragraph?
@@
 
-\item[A]
+\item[A] The authors recognize these where note sections that should not have been present in the paper.
+These sections have been removed.
+The recommended approach to abbreviations has now been adopted throughout the paper.
 \item  What is JHD? Please expand?
-\item[A]
+\item[A] This has been removed this. 
 \item Please expand OWASP. 
-\item[A]
+\item[A]Open Web Application Security Project has replaced the initial use of OWASP.
 \item 
 The references to first author in the 2nd paragraph could be avoided and the sentence rephrased to highlight only the facts and not any personal experience. 
-\item[A]
+\item[A] The references to the authors has been removed throughout.
 \item 
  What does ``three accounted for the 36 of the 44'' mean? Please clarify. I am assuming the 36 and 44 refer to number of universities. Please add that clarification, if it is so.
-\item[A]
+\item[A] This did mean universities and has been updated to indicate this.
 \item 
@@
 Instead a smattering of anecdotal information and quotes (and even more quotes in the footnotes) intended to convince the reader of a viewpoint that is never quite articulated clearly - until the end when it comes down to a small set of specific complaints about SQL Injection, XSS, and StackOverflow (which is overall a bit of a letdown as I felt as if this bloglike writing was building to a more impactful crescendo, some kind of kick-in-the-pants call to action for the community).''
-\item[A]
+\item[A] The paper has been revised to more clearly communicate the research approach employed.
+A research approach section has been added to further elicit the case study approach adopted.
+The research questions have been made more specific to more clearly communicate the intentions of the research.
+The writing style has been adjusted and a more formal approach taken.
+The conclusions have been extended and clarified.
 \end{enumerate}

### 2019-06-20 FIE2019-final-TPJHD.tex
A copy of typos resolved
bodies - body's

@@
 \item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
-\item[RQ3]Can accreditation by PSRBs enhance the provision of cybersecurity within a bodies jurisdiction?
+\item[RQ3]Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
 \end{description}
@@
 \end{description}
-\section{Policy Context}
+\subsection{Policy Context}
 
@@
 The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly in PCI DSS is commended.(\S\ref{sec:PCIDSS}).
-\item[RQ3] Can accreditation by PSRBs enhance the provision of cybersecurity within a bodies jurisdiction?
+\item[RQ3] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
 

### 2019-06-20 FIE2019-final-TPJHD.tex
Fix '82' bug, should have been 820!

@@
 Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
-Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
+Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
 Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students use of such Informal Resources.

### 2019-06-20 FIE2019-final-TPJHD.tex
A bit more shortening

@@
 
-Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
-It needs attention across the spectrum:
+Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, such as \cite{BritishAirways2018a}.
+It needs attention across the spectrum \cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}:
 \begin{quote}
 change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
-\cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}
 \end{quote}
-
+\noindent
 With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{%crick+sentance:2011,brown-et-al:sigcse2013,

### 2019-06-21 FIE2019-final-TPJHD.tex
Added author affiliations

--- 2019-06-20 FIE2019-final-TPJHD.tex
@@
 % }
-\ifanon
-\author{Anonymous}
-\else
+% \ifanon
+% \author{Anonymous}
+% \else
 % compact IEEE author format for more than three authors...
@@
 Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
-\fi
+%\fi
 \maketitle
@@
 \item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
-\item[RQ3]Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
+\item[RQ3]Can accreditation by PSRBs enhance the provision of cybersecurity within a bodies jurisdiction?
 \end{description}
@@
 \end{description}
-\subsection{Policy Context}
+\section{Policy Context}
 
@@
 The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly in PCI DSS is commended.(\S\ref{sec:PCIDSS}).
-\item[RQ3] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
+\item[RQ3] Can accreditation by PSRBs enhance the provision of cybersecurity within a bodies jurisdiction?
 

### 2019-06-21 Response-TP.tex

--- 2019-06-09 Response.tex
@@
 \item ``The research appears to be poorly structured and the analysis/argument is hard to interpret''.
-\item[A]We have done what we can, but note that the other referees disagree.
+\item[A1]The paper has been re-written to  make the structure clearer.
+In particular, the abstract has been updated to more succinctly provide the essence of the work;  the signposting has been enhanced to better guide the reader;  a research approach section to more clearly explain the approach followed has been added;  the research questions have been made more precise to communicate the intentions more clearly; and the sections have been made more self contained to better support the reading of the paper in a non-linear manner.
 \item ``The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.''
-\item[A]
+\item[A] The abstract is now two paragraphs long and  better presents the essence of the work.
 \item What do the authors mean by ``There has been a recent international working group but this has yet to report.'' 
-\item[A]
+\item[A] In parallel to the research explored in the paper, there is a working group as part of the Association for Computing Machinery (ACM) Conference on Innovation and Technology in Computer Science Education intending to develop a taxonomy of cybersecurity education.
+This has been clarified in the paper
 \item It would add more clarity to introduce what these terms DevOps vs DevSecOps represent in the paper.
 Also, please use the acronym GDPR after introducing it during the first usage such as General Data Protection Regulation (GDPR).
-\item[A]
+\item[A] Discussion related to DevOps vs DevSecOps has been removed. Abbreviations have adopted the style recommended. 
 \item There are repeated ``and'' in the sentence ``In particular the expectation of Privacy by design...'' Also, the ``by'' after ``and'' is not required.
@@
 
-\item[A]
+\item[A] These sentences have been rephrased and other typographical errors addressed.
+
 \item Please rephrase the following statement ``How might accreditation regard cybersecurity education, or help with it?'' as the sentence structure doesn't make much sense. 
-\item[A]
+\item[A] This has been rephrased to "Can accreditation by Professional, Statutory and Regulatory Bodies enhance the provision of cybersecurity within a body's jurisdiction?" The other research questions have also been rephrased to improve the clarity by which the intentions of the research are communicated.
 \item It is not clear what is the message the authors are conveying in the 2nd paragraph?
@@
 
-\item[A]
+\item[A] The authors recognize these were note sections that should not have been present in the paper.
+These sections have been removed.
+The recommended approach to abbreviations has now been adopted throughout the paper.
 \item  What is JHD? Please expand?
-\item[A]
+\item[A] This has been removed this. 
 \item Please expand OWASP. 
-\item[A]
+\item[A]Open Web Application Security Project has replaced the initial use of OWASP.
 \item 
 The references to first author in the 2nd paragraph could be avoided and the sentence rephrased to highlight only the facts and not any personal experience. 
-\item[A]
+\item[A] The references to the authors has been removed throughout and any potential conjecture removed.
 \item 
  What does ``three accounted for the 36 of the 44'' mean? Please clarify. I am assuming the 36 and 44 refer to number of universities. Please add that clarification, if it is so.
-\item[A]
+\item[A] This did mean universities and has been updated to indicate this.
 \item 

### 2019-06-21 FIE2019-final-TPJHD.tex
Slght tidy on page 1

--- 2019-06-20 FIE2019-final-TPJHD.tex
@@
 The intention of this working group is to develop a taxonomy for approaches to cybersecurity education.
-The third aim of this working group is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
+Their third aim %of this working group is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
 

### 2019-06-21 FIE2019-final-TPJHD.tex
And fix a bug in this!
Numerous tidies; no title change

@@
 The intention of this working group is to develop a taxonomy for approaches to cybersecurity education.
-Their third aim %of this working group is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
+Their third aim %of this working group
+is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
 
@@
 \subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large),%: \cite{JCNSS2018a} has only anecdotal evidence), 
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), %: \cite{JCNSS2018a} has only anecdotal evidence), 
 but also all in IT need to know \emph{some} cybersecurity.
@@
 
-A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}.
+A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}.
 This further emphasizes the need for mainstream software developers and by extension Computer Science graduates to be versed in relevant security practices.
@@
 Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
-This would seem to be particularly the case with ``agile'' processes \cite{Bartsch2011a}.
 
@@
 \item[UK]The BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
-More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
+More precisely, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
 %BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
@@
 The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
-There have been a number of international initiatives international to define curricula to support this for example Computer Science Programmes  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+There have been a number of international initiatives international to define curricula to support this for example the ACM general curriculum  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
 
@@
 
-The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterization, but a long way from being a textbook (which is not their aim).
+The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
+knowledge areas, which are useful references for the experienced educator looking for a definition or characterization, but a long way from being a textbook (which is not their aim).
 
@@
 The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}.
-These standard are linked to the ACM recommendations for curricula.
-To date the major focus of NSSC accreditation has been upon Masters degrees specializing in cybersecurity.
+These standards are linked to the ACM recommendations for curricula \cite{ACM2013a}.
+To date the major focus of NCSC accreditation has been upon Masters degrees specializing in cybersecurity.
 More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and Computer Science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
@@
 These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
-Both the IET and The BCS have a long history of expecting coverage environmental factors within the programmes they accredit.
+Both the IET and The BCS have a long history of expecting coverage of environmental factors within the programmes they accredit.
 The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}.
@@
 	\item teach ``web programming'' without teaching Cross-Site Scripting \cite[(XSS)]{OWASP2017a};
-	\item impact of GDPR upon systems development
-	\item the use of Agile Development to develop secure systems
+	\item ignore the  impact of GDPR upon systems development;
+	\item uncritically teach the use of Agile Development to develop secure systems.
 \end{itemize}
 \par
-The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly in PCI DSS is commended.(\S\ref{sec:PCIDSS}).
+The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended.(\S\ref{sec:PCIDSS}).
 \item[RQ3] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
@@
 
-\IEEEtriggeratref{45}
+\IEEEtriggeratref{44}
 

### 2019-06-22 FIE2019-final-TPJHD.tex
Tweaks to abstract

@@
 % submitted abstract (100-500 word limit):
-There are numerous dimensions to cybersecurity education: from theory to practice, hardware and software, to social and politics/policy (as well as other important facets).
-A multitude of national and international model curricula and recommendations have been presented and discussed in recent years, with varying levels of impact on education and practice.
+There are numerous dimensions to cybersecurity education: hardware and software, from theory to practice (and vice versa), to social and politics/policy (as well as other important facets).
+A multitude of national and international model curricula and recommendations have been presented and discussed in recent years, with varying levels of impact on education, policy and practice.
 This paper address the key questions: ``What cybersecurity is taught and what cybersecurity should be taught to general computer science students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?'' and ``Can accreditation by professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?''.

### 2019-06-22 FIE2019-final-TPJHD.tex
Changes to title, abstract and intro

@@
 
-\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists}
+\title{Cybersecurity Education and Accreditation:\\A Case Study of the UK}
+%\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
 %\title{What Do We Mean By Cybersecurity Education?}  Tom's original
@@
 % submitted abstract (100-500 word limit):
-There are numerous dimensions to cybersecurity education: hardware and software, from theory to practice (and vice versa), to social and politics/policy (as well as other important facets).
+There are numerous dimensions to cybersecurity education: the interaction of hardware and software, from theory to practice (and vice versa), to human factors, policy and politics (as well as other important facets).
 A multitude of national and international model curricula and recommendations have been presented and discussed in recent years, with varying levels of impact on education, policy and practice.
@@
 \begin{IEEEkeywords}
-Cybersecurity, accreditation, curricula, computer science education, UK
+Cybersecurity, accreditation, curricula, computer science education, public policy, UK
 \end{IEEEkeywords}
@@
 
-Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, such as \cite{BritishAirways2018a}.
+% TC: other cyber breach examples here? e.g. Talktalk, Experian, madison dating thing, military examples?
+Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, such as~\cite{BritishAirways2018a}.
 It needs attention across the spectrum \cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}:
 \begin{quote}
-change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
+``{\emph{...change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''
 \end{quote}
 \noindent
-With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{%crick+sentance:2011,brown-et-al:sigcse2013,
-brown-et-al:toce2014,%wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,
-murphy-et-al:programming2017%,simon-et-al:sigcse2018,tryfonas+crick:petra2018
-}.
-There has been a recent Association for Computing Machinery (ACM)  working group as part of the Conference on Innovation and Technology in Computer Science Education \cite{Parrishetal2018a}, but this has yet to report.
-The intention of this working group is to develop a taxonomy for approaches to cybersecurity education.
-Their third aim %of this working group
-is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
+These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
+As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and driven by training materials, which make it harder to create a common body of knowledge.
+An international perspective is still harder, as different nations use different criteria to define local needs~\cite{schneider2013}.
+As a result, new programmes entering this space are on their own to conceptualise, design, package and market their programs, as there is no globally accepted reference model for cybersecurity to allow employers or students to understand the extent of a given cybersecurity program~\cite{conklin-et-al:2014,Parrishetal2018a}.
+
+% TC: check most recent versions of ACM curricula
+% TC: check whether the ITiCSE report will be out this year?
+With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a}.
+This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
+There has been a recent ACM working group as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, but this has yet to report.
+The intention of this working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programs; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
 
@@
 
-\subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), %: \cite{JCNSS2018a} has only anecdotal evidence), 
-but also all in IT need to know \emph{some} cybersecurity.
-This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
-In this section four case studies are briefly explored: General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
+\subsection{Cybersecurity: for all or for specialists?}
+%: \cite{JCNSS2018a} has only anecdotal evidence),
+
+In one sense, this section title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014}.
+But this is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.
+Recent evolution of cybersecurity shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
+More recent work presents cybersecurity as a ``meta-discipline''; that is, cybersecurity is used as an aggregate label for a wide variety of similar disciplines, much in the same way that the terms ``engineering'' and ``computing'' are commonly used~\cite{Parrishetal2018b}.
+
+In this section four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
 
@@
 
-%AI: GDPR essentially expects CyberSecurity to be in all coding. 
+%AI: GDPR essentially expects Cybersecurity to be in all coding. 
 In particular, for the development of software that will operate within Europe and that processes personal data, the expectation of ``Privacy by Design'' and ``Privacy by Default'' have significant legal implications.
@@
 \subsection{Current situation}
-The ACM/Institute of Electrical and Electronics Engineers (IEEE)-CS Joint Task Force on Computing Curricula,\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas(KAs) related to RQ2:
+The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
 \begin{quote}
@@
 The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
-There has  been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettreick2013}.
+There has  been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
 The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
@@
 % inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
+% PPP e.g. Carr work and IEE Comp ref, Conklin papder
+% Cyber Essentials Scheme
 
-There has been substantial Computer Science /digital curriculum reform across the
-UK~\cite{%crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,
-	brown-et-al:toce2014}%,moller+crick:jce2018}
---- but what has happened to the focus on cybersecurity?
+There has been substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,moller+crick:jce2018} --- but what has happened to the focus on cybersecurity?
 The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
@@
 %{\bf JHD to research ACM general curriculum;}
-The ACM general curriculum \cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
+The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
 
@@
 It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
-Nevertheless, it is wrong for a complete curriculum to ignore CyberSecurity issues for example
+Nevertheless, it is wrong for a complete curriculum to ignore cybersecurity issues for example
 \begin{itemize}

### 2019-06-22 FIE2019-final-TPJHD.tex
Improved discussion of CSRF advice

@@
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
+\item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
+Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks)} is silent.} that has Cross-Site Request Forgery protection.
 \item A study similar to \cite{Drop2019} is recommended for future work in Web Programming, Agile Development, Android Development and potentially other areas of the curriculum.
@@
 
-\IEEEtriggeratref{44}
+\IEEEtriggeratref{43}
 

### 2019-06-22 FIE2019-final-TPJHD.tex
More tweaks and changes

--- 2019-06-22 FIE2019-final-TPJHD.tex
@@
 
-\title{Cybersecurity Education and Accreditation:\\A Case Study of the UK}
+\title{The Future of Cybersecurity Education and Accreditation: A Case Study of the UK}
 %\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
@@
 Evaluating how cybersecurity is taught in all aspects of computer science is clearly a task of considerable size, one that is beyond the scope of this paper.
-Instead a case study-based research approach has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students.
+Instead a case study-based research approach, primarily focusing on the UK, has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students.
 Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity mean more generally for institutions and educators, and how can we teach this subject more effectively?
@@
 These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
-As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and driven by training materials, which make it harder to create a common body of knowledge.
-An international perspective is still harder, as different nations use different criteria to define local needs~\cite{schneider2013}.
-As a result, new programmes entering this space are on their own to conceptualise, design, package and market their programs, as there is no globally accepted reference model for cybersecurity to allow employers or students to understand the extent of a given cybersecurity program~\cite{conklin-et-al:2014,Parrishetal2018a}.
+As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge.
+An international perspective is harder still, as different nations use different criteria to define local needs~\cite{schneider2013}.
+As a result, new programmes entering this space are free to conceptualise, design, package and market their initiatives, as there is no globally accepted reference model for cybersecurity to allow employers or students to understand the extent or ambition of a given cybersecurity program~\cite{conklin-et-al:2014,Parrishetal2018a}.
 
@@
 Recent evolution of cybersecurity shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
-More recent work presents cybersecurity as a ``meta-discipline''; that is, cybersecurity is used as an aggregate label for a wide variety of similar disciplines, much in the same way that the terms ``engineering'' and ``computing'' are commonly used~\cite{Parrishetal2018b}.
+More recent work presents cybersecurity as a ``meta-discipline''; that is, cybersecurity should be used as an aggregate label for a wide variety of similar disciplines, much in the same way that the terms ``engineering'' and ``computing'' are commonly used~\cite{Parrishetal2018b}.
 
-In this section four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
+% this section feels someone disjoint from what we have written so far...
+To illustrate some of these challenges, four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
 
 On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
-It is hard to envisage one professioanl possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity.
-On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for expertise in this area.
+It is hard to envisage one professional possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing organisations to re-think models for governance and assurance, and to invest more in cybersecurity.
+On the other hard, the growing publicity given to these, and the growing use of electronic payments, has also led to greater demand for expertise in this area.
 % JHD what are these?
@@
 However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
+% can we evidence this?
 
 A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}.
-This further emphasizes the need for mainstream software developers and by extension Computer Science graduates to be versed in relevant security practices.
+This further emphasizes the need for mainstream software developers and by extension computer science graduates to be well-versed in relevant security practices.
 
@@
 \begin{enumerate}
-\item The person in the street --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
+\item The general public --- this is important, but there are many initiatives in this area (e.g.
+Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}, which are, rightly, largely separated from computing education.
 % ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
@@
 BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
-\item The general Computer Science graduate.
-\item The generalist Computer Science masters graduate.
-\item The specialist Computer Science graduate.
-\item The specialist Computer Science masters graduate.
+\item The general computer science graduate.
+\item The generalist computer science masters graduate.
+\item The specialist computer science graduate.
+\item The specialist computer science masters graduate.
 \end{enumerate}
@@
 \item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
-\item[RQ3]Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
+\item[RQ3]Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?
 \end{description}
@@
 \subsection{Research Approach}
-A case study based approach is adopted in this research.
-As is common in case study based research, many alternative case studies could have been chosen.
-The cases are illustrative rather than complete in terms of the available case studies.
-The cases where evaluated to articulate the progress made and highlight some opportunities for future work.
+A UK-focused case study-based approach is adopted in this research.
+As is common in case study-based research, many alternative case studies could have been chosen.
+The cases are illustrative rather than comprehensive in terms of the available case studies or challenges.
+The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.
 
@@
 
-In order to address RQ1 and RQ2 a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate Computer Science students are evaluated.
+In order to address RQ1 and RQ2 a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
 Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
 
-Finally in order to address RQ3 cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate Computer Science in one jurisdiction (the United Kingdom).
+Finally, in order to address RQ3 cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
 As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.
@@
 \begin{quote}
-The Information Assurance and Security KA is unique among the set of KAs presented here
-given the manner in which the topics are pervasive throughout other KAs
+{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
+given the manner in which the topics are pervasive throughout other KAs}}
 \end{quote}
@@
 \begin{quote}
-Universities suffer shortcomings, as well.
+{\emph{Universities suffer shortcomings, as well.
 Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity.
 There is a big catch, however.
-Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.
+Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}
 \end{quote}
@@
 But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: the paper quotes that 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping.
-As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publically available web documents.
+As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.
 
@@
 \begin{description}
-\item[UK]The BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
+\item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
 More precisely, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
@@
 \end{description}
-\subsection{Policy Context}
 
-For a few years now there has been a recognized need to build knowledge, skills and capacity in the area of cybersecurity.
+% move this section back to Intro?
+\section{Policy Context}
+
+% Add stuff about NCSC
+For a few years now there has been a recognised need to build knowledge, skills and capacity in the area of cybersecurity.
 This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
@@
 
+% should this be done at school?
 There has been substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,moller+crick:jce2018} --- but what has happened to the focus on cybersecurity?
@@
 Should we be teaching content because it's underpinning?; or because there's a skill shortage?
-
-% should we frame this as one or more case studies?
-% do we have any national survey data?
 
@@
 \section*{Acknowledgements}
-Thanks to Sally Pearce, Academic Accreditation Manager, BCS--The
-Chartered Institute for IT for supplying the summary information
-related to Accreditation. Many people, accreditors and accredited, contributed to  improving the standard of cybersecurity teaching, and spreading good practice.
+Thanks to Sally Pearce, Academic Accreditation Manager (BCS, The Chartered Institute for IT) for supplying the summary information related to Accreditation.
+Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
 
-\IEEEtriggeratref{44}
+%\IEEEtriggeratref{44}
 

### 2019-06-22 FIE2019-final-TPJHD.tex
Tidied comments on knowledge

--- 2019-06-22 FIE2019-final-TPJHD.tex
@@
 
+We demonstrate the positive effect that accreditation requirements can have, and give some recommendations both for research and for curriculum developments.
+
 %The paper proposes a number of recommendations for future cybersecurity educational initiatives, interventions and research, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
@@
 These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
-As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and driven by training materials, which make it harder to create a common body of knowledge.
+As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and driven by training materials, which make it harder to create a common body of knowledge (\cite{Bristol2019a} is still a work in progress).
 An international perspective is still harder, as different nations use different criteria to define local needs~\cite{schneider2013}.

### 2019-06-22 FIE2019-final-TPJHD.tex

--- 2019-06-22 FIE2019-final-TPJHD.tex
@@
 
-% Add stuff about NCSC
+% TC: mention NCSC
 For a few years now there has been a recognised need to build knowledge, skills and capacity in the area of cybersecurity.
@@
 
-%\IEEEtriggeratref{44}
+\IEEEtriggeratref{43}
 

### 2019-06-22 FIE2019-final-TPJHD.tex

@@
 These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
-As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge.
+As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge (for example, in the UK, \cite{Bristol2019a} is still a work in progress).
 An international perspective is harder still, as different nations use different criteria to define local needs~\cite{schneider2013}.

### 2019-06-22 FIE2019-final-TPJHD.tex
Some reorganisation

@@
 
-% this section feels someone disjoint from what we have written so far...
+%\subsection{Policy Context}
+
+This need to build knowledge, skills and capacity in the area of cybersecurity has led to the establishment of a number of initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{UKCabinetOffice} (and the setting up of the UK National Cyber Security Centre), or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.
+
+The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
+There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
+The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
+There have been a number of international initiatives international to define curricula to support this for example the ACM Computer Science Curricula Recommendations \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+
+% TC to add policy+skills context
+% inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
+% PPP e.g. Carr work and IEE Comp ref, Conklin papder
+% Cyber Essentials Scheme
+
+% should this be done at school?
+There has been substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,moller+crick:jce2018} --- but what has happened to the focus on cybersecurity?
+The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
+
+%This is basically note form: JHD suggests deleting
+%UK national economic skills priority
+%e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
+%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
+%(especially \cite{JCNSS2018a}
+%and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
+%\url{https://www.cybersecuritychallenge.org.uk/}).
+
+%There is large-scale media attention on the ``cybersecurity skills gap''
+%e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
+%and
+%\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
+%etc -- link to UK Digital Strategy and UK Industrial Strategy.
+
+% need a better title for this!
+\subsection{Some Examples}
+
 To illustrate some of these challenges, four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
@@
 
-\subsection{Research Questions}
+\section{Research Questions}
 There are various levels of specialism at which cybersecurity education can be addressed.
@@
 
-\subsection{Current situation}
+\subsection{Current State of Play}
 The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
@@
 \end{description}
-
-% move this section back to Intro?
-\section{Policy Context}
-
-% TC: mention NCSC
-For a few years now there has been a recognised need to build knowledge, skills and capacity in the area of cybersecurity.
-This need has led to the establishment of a number of initiatives from a number of national governments for example The UK Cyber Security Strategy.
-\cite{UKCabinetOffice} or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.
-
-The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
-There has  been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
-The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
-There have been a number of international initiatives international to define curricula to support this for example the ACM general curriculum  \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
-
-% TC to add policy+skills context
-% inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
-% PPP e.g. Carr work and IEE Comp ref, Conklin papder
-% Cyber Essentials Scheme
-
-% should this be done at school?
-There has been substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,moller+crick:jce2018} --- but what has happened to the focus on cybersecurity?
-The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
-
-%This is basically note form: JHD suggests deleting
-%UK national economic skills priority
-%e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
-%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
-%(especially \cite{JCNSS2018a}
-%and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
-%\url{https://www.cybersecuritychallenge.org.uk/}).
-
-%There is large-scale media attention on the ``cybersecurity skills gap''
-%e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
-%and
-%\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
-%etc -- link to UK Digital Strategy and UK Industrial Strategy.
 
@@
 
-\IEEEtriggeratref{43}
+%\IEEEtriggeratref{43}
 

### 2019-06-22 FIE2019-final-TPJHD.tex
Tweaks to abstract

@@
 % submitted abstract (100-500 word limit):
-There are numerous dimensions to cybersecurity education: the interaction of hardware and software, from theory to practice (and vice versa), to human factors, policy and politics (as well as other important facets).
+This Innovative Practice Full Paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as other important facets).
 A multitude of national and international model curricula and recommendations have been presented and discussed in recent years, with varying levels of impact on education, policy and practice.
@@
 
-We demonstrate the positive effect that accreditation requirements can have, and give some recommendations both for research and for curriculum developments.
+We demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.
 

### 2019-06-22 FIE2019-final-TPJHD.tex
New title
Tweaks to abstract

@@
 
-\title{The Future of Cybersecurity Education and Accreditation: A Case Study of the UK}
+\title{Cybersecurity Education and Accreditation:\\A Case Study of the UK}
 %\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
@@
 A multitude of national and international model curricula and recommendations have been presented and discussed in recent years, with varying levels of impact on education, policy and practice.
-This paper address the key questions: ``What cybersecurity is taught and what cybersecurity should be taught to general computer science students?'' , ``Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?'' and ``Can accreditation by professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?''.
+This paper address three key questions: {\emph{i)}} What is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?;  and {\emph{iii)}} Can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
 
@@
 \begin{IEEEkeywords}
-Cybersecurity, accreditation, curricula, computer science education, public policy, UK
+Cybersecurity, curricula, accreditation, computer science education, public policy, UK
 \end{IEEEkeywords}

### 2019-06-22 FIE2019-final-TPJHD.tex
Fixed typos

@@
 % inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
-% PPP e.g. Carr work and IEE Comp ref, Conklin papder
+% PPP e.g. Carr work and IEE Comp ref, Conklin paper
 % Cyber Essentials Scheme
@@
 A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}.
-This further emphasizes the need for mainstream software developers and by extension computer science graduates to be well-versed in relevant security practices.
+This further emphasises the need for mainstream software developers and by extension computer science graduates to be well-versed in relevant security practices.
 
@@
 The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
-knowledge areas, which are useful references for the experienced educator looking for a definition or characterization, but a long way from being a textbook (which is not their aim).
+knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
 
@@
 % Drop2019 was awarded the (unique) best paper at SIGCSE2019 - not sure what to make of this.
-Hence \cite{Drop2019} analyzed the database textbooks used by  44 of the top 50 Computer Science
+Hence \cite{Drop2019} analysed the database textbooks used by  44 of the top 50 Computer Science
 departments in the United States. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
@@
 These standards are linked to the ACM recommendations for curricula \cite{ACM2013a}.
-To date the major focus of NCSC accreditation has been upon Masters degrees specializing in cybersecurity.
+To date the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity.
 More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and Computer Science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
@@
 The ambition of this approach is to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area.)  The approach taken is not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
-In this context this is realized as an expectation cybersecurity is an inclusion in all degrees accredited by the BCS.
+In this context this is realised as an expectation cybersecurity is an inclusion in all degrees accredited by the BCS.
 e.g.
@@
 A number of HEIs are in the process of adjusting their curricula to meet the new expectations.
-In this case, the BCS notes the changes to programme design, the outputs from which will be scrutinized at the next accreditation visit.
+In this case, the BCS notes the changes to programme design, the outputs from which will be scrutinised at the next accreditation visit.
 
@@
 
+% TC: do we have big picture recommendations to make?
 \section{Conclusions}
-As regards the research questions;.
+As regards the research questions:
+
 \begin{description}
-\item[RQ1] - What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?
+\item[RQ1:] What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?
 
@@
 However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
-\item[RQ2] - Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
+\item[RQ2:] Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
 
@@
 The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended.(\S\ref{sec:PCIDSS}).
-\item[RQ3] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
+\item[RQ3:] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
 
-Accreditation (as practised by BCS) is a valuable tool in improving the standard of cybersecurity teaching, and spreading good practice, and should continue this. 
+Accreditation (as practised by BCS in the UK) is a valuable tool in improving the standard of cybersecurity teaching, and disseminating good practice, and should continue this. 
 \end{description}
 
-We have the following specific recommendations.
+% TC: do we have big picture recommendations to make?
+We have the following specific recommendations:
+
 \begin{enumerate}
@@
 
+% fix at end
 %\IEEEtriggeratref{43}

### 2019-06-22 FIE2019-final-TPJHD.tex
Added BCS roles
Fix sentence with two 'but'

@@
 In one sense, this section title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014}.
-But this is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.
+This is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.
 Recent evolution of cybersecurity shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
@@
 Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
+\par
+The first two authors are Vice-Presidents of BCS, and the third and fourth are past and present Chairs of the BCS Accreditation Committee.
 
 % fix at end
-%\IEEEtriggeratref{43}
+%\IEEEtriggeratref{40}
 

### 2019-06-22 FIE2019-final-TPJHD.tex
Added CCPA reference

@@
 
-On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
+On the one hand, GDPR (equally, the  California Consumer Privacy Act) has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
 It is hard to envisage one professional possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing organisations to re-think models for governance and assurance, and to invest more in cybersecurity.

### 2019-06-22 FIE2019-final-TPJHD.tex
Updated title and abstract

@@
 
-\title{Cybersecurity Education and Accreditation:\\A Case Study of the UK}
+\title{Cybersecurity Education and Accreditation:\\A UK Case Study}
 %\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
@@
 This Innovative Practice Full Paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as other important facets).
-A multitude of national and international model curricula and recommendations have been presented and discussed in recent years, with varying levels of impact on education, policy and practice.
+A multitude of model curricula and recommendations have been presented and discussed in international fora in recent years, with varying levels of impact on education, policy and practice.
 This paper address three key questions: {\emph{i)}} What is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?;  and {\emph{iii)}} Can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
@@
 Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity mean more generally for institutions and educators, and how can we teach this subject more effectively?
-
-We demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.
+Through this UK case study, we demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.
 

### 2019-06-22 FIE2019-final-TPJHD.tex
Further edits round PCIDSS

--- 2019-06-22 FIE2019-final-TPJHD.tex
@@
 This can be further illustrated by the demands of particular sectors for secure software.
-Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
+Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process, even after developers are trained annually in secure coding practices \cite[6.5]{PCI2018b}, cannot be trusted to deliver secure code.
 
@@
 \item The general public --- this is important, but there are many initiatives in this area (e.g.
-Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}, which are, rightly, largely separated from computing education.
+Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), which are, rightly, largely separated from computing education.
 % ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
@@
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
-ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots.
+ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots''.
 This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
@@
 By default, Spring implicitly enables protection against this.
-But all the accepted answers to CSRF-related failures simply suggested disabling they check.
+But all the accepted answers to CSRF-related failures simply suggested disabling the check.
 There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!

### 2019-06-22 FIE2019-final-TPJHD.tex
Minor tweaks
Removed 'undergraduate' for PCIDSS teaching

@@
 It is an interesting question as to whether standards such as PCI DSS should be addressed within degree courses (clearly degree courses can never cover all standards) or whether they should be addressed in professional training courses.
-However it appears the current situation is not ideal from the perspective of industry (or users of systems) and the inclusion of pertinent standards could be seen as a useful enhancement activity to the coverage of cybersecurity within generalist undergraduate computer science courses.
+However it appears the current situation is not ideal from the perspective of industry (or users of systems) and the inclusion of pertinent standards could be seen as a useful enhancement activity to the coverage of cybersecurity within generalist %undergraduate: JHD commented out, I don't think we mean this, and I teach it at MSc.
+computer science courses.
 

### 2019-06-22 FIE2019-final-TPJHD.tex
Added more examples in Intro

--- 2019-06-22 FIE2019-final-TPJHD.tex
@@
 
-\def\IntroYear{2014}  % when BCS started asking for CyberSecurity
+%\def\IntroYear{2014}  % when BCS started asking for CyberSecurity
 
@@
 
-% TC: other cyber breach examples here? e.g. Talktalk, Experian, madison dating thing, military examples?
-Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, such as~\cite{BritishAirways2018a}.
-It needs attention across the spectrum \cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}:
+Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, from credit reference agencies~\cite{equifax2017}, online dating websites~\cite{ashleymadison2015}, national airlines~\cite{BritishAirways2018a}, telecoms companies~\cite{talktalk2016}, and even between sovereign governments~\cite{ncsc2018}.
+These major breaches have had significant impact on both individual citizens and society in general, requiring attention across the spectrum for organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}:
 \begin{quote}

### 2019-06-22 FIE2019-final-TPJHD.tex
Restructured Intro and following section

@@
 Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, from credit reference agencies~\cite{equifax2017}, online dating websites~\cite{ashleymadison2015}, national airlines~\cite{BritishAirways2018a}, telecoms companies~\cite{talktalk2016}, and even between sovereign governments~\cite{ncsc2018}.
-These major breaches have had significant impact on both individual citizens and society in general, requiring attention across the spectrum for organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}:
+These major breaches have had significant impact on both individual citizens and society in general, requiring attention across the spectrum for organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: UK equivalent of NSA.}:
 \begin{quote}
-``{\emph{...change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''
+``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''
 \end{quote}
@@
 
-% TC: check most recent versions of ACM curricula
-% TC: check whether the ITiCSE report will be out this year?
 With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a}.
 This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
-There has been a recent ACM working group as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, but this has yet to report.
-The intention of this working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programs; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
+There has been a recent ACM working group as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, but this has yet to report its full findings.
+The intention of this working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
 
@@
 
-\subsection{Cybersecurity: for all or for specialists?}
-%: \cite{JCNSS2018a} has only anecdotal evidence),
+\section{Cybersecurity: for all or for specialists?}
 
-In one sense, this section title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014}.
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014}.
 This is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.
+
+This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK’s critical national infrastructure~\cite{lordscyberreport:2018}, as well as initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
+
+The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the computer science discipline for a number of years~\cite{Hentea2006,schneider2013}.
+There have been a number of international initiatives international to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
+The consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
 Recent evolution of cybersecurity shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
@@
 
-%\subsection{Policy Context}
+Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,tryfonas+crick:petra2018}, should cybersecurity be formally taught in schools, as part of a compulsory education?
+While aspects of ``e-safety'' and principles of protecting personal data are clearly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.
+The Institute of Coding~\cite{Davenportetal2019a}, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country, primarily focused on university graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
 
-This need to build knowledge, skills and capacity in the area of cybersecurity has led to the establishment of a number of initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{UKCabinetOffice} (and the setting up of the UK National Cyber Security Centre), or National Initiative for Cybersecurity Education (NICE) in the USA \cite{NICE}.
-
-The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the Computer Science discipline for a number of years \cite{Hentea2006}.
-There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
-The consensus now is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
-There have been a number of international initiatives international to define curricula to support this for example the ACM Computer Science Curricula Recommendations \cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
-
-% TC to add policy+skills context
-% inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
-% PPP e.g. Carr work and IEE Comp ref, Conklin paper
-% Cyber Essentials Scheme
-
-% should this be done at school?
-There has been substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,moller+crick:jce2018} --- but what has happened to the focus on cybersecurity?
-The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
-
-%This is basically note form: JHD suggests deleting
+% have merged into previous section
 %UK national economic skills priority
@@
 %\url{https://www.cybersecuritychallenge.org.uk/}). 
-
 %There is large-scale media attention on the ``cybersecurity skills gap''
@@
 % need a better title for this!
-\subsection{Some Examples}
+% \subsection{Some Examples}
 
-To illustrate some of these challenges, four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
+% To illustrate some of these challenges, four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
 
-On the one hand, GDPR (equally, the  California Consumer Privacy Act) has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
+% On the one hand, GDPR (equally, the  California Consumer Privacy Act) has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
 It is hard to envisage one professional possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing organisations to re-think models for governance and assurance, and to invest more in cybersecurity.
 On the other hard, the growing publicity given to these, and the growing use of electronic payments, has also led to greater demand for expertise in this area.
-% JHD what are these?
+% % JHD what are these?
 Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.
 
-%AI: GDPR essentially expects Cybersecurity to be in all coding. 
-In particular, for the development of software that will operate within Europe and that processes personal data, the expectation of ``Privacy by Design'' and ``Privacy by Default'' have significant legal implications.
+% %AI: GDPR essentially expects Cybersecurity to be in all coding. 
+% In particular, for the development of software that will operate within Europe and that processes personal data, the expectation of ``Privacy by Design'' and ``Privacy by Default'' have significant legal implications.
 However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
-% can we evidence this?
+% % can we evidence this?
 
-A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}.
+% A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}.
 This further emphasises the need for mainstream software developers and by extension computer science graduates to be well-versed in relevant security practices.
 
-This can be further illustrated by the demands of particular sectors for secure software.
+% This can be further illustrated by the demands of particular sectors for secure software.
 Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process, even after developers are trained annually in secure coding practices \cite[6.5]{PCI2018b}, cannot be trusted to deliver secure code.
@@
 \begin{enumerate}
-\item The general public --- this is important, but there are many initiatives in this area (e.g.
-Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), which are, rightly, largely separated from computing education.
+\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
 % ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
@@
     IV. ACCREDITATION:RQ3 - A. Developing Expectations &   \\ \hline
-    UK Government Cybersecurity Strategy \cite{UKCabinetOffice} & November 2011 \\ \hline
+    UK Government Cybersecurity Strategy \cite{ukcyberstrategy:2016} & November 2011 \\ \hline
     Three workshops of a consortium of industry, academia and government bodies - led by CPHC and (ISC)2 - leading to the development of Cybersecurity learning guidelines to be embedded into BCS accredited UK Computer Science and IT-related degree \cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
@@
 Internationally the expectations regarding both the breadth and depth of the expected cybersecurity coverage has been the subject of much discussion, debate and analysis.
-Like many Governments, the UK Government has been actively been seeking ways to address this \cite{UKCabinetOffice}, \cite{UKCabinetOffice2014}.
+Like many Governments, the UK Government has been actively been seeking ways to address this \cite{ukcyberstrategy:2016}, \cite{UKCabinetOffice2014}.
 In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.

### 2019-06-22 FIE2019-final.tex
Minor tweak to Acknow.

@@
 
-\title{Cybersecurity Is Too Important\\To Be Left To The Specialists}
+\title{Cybersecurity Education and Accreditation:\\A UK Case Study}
+%\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
 %\title{What Do We Mean By Cybersecurity Education?}  Tom's original
@@
     Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
-\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Education, 
+\IEEEauthorblockA{\IEEEauthorrefmark{1}%School of Education, 
 Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Dept.~Computer Science, 
+\IEEEauthorblockA{\IEEEauthorrefmark{2}%Dept.~Computer Science, 
 University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Faculty of Technology, 
+\IEEEauthorblockA{\IEEEauthorrefmark{3}%Faculty of Technology, 
 Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}Dept.~Computer\&Information Sciences, 
+\IEEEauthorblockA{\IEEEauthorrefmark{4}%Dept.~Computer\&Information Sciences, 
 Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
-\fi
+%\fi
 \maketitle
-
-\def\IntroYear{2014}  % when BCS started asking for CyberSecurity
 
@@
 % submitted abstract (100-500 word limit):
-There are numerous facets to cybersecurity education, from theory to
-practice, hardware and software, to social and technical (as well as
-other important dimensions). A multitude of national and international
-model curricula and recommendations - from national academies, learned
-societies and even governments -- have been presented and discussed in
-recent years, with varying levels of impact on policy and practice. In
-this paper we attempt to address the key questions ``what should the
-generalist computer scientist/engineer know'' and ``how well is that
-being taught''. (Cyber)security -- if thought about at all -- has
-historically been left to specialists, frequently viewed as a
-masters-level discipline; or left for development in professional
-practice, sometimes called ``information assurance'' in the UK and
-Europe, which worked, even if not well, in a sequential process model.
+This Innovative Practice Full Paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as other important facets).
+A multitude of model curricula and recommendations have been presented and discussed in international fora in recent years, with varying levels of impact on education, policy and practice.
+This paper address three key questions: {\emph{i)}} What is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?;  and {\emph{iii)}} Can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
 
-For example, in the world of software and systems development, we have
-seen the disruption of Agile (invented c.2001) and DevOps (invented
-c.2009), but not necessarily through the lens of security; however,
-more recently there has been a recognition of the need for change,
-exemplified by the emergence of ``DevSecOps''. If security is not to
-be left to the experts, then the generalist must know about it. Thus,
-in the context of widespread international computer
-science/engineering curriculum reform - both in compulsory, as well as
-post-compulsory education - what does this trend mean more generally
-for institutions and educators, and how do we teach it?
+Evaluating how cybersecurity is taught in all aspects of computer science is clearly a task of considerable size, one that is beyond the scope of this paper.
+Instead a case study-based research approach, primarily focusing on the UK, has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students.
+Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity mean more generally for institutions and educators, and how can we teach this subject more effectively?
+Through this UK case study, we demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.
 
-In this paper, we frame a key social, cultural and economic global
-challenge of cybersecurity in computing and engineering education by
-analysing the UK's national security, economic, and skills policy
-context, shaping the practical considerations for education
-professionals from schools through to universities (and into industry
-and professional practice). Through this analysis of UK educational
-policy and practice, we make a number of recommendations for future
-cybersecurity educational initiatives and interventions, including a
-number of innovative practical suggestions for educators and
-curriculum designers, with a view for potential adoption and adaption
-in other jurisdictions internationally.
+%The paper proposes a number of recommendations for future cybersecurity educational initiatives, interventions and research, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Cybersecurity, accreditation, curricula, computer science education, UK
+Cybersecurity, curricula, accreditation, computer science education, public policy, UK
 \end{IEEEkeywords}
@@
 
-Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches of one kind or another, such as \cite{BritishAirways2018a}.
-It needs attention across the spectrum:
+Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, from credit reference agencies~\cite{equifax2017}, online dating websites~\cite{ashleymadison2015}, national airlines~\cite{BritishAirways2018a}, telecoms companies~\cite{talktalk2016}, and even between sovereign governments~\cite{ncsc2018}.
+These major breaches have had significant impact on both individual citizens and society in general, requiring attention across the spectrum for organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: UK equivalent of NSA.}:
 \begin{quote}
-change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.
-\cite{Hannigan2019a}\footnote{Former Director of GCHQ: British equivalent of NSA.}
+``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''
 \end{quote}
+\noindent
+These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
+As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge (for example, in the UK, \cite{Bristol2019a} is still a work in progress).
+An international perspective is harder still, as different nations use different criteria to define local needs~\cite{schneider2013}.
+As a result, new programmes entering this space are free to conceptualise, design, package and market their initiatives, as there is no globally accepted reference model for cybersecurity to allow employers or students to understand the extent or ambition of a given cybersecurity program~\cite{conklin-et-al:2014,Parrishetal2018a}.
 
-With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, which it does both at the individual level and via recommended curricula \cite{ACM2013a} and professional accreditation requirements \cite{BCS2018a}; this is alongside a wider focus on computer science education reform, especially across the UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,brown-et-al:toce2014,wgictreview:2013,moller+crick:jce2018,davenport-et-al:latice2016,murphy-et-al:programming2017,simon-et-al:sigcse2018,tryfonas+crick:petra2018}.
-There has been a recent international working group \cite{Parrishetal2018a}, but this has yet to report.
-We note that part of its third aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these (see section \ref{sec:SQL}).
+With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a}.
+This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
+There has been a recent ACM working group as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, but this has yet to report its full findings.
+The intention of this working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
 
 Nevertheless, it is one thing to write national curricula, specifications and requirements, and
-another thing to deliver appropriate and relevant education and skills; furthermore, one could
-reasonably ask how well this is done in practice.
+another thing to deliver appropriate and relevant education and skills; furthermore, this paper asks how well this is done in practice.
 
-\subsection{CyberSecurity: for all or for specialists?}
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large: \cite{JCNSS2018a} has only anecdotal evidence), but also all in IT need to know \emph{some} cybersecurity, as the recent fashion for talking about DevSecOps rather than just DevOps exemplifies.
+\section{Cybersecurity: for all or for specialists?}
 
-This is not a brand-new concern: see \cite{Parr2014a} for concerns over five years ago, but it is a growing one.
-On the one hand, GDPR has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
-It is hard to envisage one person possessing all these attributes, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing boards to invest more in cybersecurity.
-On the other hard, the growing publicity given to these, and the growing use of electronic payment, has also led to greater demand for staff in this area.
-% JHD what are these?
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014}.
+This is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.
+
+This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK’s critical national infrastructure~\cite{lordscyberreport:2018}, as well as initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
+
+The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the computer science discipline for a number of years~\cite{Hentea2006,schneider2013}.
+There have been a number of international initiatives international to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
+The consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
+Recent evolution of cybersecurity shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
+More recent work presents cybersecurity as a ``meta-discipline''; that is, cybersecurity should be used as an aggregate label for a wide variety of similar disciplines, much in the same way that the terms ``engineering'' and ``computing'' are commonly used~\cite{Parrishetal2018b}.
+
+Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,tryfonas+crick:petra2018}, should cybersecurity be formally taught in schools, as part of a compulsory education?
+While aspects of ``e-safety'' and principles of protecting personal data are clearly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.
+The Institute of Coding~\cite{Davenportetal2019a}, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country, primarily focused on university graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
+
+% have merged into previous section
+%UK national economic skills priority
+%e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
+%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
+%(especially \cite{JCNSS2018a}
+%and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
+%\url{https://www.cybersecuritychallenge.org.uk/}). 
+%There is large-scale media attention on the ``cybersecurity skills gap''
+%e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
+%and
+%\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
+%etc -- link to UK Digital Strategy and UK Industrial Strategy.
+
+% need a better title for this!
+% \subsection{Some Examples}
+
+% To illustrate some of these challenges, four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
+
+% On the one hand, GDPR (equally, the  California Consumer Privacy Act) has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
+It is hard to envisage one professional possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing organisations to re-think models for governance and assurance, and to invest more in cybersecurity.
+On the other hard, the growing publicity given to these, and the growing use of electronic payments, has also led to greater demand for expertise in this area.
+% % JHD what are these?
 Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.
 
-%AI: GDPR essentially expects CyberSecurity to be in all coding.
+% %AI: GDPR essentially expects Cybersecurity to be in all coding. 
+% In particular, for the development of software that will operate within Europe and that processes personal data, the expectation of ``Privacy by Design'' and ``Privacy by Default'' have significant legal implications.
+However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
+% % can we evidence this?
 
-In particular the expectation of ``Privacy by Design'' and and by ``Privacy by Default'' has significant implications for all software development with a European context that processes personal data.
-Whilst both ``Privacy by Design'' and ``Privacy by Default'' have been expected good practices for a number of years, these expectations now have a legal dimension, at least in Europe.
+% A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}.
+This further emphasises the need for mainstream software developers and by extension computer science graduates to be well-versed in relevant security practices.
 
-However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
-A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426}.
-This further emphasizes the need for mainstream software developers and by extension Computer Science graduates to be versed in relevant security practices.
+% This can be further illustrated by the demands of particular sectors for secure software.
+Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process, even after developers are trained annually in secure coding practices \cite[6.5]{PCI2018b}, cannot be trusted to deliver secure code.
 
-The Payments Industry \cite[6.6]{PCI2018b} insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process even after developers are trained in secure coding practices \cite[6.5]{PCI2018b} cannot be trusted to deliver secure code.
-This would seem to be particularly the case with ``agile'' processes \cite{Bartsch2011a}.
-
-Specialist curricula abound: a recent one is \cite{ACMIEEEAISSIGSECIFIP}.
-However, given the shortage of specialists, in practice a good generalist with some cybersecurity expertise can often get a ``specialist'' job.
-\subsection{Research Questions}
+\section{Research Questions}
 There are various levels of specialism at which cybersecurity education can be addressed.
 \begin{enumerate}
-\item The person in the street --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
+\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
 % ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
@@
 BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
-\item The general CS graduate.
-\item The general CS masters graduate.
-\item The specialist CS graduate.
-\item The specialist CS masters graduate.
+\item The general computer science graduate.
+\item The generalist computer science masters graduate.
+\item The specialist computer science graduate.
+\item The specialist computer science masters graduate.
 \end{enumerate}
-The focus of this paper is on 2--3: the general CS graduate.
+The focus of this paper is on 2-3: the general Computer Science graduate.
 \begin{description}
-\item[RQ1]What should be taught to the generalist, and how?
-\item[RQ2]Should this be taught stand-alone or integrated? 
-\item[RQ3]How might accreditation regard cybersecurity Education, or help with it?
+\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science  students?
+\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
+\item[RQ3]Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?
 \end{description}
-\cite[p.~97]{ACM2013a} takes a distinct view on RQ2:
+
+\subsection{Research Approach}
+A UK-focused case study-based approach is adopted in this research.
+As is common in case study-based research, many alternative case studies could have been chosen.
+The cases are illustrative rather than comprehensive in terms of the available case studies or challenges.
+The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.
+
+The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations.
+This is followed by the evaluation of the policy context.
+Together these provide a context to the ongoing enhancement initiatives in the area of cybersecurity education.
+
+In order to address RQ1 and RQ2 a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
+Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
+
+Finally, in order to address RQ3 cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
+As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.
+
+\subsection{Current State of Play}
+The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
 \begin{quote}
-The Information Assurance and Security KA is unique among the set of KAs presented here
-given the manner in which the topics are pervasive throughout other Knowledge Areas.
+{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
+given the manner in which the topics are pervasive throughout other KAs}}
 \end{quote}
-It proposes 9 ``core'' hours and 63.5 distributed across the other Knowledge Areas.
+It proposes 9 ``core'' hours and 63.5 distributed across the other KAs.
 %\pagebreak - to fix URL issues
 \par
-Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, writes as follows
+Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, indicates
 \begin{quote}
-Universities suffer shortcomings, as well.
+{\emph{Universities suffer shortcomings, as well.
 Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity.
 There is a big catch, however.
-Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.
+Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}
 \end{quote}
 
-But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: he quotes that 61\% of UK courses offer mandatory cybersecurity content, and his research was based on web scraping, it represents a lower bound.
+But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: the paper quotes that 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping.
+As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.
 
@@
 \begin{description}
-\item[UK]BCS has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
-More precisely since 2010, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
+\item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
+More precisely, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
 %BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
-ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots.
+ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots''.
 This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
@@
 \end{description}
-\section{Policy Context}
 
-% TC to add policy+skills context
-% inc. new EU cybersecurity Act: https://ec.europa.eu/commission/news/cybersecurity-act-2018-dec-11_en
-
-There has been substantial CS/digital curriculum reform across the
-UK~\cite{crick+sentance:2011,brown-et-al:sigcse2013,wgictreview:2013,brown-et-al:toce2014,moller+crick:jce2018}
---- but what has happened to the focus on cybersecurity?
-The Institute of Coding \cite{Davenportetal2019a}, designed to improve digital skills for UK graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
-
-UK national economic skills priority
-e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
-%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
-(especially \cite{JCNSS2018a}
-and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
-\url{https://www.cybersecuritychallenge.org.uk/}).
-
-There is large-scale media
-attention on the ``cybersecurity skills gap''
-e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
-and
-\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
-etc -- link to UK Digital Strategy and UK Industrial Strategy.
-
-\section{Challenges}
+\section{Challenges: RQ1 and RQ2}
 
@@
 %{\bf JHD to research ACM general curriculum;}
-The ACM general curriculum \cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
+The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
 
-The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (2 final, 3 for comment out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
-
-Is teaching cybersecurity different? Lecturing is probably not the best way. Should we use real-life case studies (e.g. British Airways \cite{Barth2018a}).  % JHD - need's fixing.
+The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
+knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
 
 %AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
-  Should we be teaching it because it's underpinning? because there's a skill shortage?  JHD would argue both.
+  % Should we remove this?
+This raises pedagogic questions which are beyond the scope of this research to resolve.
+Should we be teaching content because it's underpinning?; or because there's a skill shortage?
 
-% should we frame this as one or more case studies?
-% do we have any national survey data?
+\subsection{Industry standards - Payments Card Industry Data Security Standards (PCI DSS)}\label{sec:PCIDSS}
+Is teaching cybersecurity different?
+Lecturing is probably not the best way.
+Should we use real-life case studies (e.g.
+British Airways \cite{Barth2018a}).
+\cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.
+There are a number of ways a more practical manner could be action, one would be by the inclusion of pertinent cybersecurity standards in the curricula.
 
-\subsection{PCI DSS}\label{sec:PCIDSS}
-The Payments Card Industry Data Security Standards \cite{PCI2018b} underpin all processing of credit/debit cards.
-Nevertheless, they are very rarely mentioned in generalist computer scientist courses --- one payments industry person commented ``I've given up even asking if recruits have heard of PCI DSS: it's so rare''.
-
-This would not matter so much if everyone handling payments data were sent by their employers on an effective\footnote{``Effective'' is important: the Processing Director at one major acquirer commented ``I just caught two developers in the elevator discussing how to meet a customer requirement, and I had to tell them this was contrary to PCI DSS.
-Their response was that the customer wanted it!''.} PCI DSS course.
-However, the payments business is now so spread across websites, often run by SMEs, or non-specialists, that some provision should be made in courses.
+The PCI DSS \cite{PCI2018b} is one such standard.
+PCI DSS underpins all processing of credit/debit cards.
+Nevertheless, they are very rarely mentioned in generalist computer scientist courses.
+This would not matter so much if everyone handling payments data were sent by their employers on an effective PCI DSS course.
+However, the payments business is now so spread across websites, often run by small and medium enterprises (SME), or non-specialists.
 Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.
 
-%Alastair to write about their Credit Fraud experience (MSc cybersecurity).  HE Academy grant (to be referenced).
+It is an interesting question as to whether standards such as PCI DSS should be addressed within degree courses (clearly degree courses can never cover all standards) or whether they should be addressed in professional training courses.
+However it appears the current situation is not ideal from the perspective of industry (or users of systems) and the inclusion of pertinent standards could be seen as a useful enhancement activity to the coverage of cybersecurity within generalist %undergraduate: JHD commented out, I don't think we mean this, and I teach it at MSc.
+computer science courses.
 
-\subsection{Educational Resources: SQL Injection}\label{sec:SQL}
+\subsection{Educational Resources}\label{sec:EDResource}
+\subsubsection{Structured Query Language (SQL) Injection}\label{sec:SQL}
 It is 15 years since \cite{Guimaraesetal2004} 
@@
 
-Clearly such a major weakness should be well-taught.
-One would like to think it is, but the first author has yet to meet a UK undergraduate who recalls being taught it \emph{outside} a specialist cybersecurity course.
-In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recomended textbooks.
+Clearly such a major weakness should be well-taught.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recommended textbooks.
 
 % Drop2019 was awarded the (unique) best paper at SIGCSE2019 - not sure what to make of this.
-Hence \cite{Drop2019} analyzed the database textbooks used by  44 of the top 50 Computer Science
-departments in the United States (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed).
-There were seven such books, but three accounted for the 36 of the 44.
+Hence \cite{Drop2019} analysed the database textbooks used by  44 of the top 50 Computer Science
+departments in the United States. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
+There were seven such books, but three books accounted for the 36 of the 44 universities.
 Five of the seven (30 of the 44) had no mention of SQL injection.
-On the other two, the more popular one has a seriously flawed discussion\footnote{``However, they imply that using parameters is equivalent to using a function to add escape characters
+On the other two, the more popular one has a seriously flawed discussion\footnote{``However, the paper implies that using parameters is equivalent to using a function to add escape characters
 around user input. This is incorrect, as using parameters allows
@@
 \par
-This blindness is not limited to textbooks: although Wikipedia has an excellent article on SQL Injection, it was not linked from the SQL page itself.\footnote{The authors intend to fix this after the paper is processed, so as not to break anonymity.}
-\subsection{Informal Resources}\label{sec:informal}
-The web abounds with informal resources: tutorials and code snippets.
-How good are these, and how good are people at using these?
-This has been looked at by \cite{Unruhetal2017a}, who took the top five results from Google for six queries.
-Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
-Searching for these fragments in PHP projects on GitHub found 82 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
-
-\subsection{The case of Java}\label{sec:Java}
-To the best of the authors' knowledge, no survey equivalent to \cite{Drop2019} has been done for Java textbooks.
+This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
+\subsubsection{The case of Java}\label{sec:Java}
+A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for  Java textbooks.
 Indeed, many such books go nowhere near security applications.
@@
 
-\subsection{Android}\label{sec:Android}
-\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. Their key finding was this.
+\subsubsection{Android}\label{sec:Android}
+A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks.
+\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
+The key finding was this.
 \begin{quote}
@@
 None of the unprimed ones implemented any security.
+\subsubsection{Informal Resources}\label{sec:informal}
+The web abounds with informal resources: tutorials and code snippets.
+How good are these, and how good are people at using these?
+This has been looked at by \cite{Unruhetal2017a}, This research took the top five results from Google for six queries.
+Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
+Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
+Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students use of such Informal Resources.
+
 \subsection{Staff}
@@
 The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}.
-These standard are linked to the ACM recommendations for curricula.
-To date the major focus of NSSC accreditation has been upon Masters degrees specializing in cybersecurity.
+These standards are linked to the ACM recommendations for curricula \cite{ACM2013a}.
+To date the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity.
 More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and Computer Science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
@@
 
+% TC: do we have big picture recommendations to make?
 \section{Conclusions}
-As regards our research questions, we can make the following comments.
+As regards the research questions:
+
 \begin{description}
-\item[RQ1]The guidelines from both ACM and BCS are good for general education.
-However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}--) resources.
-\item[RQ2]The recommendation in \cite[p.
+\item[RQ1:] What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?
+
+The guidelines from both ACM and BCS are good for general education.
+However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
+\item[RQ2:] Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
+
+The recommendation in \cite[p.
 98]{ACM2013a} that cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea.
-Indeed, a recent discussion on cybersecurity had the aerospace engineers amazed that security was a separate topic: ``for us, safety is present everywhere''.
-However, in the current state of education resources (sections \S\ref{sec:SQL}--\ref{sec:Java}) this may be a counsel of perfection.
+However, in the current state of education resources (sections \S\ref{sec:EDResource})  may be a counsel caution in this approach.
 It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
-Nevertheless, we believe it is wrong for a complete curriculum to ignore CyberSecurity issues and:
+Nevertheless, it is wrong for a complete curriculum to ignore cybersecurity issues for example
 \begin{itemize}
-	\item teach SQL without teaching SQL Injection (\cite{Drop2019} and the second author's experience);
+	\item teach SQL without teaching SQL Injection (\cite{Drop2019} );
 	\item teach ``web programming'' without teaching Cross-Site Scripting \cite[(XSS)]{OWASP2017a};
-	\item teach \verb+strcpy+ etc. without teaching the necessity of guards.
+	\item ignore the  impact of GDPR upon systems development;
+	\item uncritically teach the use of Agile Development to develop secure systems.
 \end{itemize}
 \par
-We commend the BCS-identified good practice of exploring cybersecurity via a project, possibly in PCI DSS (\S\ref{sec:PCIDSS}).
-\item[RQ3]Accreditation (as practised by BCS) is a valuable tool in improving the standard of cybersecurity teaching, and spreading good practice, and should continue this.
+The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended.(\S\ref{sec:PCIDSS}).
+\item[RQ3:] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
+
+Accreditation (as practised by BCS in the UK) is a valuable tool in improving the standard of cybersecurity teaching, and disseminating good practice, and should continue this. 
 \end{description}
 
-We have the following specific recommendations.
+% TC: do we have big picture recommendations to make?
+We have the following specific recommendations:
+
 \begin{enumerate}
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item Web Programming courses should probably do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.} that has Cross-Site Request Forgery protection.
+\item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
+Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks)} is silent.} that has Cross-Site Request Forgery protection.
+\item A study similar to \cite{Drop2019} is recommended for future work in Web Programming, Agile Development, Android Development and potentially other areas of the curriculum.
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
@@
 \section*{Acknowledgements}
-Thanks to Sally Pearce, Academic Accreditation Manager, BCS--The
-Chartered Institute for IT for supplying the summary information
-related to Accreditation. Many people, accreditors and accredited, contributed to  improving the standard of cybersecurity teaching, and spreading good practice.
+Thanks to Sally Pearce, Academic Accreditation Manager (BCS, The Chartered Institute for IT) for supplying the summary information related to accreditation of UK degree programmes.
+Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
+\par
+{\emph{N.B.}} The first two authors are current Vice-Presidents of BCS, The Chartered Institute for IT, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee.
 
-\IEEEtriggeratref{51}
+% fix at end
+%\IEEEtriggeratref{40}
 

### 2019-06-22 FIE2019-final.tex
Minor tweaks

@@
 
-Cybersecurity has increasingly been in the news over recent years, generally prompted by spectacular breaches, from credit reference agencies~\cite{equifax2017}, online dating websites~\cite{ashleymadison2015}, national airlines~\cite{BritishAirways2018a}, telecoms companies~\cite{talktalk2016}, and even between sovereign governments~\cite{ncsc2018}.
-These major breaches have had significant impact on both individual citizens and society in general, requiring attention across the spectrum for organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: UK equivalent of NSA.}:
+Cybersecurity has increasingly been a headline feature in the news over recent years, generally prompted by spectacular breaches, from major credit reference agencies~\cite{equifax2017}, telecoms companies~\cite{talktalk2016}, national airlines~\cite{BritishAirways2018a}, online dating websites~\cite{ashleymadison2015}, and even between sovereign governments~\cite{ncsc2018}.
+These major breaches have had significant impact on both individual citizens and society in general, requiring attention from organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: UK equivalent of NSA.}:
 \begin{quote}
@@
 This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
-There has been a recent ACM working group as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, but this has yet to report its full findings.
+
+A recent ACM working group, as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, has been capturing global perspectives on cybersecurity education, but has yet to report its full findings.
 The intention of this working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
-
-Nevertheless, it is one thing to write national curricula, specifications and requirements, and
-another thing to deliver appropriate and relevant education and skills; furthermore, this paper asks how well this is done in practice.
+Nevertheless, it is one thing to write national curricula, specifications and requirements, and another thing to deliver appropriate and relevant education and skills; furthermore, this paper asks how well this is done in practice.
 
@@
 
-This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK’s critical national infrastructure~\cite{lordscyberreport:2018}, as well as initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
+This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK's critical national infrastructure~\cite{lordscyberreport:2018}, as well as initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
 
 The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the computer science discipline for a number of years~\cite{Hentea2006,schneider2013}.
-There have been a number of international initiatives international to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a} and specialised cybersecurity Programmes \cite{ACMIEEEAISSIGSECIFIP}.
+There have been a number of international initiatives international to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a}, as well as specialised cybersecurity degree programmes~\cite{ACMIEEEAISSIGSECIFIP}.
 There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.

### 2019-06-22 FIE2019-final.tex
New title
Tweaks to section 2

@@
 
-\title{Cybersecurity Education and Accreditation:\\A UK Case Study}
+\title{A UK Case Study on\\Cybersecurity Education and Accreditation}
 %\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
@@
 These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
-As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge (for example, in the UK, \cite{Bristol2019a} is still a work in progress).
+As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge (for example, in the UK, the CyBOK project~\cite{Bristol2019a} is still a work in progress).
 An international perspective is harder still, as different nations use different criteria to define local needs~\cite{schneider2013}.
@@
 With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a}.
-This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
+This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
 
 A recent ACM working group, as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, has been capturing global perspectives on cybersecurity education, but has yet to report its full findings.
-The intention of this working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
+An aim of the ITiCSE working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
 Nevertheless, it is one thing to write national curricula, specifications and requirements, and another thing to deliver appropriate and relevant education and skills; furthermore, this paper asks how well this is done in practice.
@@
 
-This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK's critical national infrastructure~\cite{lordscyberreport:2018}, as well as initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}.}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
+This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK's critical national infrastructure~\cite{lordscyberreport:2018}; also initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
 
-The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the computer science discipline for a number of years~\cite{Hentea2006,schneider2013}.
-There have been a number of international initiatives international to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a}, as well as specialised cybersecurity degree programmes~\cite{ACMIEEEAISSIGSECIFIP}.
+The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the discipline of computer science for a number of years~\cite{Hentea2006,schneider2013}.
+There have been a number of international initiatives to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a}, as well as for specialised cybersecurity degree programmes~\cite{ACMIEEEAISSIGSECIFIP}.
 There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
-The consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within Computer Science and related degrees.
-Recent evolution of cybersecurity shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
+The consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within computer science and related degrees.
+Recent evolution of cybersecurity education shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
 More recent work presents cybersecurity as a ``meta-discipline''; that is, cybersecurity should be used as an aggregate label for a wide variety of similar disciplines, much in the same way that the terms ``engineering'' and ``computing'' are commonly used~\cite{Parrishetal2018b}.
 
-Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,tryfonas+crick:petra2018}, should cybersecurity be formally taught in schools, as part of a compulsory education?
-While aspects of ``e-safety'' and principles of protecting personal data are clearly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.
-The Institute of Coding~\cite{Davenportetal2019a}, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country, primarily focused on university graduates, does indeed mention cybersecurity, but merely as a sub-item in one work package (1.2a).
+Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of a compulsory education.
+While aspects of ``e-safety'' and principles of protecting personal data are increasingly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national extra-curricular competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.
+The Institute of Coding~\cite{Davenportetal2019a}, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country -- but primarily focused on university graduates -- does indeed mention cybersecurity, but merely as a sub-item in one work package.
 

### 2019-06-22 FIE2019-final.tex
Tidying up of language

@@
 There have been a number of international initiatives to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a}, as well as for specialised cybersecurity degree programmes~\cite{ACMIEEEAISSIGSECIFIP}.
-There has been a debate as to whether cybersecurity is distinct discipline from Computer Science \cite{McGettrick2013}.
+There has been a debate as to whether cybersecurity is distinct discipline from computer science \cite{McGettrick2013}.
 The consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within computer science and related degrees.
@@
 \end{enumerate}
-The focus of this paper is on 2-3: the general Computer Science graduate.
+The focus of this paper is on 2-3: the general computer science graduate.
 \begin{description}
-\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science  students?
-\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
+\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?
+\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?
 \item[RQ3]Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?
@@
 \subsection{Research Approach}
-A UK-focused case study-based approach is adopted in this research.
+A UK-focused case study-based approach is adopted in this project.
 As is common in case study-based research, many alternative case studies could have been chosen.
@@
 % Drop2019 was awarded the (unique) best paper at SIGCSE2019 - not sure what to make of this.
-Hence \cite{Drop2019} analysed the database textbooks used by  44 of the top 50 Computer Science
-departments in the United States. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
+Hence \cite{Drop2019} analysed the database textbooks used by  44 of the top 50 computer science
+departments in the USA. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
 There were seven such books, but three books accounted for the 36 of the 44 universities.
@@
 
-In the United Kingdom as in most jurisdictions, Higher Education provision addresses general Computer Science  and specialist cybersecurity courses.
-A significant number of undergraduate and postgraduate programmes are available in both the areas of Computer Science and cybersecurity (and closely related fields Computer Security, Digital / Computer Forensics, etc).
-In the UK, Universities and Colleges Admissions Service (UCAS) lists over 40 Higher Education Institutes (HEIs) providing Undergraduate qualifications related to cybersecurity for entry in September 2019.
-An even larger number of HEIs provide study opportunities related to more general Computer Science.
-UCAS lists 246 provides for undergraduate programmes related to Computer Science.
+In the UK as in most jurisdictions, higher education provision addresses general computer science and specialist cybersecurity courses.
+A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields computer security, digital/computer forensics, etc).
+In the UK, the Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions (HEIs) providing undergraduate qualifications related to cybersecurity for entry in September 2019.
+An even larger number of HEIs provide study opportunities related to more general computer science.
+UCAS lists 246 provides for undergraduate programmes related to computer science.
 
-Accreditation has evolved to directly addresses the cybersecurity challenges in both general Computer Science programmes and specialist cybersecurity programmes.
+Accreditation has evolved to directly addresses the cybersecurity challenges in both general computer science programmes and specialist cybersecurity programmes.
 In the UK, accreditation in the broad computing area is being performed by a few different agencies.
@@
 \textbf{1. Not for Profit Organisations}
-Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements.
+Tech Partnership Degrees is a not-for-profit organisation that provides endorsements to higher education programmes with specific curricula elements aimed at job market requirements.
 One of the curricula elements is related to cybersecurity.
 Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.
-Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and 5 Software Engineering for Business Programmes.
-As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education and none upon specialist cybersecurity education.
+Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and five Software Engineering for Business Programmes.
+As such, Tech Partnership Degrees currently have limited impact upon more general computer science education and none upon specialist cybersecurity education.
 
@@
 To date the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity.
-More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and Computer Science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
+More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and computer science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
 
 Currently accredited are 15 cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters cybersecurity programmes and 1 cybersecurity Degree Programme with a further 2 provisionally accredited.
-Hence the extent of accreditation is currently reasonably limited in terms of reach to Computer Science programmes.
+Hence the extent of accreditation is currently reasonably limited in terms of reach to computer science programmes.
 This appears to be a positive initiative that will hopefully further develop over time.
@@
 \textbf{3. Professional Bodies}
-The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of Computer Science and the more specialist area of cybersecurity discipline areas.
+The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas.
 The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}.
-These memoranda support the internationalising of the curriculum and promote consistency and parity in Computer Science Education globally.
+These memoranda support the internationalising of the curriculum and promote consistency and parity in computer science education globally.
 These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
@@
 In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
-In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of Computer Science Departments) and the Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
-Three workshops in 2013 and 2015 attempted to define the principles of cybersecurity education and proposed a framework for embedding these principles in UK Computing Science curricula.
-Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate Computing Science degrees.
+In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of UK computer science departments) and the Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
+Three workshops in 2013 and 2015 attempted to define the principles of cybersecurity education and proposed a framework for embedding these principles in UK computer science curricula.
+Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate computer science degrees.
 This work initially lead to a white paper related to a proposal in the form of a white paper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}.
@@
 
-The produced reference guidelines ("Cybersecurity Principles and Learning Outcomes") \cite{CPHCISC2} established a baseline of common knowledge, example learning outcome domains for cybersecurity within the Computing Science courses and guidance on embedding the concepts.
+The produced reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'') \cite{CPHCISC2} established a baseline of common knowledge, example learning outcome domains for cybersecurity within the computer science courses and guidance on embedding the concepts.
 The document provides specific guidance for embedding and enhancing relevant cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.
@@
 e.g.
-the expectation for coverage is true for Computer Science as well as cybersecurity programmes.
+the expectation for coverage is true for computer science as well as cybersecurity programmes.
 Two criteria are expected to be covered by all programmes seeking accreditation.
@@
 The agreed Cybersecurity Principles and Learning Outcome \cite{CPHCISC2} were discussed with the wider education community by a road-show led by CPHC.
-A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of Computing Science degrees.
-The workshops included case studies from universities who had embedded cybersecurity into their Computer Science curricula illustrating different approaches to implementation.
-The workshops had 102 attendees from the academic Computing Science community representing 60 UK Universities.
+A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of computer science degrees.
+The workshops included case studies from universities who had embedded cybersecurity into their computer science curricula illustrating different approaches to implementation.
+The workshops had 102 attendees from the academic computer science community representing 60 UK Universities.
 
@@
 \begin{description}
-\item[RQ1:] What cybersecurity is taught and what cybersecurity should be taught to the general Computer Science students?
+\item[RQ1:] What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?
 
@@
 However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
-\item[RQ2:] Should cybersecurity be taught stand-alone or in an integrated manner to general Computer Science students?
+\item[RQ2:] Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?
 

### 2019-06-22 FIE2019-final.tex
More tidying up of language

@@
 A multitude of model curricula and recommendations have been presented and discussed in international fora in recent years, with varying levels of impact on education, policy and practice.
-This paper address three key questions: {\emph{i)}} What is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?;  and {\emph{iii)}} Can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
+This paper address three key questions: {\emph{i)}} What is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students;  and {\emph{iii)}} Can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
 
@@
 %{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
-\section{Accreditation by PSRB: RQ3}
+\section{Accreditation by PSRBs: RQ3}
 
@@
 
-\textbf{1. Not for Profit Organisations}
+\textbf{1. Not-for-Profit Organisations}
 Tech Partnership Degrees is a not-for-profit organisation that provides endorsements to higher education programmes with specific curricula elements aimed at job market requirements.
@@
 
-The Institute of Coding is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK \cite{Davenportetal2019a}.
+The Institute of Coding is a not-for-profit entity, led by the University of Bath, that intends to enhance how digital skills are developed in higher education in the UK~\cite{Davenportetal2019a}.
 This is likely to include cybersecurity related skills.
-Like the Tech Partnership the focus is upon job market requirements.
+Like the Tech Partnership, the focus is upon job market requirements.
 Additionally the Institute is looking at potentially endorsing the demonstrable capabilities of graduates as shown by their university studies, work experience and work placements.

### 2019-06-22 FIE2019-final.tex
More tidying up of language

@@
 
-The UK's official knowledge resource, the CyBOK project \cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
+The UK's official body of knowledge resource, the CyBOK project\cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
 knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
 
-%AI noted the workshops with ISC$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
+%AI noted the workshops with (ISC)$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
   % Should we remove this?
@@
     UK Government Cybersecurity Strategy \cite{ukcyberstrategy:2016} & November 2011 \\ \hline
-    Three workshops of a consortium of industry, academia and government bodies - led by CPHC and (ISC)2 - leading to the development of Cybersecurity learning guidelines to be embedded into BCS accredited UK Computer Science and IT-related degree \cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
+    Three workshops of a consortium of industry, academia and government bodies - led by CPHC and  (ISC)$^2$ - leading to the development of Cybersecurity learning guidelines to be embedded into BCS accredited UK Computer Science and IT-related degree \cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
     UK Government report Cybersecurity Skills, Business Perspectives and Government's Next Steps Report Released \cite{UKCabinetOffice2014} & March 2014  \\ \hline
     Council of Professors and Heads of Computing (CPHC) Identifies Cybersecurity as one the top 3 concerns in Computing & April 2014 \\ \hline
-    Joint Development of White Paper from CPHC and The International Information Systems Security Certification Consortium(ISC)2 \cite{CPHCISC2014} & April -November 2014 \\ \hline
+    Joint Development of White Paper from CPHC and The International Information Systems Security Certification Consortium (ISC)$^2$ \cite{CPHCISC2014} & April -November 2014 \\ \hline
     Extended Cybersecurity Criteria included in BCS Accreditation Guidelines \cite{BCS2018a}& June 2015 \\
@@
 In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
-In 2013, an initiative was set up by (ISC)2, CPHC (the representative body of UK computer science departments) and the Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
+In 2013, an initiative was set up by (ISC)$^2$, CPHC (the representative body of UK computer science departments) and the Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
 Three workshops in 2013 and 2015 attempted to define the principles of cybersecurity education and proposed a framework for embedding these principles in UK computer science curricula.

### 2019-06-22 FIE2019-final.tex
More tidying up of language
New sentence at end of RQ3 in conclusions
Tweaked abstract to match

@@
 \documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
+%\IEEEoverridecommandlockouts
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
@@
 Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity mean more generally for institutions and educators, and how can we teach this subject more effectively?
-Through this UK case study, we demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.
+Through this UK case study, and by contrasting with the US, we demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.
 
@@
 Accreditation (as practised by BCS in the UK) is a valuable tool in improving the standard of cybersecurity teaching, and disseminating good practice, and should continue this.
+The time-lag in adoption of cybersecurity in the US seems to be caused by the accreditation differences more than any other factors.
 \end{description}

### 2019-06-22 FIE2019-final.tex
Text in III
Further at end of IB-B-5

@@
 
-It is at least plausible to attribute this difference to differences in the accreditation regimes.
+It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and learned society curricula are all similar.
 \begin{description}
@@
 Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
-Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students use of such Informal Resources.
+Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students' (and indeed others') use of such Informal Resources.
 

### 2019-06-23 FIE2019-final.tex
More tweaks

@@
 
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014}.
+In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014,davenport-et-al:latice2016}.
 This is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.
@@
 
-Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of a compulsory education.
+Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{wgictreview:2013,brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of a compulsory education.
 While aspects of ``e-safety'' and principles of protecting personal data are increasingly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national extra-curricular competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.

### 2019-06-23 FIE2019-final.tex
Loads of changes and tweaks

@@
 
-\section{Cybersecurity: for all or for specialists?}
+\section{Cybersecurity: For All, Or For Specialists?}
 
@@
 \section{Research Questions}
-There are various levels of specialism at which cybersecurity education can be addressed.
-\begin{enumerate}
+
+There are various levels of specialism at which cybersecurity education can be addressed:
+\begin{enumerate}[label=(\roman*)]
 \item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
@@
 BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
-\item The general computer science graduate.
+\item The generalist computer science graduate.
 \item The generalist computer science masters graduate.
@@
 \end{enumerate}
-The focus of this paper is on 2-3: the general computer science graduate.
+The focus of this paper is on (ii)--(iii): the general computer science graduate.
+
+This paper thus focuses on three research questions:
+
 \begin{description}
-\item[RQ1]What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?
-\item[RQ2]Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?
-\item[RQ3]Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?
+\item[RQ1] {\emph{What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?}}
+\item[RQ2] {\emph{Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?}}
+\item[RQ3] {\emph{Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?}}
 \end{description}
@@
 \subsection{Research Approach}
+
 A UK-focused case study-based approach is adopted in this project.
@@
 
-In order to address RQ1 and RQ2 a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
+In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
 Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
 
-Finally, in order to address RQ3 cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
+Finally, in order to address {\emph{RQ3}} cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
 As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.
@@
 \subsection{Current State of Play}
+
 The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
@@
 \par
-Nevertheless, the situation on the ground is different: \cite{Ackerman2019a}, describing the USA, indicates
+Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:
 \begin{quote}
@@
 
-But \cite[Table 1]{Ruiz2019a} shows that the UK situation is distinctly different: the paper quotes that 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping.
+The UK situation is distinctly different: 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping~\cite[Table 1]{Ruiz2019a}.
 As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.
 
-It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and learned society curricula are all similar.
+It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and professional body/learned society curricula are all similar.
+
 \begin{description}
 \item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
-More precisely, accredited degrees have been expected to demonstrate coverage of ``2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information  systems'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
+More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
 %BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
 \item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
-ABET is, and is requiring IAS with effect from the 2019-20 cycle (self study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``The computing topics must include: \dots{} Principles and practices for secure computing \dots''.
+ABET is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''.
 This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
@@
 
-\section{Challenges: RQ1 and RQ2}
+\section{Challenges: {\emph{RQ1}} and {\emph{RQ2}}}
 
@@
 %{\bf JHD to research ACM general curriculum;}
-The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states 3 Tier-1 and 6 Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
+The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
 
-The UK's official body of knowledge resource, the CyBOK project\cite{Bristol2019a}, has produced reference documentation for some (5  final, 7 for comment out of a planned total of 19)   % JHD updated figures 21/6/2019
+The UK's official body of knowledge resource, the CyBOK project\cite{Bristol2019a}, has produced reference documentation for some (as of June 2019: five final, seven for comment, out of a planned total of 19)   % JHD updated figures 21/6/2019
 knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
@@
 
-\subsection{Industry standards - Payments Card Industry Data Security Standards (PCI DSS)}\label{sec:PCIDSS}
+\subsection{Industry Standards: PCI DSS}\label{sec:PCIDSS}
+
 Is teaching cybersecurity different?
@@
 
-The PCI DSS \cite{PCI2018b} is one such standard.
+The Payments Card Industry Data Security Standards (PCI DSS)~\cite{PCI2018b} is one such standard.
 PCI DSS underpins all processing of credit/debit cards.
@@
 \subsection{Educational Resources}\label{sec:EDResource}
-\subsubsection{Structured Query Language (SQL) Injection}\label{sec:SQL}
+
+\subsubsection{SQL Injection}\label{sec:SQL}
+
 It is 15 years since \cite{Guimaraesetal2004} 
@@
 
-Clearly such a major weakness should be well-taught.  In general it's hard to determine what is taught, but a reasonable proxy for this is the content of recommended textbooks.
+Clearly such a major weakness should be well-taught.  In general it is hard to determine what is taught, but a reasonable proxy for this is the content of recommended textbooks.
 
@@
 Five of the seven (30 of the 44) had no mention of SQL injection.
-On the other two, the more popular one has a seriously flawed discussion\footnote{``However, the paper implies that using parameters is equivalent to using a function to add escape characters
+On the other two, the more popular one has a seriously flawed discussion\footnote{``{\emph{However, the paper implies that using parameters is equivalent to using a function to add escape characters
 around user input. This is incorrect, as using parameters allows
@@
 recommended as a sole defense since imperfect escape functions
-can easily be subverted.'' \cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``However, the fact that the first
+can easily be subverted.}}''~\cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``{\emph{However, the fact that the first
 example should not be used is not discussed until two pages after
@@
 without being aware that this is in fact an example of what they
-shouldn't do.'' \cite{Drop2019}}.
+shouldn't do.}}''~\cite{Drop2019}}.
 \par
 This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
-\subsubsection{The case of Java}\label{sec:Java}
+
+\subsubsection{The Case of Java}\label{sec:Java}
+
 A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for  Java textbooks.
@@
 \par
-This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was
+This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was:
+
 \begin{quote}
-644 out of the 1,429 inspected answer posts
+{\emph{644 out of the 1,429 inspected answer posts
 (45\%) are insecure, meaning that insecure suggestions
 popularly exist on SO. Insecure answers dominate, in
-particular, the SSL/TLS category [355 insecure versus 150 secure, i.e. $>70$\%].
+particular, the SSL/TLS category}} [355 insecure versus 150 secure, i.e. $>70$\%].
 \end{quote}
@@
 \subsubsection{Android}\label{sec:Android}
-A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks.
-\cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
-The key finding was this.
+
+A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks; \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
+The key finding was this:
 \begin{quote}
-We found that 15.4\% of all 1.3 million Android applications
+{\emph{We found that 15.4\% of all 1.3 million Android applications
 contained security-related code snippets from
 Stack Overflow. Out of these 97.9\% contain at least one
-insecure code snippet.
+insecure code snippet.}}
 \end{quote}
+
 Two caveats (in opposite directions) should be noted.
@@
 The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}.
-Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``Improve documentation: Clarify what you can(not) copy/paste''.
+Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``{\emph{Improve documentation: Clarify what you can(not) copy/paste}}''.
+
 \subsubsection{Agile}
+
 A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for ``Agile'' textbooks.
@@
 None of the unprimed ones implemented any security.
+
 \subsubsection{Informal Resources}\label{sec:informal}
+
 The web abounds with informal resources: tutorials and code snippets.
@@
 This has been looked at by \cite{Unruhetal2017a}, This research took the top five results from Google for six queries.
-Of these 30 tutorials, 6 had SQL injection weaknesses, and 3 had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
+Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
 Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
-Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students' (and indeed others') use of such Informal Resources.
+Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students' (and indeed others') use of such informal resources.
 
 \subsection{Staff}
-It is well known that cybersecurity skills are in short supply, for example \cite{Page2018a}.
+
+It is well known that cybersecurity skills are in short supply~\cite{schneider2013}. For example:
 \begin{quote}
 % JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
-Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
+{\emph{Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
 In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
 This had climbed to 51\% at the beginning of this year.
-Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
+Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.}}~\cite{Page2018a}
 \end{quote}
 The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
-%JHD possibly to write about his recruitment experiences.  ? more IoC input?
 
-In the United Kingdom there is a prominence of job adverts to recruit academic staff with specialisms in cybersecurity.
+In the UK, there is a prominence of job adverts to recruit academic staff with specialisms in cybersecurity.
 The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
 As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.
-A recent example had a professor of cybersecurity, two senior academic positions and 2 junior academic positions in one advert.
-There are examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year.
-There are also examples of  cybersecurity subject groups moving en masse from one university to another.
+Recent example have included a professor of cybersecurity, two senior academic positions and two junior academic positions in one advert.
+There are other examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year; there are also examples of cybersecurity subject groups moving en masse from one university to another.
 
-%Not sure if we want to raise something about funding of posts ??
-%Poaching of staff ??
+%Not sure if we want to raise something about funding of posts? Poaching of staff ??
 %Is this too contentious
+%TC: probably out of scope for this paper
 
 %{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
-\section{Accreditation by PSRBs: RQ3}
+\section{Accreditation by PSRBs: {\emph{RQ3}}}
 
@@
 A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields computer security, digital/computer forensics, etc).
-In the UK, the Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions (HEIs) providing undergraduate qualifications related to cybersecurity for entry in September 2019.
-An even larger number of HEIs provide study opportunities related to more general computer science.
+In the UK, the Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions providing undergraduate qualifications related to cybersecurity for entry in September 2019.
+An even larger number of institutions provide study opportunities related to more general computer science.
 UCAS lists 246 provides for undergraduate programmes related to computer science.
@@
 \textbf{1. Not-for-Profit Organisations}
+
 Tech Partnership Degrees is a not-for-profit organisation that provides endorsements to higher education programmes with specific curricula elements aimed at job market requirements.
@@
 \textbf{2. National Cyber Security Centre (NCSC)}
+
 The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
@@
 
-\textbf{3. Professional Bodies}
-The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas.
-The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}.
+\textbf{3. Professional Bodies/Learned Societies}
+
+BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas.
+The accreditation provided by these institutes are underpinned by international initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and the Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}.
 These memoranda support the internationalising of the curriculum and promote consistency and parity in computer science education globally.
 These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
-Both the IET and The BCS have a long history of expecting coverage of environmental factors within the programmes they accredit.
-The BCS has for a number of years been expecting significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}.
+Both the BCS and the IET have a long history of expecting coverage of environmental factors within the programmes they accredit.
+The BCS has for a number of years been expecting significant coverage of what it terms  {\emph{Legal, Ethical, Social and Professional Issues}}~\cite{Brooke2018}.
 Clearly cybersecurity has been and continues to be part of these expectations.
 
-In recent years the BCS has evolved its Accreditation Practices to promote and mandate the inclusion of cybersecurity within the programmes the body accredits.
+In recent years the BCS has evolved its accreditation practices to promote and mandate the inclusion of cybersecurity within the programmes the body accredits.
 The timeline which this process has followed in Table \ref{table:1}.
@@
     \hline
-    IV. ACCREDITATION:RQ3 - A. Developing Expectations &   \\ \hline
+    IV. ACCREDITATION ({\emph{RQ3}}) - A. Developing Expectations &   \\ \hline
     UK Government Cybersecurity Strategy \cite{ukcyberstrategy:2016} & November 2011 \\ \hline
@@
     \hline
-    IV. ACCREDITATION:RQ3 - B. What does the BCS tell Universities? & \\ \hline
+    IV. ACCREDITATION ({\emph{RQ3}}) - B. What does the BCS tell Universities? & \\ \hline
     Cybersecurity Principles Roadshow & March-April 2016 \\ \hline
-    IV. ACCREDITATION:RQ3 - C. Accreditation - what progress has been made? &  \\ \hline
-    All visited HEIs expected to be fully compliant 
+    IV. ACCREDITATION ({\emph{RQ3}}) - C. Accreditation - what progress has been made? &  \\ \hline
+    All visited institutions expected to be fully compliant 
  (BCS follow a 5-year Accreditation Cycle)  & September 2020\\ \hline
@@
 \subsection{Developing Expectations}
+
 Internationally the expectations regarding both the breadth and depth of the expected cybersecurity coverage has been the subject of much discussion, debate and analysis.
-Like many Governments, the UK Government has been actively been seeking ways to address this \cite{ukcyberstrategy:2016}, \cite{UKCabinetOffice2014}.
-In parallel to the work complete by the ACM \cite{ACM2013a}, in the UK considerable effort was taken to ensure Industry, Higher Education, Government and the related Professional Bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society \cite{Irons2016}.
-In 2013, an initiative was set up by (ISC)$^2$, CPHC (the representative body of UK computer science departments) and the Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
+Like many governments, the UK Government has been actively been seeking ways to address this~\cite{UKCabinetOffice2014,ukcyberstrategy:2016}.
+In parallel to the work complete by the ACM~\cite{ACM2013a}, in the UK considerable effort was taken to ensure industry, higher education, government and the relevant professional bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society~\cite{Irons2016}.
+In 2013, an initiative was set up by (ISC)$^2$, CPHC (the representative body of UK computer science departments) and the UK Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
 Three workshops in 2013 and 2015 attempted to define the principles of cybersecurity education and proposed a framework for embedding these principles in UK computer science curricula.
 Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate computer science degrees.
-This work initially lead to a white paper related to a proposal in the form of a white paper \cite{CPHCISC2014}, followed by a set of guidelines \cite{CPHCISC2}.
+This work initially lead to a white paper related to a proposal in the form of a white paper~\cite{CPHCISC2014}, followed by a set of guidelines~\cite{CPHCISC2}.
 The BCS agreed to adopt the outputs into their accreditation criteria.
 This was the first time that cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees.
-The fact that cybersecurity is included as a  component of the BCS accreditation criteria, reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards Chartered status.
+The fact that cybersecurity is included as a component of the BCS accreditation criteria, reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards chartered status.
 
@@
 \begin{itemize}
-    \item Information and risk 
-    \item Threats and attacks 
-    \item Cybersecurity architecture and operations
+    \item Information and risk;
+    \item Threats and attacks;
+    \item Cybersecurity architecture and operations;
     \item Secure systems and products; and 
-    \item Cybersecurity management
+    \item Cybersecurity management.
 \end{itemize}
@@
 
-2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices
+{\emph{2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices}}
 
-2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system
+{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system}}
 
-Additionally programmes seeking Chartered Information Technology Professional Accreditation also have to cover:
+Additionally programmes seeking Chartered Information Technology Professional (CITP) accreditation also have to cover:
 
-3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security
+{\emph{3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security}}
 
 In the context of BCS accreditation, these requirements imply an exit standard that all students on a programme must be able to demonstrate irrespective of the option choices they have made.
-This means a HEI applying for accreditation is expected to provide evidence that the criteria are taught and assessed in a non-trivial manner, to and by all students upon the programme seeking accreditation .
-A HEI is expected to provide evidence in the form of programme and module specification documentation and example assessment specifications (coursework and examinations).
+This means an institution applying for accreditation is expected to provide evidence that the criteria are taught and assessed in a non-trivial manner, to and by all students upon the programme seeking accreditation.
+An institution is expected to provide evidence in the form of programme and module specification documentation and example assessment specifications (coursework and examinations).
 These criteria and the expectation that they are taught and assessed has been present for a number of years.
 
-The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages HEIs to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
-In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals - either as a stand alone module or as a significant component(s) of other modules.
-This approach differs from the ACM approach where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
+The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages institutions to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
+In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals -- either as a stand alone module or as a significant component(s) of other modules.
+This approach differs from the ACM approach, where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
 
-\subsection{What does BCS tell Universities}
-The agreed Cybersecurity Principles and Learning Outcome \cite{CPHCISC2} were discussed with the wider education community by a road-show led by CPHC.
+\subsection{What Are Universities Told?}
+The agreed Cybersecurity Principles and Learning Outcome~\cite{CPHCISC2} were discussed with the wider education community by a road-show led by CPHC.
 A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of computer science degrees.
@@
 
-The BCS Guidelines on Course Accreditation are published upon the BCS website \cite{BCS2018a}.
-The BCS also publishes the changes that have been made \cite{BCS2018b}.
-When changes are made, the BCS communicates the changes by email and in writing to all the BCS Educational Affiliates, that is all the HEIs that seek accreditation from the BCS.
-The expectations for Cybersecurity were extended in the June 2015 version of the guidelines for consideration at Accreditation Visits that took place from September 2015 or later.
+The {\emph{BCS Guidelines on Course Accreditation}} are published online~\cite{BCS2018a}.
+The BCS also publishes the changes that have been made~\cite{BCS2018b}.
+When changes are made, the BCS communicates the changes by email and in writing to all the BCS Educational Affiliates, that is all the institutions that seek accreditation from the BCS.
+The expectations for cybersecurity were extended in the June 2015 version of the guidelines for consideration at accreditation visits that took place from September 2015 or later.
 
@@
 The accreditation process adopted by the BCS is cyclic in nature.
-Formally, the cycle is 5 years in duration.
+Formally, the cycle is five years in duration.
 The new expectations have been implemented as follows.
 To ensure continuous accreditation, accreditation visits are normally scheduled every 5 years.
-At the time of the next visit in this accreditation cycle, accreditation is conditional upon a HEI having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
+At the time of the next visit in this accreditation cycle, accreditation is conditional upon an institution having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
 It is anticipated that from 2020 the expectation will be all accredited programmes have the new expectations fully embedded.
 
-In the year prior to an Accreditation Visit, HEIs are invited to an Accreditation Briefing from the BCS.
-The intention of the briefing is to help ensure Accreditation Visits go smoothly from the perspective of both the BCS and visited HEI.
+In the year prior to an accreditation visit, insitutions are invited to a briefing from the BCS.
+The intention of the briefing is to help ensure accreditation visits are successful from the perspective of both the BCS and the institution.
 The briefings take place virtually.
-The briefing includes a summary of the process, discussion of recent changes, guidance regarding the application and a summary of common issues that are being seen in other HEIs.
+The briefing includes a summary of the process, discussion of recent changes, guidance regarding the application and a summary of common issues that are being seen in other institutions.
 Significant opportunity for seeking clarification is provided.
 One of the issues that is highlighted is not all institutions have yet evolved their programmes to fully address the increased expectation for cybersecurity.
-This is resulting in accreditation being contingent upon a HEI taking action to address this short fall or in some cases the withdrawal of accreditation.
-A number of HEIs are in the process of adjusting their curricula to meet the new expectations.
+This is resulting in accreditation being contingent upon an institution taking action to address this short fall or in some cases the withdrawal of accreditation.
+A number of institutions are in the process of adjusting their curricula to meet the new expectations.
 In this case, the BCS notes the changes to programme design, the outputs from which will be scrutinised at the next accreditation visit.
 
-\subsection{Accreditation - what progress has been made}
+\subsection{Accreditation: What Progress Has Been Made?}
 
@@
 A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines.
-What is being observed is the majority of visited HEIs have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
+What is being observed is the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
 However, a minority are requiring encouragement to do so.
 
-From the start of the Autumn 2015 term, up to and including the Autumn 2018 term, the BCS have carried out 70 accreditation visits (including 4 international visits).
-The BCS identified action was required to address concerns related to cybersecurity at 16 HEIs.
-So 54 HEIs were already delivering cybersecurity in keeping with the BCS expectations.
+From the start of the Autumn 2015 term, up to and including the Autumn 2018 term, the BCS have carried out 70 accreditation visits (including four international visits).
+The BCS identified action was required to address concerns related to cybersecurity at 16 institutions.
+So 54 institutions were already delivering cybersecurity in keeping with the BCS expectations.
 
-Long term actions were expected from 12 HEIs (6 in 2015/16, 3 in 2016/17 and 3 in the Autumn of 2018) who were awarded 'At Threshold' judgments.
-10 of these judgments were across all programmes.
-1 was specifically against a generalist masters programme only.
+Long-term actions were expected from 12 institutions (six in 2015/16, three in 2016/17 and three in Autumn 2018) who were awarded `{\emph{At Threshold}}' judgments.
+Ten of these judgments were across all programmes; one was specifically against a generalist masters programme only.
 This indicates that the BCS will expect adjustments to have taken place before the next accreditation visit.
@@
 
-Short terms 90 Day Responses where required from 4 HEIs. The outcomes of these actions were as follows:
+Short terms 90 Day Responses where required from four institutions; the outcomes of these actions were as follows:
 
-A)	Of the 11 UG programmes involved all were approved 'At Threshold'
-
-B)	Of the 9 UG programmes involved, 8 were approved and 1 refused
-
-C)	Of the 5 UG programmes involved, all approved 'At Threshold'
-
-D)	Of the 3 UG programmes involved, all 3 were refused.
+\begin{itemize}                                                                
+\item Of the 11 UG programmes involved all were approved `{\emph{At Threshold}}';
+\item Of the 9 UG programmes involved, 8 were approved and 1 refused;
+\item Of the 5 UG programmes involved, all approved `{\emph{At Threshold}}';
+\item Of the 3 UG programmes involved, all 3 were refused.
+\end{itemize}
 
@@
  \begin{quote}
-``The second year project provides an opportunity for exploring security aspects in depth with an industrial use case''.
+``{\emph{The second year project provides an opportunity for exploring security aspects in depth with an industrial use case.}}''
 \end{quote}
-In sum, this shows that many UK HEIs have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
-Clearly not all HEIs in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
 
-% TC: do we have big picture recommendations to make?
-\section{Conclusions}
-As regards the research questions:
+In sum, this shows that many UK institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
+Clearly not all institutions in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
+
+\section{Conclusions and Future Work}
+
+The work presented in this paper is a first step towards better understanding the nature, design, structure and assessment of cybersecurity education in the UK, through the lens of accreditation.
+It is clear that professional body accreditation is having a positive effect on universities, suporting wider national policy imperatives.
+We look forward to continuing this analysis, alongside other intiatives presented at the start of this paper, to provide international comparators and portability.
+
+As regards to the three research questions:
 
@@
 \begin{itemize}
-	\item teach SQL without teaching SQL Injection (\cite{Drop2019} );
+	\item teach SQL without teaching SQL Injection~\cite{Drop2019};
 	\item teach ``web programming'' without teaching Cross-Site Scripting \cite[(XSS)]{OWASP2017a};
@@
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
-One should probably use the data from \cite{Chenetal2019a}: ``If you pick up a SSL/TLS answer from Stack Overflow, there's a 70\% chance it's insecure''.
+One should probably use the data from \cite{Chenetal2019a}: ``{\emph{If you pick up a SSL/TLS answer from Stack Overflow, there's a 70\% chance it's insecure}}''.
 \end{enumerate}
@@
 \section*{Acknowledgements}
-Thanks to Sally Pearce, Academic Accreditation Manager (BCS, The Chartered Institute for IT) for supplying the summary information related to accreditation of UK degree programmes.
+Thanks to Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
 Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
 \par
-{\emph{N.B.}} The first two authors are current Vice-Presidents of BCS, The Chartered Institute for IT, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee.
+{\emph{N.B.}} The first two authors are current Vice-Presidents of BCS, The Chartered Institute for IT, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
 
-% fix at end
 %\IEEEtriggeratref{40}

### 2019-06-23 FIE2019-final.tex
Fixed typos

@@
 
-In the year prior to an accreditation visit, insitutions are invited to a briefing from the BCS.
+In the year prior to an accreditation visit, institutions are invited to a briefing from the BCS.
 The intention of the briefing is to help ensure accreditation visits are successful from the perspective of both the BCS and the institution.
@@
 The work presented in this paper is a first step towards better understanding the nature, design, structure and assessment of cybersecurity education in the UK, through the lens of accreditation.
-It is clear that professional body accreditation is having a positive effect on universities, suporting wider national policy imperatives.
-We look forward to continuing this analysis, alongside other intiatives presented at the start of this paper, to provide international comparators and portability.
+It is clear that professional body accreditation is having a positive effect on universities, supporting wider national policy imperatives.
+We look forward to continuing this analysis, alongside other initiatives presented at the start of this paper, to provide international comparators and portability.
 
@@
 \par
-The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended.(\S\ref{sec:PCIDSS}).
+The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended (\S\ref{sec:PCIDSS}).
 \item[RQ3:] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
@@
 
-% TC: do we have big picture recommendations to make?
-We have the following specific recommendations:
+We also have the following specific recommendations:
 

### 2019-06-23 FIE2019-final.tex
Final changes before submission

@@
 It is clear that professional body accreditation is having a positive effect on universities, supporting wider national policy imperatives.
-We look forward to continuing this analysis, alongside other initiatives presented at the start of this paper, to provide international comparators and portability.
+It is also clear that there is a significant need for mobilising the international computer science academic community -- alongside some of the existing initiatives presented at the start of this paper -- to continue this focus on cybersecurity education, and provide international comparators, portability and sharing of best practice.
 

### 2019-06-23 Response.tex
Final comments addressed

--- 2019-06-09 Response.tex
@@
 \def\r#1#2{``#1''$\rightarrow$``#2''}
-\documentclass{article}
+\documentclass[a4paper,11pt]{article}
+\usepackage[top=1.8cm,bottom=1.8cm,left=1.8cm,right=1.8cm,asymmetric]{geometry}
 \bibliographystyle{alphaurl}
@@
 \author{Crick/Davenport/Irons/Prickett}
-\title{Response to Referees}
-\date{June 2019}
+\title{Response to Referees for FIE 2019 Paper \#619 (1570524395):\\``A
+  UK Case Study on Cybersecurity Education and Accreditation''}
+\date{22 June 2019}
 \maketitle
 
-\section{Referee 1}
-No comments that require addressing in detail.
-\section{Referee 2}
-No comments that require addressing in detail.
-\section{Referee 3}
+\section*{Summary}
+
+In general, alongside addressing the main points from the referees,
+the paper has been substantially restructured and refined for clarity
+of its main message, as well as overall presentation and
+readability. This has included a new title, refreshed abstract,
+restructured introductory section, restructured conclusions, as well
+as updating/tidying of the references.
+
+\section*{Referee 1}
+
+All addressed/no comments that require addressing in detail.
+
+\section*{Referee 2}
+
+All addressed/no comments that require addressing in detail.
+
+\section*{Referee 3}
+
 \begin{enumerate}
-\item ``The research appears to be poorly structured and the analysis/argument is hard to interpret''.
-\item[A]We have done what we can, but note that the other referees disagree.
-\item ``The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.''
-\item[A]
+\item ``{\emph{The research appears to be poorly structured and the analysis/argument is hard to interpret}}''.
+\item[A]The paper has been re-written to make the structure
+clearer. In particular, the abstract has been updated to more
+succinctly provide the essence of the work; the signposting has been
+enhanced to better guide the reader; a research approach section to
+more clearly explain the approach followed has been added; the
+research questions have been made more precise to communicate the
+intentions more clearly; and the sections have been made more
+self-contained and logically structured to better support the reading
+of the paper in a non-linear manner.
+
+\item ``{\emph{The abstract is too long and could be made concise. For example, the 2nd paragraph can be avoided as the main idea has been conveyed in the 1st paragraph as well.}}''
+\item[A] The abstract is now two paragraphs long and better represents
+  the main thrust of the work.
+
+\item {\emph{What do the authors mean by ``There has been a recent international working group but this has yet to report.''}} 
+\item[A] In parallel to the research explored in the paper, there is a
+  working group as part of the ACM conference series on Innovation and Technology in Computer Science
+  Education (ITiCSE) intending to develop a taxonomy of cybersecurity
+  education. This has been clarified in Section 1 of the paper.
+
+\item {\emph{It would add more clarity to introduce what these terms DevOps vs DevSecOps represent in the paper.
+Also, please use the acronym GDPR after introducing it during the first usage such as General Data Protection Regulation (GDPR).}}
+\item[A] Discussion related to DevOps vs DevSecOps has been
+  removed, as well as the early section on GDPR. Abbreviations have
+  adopted the style recommended throughout.
+
+\item {\emph{There are repeated ``and'' in the sentence ``In particular the expectation of Privacy by design...'' Also, the ``by'' after ``and'' is not required.}}
+
+{\emph{Please consider rephrasing to ``While both Privacy by design and privacy by defualt have been expected to be good practices''.}}
+
+\item[A] These sentences have been rephrased and other typographical errors addressed.
+
+\item {\emph{Please rephrase the following statement ``How might
+  accreditation regard cybersecurity education, or help with it?'' as
+  the sentence structure doesn't make much sense.}}
+
+\item[A] This has been rephrased to ``{\emph{Can accreditation by
+      Professional, Statutory and Regulatory Bodies enhance the
+      provision of cybersecurity within a body's jurisdiction?}}'' The
+  other research questions have also been rephrased to improve the
+  clarity by which the intentions of the research are communicated.
+
+\item {\emph{It is not clear what is the message the authors are
+      conveying in the 2nd paragraph? It lists three different web
+      links and a reference. Also, the web links in 2nd and 3rd
+      paragraphs could be moved to the reference and they could be
+      referenced with a number in this paragraph. Again, please use
+      the acronym KA after introducing it as Knowledge Areas (KA) at
+      the first instance of use.}}
+
+\item[A] The authors apologise that these were note sections and should
+  not have been present in the paper; these sections have been
+  removed. The recommended approach to abbreviations has now been
+  adopted throughout the paper.
+
+\item  {\emph{What is JHD? Please expand?}}
+
+\item[A] This has been removed this.
+
+\item {\emph{Please expand OWASP.}}
+
+\item[A]Open Web Application Security Project has replaced the initial use of OWASP.
+
+\item {\emph{The references to first author in the 2nd paragraph could
+      be avoided and the sentence rephrased to highlight only the
+      facts and not any personal experience.}}
+
+\item[A] The references to the authors has been removed throughout and
+  any potential conjecture removed.
+
+\item {\emph{What does ``three accounted for the 36 of the 44'' mean?
+      Please clarify. I am assuming the 36 and 44 refer to number of
+      universities. Please add that clarification, if it is so.}}
+
+\item[A] This did mean universities and has been updated to indicate
+  this.
+
+\item {\emph{``If there are only 82 instances of these fragments, how
+      can 117 of them be verified?'' Please add clarification to this
+      discussion.}}
+
+\item[A] Typo on our part: this should have been 820, not 82.
+\end{enumerate}
+
+\section*{Referee 4}
+
+\begin{enumerate}
+\item ``{\emph{The problem is it is not a research paper in even the
+      broadest interpretation of that sense. The RQs are invented
+      based on topic areas the authors wish to riff on, only RQ3 comes
+      (somewhat) close to a RQ with an associated presentation based
+      on the literature and some analysis. The 1st 2 RQs, as well as
+      the Introduction sections, are selectively argued,
+      stream-of-consciousness almost writing. Entertaining,
+      informative, arguable, debatable - but not research. No
+      methodology is applied, no systematic review, etc. Instead a
+      smattering of anecdotal information and quotes (and even more
+      quotes in the footnotes) intended to convince the reader of a
+      viewpoint that is never quite articulated clearly - until the
+      end when it comes down to a small set of specific complaints
+      about SQL Injection, XSS, and StackOverflow (which is overall a
+      bit of a letdown as I felt as if this bloglike writing was
+      building to a more impactful crescendo, some kind of
+      kick-in-the-pants call to action for the community).}}''
+
+\item[A] The paper has been revised to more clearly communicate the
+  research approach employed, as well as the overall aims and
+  intention of the paper (for example, new title, abstract and
+  introductory framing). A research approach section has been
+  added to further elicit the case study approach adopted. The
+  research questions have been made more specific to more clearly
+  communicate the intentions of the research. The writing style has
+  been adjusted and a more formal approach taken. The conclusions have
+  been extended and clarified.
+
+\end{enumerate}
+
+\section*{Programme Committee}
+
+All addressed/no comments that require addressing in detail.
+
+\end{document}
 \item 

### 2019-06-23 FIE2019-final.tex
Removed hyperref for IEEE Xplore
Final tidying

@@
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
 
@@
 
-This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK's critical national infrastructure~\cite{lordscyberreport:2018}; also initiatives such as Cyber Essentials\footnote{Cyber Essentials is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
+This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK's critical national infrastructure~\cite{lordscyberreport:2018}; also initiatives such as {\emph{Cyber Essentials}}\footnote{{\emph{Cyber Essentials}} is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
 
@@
 \begin{enumerate}[label=(\roman*)]
-\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
+\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education;
 % ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
@@
 BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
-\item The generalist computer science graduate.
-\item The generalist computer science masters graduate.
-\item The specialist computer science graduate.
+\item The generalist computer science graduate;
+\item The generalist computer science masters graduate;
+\item The specialist computer science graduate;
 \item The specialist computer science masters graduate.
 \end{enumerate}
-The focus of this paper is on (ii)--(iii): the general computer science graduate.
 
-This paper thus focuses on three research questions:
+The focus of this paper is on (ii)--(iii): the general computer science graduate. We thus focus on three research questions:
 
@@
 It is 15 years since \cite{Guimaraesetal2004} 
-wrote ``All the topics listed above should be presented in the first
-Database Course'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}.
-SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon (\url{https://xkcd.com/327/}, dating back to 2007 according to the Internet Archive) and website (\url{http://bobby-tables.com/}).
+wrote ``{\emph{All the topics listed above should be presented in the first
+Database Course}}'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}.
+SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon\footnote{\url{https://xkcd.com/327/}, dating back to 2007, according to the Internet Archive} and website\footnote{\url{http://bobby-tables.com/}}.
 Nevertheless SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.
-\cite[the UK's definitive reference]{Bristol2018a} states ``a wide range of attack techniques for exploiting SQL
-injection or script injection are known and documented.''.
+\cite[the UK's definitive reference]{Bristol2018a} states ``{\emph{\ldots a wide range of attack techniques for exploiting SQL
+injection or script injection are known and documented.}}''.
 
@@
 shouldn't do.}}''~\cite{Drop2019}}.
-\par
 This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
@@
 But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources.
-\cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack overflow (\url{https://stackoverflow.com}) resource.
-53\% were about the Spring Security framework (\url{https://projects.spring.io/spring-security/}), dominated by authentication (45\%).
+\cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack Overflow online resource.
+53\% were about the Spring Security framework\footnote{\url{https://projects.spring.io/spring-security/}}, dominated by authentication (45\%).
 The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.
@@
 There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
-solved the issue!!! I have no idea why it was enabled by default''. As of writing (16 January 2018) there were no negative comments about this disabling of a vital security feature.
-\par
-This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}). Their first finding was:
+solved the issue!!!
+I have no idea why it was enabled by default''.
+As of writing (22 June 2019) there were no negative comments about this disabling of a vital security feature.
+This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}).
+Their first finding was:
 
@@
 It is well known that cybersecurity skills are in short supply~\cite{schneider2013}. For example:
+
 \begin{quote}
@@
 \end{quote}
+
 The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
-
 In the UK, there is a prominence of job adverts to recruit academic staff with specialisms in cybersecurity.
@@
 The document provides specific guidance for embedding and enhancing relevant cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.
-The document suggested 5 areas of coverage
+The document suggested five areas of coverage
 
@@
 
-The BCS accreditation (2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages institutions to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
+The BCS accreditation (criteria 2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages institutions to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
 In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals -- either as a stand alone module or as a significant component(s) of other modules.
@@
 The new expectations have been implemented as follows.
-To ensure continuous accreditation, accreditation visits are normally scheduled every 5 years.
+To ensure continuous accreditation, accreditation visits are normally scheduled every five years.
 At the time of the next visit in this accreditation cycle, accreditation is conditional upon an institution having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
@@
 
-In sum, this shows that many UK institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
+In summary, this shows that many UK institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
 Clearly not all institutions in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
@@
 \begin{description}
-\item[RQ1:] What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?
+\item[RQ1:] {\emph{What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?}}
 
@@
 However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
-\item[RQ2:] Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?
+
+\item[RQ2:] {\emph{Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?}}
 
@@
 The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended (\S\ref{sec:PCIDSS}).
-\item[RQ3:] Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?
+
+\item[RQ3:] {\emph{Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?}}
 
@@
 \item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
-Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks)} is silent.} that has Cross-Site Request Forgery protection.
+Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks}) is silent.} that has Cross-Site Request Forgery protection.
 \item A study similar to \cite{Drop2019} is recommended for future work in Web Programming, Agile Development, Android Development and potentially other areas of the curriculum.

### 2019-07-18 FIE2019-final.tex
Added ack to IoC

@@
 It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
-Nevertheless, it is wrong for a complete curriculum to ignore cybersecurity issues for example
+Nevertheless, it is wrong for a complete curriculum to ignore cybersecurity issues for example:
 \begin{itemize}
@@
 Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
+All authors' institutions are members of the Institute of Coding (\url{https://instituteofcoding.org/}), an initiative funded by the Office for Students (England) and the Higher Education Funding Council for Wales.
 \par

### 2019-07-24 FIE2019-final.tex
Tidying
Final tweaks to abstract

@@
-\newif\ifanon
-\anontrue
 \documentclass[conference]{IEEEtran}
@@
 \usepackage{enumitem}
+\usepackage[noadjust]{cite}
 \usepackage[hyphenbreaks]{breakurl}
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
 \usepackage{booktabs} % For formal tables
@@
 \title{A UK Case Study on\\Cybersecurity Education and Accreditation}
-%\title{Cybersecurity Education Is Too Important\\To Be Left To The Specialists} Save this for a future paper?
-%\title{What Do We Mean By Cybersecurity Education?}  Tom's original
 
@@
 % }
-% \ifanon
-% \author{Anonymous}
-% \else
+
 % compact IEEE author format for more than three authors...
@@
 Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
-%\fi
+
 \maketitle
@@
 \begin{abstract}
-% submitted abstract (100-500 word limit):
-This Innovative Practice Full Paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as other important facets).
+This Innovative Practice Full Paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are prioritised, implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as various other important facets).
 A multitude of model curricula and recommendations have been presented and discussed in international fora in recent years, with varying levels of impact on education, policy and practice.
-This paper address three key questions: {\emph{i)}} What is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students;  and {\emph{iii)}} Can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
+This paper address three key questions: {\emph{i)}} what is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} should cybersecurity be taught stand-alone or in an integrated manner to general computer science students; and {\emph{iii)}} can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
 
 Evaluating how cybersecurity is taught in all aspects of computer science is clearly a task of considerable size, one that is beyond the scope of this paper.
-Instead a case study-based research approach, primarily focusing on the UK, has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students.
-Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity mean more generally for institutions and educators, and how can we teach this subject more effectively?
-Through this UK case study, and by contrasting with the US, we demonstrate the positive effect that national accreditation requirements can have, and give some recommendations both for future research and curriculum developments.
-
-%The paper proposes a number of recommendations for future cybersecurity educational initiatives, interventions and research, including a number of innovative practical suggestions for educators and curriculum designers, with a view for potential adoption and adaption in other jurisdictions internationally.
+Instead a case study-based research approach -- primarily focusing on the UK -- has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students.
+Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity knowledge and skills mean more generally for institutions and educators, and how can we teach this subject more effectively?
+Through this UK case study, and by contrasting with related initiatives in the US, we demonstrate the positive effect that national accreditation requirements can have, and offer some recommendations both for future research and curriculum developments.
 \end{abstract}

### 2019-07-24 FIE2019-final.tex
Changes to Intro

@@
 
-Cybersecurity has increasingly been a headline feature in the news over recent years, generally prompted by spectacular breaches, from major credit reference agencies~\cite{equifax2017}, telecoms companies~\cite{talktalk2016}, national airlines~\cite{BritishAirways2018a}, online dating websites~\cite{ashleymadison2015}, and even between sovereign governments~\cite{ncsc2018}.
-These major breaches have had significant impact on both individual citizens and society in general, requiring attention from organisations of all sizes~\cite{Hannigan2019a}\footnote{Former Director of GCHQ: UK equivalent of NSA.}:
+Cybersecurity has increasingly been a headline feature in the news over recent years, generally prompted by spectacular breaches, including major credit reference agencies~\cite{equifax2017}, telecoms companies~\cite{talktalk2016}, national airlines~\cite{BritishAirways2018a}, online dating websites~\cite{ashleymadison2015}, and even between sovereign governments~\cite{ncsc2018}.
+These major breaches have had significant impact on both individual citizens and society in general, requiring attention from organisations of all sizes:
 \begin{quote}
-``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''
+``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''\\
+Robert Hannigan, former Director of GCHQ~\cite{Hannigan2019a}'
 \end{quote}
-\noindent
+
 These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
@@
 
-With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as post-compulsory -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a}.
+With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as tertiary -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a,NCSC2017}.
 This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
 
-A recent ACM working group, as part of the Innovation and Technology in Computer Science Education (ITiCSE) conference series~\cite{Parrishetal2018a}, has been capturing global perspectives on cybersecurity education, but has yet to report its full findings.
-An aim of the ITiCSE working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``catalog existing [\dots] knowledge materials'', but there is no mention of any quality control over these.
-Nevertheless, it is one thing to write national curricula, specifications and requirements, and another thing to deliver appropriate and relevant education and skills; furthermore, this paper asks how well this is done in practice.
+A recent ACM working group~\cite{Parrishetal2018a}, as part of the {\emph{Innovation and Technology in Computer Science Education}} (ITiCSE) conference series, has been capturing global perspectives on cybersecurity education, but as of July 2019 has yet to report its full findings.
+An aim of the ITiCSE working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``{\emph{catalog existing [\dots] knowledge materials}}'', but there is no mention of any quality control over these.
+Nevertheless, it is one thing to write national curricula, specifications and requirements, and another thing to deliver appropriate and relevant education and skills; this paper asks how well this is done in practice.
 
@@
 The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
-The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}.
+The NCSC publishes and accredits to a number of cybersecurity standards~\cite{NCSC2017}.
 These standards are linked to the ACM recommendations for curricula \cite{ACM2013a}.

### 2019-07-25 FIE2019-final.tex
Changes to Cyber for All? section

@@
 These major breaches have had significant impact on both individual citizens and society in general, requiring attention from organisations of all sizes:
+
 \begin{quote}
 ``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''\\
-Robert Hannigan, former Director of GCHQ~\cite{Hannigan2019a}'
+\hfill Robert Hannigan~\cite{Hannigan2019a}, former Director of GCHQ
 \end{quote}
@@
 
-A recent ACM working group~\cite{Parrishetal2018a}, as part of the {\emph{Innovation and Technology in Computer Science Education}} (ITiCSE) conference series, has been capturing global perspectives on cybersecurity education, but as of July 2019 has yet to report its full findings.
+An ACM working group~\cite{Parrishetal2018a}, established in 2018 as part of the {\emph{Innovation and Technology in Computer Science Education}} (ITiCSE) conference series, has been capturing global perspectives on cybersecurity education, but as of July 2019 has yet to report its full findings.
 An aim of the ITiCSE working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``{\emph{catalog existing [\dots] knowledge materials}}'', but there is no mention of any quality control over these.
@@
 
-\section{Cybersecurity: For All, Or For Specialists?}
+\section{Cybersecurity For All, Or Just For Specialists?}
 
 In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014,davenport-et-al:latice2016}.
-This is not a new concern: see~\cite{Parr2014a} for concerns over five years ago, but it is a growing one.
+This is not a new concern~\cite{Parr2014a}, but it is a growing one.
 
-This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (as well as the setting up of the UK National Cyber Security Centre, with increased scrutiny of what this means for the UK's critical national infrastructure~\cite{lordscyberreport:2018}; also initiatives such as {\emph{Cyber Essentials}}\footnote{{\emph{Cyber Essentials}} is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity -- ENISA -- the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
+This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the publication in 2016 of the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (along with the setting up of the National Cyber Security Centre, as well as increased scrutiny of the resilience of the UK's critical national infrastructure~\cite{lordscyberreport:2018}; also industry-focused initiatives such as {\emph{Cyber Essentials}}\footnote{{\emph{Cyber Essentials}} is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity: ENISA, the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
 
 The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the discipline of computer science for a number of years~\cite{Hentea2006,schneider2013}.
-There have been a number of international initiatives to define curricula to support this, for example the ACM Computer Science Curricula Recommendations~\cite[which added ``Information Assurance and Security'' for the first time]{ACM2013a}, as well as for specialised cybersecurity degree programmes~\cite{ACMIEEEAISSIGSECIFIP}.
-There has been a debate as to whether cybersecurity is distinct discipline from computer science \cite{McGettrick2013}.
-The consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within computer science and related degrees.
+There have been a number of international initiatives to define curricula to support this, for example the 2013 ACM Computer Science Curricula Recommendations~\cite[which added ``{\emph{Information Assurance and Security}}'' for the first time]{ACM2013a}, as well as for specialised cybersecurity degree programmes~\cite{ACMIEEEAISSIGSECIFIP}.
+There has been debate as to whether cybersecurity is distinct discipline from computer science \cite{McGettrick2013}; the consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within computer science and related degrees.
 Recent evolution of cybersecurity education shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
@@
 
-Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{wgictreview:2013,brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of a compulsory education.
+Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{wgictreview:2013,brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of compulsory school-level education.
 While aspects of ``e-safety'' and principles of protecting personal data are increasingly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national extra-curricular competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.
-The Institute of Coding~\cite{Davenportetal2019a}, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country -- but primarily focused on university graduates -- does indeed mention cybersecurity, but merely as a sub-item in one work package.
+The Institute of Coding, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country~\cite{Davenportetal2019a} -- but primarily focused on university graduates -- does indeed mention cybersecurity, but merely as a sub-item in one work package.
 

### 2019-07-26 FIE2019-final.tex
Changes to Methodology section

@@
 
-\section{Research Questions}
+\section{Methodology}
 
-There are various levels of specialism at which cybersecurity education can be addressed:
+\subsection{Research Questions}
+
+There are various levels of specialism at which cybersecurity education and skills can be addressed; for example:
 \begin{enumerate}[label=(\roman*)]
-\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education;
 % ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
@@
 \item The specialist computer science graduate;
-\item The specialist computer science masters graduate.
+\item The specialist computer science masters graduate;
+\item The reskilling/upskilling/professional development of the IT industry and the wider workforce;
+\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
 \end{enumerate}
 
-The focus of this paper is on (ii)--(iii): the general computer science graduate. We thus focus on three research questions:
+The focus of this paper is on (i)--(ii): the general computer science graduate. We thus focus on three research questions:
 
@@
 
-A UK-focused case study-based approach is adopted in this project.
-As is common in case study-based research, many alternative case studies could have been chosen.
-The cases are illustrative rather than comprehensive in terms of the available case studies or challenges.
+A UK-focused case study-based approach has been adopted for this project.
+As is common in case study-based research, many alternative case studies could have been chosen; furthermore, the cases are illustrative rather than comprehensive in terms of the available case studies or challenges.
 The cases were evaluated to articulate the progress made and highlight opportunities for future developments in education and practice.
 
-The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations.
-This is followed by the evaluation of the policy context.
+The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations; this is contextualised by the wider policy context.
 Together these provide a context to the ongoing enhancement initiatives in the area of cybersecurity education.
 
-In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
-Namely, an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
+In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of specific case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
+Namely: an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
 
-Finally, in order to address {\emph{RQ3}} cases studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
-As discussed in the current situation section, in this jurisdiction mandating cybersecurity within PSRB accreditation is at a reasonably mature stage.
+Finally, in order to address {\emph{RQ3}}, case studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
+As discussed, mandating cybersecurity within PSRB accreditation in this jurisdiction is at a reasonably mature stage.
 
@@
 
-The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to RQ2:
+The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to {\emph{RQ2}}:
+
 \begin{quote}
-{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
-given the manner in which the topics are pervasive throughout other KAs}}
+``{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
+given the manner in which the topics are pervasive throughout other KAs.}}''
 \end{quote}
-It proposes 9 ``core'' hours and 63.5 distributed across the other KAs.
-%\pagebreak - to fix URL issues
-\par
-Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:
+
+It proposes nine ``core'' hours and 63.5 distributed across the other KAs. Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:
+
 \begin{quote}
-{\emph{Universities suffer shortcomings, as well.
+``{\emph{Universities suffer shortcomings, as well.
 Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity.
 There is a big catch, however.
-Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}
+Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}''
 \end{quote}
@@
 As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.
-
 It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and professional body/learned society curricula are all similar.
@@
 \begin{description}
-\item[UK]BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
+\item[UK] BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (due to the five-year cycle).
 More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
-%BCS has had cybersecurity in the curriculum since 2013, and required cybersecurity in accreditation since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (because of the five-year cycle).
-\item[US]ACM has equally had cybersecurity (``Information Assurance and Security'' --- IAS) in the curriculum since 2013 \cite{ACM2013a}, but it is not the accrediting body.
-ABET is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''.
-This should mean that  all accredited universities should be compliant by 2025 (because of the six-year cycle).
-%\cite[chapter 7]{Davenportminor2019a}.  % need a better reference
+\item[USA] The Association of Computing Machinery (ACM) has equally had cybersecurity (IAS: ``Information Assurance and Security'') in the curriculum since 2013~\cite{ACM2013a}, but it is not the accrediting body.
+The Accreditation Board for Engineering and Technology (ABET) is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''.
+This should mean that  all accredited universities should be compliant by 2025 (due to their six-year cycle).
 \end{description}
@@
 %What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
-%{\bf JHD to research ACM general curriculum;}
 The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.

### 2019-07-26 FIE2019-final.tex
Changes to Challenges section

@@
 
-\section{Challenges: {\emph{RQ1}} and {\emph{RQ2}}}
+\section{Challenges: ({\emph{RQ1}} and {\emph{RQ2}})}
 
 %What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
-The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 Hours for IAS topics embedded in other Knowledge Areas.
+The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 hours for IAS topics embedded in other Knowledge Areas.
 
-The UK's official body of knowledge resource, the CyBOK project\cite{Bristol2019a}, has produced reference documentation for some (as of June 2019: five final, seven for comment, out of a planned total of 19)   % JHD updated figures 21/6/2019
-knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
+The UK's official body of knowledge resource, the CyBOK project~\cite{Bristol2019a}, has produced reference documentation for some (as of June 2019: five final, seven for comment, out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
+% JHD updated figures 21/6/2019
 
@@
 
-Is teaching cybersecurity different?
-Lecturing is probably not the best way.
-Should we use real-life case studies (e.g.
-British Airways \cite{Barth2018a}).
-\cite{Weiss:2013:THC:2527148.2527180} suggests that there are benefits from teaching this discipline area in a more practical manner.
-There are a number of ways a more practical manner could be action, one would be by the inclusion of pertinent cybersecurity standards in the curricula.
+Is teaching cybersecurity different from other computing topics?
+Lecturing is probably not the best way; how should we use real-life case studies (e.g.
+British Airways \cite{Barth2018a}) and authentic assessment?
+Recent work suggests that there are benefits from teaching this discipline area in a more practical manner \cite{Weiss:2013:THC:2527148.2527180}; one would be by the inclusion of pertinent cybersecurity standards in the curricula.
 
-The Payments Card Industry Data Security Standards (PCI DSS)~\cite{PCI2018b} is one such standard.
-PCI DSS underpins all processing of credit/debit cards.
+The {\emph{Payments Card Industry Data Security Standards}} (PCI DSS)~\cite{PCI2018b} is one such standard, underpinning all processing of credit/debit cards globally.
 Nevertheless, they are very rarely mentioned in generalist computer scientist courses.
@@
 However, the payments business is now so spread across websites, often run by small and medium enterprises (SME), or non-specialists.
-Even larger enterprises are not immune: \cite{Barth2018a} reports that the recent British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance.
+Even larger enterprises are not immune: the 2018 British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance~\cite{Barth2018a}.
 
 It is an interesting question as to whether standards such as PCI DSS should be addressed within degree courses (clearly degree courses can never cover all standards) or whether they should be addressed in professional training courses.
-However it appears the current situation is not ideal from the perspective of industry (or users of systems) and the inclusion of pertinent standards could be seen as a useful enhancement activity to the coverage of cybersecurity within generalist %undergraduate: JHD commented out, I don't think we mean this, and I teach it at MSc.
-computer science courses.
+However it appears the current situation is not ideal from the perspective of industry (or users of systems) and the inclusion of key standards could be seen as a valuable enhancement activity to the coverage of cybersecurity within generalist computer science courses.
 
@@
 
-It is 15 years since \cite{Guimaraesetal2004} 
-wrote ``{\emph{All the topics listed above should be presented in the first
-Database Course}}'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}.
-SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon\footnote{\url{https://xkcd.com/327/}, dating back to 2007, according to the Internet Archive} and website\footnote{\url{http://bobby-tables.com/}}.
+It is 15 years since \cite{Guimaraesetal2004} wrote ``{\emph{All the topics listed above should be presented in the first Database Course}}'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}.
+SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon and website\footnote{See: \url{https://xkcd.com/327/} (dating back to 2007, according to the Internet Archive) and \url{http://bobby-tables.com/}}.
 Nevertheless SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.
-\cite[the UK's definitive reference]{Bristol2018a} states ``{\emph{\ldots a wide range of attack techniques for exploiting SQL
-injection or script injection are known and documented.}}''.
+CyBOK states ``{\emph{\ldots a wide range of attack techniques for exploiting SQL injection or script injection are known and documented.}}''~\cite{Bristol2018a}.
 
-Clearly such a major weakness should be well-taught.  In general it is hard to determine what is taught, but a reasonable proxy for this is the content of recommended textbooks.
-
-% Drop2019 was awarded the (unique) best paper at SIGCSE2019 - not sure what to make of this.
-Hence \cite{Drop2019} analysed the database textbooks used by  44 of the top 50 computer science
-departments in the USA. % (using \cite{StangerMartin2015a}'s list, the other six didn't have a book listed). 
+Clearly such a major weakness should be well-taught; in general it is hard to determine what is actually delivered as part of a specific degree programme, but a reasonable proxy for this is the content of recommended textbooks.
+This was the rationale for a 2019 analysis of database textbooks used by 44 of the top 50 computer science departments in the USA\cite{Drop2019}.
 There were seven such books, but three books accounted for the 36 of the 44 universities.
 Five of the seven (30 of the 44) had no mention of SQL injection.
-On the other two, the more popular one has a seriously flawed discussion\footnote{``{\emph{However, the paper implies that using parameters is equivalent to using a function to add escape characters
-around user input. This is incorrect, as using parameters allows
-SQL statements to be pre-compiled, and prevents any user input
-from being interpreted as code, while escaping user input is not
-recommended as a sole defense since imperfect escape functions
-can easily be subverted.}}''~\cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``{\emph{However, the fact that the first
-example should not be used is not discussed until two pages after
-the example in the text, and is not mentioned at all in the caption or
-on the page where the figure appears. This means a student who is
-skimming the text looking for an example to modify for their own
-code could simply copy the code that first appears in the example,
-without being aware that this is in fact an example of what they
-shouldn't do.}}''~\cite{Drop2019}}.
-This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself.\footnote{Fixing this is a suggestion for future work.}
+On the other two, the more popular one has a seriously flawed discussion\footnote{``{\emph{However, the paper implies that using parameters is equivalent to using a function to add escape characters around user input.
+This is incorrect, as using parameters allows SQL statements to be pre-compiled, and prevents any user input from being interpreted as code, while escaping user input is not recommended as a sole defense since imperfect escape functions can easily be subverted.}}''~\cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``{\emph{However, the fact that the first example should not be used is not discussed until two pages after the example in the text, and is not mentioned at all in the caption or on the page where the figure appears.
+This means a student who is skimming the text looking for an example to modify for their own code could simply copy the code that first appears in the example, without being aware that this is in fact an example of what they shouldn't do.}}''~\cite{Drop2019}}.
+This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself at the time of writing.%\footnote{Fixing this is a suggestion for future work.}
 
@@
 But all the accepted answers to CSRF-related failures simply suggested disabling the check.
-There were no negative comments about this, and indeed a typical response is ``Adding \verb!csrf().disable()!
-solved the issue!!!
-I have no idea why it was enabled by default''.
-As of writing (22 June 2019) there were no negative comments about this disabling of a vital security feature.
+There were no negative comments about this, and indeed a typical response is ``{\emph{Adding}} \verb!csrf().disable()!
+{\emph{{solved the issue!!!
+I have no idea why it was enabled by default}}''.
+As of writing, there were no negative comments about this disabling of a vital security feature.
 This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}).
@@
 \begin{quote}
-{\emph{644 out of the 1,429 inspected answer posts
+``{\emph{644 out of the 1,429 inspected answer posts
 (45\%) are insecure, meaning that insecure suggestions
 popularly exist on SO. Insecure answers dominate, in
-particular, the SSL/TLS category}} [355 insecure versus 150 secure, i.e. $>70$\%].
+particular, the SSL/TLS category}}''\newline[355 insecure versus 150 secure, i.e. $>70$\%].
 \end{quote}
@@
 
-A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks; \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
-The key finding was this:
+Another recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks; \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
+The key finding was:
+
 \begin{quote}
-{\emph{We found that 15.4\% of all 1.3 million Android applications
+``{\emph{We found that 15.4\% of all 1.3 million Android applications
 contained security-related code snippets from
 Stack Overflow. Out of these 97.9\% contain at least one
-insecure code snippet.}}
+insecure code snippet.}}''
 \end{quote}
@@
 On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
-\par
 The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}.
-Their key recommendation  \cite[Slide 32]{Votipkaetal2019a} is ``{\emph{Improve documentation: Clarify what you can(not) copy/paste}}''.
+Their key recommendation~\cite[Slide 32]{Votipkaetal2019a} was ``{\emph{Improve documentation: Clarify what you can(not) copy/paste}}''.
 
@@
 
-A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for ``Agile'' textbooks.
+A further recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for ``Agile'' textbooks.
 Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
-\par
-Agile's preference for functionality over non-functional requirements is clearly displayed in practice. \cite{Naiakshinaetal2017a} asked 20 student developers to imagine they
-were part of a team working on creating a social networking site for
-our university and to implement a password storage mechanism for this.
+Agile's preference for functionality over non-functional requirements is clearly displayed in practice.
+\cite{Naiakshinaetal2017a} asked 20 student developers to imagine they were part of a team working on creating a social networking site for our university and to implement a password storage mechanism for this.
 10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not.
 None of the unprimed ones implemented any security.
+This would also reaffirm the point raised in section~\ref{sec:PCIDSS}, regarding the effective use of industrial case studies and authentic assessment approaches.
 
@@
 
-The web abounds with informal resources: tutorials and code snippets.
+The web abounds with informal resources, such as tutorials and code snippets.
 How good are these, and how good are people at using these?
-This has been looked at by \cite{Unruhetal2017a}, This research took the top five results from Google for six queries.
+This has been looked at by \cite{Unruhetal2017a}, taking the top five search results from Google for six queries.
 Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
 Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
-Some students clearly make use of these resources, a recommendation of future work is to explore and evaluate students' (and indeed others') use of such informal resources.
+Some students clearly make use of these resources; thus a recommendation of future work is to explore and evaluate students' (and indeed others') use of such informal resources.
 
-\subsection{Staff}
+\subsection{Faculty}
 
-It is well known that cybersecurity skills are in short supply~\cite{schneider2013}. For example:
-
-\begin{quote}
-% JHD: I am in touch with ESG to try to find more, and someUK-specific numbers.
-{\emph{Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
-In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
-This had climbed to 51\% at the beginning of this year.
-Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.}}~\cite{Page2018a}
-\end{quote}
+It is well known that cybersecurity skills are in short supply, in both industry~\cite{Ackerman2019a} and academia~\cite{schneider2013}.
+For example, research into the state of IT conducted annually by Enterprise Strategy Group (ESG) has revealed that the skills gap in information security continues to widen and has doubled in the past five years; in 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills -- this had climbed to 51\% at the beginning of 2018~\cite{ESG:2018}.
+Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
 
 The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
-In the UK, there is a prominence of job adverts to recruit academic staff with specialisms in cybersecurity.
+In the UK, there has been a resurgence of job adverts to recruit academic staff with specialisms in cybersecurity over the past three years.
 The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
@@
 Recent example have included a professor of cybersecurity, two senior academic positions and two junior academic positions in one advert.
-There are other examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year; there are also examples of cybersecurity subject groups moving en masse from one university to another.
+There are other examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year; there are also examples of cybersecurity research groups moving en masse from one university to another.
 
-%Not sure if we want to raise something about funding of posts? Poaching of staff ??
-%Is this too contentious
-%TC: probably out of scope for this paper
+% \begin{quote}
+% ``{\emph{Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
+In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
+This had climbed to 51\% at the beginning of this year.
+Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.}}''~\cite{Page2018a}
+% \end{quote}
 

### 2019-07-26 FIE2019-final.tex
Changes to Challenges section

@@
 For example, research into the state of IT conducted annually by Enterprise Strategy Group (ESG) has revealed that the skills gap in information security continues to widen and has doubled in the past five years; in 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills -- this had climbed to 51\% at the beginning of 2018~\cite{ESG:2018}.
-Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.
+Clearly, cybersecurity is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom~\cite{Ackerman2019a}.
 

### 2019-07-26 FIE2019-final.tex
Changes to Accreditation section, plus new case study section

@@
 %{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
-\section{Accreditation by PSRBs: {\emph{RQ3}}}
+\section{Accreditation by PSRBs ({\emph{RQ3}})}
 
-In the UK as in most jurisdictions, higher education provision addresses general computer science and specialist cybersecurity courses.
-A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields computer security, digital/computer forensics, etc).
-In the UK, the Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions providing undergraduate qualifications related to cybersecurity for entry in September 2019.
-An even larger number of institutions provide study opportunities related to more general computer science.
-UCAS lists 246 provides for undergraduate programmes related to computer science.
+In the UK --- as in most jurisdictions --- higher education provision addresses general computer science and specialist cybersecurity courses.
+A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields, such as computer security, digital/computer forensics, etc).
+The UK's Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions providing undergraduate qualifications related to cybersecurity for entry in September 2019.
+An even larger number of UK institutions provide study opportunities related to more general undergraduate computer science: UCAS lists 246 providers for undergraduate programmes for the 2019/2020 academic year.
 
 Accreditation has evolved to directly addresses the cybersecurity challenges in both general computer science programmes and specialist cybersecurity programmes.
-In the UK, accreditation in the broad computing area is being performed by a few different agencies.
+In the UK, accreditation in the broad computing area is being performed by a few organisations with overlapping interests and priorities.
 These include:
 
-\textbf{1. Not-for-Profit Organisations}
+\subsection{National Cyber Security Centre (NCSC)}
+
+The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
+The NCSC publishes and accredits to a number of cybersecurity standards~\cite{NCSC2017}; these standards are linked to the ACM recommendations for curricula~\cite{ACM2013a}.
+To date, the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity.
+More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity, and computer science degrees with a significant cybersecurity focus.
+
+The NCSC have accredited 15 cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters cybersecurity programmes and 1 cybersecurity degree programme, with a further 2 provisionally accredited.
+Hence the extent of accreditation is currently reasonably limited in terms of reach to computer science programmes.
+This appears to be a positive initiative that will further develop over the next few years, especially with increasing industry engagement in this area.
+
+\subsection{Not-for-Profit Organisations}
 
@@
 One of the curricula elements is related to cybersecurity.
-Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.
+Tech Partnership Degrees have a specialist scope, endorsing programmes in the named areas of ``IT Management for Business'' and ``Software Engineering for Business''.
 Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and five Software Engineering for Business Programmes.
@@
 
-The Institute of Coding is a not-for-profit entity, led by the University of Bath, that intends to enhance how digital skills are developed in higher education in the UK~\cite{Davenportetal2019a}.
+The Institute of Coding\footnote{\url{https://instituteofcoding.org}} is a not-for-profit entity, led by the University of Bath, that intends to enhance how digital skills are developed in higher education in the UK~\cite{Davenportetal2019a}.
 This is likely to include cybersecurity related skills.
@@
 
-\textbf{2. National Cyber Security Centre (NCSC)}
+\subsection{Professional Bodies and Learned Societies}
 
-The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
-The NCSC publishes and accredits to a number of cybersecurity standards~\cite{NCSC2017}.
-These standards are linked to the ACM recommendations for curricula \cite{ACM2013a}.
-To date the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity.
-More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and computer science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
-
-Currently accredited are 15 cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters cybersecurity programmes and 1 cybersecurity Degree Programme with a further 2 provisionally accredited.
-Hence the extent of accreditation is currently reasonably limited in terms of reach to computer science programmes.
-This appears to be a positive initiative that will hopefully further develop over time.
-
-\textbf{3. Professional Bodies/Learned Societies}
-
-BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) both accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas.
+Both BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas.
 The accreditation provided by these institutes are underpinned by international initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and the Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}.
 These memoranda support the internationalising of the curriculum and promote consistency and parity in computer science education globally.
-These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
+These professional bodies are also UK charities and hence have responsibilities for public good which extends beyond short term job market needs~\cite{Stensaker2006,Mutereko2017}.
 Both the BCS and the IET have a long history of expecting coverage of environmental factors within the programmes they accredit.
-The BCS has for a number of years been expecting significant coverage of what it terms  {\emph{Legal, Ethical, Social and Professional Issues}}~\cite{Brooke2018}.
+The BCS has for a number of years been expecting significant coverage of what it terms  {\emph{Legal, Social, Ethical and Professional Issues}}~\cite{Brooke2018}.
 Clearly cybersecurity has been and continues to be part of these expectations.
 
+\section{Case Study: BCS Accreditation}
+
 In recent years the BCS has evolved its accreditation practices to promote and mandate the inclusion of cybersecurity within the programmes the body accredits.
-The timeline which this process has followed in Table \ref{table:1}.
+The timeline which this process has followed in presented in Table~\ref{table:1}, and will be discussed in more detail in the following case study.
+
 %\begin{center}
@@
     UK Government Cybersecurity Strategy \cite{ukcyberstrategy:2016} & November 2011 \\ \hline
-    Three workshops of a consortium of industry, academia and government bodies - led by CPHC and  (ISC)$^2$ - leading to the development of Cybersecurity learning guidelines to be embedded into BCS accredited UK Computer Science and IT-related degree \cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
+    Three workshops of a consortium of industry, academia and government bodies -- led by CPHC and  (ISC)$^2$ -- leading to the development of cybersecurity learning guidelines to be embedded into BCS accredited UK computer science and IT-related degrees~\cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
     UK Government report Cybersecurity Skills, Business Perspectives and Government's Next Steps Report Released \cite{UKCabinetOffice2014} & March 2014  \\ \hline
@@
     All visited institutions expected to be fully compliant 
- (BCS follow a 5-year Accreditation Cycle)  & September 2020\\ \hline
+ (BCS follow a five-year accreditation cycle)  & September 2020\\ \hline
   \end{tabular}
@@
 Internationally the expectations regarding both the breadth and depth of the expected cybersecurity coverage has been the subject of much discussion, debate and analysis.
-Like many governments, the UK Government has been actively been seeking ways to address this~\cite{UKCabinetOffice2014,ukcyberstrategy:2016}.
-In parallel to the work complete by the ACM~\cite{ACM2013a}, in the UK considerable effort was taken to ensure industry, higher education, government and the relevant professional bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society~\cite{Irons2016}.
+Like many governments, the UK Government has actively been seeking ways to address this~\cite{UKCabinetOffice2014,ukcyberstrategy:2016}.
+In parallel to the work completed by the ACM~\cite{ACM2013a} in the USA, considerable effort in the UK have been taken to ensure industry, higher education, government and the relevant professional bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society~\cite{Irons2016}.
 In 2013, an initiative was set up by (ISC)$^2$, CPHC (the representative body of UK computer science departments) and the UK Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
 Three workshops in 2013 and 2015 attempted to define the principles of cybersecurity education and proposed a framework for embedding these principles in UK computer science curricula.
-Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 Universities that offer undergraduate computer science degrees.
-This work initially lead to a white paper related to a proposal in the form of a white paper~\cite{CPHCISC2014}, followed by a set of guidelines~\cite{CPHCISC2}.
+Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 universities that offer undergraduate computer science degrees.
+This work initially lead to a white paper related to a proposal in the form of a white paper in 2014~\cite{CPHCISC2014}, followed by a set of guidelines in 2015~\cite{CPHCISC2}.
 The BCS agreed to adopt the outputs into their accreditation criteria.
 This was the first time that cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees.
-The fact that cybersecurity is included as a component of the BCS accreditation criteria, reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards chartered status.
+The fact that cybersecurity is included as a component of the BCS accreditation criteria reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards chartered status.
 
@@
 
-The ambition of this approach is to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area.)  The approach taken is not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
+The ambition of this approach was to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area); it was not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
 In this context this is realised as an expectation cybersecurity is an inclusion in all degrees accredited by the BCS.
@@
 Two criteria are expected to be covered by all programmes seeking accreditation.
-These are \cite{BCS2018a}:
+These are:
 
-{\emph{2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices}}
+``{\emph{2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices}}''
 
-{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system}}
+``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system}}''~\cite{BCS2018a}
 
-Additionally programmes seeking Chartered Information Technology Professional (CITP) accreditation also have to cover:
+Additionally, programmes seeking Chartered Information Technology Professional (CITP) accreditation also have to cover:
 
-{\emph{3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security}}
+``{\emph{3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security}}''~\cite{BCS2018a}
 
@@
 
-The BCS accreditation (criteria 2.1.6, 2.1.9 and 3.1.2) is not prescriptive, but encourages institutions to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
+The relevant BCS accreditation (criteria 2.1.6, 2.1.9 and 3.1.2, as presented above) is not prescriptive, but encourages institutions to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
 In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals -- either as a stand alone module or as a significant component(s) of other modules.
-This approach differs from the ACM approach, where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
+This approach differs from the ACM approach~\cite{ACM2013a}, where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
 
 \subsection{What Are Universities Told?}
-The agreed Cybersecurity Principles and Learning Outcome~\cite{CPHCISC2} were discussed with the wider education community by a road-show led by CPHC.
+
+The agreed {\emph{Cybersecurity Principles and Learning Outcomes}}~\cite{CPHCISC2} were discussed with the wider education community, led by CPHC.
 A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of computer science degrees.
@@
 
-The {\emph{BCS Guidelines on Course Accreditation}} are published online~\cite{BCS2018a}.
-The BCS also publishes the changes that have been made~\cite{BCS2018b}.
+The {\emph{BCS Guidelines on Course Accreditation}} are published online~\cite{BCS2018a}, with the BCS publishing any changes that have been made~\cite{BCS2018b}.
 When changes are made, the BCS communicates the changes by email and in writing to all the BCS Educational Affiliates, that is all the institutions that seek accreditation from the BCS.
@@
 This change to the accreditation guidelines is now in an implementation period.
-The accreditation process adopted by the BCS is cyclic in nature.
-Formally, the cycle is five years in duration.
-The new expectations have been implemented as follows.
-To ensure continuous accreditation, accreditation visits are normally scheduled every five years.
-At the time of the next visit in this accreditation cycle, accreditation is conditional upon an institution having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
+The accreditation process adopted by the BCS is cyclic in nature; formally, the cycle is five years in duration.
+The new expectations have been implemented as follows: to ensure continuous accreditation, accreditation visits are normally scheduled every five years; at the time of the next visit in this accreditation cycle, accreditation is conditional upon an institution having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
 It is anticipated that from 2020 the expectation will be all accredited programmes have the new expectations fully embedded.
@@
 The intention of the briefing is to help ensure accreditation visits are successful from the perspective of both the BCS and the institution.
-The briefings take place virtually.
-The briefing includes a summary of the process, discussion of recent changes, guidance regarding the application and a summary of common issues that are being seen in other institutions.
+The briefings take place virtually and includes a summary of the process, discussion of recent changes, guidance regarding the application and a summary of common issues that are being seen in other institutions.
 Significant opportunity for seeking clarification is provided.
-One of the issues that is highlighted is not all institutions have yet evolved their programmes to fully address the increased expectation for cybersecurity.
+One of the issues often highlighted is that not all institutions have yet evolved their programmes to fully address the increased expectation for cybersecurity.
 This is resulting in accreditation being contingent upon an institution taking action to address this short fall or in some cases the withdrawal of accreditation.
@@
 This initiative is a collective attempt to formally include cybersecurity in all BCS Accredited programmes.
-Some of these programmes will be specialist cybersecurity programmes, however the majority will take a different emphasis.
-This is a work in progress.
-A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines.
+Some of these programmes will be specialist cybersecurity programmes, however the majority will take a different emphasis; this is a work in progress.
+A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines~\cite{BCS2018b}.
 What is being observed is the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
@@
 From the start of the Autumn 2015 term, up to and including the Autumn 2018 term, the BCS have carried out 70 accreditation visits (including four international visits).
-The BCS identified action was required to address concerns related to cybersecurity at 16 institutions.
-So 54 institutions were already delivering cybersecurity in keeping with the BCS expectations.
+The BCS identified action was required to address concerns related to cybersecurity at 16 institutions; thus, 54 institutions were already delivering cybersecurity in keeping with the BCS expectations.
 
@@
 
-Short terms 90 Day Responses where required from four institutions; the outcomes of these actions were as follows:
+Short terms 90 Day Responses where required from four institutions; the outcomes of these actions were as follows: ({\emph{i}}) of the 11 UG programmes involved all were approved `{\emph{At Threshold}}'; ({\emph{ii}}) of the nine UG programmes involved, eight were approved and one refused; ({\emph{iii}}) of the five UG programmes involved, all approved `{\emph{At Threshold}}'; and ({\emph{iv}}) of the 3 UG programmes involved, all 3 were refused.
 
-\begin{itemize}                                                                
-\item Of the 11 UG programmes involved all were approved `{\emph{At Threshold}}';
-\item Of the 9 UG programmes involved, 8 were approved and 1 refused;
-\item Of the 5 UG programmes involved, all approved `{\emph{At Threshold}}';
-\item Of the 3 UG programmes involved, all 3 were refused.
-\end{itemize}
+% \begin{itemize}                                                                
+% \item Of the 11 UG programmes involved all were approved `{\emph{At Threshold}}';
+% \item Of the 9 UG programmes involved, 8 were approved and 1 refused;
+% \item Of the 5 UG programmes involved, all approved `{\emph{At Threshold}}';
+% \item Of the 3 UG programmes involved, all 3 were refused.
+% \end{itemize}
 
 Good practice was identified at one university by the commendation:
+
  \begin{quote}
@@
 In summary, this shows that many UK institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
-Clearly not all institutions in the UK necessarily have to apply for accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
+Clearly not all institutions in the UK necessarily have to apply for PSRB accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
 
@@
 Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
-All authors' institutions are members of the Institute of Coding (\url{https://instituteofcoding.org/}), an initiative funded by the Office for Students (England) and the Higher Education Funding Council for Wales.
+All authors' institutions are members of the Institute of Coding, an initiative funded by the Office for Students (England) and the Higher Education Funding Council for Wales.
 \par

### 2019-07-26 FIE2019-final.tex
Changes to Conclusions

@@
 The work presented in this paper is a first step towards better understanding the nature, design, structure and assessment of cybersecurity education in the UK, through the lens of accreditation.
-It is clear that professional body accreditation is having a positive effect on universities, supporting wider national policy imperatives.
+It is clear that PSRB accreditation is having a positive effect on universities, supporting wider national policy imperatives.
 It is also clear that there is a significant need for mobilising the international computer science academic community -- alongside some of the existing initiatives presented at the start of this paper -- to continue this focus on cybersecurity education, and provide international comparators, portability and sharing of best practice.
-
-As regards to the three research questions:
+With regards to the three research questions:
 
@@
 Nevertheless, it is wrong for a complete curriculum to ignore cybersecurity issues for example:
+
 \begin{itemize}
@@
 	\item ignore the  impact of GDPR upon systems development;
-	\item uncritically teach the use of Agile Development to develop secure systems.
+	\item uncritically teach the use of Agile development to develop secure systems.
 \end{itemize}
-\par
+
 The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended (\S\ref{sec:PCIDSS}).
@@
 Accreditation (as practised by BCS in the UK) is a valuable tool in improving the standard of cybersecurity teaching, and disseminating good practice, and should continue this.
-The time-lag in adoption of cybersecurity in the US seems to be caused by the accreditation differences more than any other factors.
+The time-lag in adoption of cybersecurity in the USA seems to be caused by the accreditation differences more than any other factors.
 \end{description}
@@
 Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks}) is silent.} that has Cross-Site Request Forgery protection.
-\item A study similar to \cite{Drop2019} is recommended for future work in Web Programming, Agile Development, Android Development and potentially other areas of the curriculum.
+\item A study similar to \cite{Drop2019} is recommended for future work focusing on Java development, Android development and Agile development, and potentially other areas of the curriculum, to better understand how modern textbooks support key areas of the curriculum.
 \item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
 One should probably use the data from \cite{Chenetal2019a}: ``{\emph{If you pick up a SSL/TLS answer from Stack Overflow, there's a 70\% chance it's insecure}}''.
+\item Finally, further work exploring and evaluating effective pedagogical and assessment approaches for cybersecurity education, in a similar vein to what has been seen with computer science education more generally (e.g.~\cite{davenport-et-al:latice2016}).
 \end{enumerate}

### 2019-07-26 FIE2019-final.tex
Changes to Acks

@@
 \section*{Acknowledgements}
-Thanks to Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
+
+The authors wish to thank Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
 Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
 All authors' institutions are members of the Institute of Coding, an initiative funded by the Office for Students (England) and the Higher Education Funding Council for Wales.
-\par
-{\emph{N.B.}} The first two authors are current Vice-Presidents of BCS, The Chartered Institute for IT, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
+
+{\emph{N.B.}} The first two authors are current Vice-Presidents of the BCS, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
 

### 2019-07-26 FIE2019-final.tex
Back to nine pages

@@
 
-Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{wgictreview:2013,brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of compulsory school-level education.
+Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{crick+sentance:2011,wgictreview:2013,brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of compulsory school-level education.
 While aspects of ``e-safety'' and principles of protecting personal data are increasingly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national extra-curricular competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.
@@
 Clearly such a major weakness should be well-taught; in general it is hard to determine what is actually delivered as part of a specific degree programme, but a reasonable proxy for this is the content of recommended textbooks.
-This was the rationale for a 2019 analysis of database textbooks used by 44 of the top 50 computer science departments in the USA\cite{Drop2019}.
+This was the rationale for a 2019 analysis of database textbooks used by 44 of the top 50 computer science departments in the USA~\cite{Drop2019}.
 There were seven such books, but three books accounted for the 36 of the 44 universities.
 Five of the seven (30 of the 44) had no mention of SQL injection.
-On the other two, the more popular one has a seriously flawed discussion\footnote{``{\emph{However, the paper implies that using parameters is equivalent to using a function to add escape characters around user input.
-This is incorrect, as using parameters allows SQL statements to be pre-compiled, and prevents any user input from being interpreted as code, while escaping user input is not recommended as a sole defense since imperfect escape functions can easily be subverted.}}''~\cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``{\emph{However, the fact that the first example should not be used is not discussed until two pages after the example in the text, and is not mentioned at all in the caption or on the page where the figure appears.
-This means a student who is skimming the text looking for an example to modify for their own code could simply copy the code that first appears in the example, without being aware that this is in fact an example of what they shouldn't do.}}''~\cite{Drop2019}}.
-This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself at the time of writing.%\footnote{Fixing this is a suggestion for future work.}
+On the other two, the more popular one has a seriously flawed discussion, and the other, while generally excellent, had a presentational problem~\cite{Drop2019}.
+This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself at the time of writing.
+
+%\footnote{Fixing this is a suggestion for future work.}
+
+% first footnote removed:
+% \footnote{``{\emph{However, the paper implies that using parameters is equivalent to using a function to add escape characters around user input.
+This is incorrect, as using parameters allows SQL statements to be pre-compiled, and prevents any user input from being interpreted as code, while escaping user input is not recommended as a sole defense since imperfect escape functions can easily be subverted.}}''~\cite{Drop2019}}
+
+% second footnote removed:
+% \footnote{``{\emph{However, the fact that the first example should not be used is not discussed until two pages after the example in the text, and is not mentioned at all in the caption or on the page where the figure appears.
+This means a student who is skimming the text looking for an example to modify for their own code could simply copy the code that first appears in the example, without being aware that this is in fact an example of what they shouldn't do.}}''~\cite{Drop2019}}
 
@@
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
-Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks}) is silent.} that has Cross-Site Request Forgery protection.
+\item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework that has Cross-Site Request Forgery protection.
+% \footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
+Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks}) is silent.}
 \item A study similar to \cite{Drop2019} is recommended for future work focusing on Java development, Android development and Agile development, and potentially other areas of the curriculum, to better understand how modern textbooks support key areas of the curriculum.

### 2019-07-30 ITNow.tex
Added ITNow skeleton

--- 2019-07-26 FIE2019-final.tex
@@
-\documentclass[conference]{IEEEtran}
-%\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\documentclass[a4paper,11pt]{article}
+\usepackage[top=1.8cm,bottom=1.8cm,left=1.8cm,right=1.8cm,asymmetric]{geometry}
+\usepackage{url}
 \usepackage{paralist}
-\usepackage{url}
 \usepackage{enumitem}
@@
 \usepackage[hyphenbreaks]{breakurl}
-%\usepackage[T1]{fontenc}
+\usepackage{authblk}
+\usepackage{tabularx}
 \usepackage{booktabs} % For formal tables
+\usepackage[colorlinks=true,hyperfootnotes=true]{hyperref}
 
-%\usepackage{hyperref}
+\title{Cybersecurity Education and Accreditation in the UK}
 
+% authors
+\author[1]{Tom Crick}
+\author[2]{James H. Davenport}
+\author[3]{Alastair Irons}
+\author[4]{Tom Prickett}
+% institutions
+\affil[1]{Swansea University, Swansea, UK}
+\affil[2]{University of Bath, Bath, UK}
+\affil[3]{Sunderland University, Sunderland, UK}
+\affil[4]{Northumbria University, Newcastle upon Tyne, UK}
+% emails
+\affil[1]{\url{thomas.crick@swansea.ac.uk}}
+\affil[2]{\url{j.h.davenport@bath.ac.uk}}
+\affil[3]{\url{alastair.irons@sunderland.ac.uk}}
+\affil[4]{\url{tom.prickett@northumbria.ac.uk}}
+
+\renewcommand\Authands{ and }
 \def\UrlBreaks{\do\/\do-}
 
+\date{August 2019}
+
 \begin{document}
-
-\title{A UK Case Study on\\Cybersecurity Education and Accreditation}
-
-% \author{\IEEEauthorblockN{Tom Crick}
-% \IEEEauthorblockA{\textit{School of Education} \\
-% \textit{Swansea University}\\
-% Swansea, UK \\
-% thomas.crick@swansea.ac.uk}
-% \and
-% \IEEEauthorblockN{James H. Davenport}
-% \IEEEauthorblockA{\textit{Department of Computer Science} \\
-% \textit{University of Bath}\\
-% Bath, UK \\
-% j.h.davenport@bath.ac.uk}
-% \and
-% \IEEEauthorblockN{Alastair Irons}
-% \IEEEauthorblockA{\textit{Faculty of Technology} \\
-% \textit{name of organization (of Aff.)}\\
-% Sunderland University, Sunderland, UK\\
-% alastair.irons@sunderland.ac.uk}
-% \and
-% \IEEEauthorblockN{Tom Prickett}
-% \IEEEauthorblockA{\textit{Department of Computing and Information Sciences} \\
-% \textit{name of organization (of Aff.)}\\
-% Northumbria University, Newcastle upon Tyne, UK \\
-% tom.prickett@northumbria.ac.uk}
-% }
-
-% compact IEEE author format for more than three authors...
-\author{\IEEEauthorblockN{Tom Crick\IEEEauthorrefmark{1}, James
-    H. Davenport\IEEEauthorrefmark{2}, Alastair
-    Irons\IEEEauthorrefmark{3} and Tom Prickett\IEEEauthorrefmark{4}} 
-\IEEEauthorblockA{\IEEEauthorrefmark{1}%School of Education, 
-Swansea University, Swansea, UK; Email: thomas.crick@swansea.ac.uk} 
-\IEEEauthorblockA{\IEEEauthorrefmark{2}%Dept.~Computer Science, 
-University of Bath, Bath, UK; Email: j.h.davenport@bath.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}%Faculty of Technology, 
-Sunderland University, Sunderland, UK; Email: alastair.irons@sunderland.ac.uk}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}%Dept.~Computer\&Information Sciences, 
-Northumbria University, Newcastle upon Tyne, UK; Email: tom.prickett@northumbria.ac.uk}}
-
 \maketitle
 
+% TC: merge in abstract into main body -- perhaps parts useful as opening preamble?
 \begin{abstract}
-This Innovative Practice Full Paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are prioritised, implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as various other important facets).
+This paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are prioritised, implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as various other important facets).
 A multitude of model curricula and recommendations have been presented and discussed in international fora in recent years, with varying levels of impact on education, policy and practice.
@@
 
-\begin{IEEEkeywords}
-Cybersecurity, curricula, accreditation, computer science education, public policy, UK
-\end{IEEEkeywords}
+% \noindent {\footnotesize{{\textbf{Keywords}}: Cybersecurity, curricula, accreditation, computer science education, public policy, UK}}
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 \begin{quote}
-``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''\\
-\hfill Robert Hannigan~\cite{Hannigan2019a}, former Director of GCHQ
+``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}'' Robert Hannigan~\cite{Hannigan2019a}, former Director of GCHQ
 \end{quote}
@@
 
-\section{Cybersecurity For All, Or Just For Specialists?}
+\section*{Cybersecurity For All, Or Just For Specialists?}
 
@@
 
-% have merged into previous section
-%UK national economic skills priority
-%e.g. \url{https://www.gov.uk/government/publications/national-cyber-security-strategy-2016-to-2021}
-%(esp. \url{https://publications.parliament.uk/pa/jt201719/jtselect/jtnatsec/706/70605.htm}
-%(especially \cite{JCNSS2018a}
-%and \url{https://www.ncsc.gov.uk/blog/skills-and-training} and
-%\url{https://www.cybersecuritychallenge.org.uk/}). 
-%There is large-scale media attention on the ``cybersecurity skills gap''
-%e.g. \url{https://www.contracts.mod.uk/do-features-and-articles/digital-skills-shortage-threatens-uk-cyber-security/}
-%and
-%\url{https://www.itpro.co.uk/cyber-security/31554/uk-government-lacks-urgency-in-tackling-cyber-security-skills-gap}
-%etc -- link to UK Digital Strategy and UK Industrial Strategy.
+% \section*{Methodology}
 
-% need a better title for this!
-% \subsection{Some Examples}
+% \subsection*{Research Questions}
 
-% To illustrate some of these challenges, four examples are briefly explored: the European Union's General Data Protection Regulation (GDPR); Agile development; one industry sector (with rigorous demands for secure software); and impact of these cases upon one another.
+% There are various levels of specialism at which cybersecurity education and skills can be addressed; for example:
 
-% On the one hand, GDPR (equally, the  California Consumer Privacy Act) has increased the corporate penalties for failure, and therefore the demand for various forms of cybersecurity specialists, both explicitly by requiring Data Protection Officers\footnote{\cite{EU292016a} places quite strong requirements on these: ``expertise in national and European data protection laws and practices and an in-depth understanding of the GDPR'', as well as ``principles of data processing'', ``data protection by design and by default'' and ``security of processing''.
-It is hard to envisage one professional possessing all these attributes and capabilities, and indeed ``DPO as a service'' is being promoted \cite{McCreanor2018d}.} and implicitly by causing organisations to re-think models for governance and assurance, and to invest more in cybersecurity.
-On the other hard, the growing publicity given to these, and the growing use of electronic payments, has also led to greater demand for expertise in this area.
-% % JHD what are these?
-Also the next paragraph needs thinking about - TP - I added the next paragraph but now wonder if it should be deleted - I wonder if there is a point to be made about ''privacy by design and ''privacy by default'' but agile is covered better later.
+% \begin{enumerate}[label=(\roman*)]
+% \item The generalist computer science graduate;
+% \item The generalist computer science masters graduate;
+% \item The specialist computer science graduate;
+% \item The specialist computer science masters graduate;
+% \item The reskilling/upskilling/professional development of the IT industry and the wider workforce;
+% \item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
+% \end{enumerate}
 
-% %AI: GDPR essentially expects Cybersecurity to be in all coding. 
-% In particular, for the development of software that will operate within Europe and that processes personal data, the expectation of ``Privacy by Design'' and ``Privacy by Default'' have significant legal implications.
-However, ``Privacy by Design'' is no good without ``Security by Implementation'', and many security weaknesses are implementation weaknesses.
-% % can we evidence this?
+% The focus of this paper is on {\emph{(i)}}--{\emph{(ii)}}: the general computer science graduate. We thus focus on three research questions:
 
-% A number of studies have documented the challenges of aligning security methods in Agile Practices \cite{vanderHeijden:2018:EPS:3239235.3267426,Bartsch2011a}.
-This further emphasises the need for mainstream software developers and by extension computer science graduates to be well-versed in relevant security practices.
+% \begin{description}
+% \item[RQ1] {\emph{What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?}}
+% \item[RQ2] {\emph{Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?}}
+% \item[RQ3] {\emph{Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?}}
+% \end{description}
 
-% This can be further illustrated by the demands of particular sectors for secure software.
-Considering the case of the Payments Industry \cite[6.6]{PCI2018b} which insists on either independent code reviews by security specialists or a Web Application Firewall: essentially saying that the normal development process, even after developers are trained annually in secure coding practices \cite[6.5]{PCI2018b}, cannot be trusted to deliver secure code.
+% \subsection*{Research Approach}
 
-\section{Methodology}
-
-\subsection{Research Questions}
-
-There are various levels of specialism at which cybersecurity education and skills can be addressed; for example:
-\begin{enumerate}[label=(\roman*)]
-% ({\emph{interesting point from a public engagement/science communication perspective -- aren't ``we'' trying to create a digitally competent and capable citizenry?
-Minor point but maybe worth mentioning e.g.
-being safe online, accessing public services, etc}})
-%\item[JHD]This is important, agreed, but I think this is being a user: think car driving rather than car design.
-BUT the point I heard from the WC Engineers ``how can security be an option --- safety certainly isn't'' is also relevant.
-\item The generalist computer science graduate;
-\item The generalist computer science masters graduate;
-\item The specialist computer science graduate;
-\item The specialist computer science masters graduate;
-\item The reskilling/upskilling/professional development of the IT industry and the wider workforce;
-\item The general public --- this is important, but there are many initiatives in this area, which are, rightly, largely separated from computing education.
-\end{enumerate}
-
-The focus of this paper is on (i)--(ii): the general computer science graduate. We thus focus on three research questions:
-
-\begin{description}
-\item[RQ1] {\emph{What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?}}
-\item[RQ2] {\emph{Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?}}
-\item[RQ3] {\emph{Can accreditation by professional, statutory and regulatory bodies (PSRBs) enhance the provision of cybersecurity within a body's jurisdiction?}}
-\end{description}
-
-\subsection{Research Approach}
-
-A UK-focused case study-based approach has been adopted for this project.
+% A UK-focused case study-based approach has been adopted for this project.
 As is common in case study-based research, many alternative case studies could have been chosen; furthermore, the cases are illustrative rather than comprehensive in terms of the available case studies or challenges.
@@
 
-The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations; this is contextualised by the wider policy context.
+% The first set of cases that are considered are the current situation in terms of recommendations from a sample of relevant PSRBs, together with the published evidence regarding compliance with these recommendations; this is contextualised by the wider policy context.
 Together these provide a context to the ongoing enhancement initiatives in the area of cybersecurity education.
 
-In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of specific case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
+% In order to address {\emph{RQ1}} and {\emph{RQ2}} a number of specific case studies pertaining to challenges of delivering cybersecurity to undergraduate computer science students are evaluated.
 Namely: an industry problem with evident cybersecurity implications; the current state of educational resources with respect to cybersecurity; and the challenges evident in the UK related to the recruitment and retention of suitably qualified academic staff in the cyber security area.
 
-Finally, in order to address {\emph{RQ3}}, case studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
+% Finally, in order to address {\emph{RQ3}}, case studies are evaluated related to the challenges and successes of PSRB accreditation of cybersecurity in undergraduate computer science in one jurisdiction (the UK).
 As discussed, mandating cybersecurity within PSRB accreditation in this jurisdiction is at a reasonably mature stage.
 
-\subsection{Current State of Play}
+\subsection*{Current State of Play}
 
@@
 \begin{description}
-\item[UK] BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015 (Table \ref{table:1}), with the result that all accredited universities should be compliant by 2020 (due to the five-year cycle).
+\item[UK] BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015, with the result that all accredited universities should be compliant by 2020 (due to the five-year cycle).
 More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
@@
 
-\section{Challenges: ({\emph{RQ1}} and {\emph{RQ2}})}
+\section*{Challenges: ({\emph{RQ1}} and {\emph{RQ2}})}
 
@@
 
-%AI noted the workshops with (ISC)$^2$, with the ``this is a fad'' concept.  {\bf AI to write up}.
-  % Should we remove this?
-This raises pedagogic questions which are beyond the scope of this research to resolve.
-Should we be teaching content because it's underpinning?; or because there's a skill shortage?
-
-\subsection{Industry Standards: PCI DSS}\label{sec:PCIDSS}
+\subsection*{Industry Standards: PCI DSS}\label{sec:PCIDSS}
 
@@
 
-\subsection{Educational Resources}\label{sec:EDResource}
+% TC: pick a couple of these as examples?
+\subsection*{Educational Resources}\label{sec:EDResource}
 
-\subsubsection{SQL Injection}\label{sec:SQL}
+\subsubsection*{SQL Injection}\label{sec:SQL}
 
@@
 Clearly such a major weakness should be well-taught; in general it is hard to determine what is actually delivered as part of a specific degree programme, but a reasonable proxy for this is the content of recommended textbooks.
-This was the rationale for a 2019 analysis of database textbooks used by 44 of the top 50 computer science departments in the USA\cite{Drop2019}.
+This was the rationale for a 2019 analysis of database textbooks used by 44 of the top 50 computer science departments in the USA~\cite{Drop2019}.
 There were seven such books, but three books accounted for the 36 of the 44 universities.
@@
 This means a student who is skimming the text looking for an example to modify for their own code could simply copy the code that first appears in the example, without being aware that this is in fact an example of what they shouldn't do.}}''~\cite{Drop2019}}.
-This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself at the time of writing.%\footnote{Fixing this is a suggestion for future work.}
+This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself at the time of writing\footnote{Fixing this is a suggestion for future work.}.
 
-\subsubsection{The Case of Java}\label{sec:Java}
+\subsubsection*{The Case of Java}\label{sec:Java}
 
@@
 popularly exist on SO. Insecure answers dominate, in
-particular, the SSL/TLS category}}''\newline[355 insecure versus 150 secure, i.e. $>70$\%].
+particular, the SSL/TLS category}}'' [355 insecure versus 150 secure, i.e. $>70$\%].
 \end{quote}
 
-\subsubsection{Android}\label{sec:Android}
+\subsubsection*{Android}\label{sec:Android}
 
@@
 
-\subsubsection{Agile}
+\subsubsection*{Agile}
 
@@
 
-\subsubsection{Informal Resources}\label{sec:informal}
+\subsubsection*{Informal Resources}\label{sec:informal}
 
@@
 
-\subsection{Faculty}
+\subsection*{Faculty}
 
-It is well known that cybersecurity skills are in short supply, in both industry~\cite{Ackerman2019a} and academia~\cite{schneider2013}.
-For example, research into the state of IT conducted annually by Enterprise Strategy Group (ESG) has revealed that the skills gap in information security continues to widen and has doubled in the past five years; in 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills -- this had climbed to 51\% at the beginning of 2018~\cite{ESG:2018}.
-Clearly, cybersecurity is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom~\cite{Ackerman2019a}.
+It is well known that cybersecurity skills are in short supply, in both industry~\cite{Ackerman2019a} and academia~\cite{schneider2013}. For example:
+
+\begin{quote}
+{\emph{Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)~\cite{ESG:2018} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
+In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
+This had climbed to 51\% at the beginning of this year.
+Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.}}~\cite{Page2018a}
+\end{quote}
 
@@
 
-% \begin{quote}
-% ``{\emph{Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)\footnote{Apparently \url{https://www.esg-global.com/research/esg-brief-2018-cybersecurity-spending-trends}.} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
-In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
-This had climbed to 51\% at the beginning of this year.
-Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.}}''~\cite{Page2018a}
-% \end{quote}
-
-%{\bf JHD to ? Poll CPHC, and/or survey job adverts.}
-\section{Accreditation by PSRBs ({\emph{RQ3}})}
+\section*{Accreditation by PSRBs ({\emph{RQ3}})}
 
@@
 
-\subsection{National Cyber Security Centre (NCSC)}
+\subsection*{National Cyber Security Centre (NCSC)}
 
@@
 
-\subsection{Not-for-Profit Organisations}
+\subsection*{Not-for-Profit Organisations}
 
@@
 
-\subsection{Professional Bodies and Learned Societies}
+\subsection*{Professional Bodies and Learned Societies}
 
@@
 
-\section{Case Study: BCS Accreditation}
+% TC: main case study
+\section*{Case Study: BCS Accreditation}
 
 In recent years the BCS has evolved its accreditation practices to promote and mandate the inclusion of cybersecurity within the programmes the body accredits.
-The timeline which this process has followed in presented in Table~\ref{table:1}, and will be discussed in more detail in the following case study.
+
+% The timeline which this process has followed in presented in Table~\ref{table:1}, and will be discussed in more detail in the following case study.
 
 %\begin{center}
-  \begin{table}[h!]
-  \caption{Timeline of the development of Cybersecurity expectations}
-  \label{table:1}
-  \begin{tabular}{ | p{6cm} |p{1.5cm} |}
-    \hline
-    IV. ACCREDITATION ({\emph{RQ3}}) - A. Developing Expectations &  \\ \hline
-    UK Government Cybersecurity Strategy \cite{ukcyberstrategy:2016} & November 2011 \\ \hline
-    Three workshops of a consortium of industry, academia and government bodies -- led by CPHC and  (ISC)$^2$ -- leading to the development of cybersecurity learning guidelines to be embedded into BCS accredited UK computer science and IT-related degrees~\cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
-    UK Government report Cybersecurity Skills, Business Perspectives and Government's Next Steps Report Released \cite{UKCabinetOffice2014} & March 2014  \\ \hline
-    Council of Professors and Heads of Computing (CPHC) Identifies Cybersecurity as one the top 3 concerns in Computing & April 2014 \\ \hline
-    Joint Development of White Paper from CPHC and The International Information Systems Security Certification Consortium (ISC)$^2$ \cite{CPHCISC2014} & April -November 2014 \\ \hline
-    Extended Cybersecurity Criteria included in BCS Accreditation Guidelines \cite{BCS2018a}& June 2015 \\
-    \hline
-    IV. ACCREDITATION ({\emph{RQ3}}) - B. What does the BCS tell Universities? & \\ \hline
-    Cybersecurity Principles Roadshow & March-April 2016 \\ \hline
-    IV. ACCREDITATION ({\emph{RQ3}}) - C. Accreditation - what progress has been made? &  \\ \hline
-    All visited institutions expected to be fully compliant 
- (BCS follow a five-year accreditation cycle)  & September 2020\\ \hline
-  \end{tabular}
-  \end{table}
+%   \begin{table}[h!]
+%   \begin{tabularx}{\textwidth}{ |X|X| }
+%     \hline
+%     IV. ACCREDITATION ({\emph{RQ3}}) - A. Developing Expectations &  \\ \hline
+%     UK Government Cybersecurity Strategy \cite{ukcyberstrategy:2016} & November 2011 \\ \hline
+%     Three workshops of a consortium of industry, academia and government bodies -- led by CPHC and  (ISC)$^2$ -- leading to the development of cybersecurity learning guidelines to be embedded into BCS accredited UK computer science and IT-related degrees~\cite{CPHCISC2}  & 2013 to June 2015 \\ \hline
+%     UK Government report Cybersecurity Skills, Business Perspectives and Government's Next Steps Report Released \cite{UKCabinetOffice2014} & March 2014  \\ \hline
+%     Council of Professors and Heads of Computing (CPHC) Identifies Cybersecurity as one the top 3 concerns in Computing & April 2014 \\ \hline
+%     Joint Development of White Paper from CPHC and The International Information Systems Security Certification Consortium (ISC)$^2$ \cite{CPHCISC2014} & April -November 2014 \\ \hline
+%     Extended Cybersecurity Criteria included in BCS Accreditation Guidelines \cite{BCS2018a}& June 2015 \\
+%     \hline
+%     IV. ACCREDITATION ({\emph{RQ3}}) - B. What does the BCS tell Universities? & \\ \hline
+%     Cybersecurity Principles Roadshow & March-April 2016 \\ \hline
+%     IV. ACCREDITATION ({\emph{RQ3}}) - C. Accreditation - what progress has been made? &  \\ \hline
+%     All visited institutions expected to be fully compliant 
+%  (BCS follow a five-year accreditation cycle)  & September 2020\\ \hline
+%   \end{tabularx}
+% \caption{Timeline of the development of cybersecurity expectations in the UK}
+%   \label{table:1}
+%   \end{table}
 %\end{center}
 
-\subsection{Developing Expectations}
+\subsection*{Developing Expectations}
 
@@
 
-``{\emph{2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices}}''
+\begin{quote}
+``{\emph{2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices}}'''~\cite{BCS2018a}
+\end{quote}
 
+\begin{quote}
 ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system}}''~\cite{BCS2018a}
+\end{quote}
 
@@
 
+\begin{quote}
 ``{\emph{3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security}}''~\cite{BCS2018a}
+\end{quote}
 
@@
 
-\subsection{What Are Universities Told?}
+\subsection*{What Are Universities Told?}
 
@@
 
-\subsection{Accreditation: What Progress Has Been Made?}
+\subsection*{Accreditation: What Progress Has Been Made?}
 
@@
 
-\section{Conclusions and Future Work}
+\section*{Looking Ahead}
 
@@
 \item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item Web Programming courses should  do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
-Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks}) is silent.} that has Cross-Site Request Forgery protection.
+\item Web Programming courses should do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework that has Cross-Site Request Forgery protection\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
+Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks}) is silent.}.
 \item A study similar to \cite{Drop2019} is recommended for future work focusing on Java development, Android development and Agile development, and potentially other areas of the curriculum, to better understand how modern textbooks support key areas of the curriculum.
@@
 
-\section*{Acknowledgements}
+% \section*{Acknowledgements}
 
-The authors wish to thank Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
+% The authors wish to thank Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
 Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
@@
 
-{\emph{N.B.}} The first two authors are current Vice-Presidents of the BCS, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
+% {\emph{N.B.}} The first two authors are current Vice-Presidents of the BCS, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
 
-%\IEEEtriggeratref{40}
-
-\bibliographystyle{IEEEtran}
+% bib
+\bibliographystyle{unsrt}
 \bibliography{FIE2019}

### 2019-08-19 ITNow_TP.tex
Pruned version of ITNow article
1798 words (without author info) - Acknowledgements is 118 words

--- 2019-07-30 ITNow.tex
@@
 
-\title{Cybersecurity Education and Accreditation in the UK}
+\title{Cybersecurity in UK Higher Education?}
 
@@
 % TC: merge in abstract into main body -- perhaps parts useful as opening preamble?
-\begin{abstract}
-This paper presents a national case study-based analysis of the numerous dimensions to cybersecurity education and how they are prioritised, implemented and accredited; from understanding the interaction of hardware and software, moving from theory to practice (and vice versa), to human factors, policy and politics (as well as various other important facets).
-A multitude of model curricula and recommendations have been presented and discussed in international fora in recent years, with varying levels of impact on education, policy and practice.
-This paper address three key questions: {\emph{i)}} what is taught and what should be taught for cybersecurity to general computer science students; {\emph{ii)}} should cybersecurity be taught stand-alone or in an integrated manner to general computer science students; and {\emph{iii)}} can accreditation by national professional, statutory and regulatory bodies enhance the provision of cybersecurity within a body's jurisdiction?
 
-Evaluating how cybersecurity is taught in all aspects of computer science is clearly a task of considerable size, one that is beyond the scope of this paper.
-Instead a case study-based research approach -- primarily focusing on the UK -- has been adopted to evaluate the evidence of the teaching of cybersecurity within general computer science to university-level students.
-Thus, in the context of widespread international computer science/engineering curriculum reform, what does this need to embed cybersecurity knowledge and skills mean more generally for institutions and educators, and how can we teach this subject more effectively?
-Through this UK case study, and by contrasting with related initiatives in the US, we demonstrate the positive effect that national accreditation requirements can have, and offer some recommendations both for future research and curriculum developments.
-\end{abstract}
+\begin{strapline}
+The BCS has been promoting further inclusion of Cybersecurity within general Computer Science University Education since 2010. We explore progress and some of the outstanding related issues.
+
+\end{strapline}
 
@@
 
-Cybersecurity has increasingly been a headline feature in the news over recent years, generally prompted by spectacular breaches, including major credit reference agencies~\cite{equifax2017}, telecoms companies~\cite{talktalk2016}, national airlines~\cite{BritishAirways2018a}, online dating websites~\cite{ashleymadison2015}, and even between sovereign governments~\cite{ncsc2018}.
-These major breaches have had significant impact on both individual citizens and society in general, requiring attention from organisations of all sizes:
+Cybersecurity has increasingly been a headline feature in the news over recent years, generally prompted by spectacular breaches, including major credit reference agencies (Equifax), telecoms companies (Talk Talk), national airlines BritishAirways, online dating websites (Ashley Madison), and even between sovereign governments (The UK National Cyber Security  Centre has reported Rusian activity).
 
-\begin{quote}
-``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}'' Robert Hannigan~\cite{Hannigan2019a}, former Director of GCHQ
-\end{quote}
-
-These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals~\cite{McGettrick2013}.
-As no shared framework for ``cybersecurity'' as an academic discipline exists, growth has been unfocused and largely driven by training materials, which makes it harder to establish a common body of knowledge (for example, in the UK, the CyBOK project~\cite{Bristol2019a} is still a work in progress).
-An international perspective is harder still, as different nations use different criteria to define local needs~\cite{schneider2013}.
-As a result, new programmes entering this space are free to conceptualise, design, package and market their initiatives, as there is no globally accepted reference model for cybersecurity to allow employers or students to understand the extent or ambition of a given cybersecurity program~\cite{conklin-et-al:2014,Parrishetal2018a}.
-
-With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as tertiary -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a,NCSC2017}.
-This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}.
-
-An ACM working group~\cite{Parrishetal2018a}, established in 2018 as part of the {\emph{Innovation and Technology in Computer Science Education}} (ITiCSE) conference series, has been capturing global perspectives on cybersecurity education, but as of July 2019 has yet to report its full findings.
-An aim of the ITiCSE working group is to develop a taxonomy for approaches to cybersecurity education, resulting in improved standards and goals for many different types of cybersecurity programmes; a further aim is to ``{\emph{catalog existing [\dots] knowledge materials}}'', but there is no mention of any quality control over these.
-Nevertheless, it is one thing to write national curricula, specifications and requirements, and another thing to deliver appropriate and relevant education and skills; this paper asks how well this is done in practice.
+These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals.
+With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as tertiary -- to respond to this situation, at the individual level and via recommended curricula and professional accreditation requirements.
+This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK.
 
@@
 
-In one sense, this title is a false dichotomy: there is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth~\cite{manson+pike:2014,davenport-et-al:latice2016}.
-This is not a new concern~\cite{Parr2014a}, but it is a growing one.
-
-This need to build knowledge, skills and capacity in the area of cybersecurity has also led to the establishment of a number of strategic policy initiatives from a number of national governments, for example the publication in 2016 of the UK's Cyber Security Strategy~\cite{ukcyberstrategy:2016} (along with the setting up of the National Cyber Security Centre, as well as increased scrutiny of the resilience of the UK's critical national infrastructure~\cite{lordscyberreport:2018}; also industry-focused initiatives such as {\emph{Cyber Essentials}}\footnote{{\emph{Cyber Essentials}} is a UK Government-backed, industry-supported scheme to help organisations protect themselves against common online threats: \url{https://www.cyberessentials.ncsc.gov.uk}}), the EU Cybersecurity Act~\cite{eucyber2018} (which reinforces the mandate of the EU Agency for Cybersecurity: ENISA, the European Union Agency for Network and Information and Security), or the National Initiative for Cybersecurity Education (NICE) in the USA~\cite{NICE}.
-
-The teaching of cybersecurity in higher education pre-dates these initiatives and there has been recognition of the need for the inclusion of cybersecurity as part of the discipline of computer science for a number of years~\cite{Hentea2006,schneider2013}.
-There have been a number of international initiatives to define curricula to support this, for example the 2013 ACM Computer Science Curricula Recommendations~\cite[which added ``{\emph{Information Assurance and Security}}'' for the first time]{ACM2013a}, as well as for specialised cybersecurity degree programmes~\cite{ACMIEEEAISSIGSECIFIP}.
-There has been debate as to whether cybersecurity is distinct discipline from computer science \cite{McGettrick2013}; the consensus increasingly is that cybersecurity is both a discipline in its own right and that cybersecurity should be taught within computer science and related degrees.
-Recent evolution of cybersecurity education shows that it has begun to take shape as a true academic perspective, as opposed to simply being a training domain for certain specialised jobs.
-More recent work presents cybersecurity as a ``meta-discipline''; that is, cybersecurity should be used as an aggregate label for a wide variety of similar disciplines, much in the same way that the terms ``engineering'' and ``computing'' are commonly used~\cite{Parrishetal2018b}.
-
-Further to the substantial computer science and digital skills curriculum reform across the UK~\cite{crick+sentance:2011,wgictreview:2013,brown-et-al:toce2014,tryfonas+crick:petra2018}, the question is raised whether cybersecurity should be formally taught in schools, as part of compulsory school-level education.
-While aspects of ``e-safety'' and principles of protecting personal data are increasingly visible in formal curricula~\cite{moller+crick:jce2018}, the majority of UK schoolchildren's exposure to cybersecurity skills is through national extra-curricular competitions, for example Cyber Security Challenge UK\footnote{\url{https://www.cybersecuritychallenge.org.uk}}.
-The Institute of Coding, a \pounds40m+ initiative by the UK Government to transform the digital skills profile of the country~\cite{Davenportetal2019a} -- but primarily focused on university graduates -- does indeed mention cybersecurity, but merely as a sub-item in one work package.
+There is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth.
+There has been recognition of the need for the inclusion of cybersecurity as part of the discipline of computer science for a number of years.
 
@@
 
-\subsection*{Current State of Play}
-
-The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs) related to {\emph{RQ2}}:
-
-\begin{quote}
-``{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
-given the manner in which the topics are pervasive throughout other KAs.}}''
-\end{quote}
-
-It proposes nine ``core'' hours and 63.5 distributed across the other KAs. Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:
-
-\begin{quote}
-``{\emph{Universities suffer shortcomings, as well.
-Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity.
-There is a big catch, however.
-Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}''
-\end{quote}
-
-The UK situation is distinctly different: 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping~\cite[Table 1]{Ruiz2019a}.
-As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.
-It is at least plausible to attribute this difference to differences in the accreditation regimes, as the external circumstances, governmental pressures, and professional body/learned society curricula are all similar.
-
-\begin{description}
-\item[UK] BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015, with the result that all accredited universities should be compliant by 2020 (due to the five-year cycle).
-More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
-\item[USA] The Association of Computing Machinery (ACM) has equally had cybersecurity (IAS: ``Information Assurance and Security'') in the curriculum since 2013~\cite{ACM2013a}, but it is not the accrediting body.
-The Accreditation Board for Engineering and Technology (ABET) is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''.
-This should mean that  all accredited universities should be compliant by 2025 (due to their six-year cycle).
-\end{description}
-
-\section*{Challenges: ({\emph{RQ1}} and {\emph{RQ2}})}
-
-%What should be in the general curriculum.  Pervasive (as in Maths or LSEPI) rather than specialist option.  
-The ACM Computer Science Curricula Recommendations~\cite{ACM2013a} states three Tier-1 and six Tier-2 hours for ``Information Assurance and Security'', but this is the ``IAS-only'' topics, and ACM expects 32 Tier-1 and 31.5 Tier-2 hours for IAS topics embedded in other Knowledge Areas.
-
-The UK's official body of knowledge resource, the CyBOK project~\cite{Bristol2019a}, has produced reference documentation for some (as of June 2019: five final, seven for comment, out of a planned total of 19) knowledge areas, which are useful references for the experienced educator looking for a definition or characterisation, but a long way from being a textbook (which is not their aim).
-% JHD updated figures 21/6/2019
-
-\subsection*{Industry Standards: PCI DSS}\label{sec:PCIDSS}
-
-Is teaching cybersecurity different from other computing topics?
-Lecturing is probably not the best way; how should we use real-life case studies (e.g.
-British Airways \cite{Barth2018a}) and authentic assessment?
-Recent work suggests that there are benefits from teaching this discipline area in a more practical manner \cite{Weiss:2013:THC:2527148.2527180}; one would be by the inclusion of pertinent cybersecurity standards in the curricula.
-
-The {\emph{Payments Card Industry Data Security Standards}} (PCI DSS)~\cite{PCI2018b} is one such standard, underpinning all processing of credit/debit cards globally.
-Nevertheless, they are very rarely mentioned in generalist computer scientist courses.
-This would not matter so much if everyone handling payments data were sent by their employers on an effective PCI DSS course.
-However, the payments business is now so spread across websites, often run by small and medium enterprises (SME), or non-specialists.
-Even larger enterprises are not immune: the 2018 British Airways breach was caused by a failure to adhere to PCI DSS in the website maintenance~\cite{Barth2018a}.
-
-It is an interesting question as to whether standards such as PCI DSS should be addressed within degree courses (clearly degree courses can never cover all standards) or whether they should be addressed in professional training courses.
-However it appears the current situation is not ideal from the perspective of industry (or users of systems) and the inclusion of key standards could be seen as a valuable enhancement activity to the coverage of cybersecurity within generalist computer science courses.
+\section*{Challenges} \label{sec}
 
@@
 
-\subsubsection*{SQL Injection}\label{sec:SQL}
+SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 , and has been in the Top 10 since at least 2003.
 
-It is 15 years since \cite{Guimaraesetal2004} wrote ``{\emph{All the topics listed above should be presented in the first Database Course}}'', and the first such topic was SQL injection \cite{SPIDynamics2002,Anonymous2018b}.
-SQL injection as an attack has been around for twenty years \cite{HornerHyslip2017a}, has its own cartoon and website\footnote{See: \url{https://xkcd.com/327/} (dating back to 2007, according to the Internet Archive) and \url{http://bobby-tables.com/}}.
-Nevertheless SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 \cite{OWASP2017a}, and has been in the Top 10 since at least 2003.
-CyBOK states ``{\emph{\ldots a wide range of attack techniques for exploiting SQL injection or script injection are known and documented.}}''~\cite{Bristol2018a}.
-
-Clearly such a major weakness should be well-taught; in general it is hard to determine what is actually delivered as part of a specific degree programme, but a reasonable proxy for this is the content of recommended textbooks.
+It is hard to determine what is actually delivered as part of a specific degree programme, but a reasonable proxy for this is the content of recommended textbooks.
 This was the rationale for a 2019 analysis of database textbooks used by 44 of the top 50 computer science departments in the USA~\cite{Drop2019}.
@@
 Five of the seven (30 of the 44) had no mention of SQL injection.
-On the other two, the more popular one has a seriously flawed discussion\footnote{``{\emph{However, the paper implies that using parameters is equivalent to using a function to add escape characters around user input.
-This is incorrect, as using parameters allows SQL statements to be pre-compiled, and prevents any user input from being interpreted as code, while escaping user input is not recommended as a sole defense since imperfect escape functions can easily be subverted.}}''~\cite{Drop2019}}, and the other, while generally excellent, had a presentational problem\footnote{``{\emph{However, the fact that the first example should not be used is not discussed until two pages after the example in the text, and is not mentioned at all in the caption or on the page where the figure appears.
-This means a student who is skimming the text looking for an example to modify for their own code could simply copy the code that first appears in the example, without being aware that this is in fact an example of what they shouldn't do.}}''~\cite{Drop2019}}.
-This oversight is not limited to textbooks: although Wikipedia has an article on SQL Injection, it was not linked from the SQL page itself at the time of writing\footnote{Fixing this is a suggestion for future work.}.
+On the other two, the more popular one has a seriously flawed discussion.
 
@@
 
-A recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for  Java textbooks.
-Indeed, many such books go nowhere near security applications.
-But this means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources.
+Many Java books go nowhere near security applications.
+This means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources.
 \cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack Overflow online resource.
-53\% were about the Spring Security framework\footnote{\url{https://projects.spring.io/spring-security/}}, dominated by authentication (45\%).
+53\% were about the Spring Security framework, dominated by authentication (45\%).
 The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.
@@
 There were no negative comments about this, and indeed a typical response is ``{\emph{Adding}} \verb!csrf().disable()!
-{\emph{{solved the issue!!!
-I have no idea why it was enabled by default}}''.
-As of writing, there were no negative comments about this disabling of a vital security feature.
-This research was further developed by \cite{Chenetal2019a}  (and popularised in a security community in \cite{Zorz2019a}).
-Their first finding was:
-
-\begin{quote}
-``{\emph{644 out of the 1,429 inspected answer posts
-(45\%) are insecure, meaning that insecure suggestions
-popularly exist on SO. Insecure answers dominate, in
-particular, the SSL/TLS category}}'' [355 insecure versus 150 secure, i.e. $>70$\%].
-\end{quote}
+{\emph{{solved the issue!!! I have no idea why it was enabled by default}}''. As of writing, there were no negative comments about this disabling of a vital security feature.}
 
@@
 
-Another recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for Android textbooks; \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications.
-The key finding was:
+For Android textbooks; \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. The key finding was:
 
@@
 
-Two caveats (in opposite directions) should be noted.
-The labelling was conservative, in that snippets were only labelled as insecure if that was demonstrable, and, for example, mere use of outdated SSL/TLS was not automatically deemed insecure.
-On the other hand, the insecure snippet might have been used in a way that did not expose the insecurity.
-The uncritical reading of Stack Overflow was also noted in \cite[Slide 29]{Votipkaetal2019a}.
-Their key recommendation~\cite[Slide 32]{Votipkaetal2019a} was ``{\emph{Improve documentation: Clarify what you can(not) copy/paste}}''.
-
-\subsubsection*{Agile}
-
-A further recommendation for future work is a comprehensive survey equivalent to \cite{Drop2019} for ``Agile'' textbooks.
-Many authors have found disconnects between Agile practices and secure software development: notably \cite{Bartsch2011a} for small projects and \cite{vanderHeijden:2018:EPS:3239235.3267426} for large projects.
-Agile's preference for functionality over non-functional requirements is clearly displayed in practice.
-\cite{Naiakshinaetal2017a} asked 20 student developers to imagine they were part of a team working on creating a social networking site for our university and to implement a password storage mechanism for this.
-10 (``primed'') were explicitly told that the storage had to be secure and 10 (``unprimed'') were not.
-None of the unprimed ones implemented any security.
-This would also reaffirm the point raised in section~\ref{sec:PCIDSS}, regarding the effective use of industrial case studies and authentic assessment approaches.
-
 \subsubsection*{Informal Resources}\label{sec:informal}
@@
 This has been looked at by \cite{Unruhetal2017a}, taking the top five search results from Google for six queries.
-Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting\footnote{Number 7 in OWASP's Top Ten \cite{OWASP2017a}.} weaknesses.
+Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting (Number 7 in OWASP's Top Ten) weaknesses.
 Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
-Some students clearly make use of these resources; thus a recommendation of future work is to explore and evaluate students' (and indeed others') use of such informal resources.
 
-\subsection*{Faculty}
+\subsection*{Who will teach the Cybersecurity?}\label{sec:staffing}
 
-It is well known that cybersecurity skills are in short supply, in both industry~\cite{Ackerman2019a} and academia~\cite{schneider2013}. For example:
-
-\begin{quote}
-{\emph{Research into the state of IT conducted annually by Enterprise Strategy Group (ESG)~\cite{ESG:2018} has revealed that the skills gap in information security continues to widen and has doubled in the past five years.
-In 2014, 23\% of respondents to the survey stated that their organisation had a problematic shortage of information security skills.
-This had climbed to 51\% at the beginning of this year.
-Clearly, this is an issue which is being felt across many industries and organisations, and is a concern which extends beyond IT leadership into the boardroom.}}~\cite{Page2018a}
-\end{quote}
-
-The ESG survey is international, but ESG have confirmed that the UK figures are very similar.
-In the UK, there has been a resurgence of job adverts to recruit academic staff with specialisms in cybersecurity over the past three years.
+It is well known that cybersecurity skills are in short supply, in both industry and academia.
 The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
 As universities expand their cybersecurity provision it is not uncommon to find multiple jobs advertised at the same time.
-Recent example have included a professor of cybersecurity, two senior academic positions and two junior academic positions in one advert.
-There are other examples in the UK of cybersecurity lecturing jobs remaining unfilled for longer than a year; there are also examples of cybersecurity research groups moving en masse from one university to another.
 
-\section*{Accreditation by PSRBs ({\emph{RQ3}})}
+\section*{Accreditation of degree courses }
 
@@
 A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields, such as computer security, digital/computer forensics, etc).
-The UK's Universities and Colleges Admissions Service (UCAS) lists over 40 higher education institutions providing undergraduate qualifications related to cybersecurity for entry in September 2019.
-An even larger number of UK institutions provide study opportunities related to more general undergraduate computer science: UCAS lists 246 providers for undergraduate programmes for the 2019/2020 academic year.
-
-Accreditation has evolved to directly addresses the cybersecurity challenges in both general computer science programmes and specialist cybersecurity programmes.
-In the UK, accreditation in the broad computing area is being performed by a few organisations with overlapping interests and priorities.
-These include:
-
-\subsection*{National Cyber Security Centre (NCSC)}
-
-The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
-The NCSC publishes and accredits to a number of cybersecurity standards~\cite{NCSC2017}; these standards are linked to the ACM recommendations for curricula~\cite{ACM2013a}.
-To date, the major focus of NCSC accreditation has been upon Masters degrees specialising in cybersecurity.
-More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity, and computer science degrees with a significant cybersecurity focus.
-
-The NCSC have accredited 15 cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters cybersecurity programmes and 1 cybersecurity degree programme, with a further 2 provisionally accredited.
-Hence the extent of accreditation is currently reasonably limited in terms of reach to computer science programmes.
-This appears to be a positive initiative that will further develop over the next few years, especially with increasing industry engagement in this area.
-
-\subsection*{Not-for-Profit Organisations}
-
-Tech Partnership Degrees is a not-for-profit organisation that provides endorsements to higher education programmes with specific curricula elements aimed at job market requirements.
-One of the curricula elements is related to cybersecurity.
-Tech Partnership Degrees have a specialist scope, endorsing programmes in the named areas of ``IT Management for Business'' and ``Software Engineering for Business''.
-Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and five Software Engineering for Business Programmes.
-As such, Tech Partnership Degrees currently have limited impact upon more general computer science education and none upon specialist cybersecurity education.
-
-The Institute of Coding\footnote{\url{https://instituteofcoding.org}} is a not-for-profit entity, led by the University of Bath, that intends to enhance how digital skills are developed in higher education in the UK~\cite{Davenportetal2019a}.
-This is likely to include cybersecurity related skills.
-Like the Tech Partnership, the focus is upon job market requirements.
-Additionally the Institute is looking at potentially endorsing the demonstrable capabilities of graduates as shown by their university studies, work experience and work placements.
-Given the size of this initiative this potentially has a significant role to play however at time of writing it is clearly a work in progress.
-
-\subsection*{Professional Bodies and Learned Societies}
-
-Both BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) accredit programmes in the general area of computer science and the more specialist area of cybersecurity discipline areas.
-The accreditation provided by these institutes are underpinned by international initiatives such as the Washington Accord\footnote{\url{http://www.ieagreements.org/accords/washington/}.} and the Seoul Accord\footnote{\url{https://www.seoulaccord.org/}.}.
-These memoranda support the internationalising of the curriculum and promote consistency and parity in computer science education globally.
-These professional bodies are also UK charities and hence have responsibilities for public good which extends beyond short term job market needs~\cite{Stensaker2006,Mutereko2017}.
-Both the BCS and the IET have a long history of expecting coverage of environmental factors within the programmes they accredit.
-The BCS has for a number of years been expecting significant coverage of what it terms  {\emph{Legal, Social, Ethical and Professional Issues}}~\cite{Brooke2018}.
-Clearly cybersecurity has been and continues to be part of these expectations.
 
 % TC: main case study
-\section*{Case Study: BCS Accreditation}
+\section*{BCS Accreditation}
 
-In recent years the BCS has evolved its accreditation practices to promote and mandate the inclusion of cybersecurity within the programmes the body accredits.
+Since 2010 the BCS has evolved its accreditation practices to promote and mandate the inclusion of of first information security (2010) and then cybersecurity (2013) within the programmes the body accredits.
 
@@
 Internationally the expectations regarding both the breadth and depth of the expected cybersecurity coverage has been the subject of much discussion, debate and analysis.
-Like many governments, the UK Government has actively been seeking ways to address this~\cite{UKCabinetOffice2014,ukcyberstrategy:2016}.
-In parallel to the work completed by the ACM~\cite{ACM2013a} in the USA, considerable effort in the UK have been taken to ensure industry, higher education, government and the relevant professional bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society~\cite{Irons2016}.
-In 2013, an initiative was set up by (ISC)$^2$, CPHC (the representative body of UK computer science departments) and the UK Cabinet Office to examine embedding cybersecurity into undergraduate degrees in the UK.
-Three workshops in 2013 and 2015 attempted to define the principles of cybersecurity education and proposed a framework for embedding these principles in UK computer science curricula.
-Attendees at the workshops included industry, professional bodies, UK government departments and more than 30 universities that offer undergraduate computer science degrees.
-This work initially lead to a white paper related to a proposal in the form of a white paper in 2014~\cite{CPHCISC2014}, followed by a set of guidelines in 2015~\cite{CPHCISC2}.
-The BCS agreed to adopt the outputs into their accreditation criteria.
-This was the first time that cybersecurity has been extensively referenced within accreditation criteria for computing and IT-related degrees.
-The fact that cybersecurity is included as a component of the BCS accreditation criteria reflects the importance placed on cybersecurity and the expectation that all computing graduates should have knowledge and skills in cybersecurity as they move towards chartered status.
+Like many governments, the UK Government has actively been seeking ways to address this.
+In parallel to the work completed by the ACM in the USA, considerable effort in the UK have been taken to ensure industry, higher education, government and the relevant professional bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society as discussed in a previous issue of ITNow~\cite{Irons2016}.
 
-The produced reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'') \cite{CPHCISC2} established a baseline of common knowledge, example learning outcome domains for cybersecurity within the computer science courses and guidance on embedding the concepts.
+The produced reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'')  established a baseline of common knowledge, example learning outcome domains for cybersecurity within the computer science courses and guidance on embedding the concepts.
 The document provides specific guidance for embedding and enhancing relevant cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.
-The document suggested five areas of coverage
-
-\begin{itemize}
-    \item Information and risk;
-    \item Threats and attacks;
-    \item Cybersecurity architecture and operations;
-    \item Secure systems and products; and 
-    \item Cybersecurity management.
-\end{itemize}
-
-The ambition of this approach was to influence the curricula of all programmes seeking accreditation (regardless of the precise discipline area); it was not intended to be prescriptive or stifle innovation, however it is intended to promote curricula that would benefit the students upon programmes, their future employers and wider society.
-In this context this is realised as an expectation cybersecurity is an inclusion in all degrees accredited by the BCS.
-e.g.
-the expectation for coverage is true for computer science as well as cybersecurity programmes.
-Two criteria are expected to be covered by all programmes seeking accreditation.
-These are:
-
-\begin{quote}
-``{\emph{2.1.6 Recognise the legal, social, ethical and professional issues involved in the exploitation of computer technology and be guided by the adoption of appropriate professional, ethical and legal practices}}'''~\cite{BCS2018a}
-\end{quote}
-
-\begin{quote}
-``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information system}}''~\cite{BCS2018a}
-\end{quote}
-
-Additionally, programmes seeking Chartered Information Technology Professional (CITP) accreditation also have to cover:
-
-\begin{quote}
-``{\emph{3.1.2 Knowledge and understanding of methods, techniques and tools for information modelling, management and security}}''~\cite{BCS2018a}
-\end{quote}
-
-In the context of BCS accreditation, these requirements imply an exit standard that all students on a programme must be able to demonstrate irrespective of the option choices they have made.
-This means an institution applying for accreditation is expected to provide evidence that the criteria are taught and assessed in a non-trivial manner, to and by all students upon the programme seeking accreditation.
-An institution is expected to provide evidence in the form of programme and module specification documentation and example assessment specifications (coursework and examinations).
-These criteria and the expectation that they are taught and assessed has been present for a number of years.
-
-The relevant BCS accreditation (criteria 2.1.6, 2.1.9 and 3.1.2, as presented above) is not prescriptive, but encourages institutions to embed cybersecurity teaching across a range of subject areas in the computer science curriculum such as programming, software design, databases, networking, architecture.
-In addition there an expectation that there is significant coverage of cybersecurity principles and fundamentals -- either as a stand alone module or as a significant component(s) of other modules.
-This approach differs from the ACM approach~\cite{ACM2013a}, where the expectations are more explicit and the curriculum expectations are specified at a more granular level.
-
-\subsection*{What Are Universities Told?}
-
-The agreed {\emph{Cybersecurity Principles and Learning Outcomes}}~\cite{CPHCISC2} were discussed with the wider education community, led by CPHC.
-A series of workshops took place in 2015 which presented the rationale for embedding cybersecurity in the curriculum of computer science degrees.
-The workshops included case studies from universities who had embedded cybersecurity into their computer science curricula illustrating different approaches to implementation.
-The workshops had 102 attendees from the academic computer science community representing 60 UK Universities.
-
-The {\emph{BCS Guidelines on Course Accreditation}} are published online~\cite{BCS2018a}, with the BCS publishing any changes that have been made~\cite{BCS2018b}.
-When changes are made, the BCS communicates the changes by email and in writing to all the BCS Educational Affiliates, that is all the institutions that seek accreditation from the BCS.
-The expectations for cybersecurity were extended in the June 2015 version of the guidelines for consideration at accreditation visits that took place from September 2015 or later.
-
-This change to the accreditation guidelines is now in an implementation period.
-The accreditation process adopted by the BCS is cyclic in nature; formally, the cycle is five years in duration.
-The new expectations have been implemented as follows: to ensure continuous accreditation, accreditation visits are normally scheduled every five years; at the time of the next visit in this accreditation cycle, accreditation is conditional upon an institution having considered the guidelines and either adjusted the curriculum to meet the new expectations or have a formal plan in place for when and how adjustments will be made.
-It is anticipated that from 2020 the expectation will be all accredited programmes have the new expectations fully embedded.
-
-In the year prior to an accreditation visit, institutions are invited to a briefing from the BCS.
-The intention of the briefing is to help ensure accreditation visits are successful from the perspective of both the BCS and the institution.
-The briefings take place virtually and includes a summary of the process, discussion of recent changes, guidance regarding the application and a summary of common issues that are being seen in other institutions.
-Significant opportunity for seeking clarification is provided.
-One of the issues often highlighted is that not all institutions have yet evolved their programmes to fully address the increased expectation for cybersecurity.
-This is resulting in accreditation being contingent upon an institution taking action to address this short fall or in some cases the withdrawal of accreditation.
-A number of institutions are in the process of adjusting their curricula to meet the new expectations.
-In this case, the BCS notes the changes to programme design, the outputs from which will be scrutinised at the next accreditation visit.
 
@@
 Some of these programmes will be specialist cybersecurity programmes, however the majority will take a different emphasis; this is a work in progress.
-A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines~\cite{BCS2018b}.
+A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines.
 What is being observed is the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
@@
 In summary, this shows that many UK institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
-Clearly not all institutions in the UK necessarily have to apply for PSRB accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
+Clearly not all institutions in the UK necessarily have to apply for BCS accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
 
@@
 
-The work presented in this paper is a first step towards better understanding the nature, design, structure and assessment of cybersecurity education in the UK, through the lens of accreditation.
-It is clear that PSRB accreditation is having a positive effect on universities, supporting wider national policy imperatives.
-It is also clear that there is a significant need for mobilising the international computer science academic community -- alongside some of the existing initiatives presented at the start of this paper -- to continue this focus on cybersecurity education, and provide international comparators, portability and sharing of best practice.
-With regards to the three research questions:
+The work presented in this article is a first step towards better understanding the nature, design, structure and assessment of cybersecurity education in the UK It is clear that we the BCS are having a positive effect on universities via accreditation, supporting wider national policy imperatives.
+However:
 
 \begin{description}
-\item[RQ1:] {\emph{What cybersecurity is taught and what cybersecurity should be taught to the general computer science students?}}
+\item  It  would seem to be necessary to adopt an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
 
-The guidelines from both ACM and BCS are good for general education.
-However, the most important item would seem to be an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
+\item Cybersecurity can be taught largely through other Knowledge Areas is a good idea.
+However, in the current state of education resources (sections \S\ref{sec:EDResource}) may be a counsel caution in this approach.
 
-\item[RQ2:] {\emph{Should cybersecurity be taught stand-alone or in an integrated manner to general computer science students?}}
+\item 
+Our Accreditation  is a valuable tool in improving the standard of cybersecurity teaching, and disseminating good practice, and should continue this.
 
-The recommendation in \cite[p.
-98]{ACM2013a} that cybersecurity be taught largely through other Knowledge Areas is, in abstract, a good idea.
-However, in the current state of education resources (sections \S\ref{sec:EDResource}) may be a counsel caution in this approach.
-It is more important that issues like SQL injection \cite{Drop2019} or correct use of SSL/TLS \cite{Chenetal2019a} be taught somewhere than that they not be taught at all.
-Nevertheless, it is wrong for a complete curriculum to ignore cybersecurity issues for example:
+\item
+This article focused upon University Education.
+What cybersecurity capability (and development needs) exists in terms of the existing workforce,  the person in the street, etc could benefit from a similar exploration.
+\end {description}
 
-\begin{itemize}
-	\item teach SQL without teaching SQL Injection~\cite{Drop2019};
-	\item teach ``web programming'' without teaching Cross-Site Scripting \cite[(XSS)]{OWASP2017a};
-	\item ignore the  impact of GDPR upon systems development;
-	\item uncritically teach the use of Agile development to develop secure systems.
-\end{itemize}
+\section*{Acknowledgements}
 
-The BCS-identified good practice of exploring cybersecurity via a project referencing Industry Standards, possibly PCI DSS, is commended (\S\ref{sec:PCIDSS}).
-
-\item[RQ3:] {\emph{Can accreditation by PSRBs enhance the provision of cybersecurity within a body's jurisdiction?}}
-
-Accreditation (as practised by BCS in the UK) is a valuable tool in improving the standard of cybersecurity teaching, and disseminating good practice, and should continue this.
-The time-lag in adoption of cybersecurity in the USA seems to be caused by the accreditation differences more than any other factors.
-\end{description}
-
-We also have the following specific recommendations:
-
-\begin{enumerate}
-\item Database courses should look carefully at the security aspects of the texts they use, and the examples they quote, on the lines of \cite{Drop2019}.
-\item Web Programming courses should do the same, with emphasis on the avoidance of Cross-Site Scripting, and, for production use, the use of a suitable framework that has Cross-Site Request Forgery protection\footnote{Most modern frameworks do, but this is not often discussed when looking at the advantages of frameworks.
-Even Mozilla's tutorial (\url{https://developer.mozilla.org/en-US/docs/Learn/Server-side/First_steps/Web_frameworks}) is silent.}.
-\item A study similar to \cite{Drop2019} is recommended for future work focusing on Java development, Android development and Agile development, and potentially other areas of the curriculum, to better understand how modern textbooks support key areas of the curriculum.
-\item All computer science courses should emphasise that informal resources should come with a ``security health warning'': see sections \ref{sec:informal} and \ref{sec:Java}.
-One should probably use the data from \cite{Chenetal2019a}: ``{\emph{If you pick up a SSL/TLS answer from Stack Overflow, there's a 70\% chance it's insecure}}''.
-\item Finally, further work exploring and evaluating effective pedagogical and assessment approaches for cybersecurity education, in a similar vein to what has been seen with computer science education more generally (e.g.~\cite{davenport-et-al:latice2016}).
-\end{enumerate}
-
-% \section*{Acknowledgements}
-
-% The authors wish to thank Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
+ The authors wish to thank Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
 Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
@@
 
-% {\emph{N.B.}} The first two authors are current Vice-Presidents of the BCS, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
+ {\emph{N.B.}} The first two authors are current Vice-Presidents of the BCS, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
 

### 2019-09-03 ITNow_TP.tex
Further reduction
Does not reference FIE yet

@@
 
-Cybersecurity has increasingly been a headline feature in the news over recent years, generally prompted by spectacular breaches, including major credit reference agencies (Equifax), telecoms companies (Talk Talk), national airlines BritishAirways, online dating websites (Ashley Madison), and even between sovereign governments (The UK National Cyber Security  Centre has reported Rusian activity).
+Cybersecurity matters, breaches are legion and sadly are a common feature in the news.
+Universities are  addressing the demand for educated cybersecurity professionals.
+There are calls for formal education to respond to this situation.
+This is reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK.
 
-These global cybersecurity crises have compelled academic institutions to address the demand for educated cybersecurity professionals.
-With this significant economic and societal focus on cybersecurity, there are calls for formal education -- school-level as well as tertiary -- to respond to this situation, at the individual level and via recommended curricula and professional accreditation requirements.
-This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK.
+This raises the question, should cybersecurity be of importance for all or just for specialists?
+There is a serious need for cybersecurity specialists, but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for both.
 
-\section*{Cybersecurity For All, Or Just For Specialists?}
-
-There is a serious need for cybersecurity specialists (estimates vary, but are always large), but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for depth as well as breadth.
-There has been recognition of the need for the inclusion of cybersecurity as part of the discipline of computer science for a number of years.
+In this article we explore some of the challenges related to cybersecurity education of general computer scientists (including information technology and other related degrees), and what we, the BCS, have been doing to enhance cybersecurity in education at university.
 
@@
 \section*{Challenges} \label{sec}
-
 % TC: pick a couple of these as examples?
 \subsection*{Educational Resources}\label{sec:EDResource}
-
-SQL injection is still a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 , and has been in the Top 10 since at least 2003.
-
-It is hard to determine what is actually delivered as part of a specific degree programme, but a reasonable proxy for this is the content of recommended textbooks.
-This was the rationale for a 2019 analysis of database textbooks used by 44 of the top 50 computer science departments in the USA~\cite{Drop2019}.
-There were seven such books, but three books accounted for the 36 of the 44 universities.
+It is hard to determine what is actually delivered as part of a degree programme, but a reasonable proxy for this are recommended textbooks and other supporting resources.
+\subsubsection*{The Case of SQL Injection}\label{sec:SQL}
+SQL injection is a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 , and has been in the Top 10 since at least 2003.
+Database textbooks where analysed in 2019 ~\cite{Drop2019}.
+There were seven such books, but three books accounted for the 36 of 44 universities.
 Five of the seven (30 of the 44) had no mention of SQL injection.
@@
 
-Many Java books go nowhere near security applications.
-This means that the programmer who has to implement security is left to the documentation of the package/API being used, and to informal resources.
+Many Java books do not cover security.
+If you want to know about security you need to look at the documentation of the package/API being used, and/or  informal resources.
 \cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack Overflow online resource.
 53\% were about the Spring Security framework, dominated by authentication (45\%).
-The discussion \cite[\S4.3.1]{Mengetal2018a} of cross-site request forgery (CSRF) is especially worrying.
-By default, Spring implicitly enables protection against this.
+One example being, by default, Spring enables protection against cross-site request forgery (CSRF).
 But all the accepted answers to CSRF-related failures simply suggested disabling the check.
 There were no negative comments about this, and indeed a typical response is ``{\emph{Adding}} \verb!csrf().disable()!
-{\emph{{solved the issue!!! I have no idea why it was enabled by default}}''. As of writing, there were no negative comments about this disabling of a vital security feature.}
+{\emph{{solved the issue!!! I have no idea why it was enabled by default}}''. }
 
@@
 
-For Android textbooks; \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow in Android applications. The key finding was:
+For Android, \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow. The key finding was:
 
@@
 The web abounds with informal resources, such as tutorials and code snippets.
-How good are these, and how good are people at using these?
-This has been looked at by \cite{Unruhetal2017a}, taking the top five search results from Google for six queries.
+How good are these?
+One invesigation \cite{Unruhetal2017a}, took the top five search results from Google for six queries.
 Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting (Number 7 in OWASP's Top Ten) weaknesses.
@@
 
-It is well known that cybersecurity skills are in short supply, in both industry and academia.
+Cybersecurity skills are in short supply, in both industry and academia.
 The demand for cybersecurity skills in industry makes it difficult for academia to attract academics with knowledge, practical experience, research background and academic aspirations.
@@
 
-\section*{Accreditation of degree courses }
+\section*{What has the BCS been doing to promote the inclusion of cyber security in general computer science degrees? }
 
-In the UK --- as in most jurisdictions --- higher education provision addresses general computer science and specialist cybersecurity courses.
-A significant number of undergraduate and postgraduate programmes are available in both the areas of computer science and cybersecurity (and closely related fields, such as computer security, digital/computer forensics, etc).
-
-% TC: main case study
-\section*{BCS Accreditation}
-
-Since 2010 the BCS has evolved its accreditation practices to promote and mandate the inclusion of of first information security (2010) and then cybersecurity (2013) within the programmes the body accredits.
+ The BCS has been using our accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) within the programmes we accredit.
 
@@
 
-\subsection*{Developing Expectations}
+Industry, higher education, government and the relevant professional bodies have collaborated on the production of a set of guidelines which are to the benefit of the various stakeholders and wider society as discussed in a previous issue of ITNow~\cite{Irons2016}.
+This work led to the production of reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'')  and established a baseline of common knowledge, example learning outcome domains for cybersecurity within the degree courses and guidance on embedding the concepts.
 
-Internationally the expectations regarding both the breadth and depth of the expected cybersecurity coverage has been the subject of much discussion, debate and analysis.
-Like many governments, the UK Government has actively been seeking ways to address this.
-In parallel to the work completed by the ACM in the USA, considerable effort in the UK have been taken to ensure industry, higher education, government and the relevant professional bodies collaborated on a set of guidelines which are to the benefit of the various stakeholders and wider society as discussed in a previous issue of ITNow~\cite{Irons2016}.
-
-The produced reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'')  established a baseline of common knowledge, example learning outcome domains for cybersecurity within the computer science courses and guidance on embedding the concepts.
-The document provides specific guidance for embedding and enhancing relevant cybersecurity principles, concepts and learning outcomes within their undergraduate curricula.
-
-\subsection*{Accreditation: What Progress Has Been Made?}
-
-This initiative is a collective attempt to formally include cybersecurity in all BCS Accredited programmes.
-Some of these programmes will be specialist cybersecurity programmes, however the majority will take a different emphasis; this is a work in progress.
-A full cycle of accreditation visits has not yet taken place following the adjustment to the BCS Guidelines.
+Since 2015, we have been expecting accredited degrees to be compliant with the cybersecurity guidelines.
+A full cycle of accreditation visits has not yet taken place following this inclusion in requirements.
 What is being observed is the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
@@
 
-From the start of the Autumn 2015 term, up to and including the Autumn 2018 term, the BCS have carried out 70 accreditation visits (including four international visits).
-The BCS identified action was required to address concerns related to cybersecurity at 16 institutions; thus, 54 institutions were already delivering cybersecurity in keeping with the BCS expectations.
+From the start of the Autumn 2015 term, up to and including the Summer 2019 term, the BCS have carried out 82 accreditation visits (including five international visits (2 in South Africa and 1 in Brunei, Cyprus, and Ireland)) .
+The BCS identified action was required to address concerns related to cybersecurity at 23 institutions; thus, 59 institutions were already delivering cybersecurity in keeping with the BCS expectations.
 
-Long-term actions were expected from 12 institutions (six in 2015/16, three in 2016/17 and three in Autumn 2018) who were awarded `{\emph{At Threshold}}' judgments.
-Ten of these judgments were across all programmes; one was specifically against a generalist masters programme only.
+Long-term actions (`{\emph{At Threshold}}' judgements) were expected from 14 institutions (six in 2015/16, three in 2016/17 and five in 2018/19).
+Thirteen of these judgments were across all programmes; one was specifically against a generalist masters programme only.
 This indicates that the BCS will expect adjustments to have taken place before the next accreditation visit.
-As indicated earlier, this was commonly the case that adjustments had been made to approved programmes of study, however the adjusted modules had not yet been delivered so the evidence base was incomplete in terms of how cybersecurity was assessed.
+It was commonly the case that adjustments had been made to design of the programmes of study, however the adjusted programme had not yet been delivered so the evidence base was incomplete in terms of how cybersecurity was assessed.
 
-Short terms 90 Day Responses where required from four institutions; the outcomes of these actions were as follows: ({\emph{i}}) of the 11 UG programmes involved all were approved `{\emph{At Threshold}}'; ({\emph{ii}}) of the nine UG programmes involved, eight were approved and one refused; ({\emph{iii}}) of the five UG programmes involved, all approved `{\emph{At Threshold}}'; and ({\emph{iv}}) of the 3 UG programmes involved, all 3 were refused.
+Short term actions were required from nine institutions; the outcomes of these actions were as follows: ({\emph{i}}) of the 11 UG programmes involved all were approved `{\emph{At Threshold}}'; ({\emph{ii}}) of the nine UG programmes involved, eight were approved and one refused; ({\emph{iii}}) of the five UG programmes involved, all approved `{\emph{At Threshold}}'; and ({\emph{iv}}) of the 3 UG programmes involved, all 3 were refused; and ({\emph{v}}) a further 5 which at the time of writing the outcome is not known
 
@@
 
-Good practice was identified at one university by the commendation:
+Good practice was identified at three universities by the commendation:
 
@@
 \end{quote}
+\begin{quote}
+``{\emph{Hacktivity and related learning and teaching approaches}}''
+\end{quote}
+\begin{quote}
+``{\emph{Cyber Security Centre which permeates both the course and supports external links and opportunities for students.}}''
+\end{quote}
 
-In summary, this shows that many UK institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
-Clearly not all institutions in the UK necessarily have to apply for BCS accreditation, or apply for accreditation for all their programmes, but even so this is significant evidence of inclusion of cybersecurity to an agreed standard.
+In summary, this shows that many accredited institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
 
@@
 
-The work presented in this article is a first step towards better understanding the nature, design, structure and assessment of cybersecurity education in the UK It is clear that we the BCS are having a positive effect on universities via accreditation, supporting wider national policy imperatives.
-However:
+The BCS are improving the cybersecurity capacity in the UK by enhancing the education experience of the graduates on BCS Accredited degree programmes.
+If you want to reassured the graduates you recruit have some cyber knowledge then seek a BCS accredited degree.
+There are challenges ahead in terms of enhancing the resources upon which the degree provision depends.
 
-\begin{description}
-\item  It  would seem to be necessary to adopt an attitude of caution with respect to both offline (\S\ref{sec:SQL}) and online (\S\ref{sec:Java}) resources.
-
-\item Cybersecurity can be taught largely through other Knowledge Areas is a good idea.
-However, in the current state of education resources (sections \S\ref{sec:EDResource}) may be a counsel caution in this approach.
-
-\item 
-Our Accreditation  is a valuable tool in improving the standard of cybersecurity teaching, and disseminating good practice, and should continue this.
-
-\item
-This article focused upon University Education.
-What cybersecurity capability (and development needs) exists in terms of the existing workforce,  the person in the street, etc could benefit from a similar exploration.
-\end {description}
-
-\section*{Acknowledgements}
-
- The authors wish to thank Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes.
-Many people, accreditors and accredited, have contributed to improving the standard of cybersecurity teaching in the UK, and spreading good practice.
-All authors' institutions are members of the Institute of Coding, an initiative funded by the Office for Students (England) and the Higher Education Funding Council for Wales.
-
- {\emph{N.B.}} The first two authors are current Vice-Presidents of the BCS, and the third and fourth are past and present Chairs of the BCS Academic Accreditation Committee, providing significant personal insight into the UK's national accreditation policy and procedures.
+% This article focused upon University Education. What cybersecurity capability (and 
+%development needs) exists in terms of the existing workforce,  the person in the street, etc % could benefit from a similar exploration.
 

### 2019-09-11 JHDNotes.tex
First Thoughts

\font\manual=manfnt
\def\dbend{{\manual\char127}}
\def\eqq{{\buildrel?\over=}}
\def\C{\mathbf{ C}}
\def\N{\mathbf{ N}}
\def\Q{\mathbf{ Q}}
\def\R{\mathbf{ R}}
\def\Z{\mathbf{Z}}
\def\F{\mathbf{F}}
\def\softO{\tilde{\cal O}}
\def\O{{\cal O}}
%\def\N{{\bf N}}
%\def\Q{{\bf Q}}
%\def\R{{\bf R}}
%\def\Z{{\bf Z}}
\def\action#1{\hfil\rlap{\bf#1}\break}
\def\pcite#1{[#1]}
\def\Res{\mathop{\rm Res}\nolimits}
\def\r#1#2{``#1''$\rightarrow$``#2''}
\documentclass{article}
\bibliographystyle{alphaurl}
\usepackage[hyphens]{url}
%\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{combelow}
\usepackage[show]{ed}
\usepackage{graphicx}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{notation}{Notation}
\newtheorem{example}{Example}
\newtheorem{problem}{Problem}
\def\decision#1{\par{\bf #1}}
\pagestyle{empty}
\begin{document}
\author{Notes by J.H.Davenport}
\title{Accreditation Thoughts}
\date{11 September 2019}
\maketitle
Note \url{https://hbr.org/2019/08/every-computer-science-degree-should-require-a-course-in-cybersecurity}
\section{Possible Questions for Paul to ask Seoul 8+6}
\begin{itemize}
\item[0)]By CyberSecurity we include what might be called "Information Security, IT Security etc."
\item Is there any CyberSecurity in your professional requirements currently?
\item  Do you have plans to change this
\item Are there any specific detailed requirements
\item Is there any quantification of how much CyberSecurity there should be.
\item Any other comments
\end{itemize}

\section{}
\end{document}
\section{}
\begin{description}
\item[Q]
\item[A]
\item[Q]
\item[A]
\end{description}

\end{document}
\bibliography{../../../../jhd}
\section{}
\begin{description}
\item
\end{description}
\begin{itemize}
\item
\end{itemize}
\begin{example}
\end{example}
\begin{definition}
\end{definition}
\begin{theorem}
\end{theorem}
\begin{enumerate}
\end{enumerate}
\begin{description}
\item[Theme 1]
\item[Theme 2]
\item[Theme 3]
\item[Theme 4]
\item[Theme 5]
\end{description}

### 2019-09-11 ITNow_TP.tex

@@
 
-Cybersecurity matters, breaches are legion and sadly are a common feature in the news.
+Cybersecurity matters!
+Breaches are legion and sadly are a common feature in the news.
 Universities are  addressing the demand for educated cybersecurity professionals.
@@
 \subsubsection*{The Case of SQL Injection}\label{sec:SQL}
-SQL injection is a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 , and has been in the Top 10 since at least 2003.
-Database textbooks where analysed in 2019 ~\cite{Drop2019}.
+SQL injection is a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10, and has been in the Top 10 since at least 2003.
+Database textbooks were analysed in 2019 ~\cite{Drop2019}.
 There were seven such books, but three books accounted for the 36 of 44 universities.
@@
 How good are these?
-One invesigation \cite{Unruhetal2017a}, took the top five search results from Google for six queries.
+One investigation \cite{Unruhetal2017a}, took the top five search results from Google for six queries.
 Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting (Number 7 in OWASP's Top Ten) weaknesses.
@@
 
-\section*{What has the BCS been doing to promote the inclusion of cyber security in general computer science degrees? }
+\section*{What has the BCS been doing to promote the inclusion of cybersecurity in general computer science degrees? }
 
- The BCS has been using our accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) within the programmes we accredit.
+ The BCS has been using the accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) within the programmes we accredit.
 
@@
 Industry, higher education, government and the relevant professional bodies have collaborated on the production of a set of guidelines which are to the benefit of the various stakeholders and wider society as discussed in a previous issue of ITNow~\cite{Irons2016}.
-This work led to the production of reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'')  and established a baseline of common knowledge, example learning outcome domains for cybersecurity within the degree courses and guidance on embedding the concepts.
+This work led to the production of reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'')  and established a baseline of common knowledge, example learning outcome domains for cybersecurity within  degree programmes and guidance on embedding the concepts.
 
 Since 2015, we have been expecting accredited degrees to be compliant with the cybersecurity guidelines.
-A full cycle of accreditation visits has not yet taken place following this inclusion in requirements.
-What is being observed is the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
+Universties are visited on a quinquennial basis and a full cycle of accreditation visits has not yet taken place following this inclusion in requirements.
+What is being observed is that the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so.
 However, a minority are requiring encouragement to do so.
 
-From the start of the Autumn 2015 term, up to and including the Summer 2019 term, the BCS have carried out 82 accreditation visits (including five international visits (2 in South Africa and 1 in Brunei, Cyprus, and Ireland)) .
-The BCS identified action was required to address concerns related to cybersecurity at 23 institutions; thus, 59 institutions were already delivering cybersecurity in keeping with the BCS expectations.
+From the start of the Autumn 2015 term, up to and including the Summer 2019 term, the BCS has carried out 82 accreditation visits including five international visits (2 in South Africa and 1 in Brunei, Cyprus, and Ireland) .
+The BCS identified action was required to address concerns related to cybersecurity at 23 institutions; thus, 59 institutions were already delivering cybersecurity in line with the BCS expectations.
 
-Long-term actions (`{\emph{At Threshold}}' judgements) were expected from 14 institutions (six in 2015/16, three in 2016/17 and five in 2018/19).
-Thirteen of these judgments were across all programmes; one was specifically against a generalist masters programme only.
+Long-term actions (`{\emph{At Threshold }}' judgements) were expected from 14 institutions (six in 2015/16, three in 2016/17 and five in 2018/19).
+13 of these judgments were across all programmes; one was specifically against a generalist masters programme only.
 This indicates that the BCS will expect adjustments to have taken place before the next accreditation visit.
-It was commonly the case that adjustments had been made to design of the programmes of study, however the adjusted programme had not yet been delivered so the evidence base was incomplete in terms of how cybersecurity was assessed.
+It was commonly the case that adjustments had been made to design of the programmes of study, however, the adjusted programme had not yet been delivered so the evidence base was incomplete in terms of how cybersecurity was assessed.
 
-Short term actions were required from nine institutions; the outcomes of these actions were as follows: ({\emph{i}}) of the 11 UG programmes involved all were approved `{\emph{At Threshold}}'; ({\emph{ii}}) of the nine UG programmes involved, eight were approved and one refused; ({\emph{iii}}) of the five UG programmes involved, all approved `{\emph{At Threshold}}'; and ({\emph{iv}}) of the 3 UG programmes involved, all 3 were refused; and ({\emph{v}}) a further 5 which at the time of writing the outcome is not known
+Short term actions were required from nine institutions; the outcomes of these actions were as follows: ({\emph{i}}) of the eleven UG programmes involved all were approved `{\emph{At Threshold}}'; ({\emph{ii}}) of the nine UG programmes involved, eight were approved and one refused; ({\emph{iii}}) of the five UG programmes involved, all were approved `{\emph{At Threshold}}'; and ({\emph{iv}}) of the three UG programmes involved, all were refused; and ({\emph{v}}) a further five which at the time of writing the outcome is not known
 
@@
  \begin{quote}
-``{\emph{The second year project provides an opportunity for exploring security aspects in depth with an industrial use case.}}''
+``{\emph{The second-year project provides an opportunity for exploring security aspects in depth with an industrial use case.}}''
 \end{quote}
@@
 
-The BCS are improving the cybersecurity capacity in the UK by enhancing the education experience of the graduates on BCS Accredited degree programmes.
-If you want to reassured the graduates you recruit have some cyber knowledge then seek a BCS accredited degree.
+The BCS is improving the cybersecurity capacity in the UK by enhancing the education experience of the students on BCS accredited degree programmes.
+If you want to be reassured that the graduates you recruit have some cyber knowledge, then seek a BCS accredited degree.
 There are challenges ahead in terms of enhancing the resources upon which the degree provision depends.

### 2019-09-11 JHDNotes.tex
Added Comments
Added some dates
Added CEP etc.

@@
 \section{Possible Questions for Paul to ask Seoul 8+6}
-\begin{itemize}
-\item[0)]By CyberSecurity we include what might be called "Information Security, IT Security etc."
+\begin{enumerate}
+\item[0.]By CyberSecurity we include what might be called ``Information Security, IT Security etc.''
 \item Is there any CyberSecurity in your professional requirements currently?
-\item  Do you have plans to change this
+\item  Do you have plans to change this??
 \item Are there any specific detailed requirements
-\item Is there any quantification of how much CyberSecurity there should be.
-\item Any other comments
-\end{itemize}
+\item Is there any quantification of how much CyberSecurity there should be?
+\item Any other comments.
+\end{enumerate}
 
-\section{}
+\section{Publishing Venues}
+\begin{description}
+\item[IT Now]Being pursued, let by TomP.
+\item[CEP]A UK focused conference (not necessarily a bad thing).
+\item[\qquad11 October 2019]Deadline for abstracts
+\item[\qquad18 October 2019]Deadline for full papers
+\item[SIGCSE 2020]Main papers are closed, but there may be posters.
+\item[ITiCSE]See \url{https://iticse.acm.org/ITiCSE2020/} --- apparently no dates.
+\item[Koli]2019 closed 5 August. Presumbaly 2020 will be similar.
+\item[Others]??
+\end{description}
 \end{document}

### 2019-09-11 JHDNotes.tex
Added refereences to SA

@@
 \begin{enumerate}
-\item[0.]By CyberSecurity we include what might be called ``Information Security, IT Security etc.''
+\item[0.]By CyberSecurity we include what might be called ``Information Security, IT Security etc.'', in fact security aspects of Computing as characterised in footnote 1 of Appendix D.
 \item Is there any CyberSecurity in your professional requirements currently?
@@
 \end{enumerate}
-
+NB: bottom of age 3 reads ``taking into account all relevant technical, legal, social, cultural, economic, and environmental issues and respecting the need for sustainability'' --- in the light of GDPR, this would seem to imply a certain amount of CyberSecurity!
 \section{Publishing Venues}

### 2019-09-15 ITNow_TP_v2.tex
Attempt to change audience

--- 2019-09-11 ITNow_TP.tex
@@
 \author[3]{Alastair Irons}
-\author[4]{Tom Prickett}
+\author[4]{Sally Pearce}
+\author[5]{Tom Prickett}
 % institutions
@@
 \affil[3]{Sunderland University, Sunderland, UK}
-\affil[4]{Northumbria University, Newcastle upon Tyne, UK}
+\affil[4]{Accreditation Team Manager,BCS}
+\affil[5]{Northumbria University, Newcastle upon Tyne, UK}
 % emails
@@
 \affil[3]{\url{alastair.irons@sunderland.ac.uk}}
-\affil[4]{\url{tom.prickett@northumbria.ac.uk}}
+\affil[3]{\url{sally.pearce@BCS.uk}}
+\affil[5]{\url{tom.prickett@northumbria.ac.uk}}
 
@@
 
-\date{August 2019}
+\date{September 2019}
 
@@
 \begin{strapline}
-The BCS has been promoting further inclusion of Cybersecurity within general Computer Science University Education since 2010. We explore progress and some of the outstanding related issues.
-
+"Every Computer Science Degree Should Require a Course in Cybersecurity". We explore the progress to date.
 \end{strapline}
@@
 
-Cybersecurity matters!
-Breaches are legion and sadly are a common feature in the news.
-Universities are  addressing the demand for educated cybersecurity professionals.
-There are calls for formal education to respond to this situation.
-This is reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK.
+On the 27th August of this year the Harvard Business Review reported "Every Computer Science Degree Should Require a Course in Cybersecurity".
+How could you argue with the intent of this suggestion?
+The BCS has been promoting this for a number of years and how fantastic to see our concerns highlighted to such an audience.
+But, and this is a subtly but significant \textit{'but'}, is this the best approach?
+We feel not.
+As our recent research {\cite{Cricketal2019}} has argued cybersecurity is too important to be left to specialists and needs to be everyone's responsibility and addressed across a course.
 
-This raises the question, should cybersecurity be of importance for all or just for specialists?
-There is a serious need for cybersecurity specialists, but also all in IT need to know \emph{some} cybersecurity -- thus, there is a case for both.
-
-In this article we explore some of the challenges related to cybersecurity education of general computer scientists (including information technology and other related degrees), and what we, the BCS, have been doing to enhance cybersecurity in education at university.
+This article explores some challenges regarding how cybersecurity is taught which informs this view and provides a progress report regarding BCS efforts to promote the inclusion of cybersecurity in degree course.
 
@@
 \subsection*{Educational Resources}\label{sec:EDResource}
-It is hard to determine what is actually delivered as part of a degree programme, but a reasonable proxy for this are recommended textbooks and other supporting resources.
-\subsubsection*{The Case of SQL Injection}\label{sec:SQL}
+%It is hard to determine what is actually delivered as part of a degree programme, but a reasonable proxy for this are recommended textbooks and other supporting resources.
+Lets start by looking at the books and other resources that are used to teach degrees.
+Do these support the promotion of good cybersecurity practices?
+As you will see, a single course on the subject may not be enough.
+We think that would be like \textit{'shutting the proverbial stable door after the horse has bolted'}.
+\subsubsection*{Databases}\label{sec:SQL}
+A core part of any computer science degree.
 SQL injection is a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10, and has been in the Top 10 since at least 2003.
 Database textbooks were analysed in 2019 ~\cite{Drop2019}.
-There were seven such books, but three books accounted for the 36 of 44 universities.
-Five of the seven (30 of the 44) had no mention of SQL injection.
-On the other two, the more popular one has a seriously flawed discussion.
+This work shows this flaw is not being addressed.
+Small wonder it continues to be an issue!
 
 \subsubsection*{The Case of Java}\label{sec:Java}
-
+Our research shows Java is still the most commonly taught programme language at universities in the UK.
 Many Java books do not cover security.
 If you want to know about security you need to look at the documentation of the package/API being used, and/or  informal resources.
-\cite{Mengetal2018a} analysed 503 cybersecurity-related postings on the popular Stack Overflow online resource.
-53\% were about the Spring Security framework, dominated by authentication (45\%).
+\cite{Mengetal2018a} analysed this and Stack Overflow in particular.
+Looking at the Spring Framework 53\% of questions related to cybersecurity were dominated by authentication (45\%).
 One example being, by default, Spring enables protection against cross-site request forgery (CSRF).
@@
 There were no negative comments about this, and indeed a typical response is ``{\emph{Adding}} \verb!csrf().disable()!
-{\emph{{solved the issue!!! I have no idea why it was enabled by default}}''. }
-
-\subsubsection*{Android}\label{sec:Android}
-
-For Android, \cite{Fischeretal2017a} looked specifically at the use of resources from Stack Overflow. The key finding was:
-
-\begin{quote}
-``{\emph{We found that 15.4\% of all 1.3 million Android applications
-contained security-related code snippets from
-Stack Overflow. Out of these 97.9\% contain at least one
-insecure code snippet.}}''
-\end{quote}
+{\emph{{solved the issue!!! I have no idea why it was enabled by default}}''. } Quite concerning!
 
 \subsubsection*{Informal Resources}\label{sec:informal}
-
 The web abounds with informal resources, such as tutorials and code snippets.
 How good are these?
-One investigation \cite{Unruhetal2017a}, took the top five search results from Google for six queries.
+One investigation \cite{Unruhetal2017a} considered 30 popular tutorials and found six had SQL injection weaknesses, and three had Cross-Site Scripting (Number 7 in OWASP's Top Ten) weaknesses.
+Looking for related code fragments on GitHub, found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.
+Hardly encouraging.
+
+%took the top five search results from Google for six queries.
 Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting (Number 7 in OWASP's Top Ten) weaknesses.
@@
 \subsection*{Who will teach the Cybersecurity?}\label{sec:staffing}
-
+Enough about the books, who is going to teach this?
 Cybersecurity skills are in short supply, in both industry and academia.
@@
 
+Enough of the challenges.
+What have we been doing about it?
 The BCS has been using the accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) within the programmes we accredit.
@@
 
+Is it sufficient to leave to cybersecurity to Computer Scientist? We hope this is conversation that will begin shortly.
+
 % This article focused upon University Education. What cybersecurity capability (and 

### 2019-09-16 ITNow_TP_v2.tex

@@
 \begin{strapline}
-"Every Computer Science Degree Should Require a Course in Cybersecurity". We explore the progress to date.
+Including cybersecurity in university education. We explore the progress to date.
 \end{strapline}
@@
 The BCS has been promoting this for a number of years and how fantastic to see our concerns highlighted to such an audience.
-But, and this is a subtly but significant \textit{'but'}, is this the best approach?
-We feel not.
-As our recent research {\cite{Cricketal2019}} has argued cybersecurity is too important to be left to specialists and needs to be everyone's responsibility and addressed across a course.
+As our recent research {\cite{Cricketal2019}} has argued cybersecurity is too important to be left to specialists and hence is an essential inclusion in Computer Science, Information Technology and all other related degree programmes.
 
-This article explores some challenges regarding how cybersecurity is taught which informs this view and provides a progress report regarding BCS efforts to promote the inclusion of cybersecurity in degree course.
+This article explores some challenges related to the teaching of cybersecurity and provides a progress report regarding BCS efforts to promote the inclusion of cybersecurity in accredited degree programmes.
 
@@
 %It is hard to determine what is actually delivered as part of a degree programme, but a reasonable proxy for this are recommended textbooks and other supporting resources.
-Lets start by looking at the books and other resources that are used to teach degrees.
+Let's start by looking at the books and other resources that are used to teach degrees.
 Do these support the promotion of good cybersecurity practices?
-As you will see, a single course on the subject may not be enough.
-We think that would be like \textit{'shutting the proverbial stable door after the horse has bolted'}.
+As you will see, that whilst progress has been made, there still appears to be work to do.
 \subsubsection*{Databases}\label{sec:SQL}
@@
 Database textbooks were analysed in 2019 ~\cite{Drop2019}.
-This work shows this flaw is not being addressed.
+This work shows this flaw is not commonly being addressed in database textbooks.
 Small wonder it continues to be an issue!
@@
 \subsubsection*{The Case of Java}\label{sec:Java}
-Our research shows Java is still the most commonly taught programme language at universities in the UK.
+Our research shows Java is still the most commonly taught programming language at universities in the UK.
 Many Java books do not cover security.
@@
 \cite{Mengetal2018a} analysed this and Stack Overflow in particular.
+We also know our students love resources like Stack Overflow.
 Looking at the Spring Framework 53\% of questions related to cybersecurity were dominated by authentication (45\%).
@@
 The web abounds with informal resources, such as tutorials and code snippets.
+We know these are very popular with the students we teach.
 How good are these?
@@
 \subsection*{Who will teach the Cybersecurity?}\label{sec:staffing}
-Enough about the books, who is going to teach this?
+Who is going to teach this?
 Cybersecurity skills are in short supply, in both industry and academia.
@@
 What have we been doing about it?
-The BCS has been using the accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) within the programmes we accredit.
+The BCS has been using the accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) in accredited programmes.
 
@@
 
-In summary, this shows that many accredited institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
+%In summary, this shows that many accredited institutions have now embedded cybersecurity in their provision, a number are in the process of doing so and a minority have chosen not to.
 
@@
 The BCS is improving the cybersecurity capacity in the UK by enhancing the education experience of the students on BCS accredited degree programmes.
-If you want to be reassured that the graduates you recruit have some cyber knowledge, then seek a BCS accredited degree.
-There are challenges ahead in terms of enhancing the resources upon which the degree provision depends.
+If you want to be reassured that the graduates you recruit have some cybersecurity knowledge, then seek a BCS accredited degree.
+There are challenges ahead in terms of enhancing the resources upon which the degree provision depends and developing the cybersecurity capacity within universities.
 
-Is it sufficient to leave to cybersecurity to Computer Scientist? We hope this is conversation that will begin shortly.
+As a final rhetorical point, is it sufficient to leave to cybersecurity to Computer Scientist?
+Should it be solely the preserve of those who work with Tech?
+Or should it be everyone's responsibility?
+We hope this is a conversation that will begin shortly.
 

### 2019-09-29 CEP20/CEP_accreditationpostshadbolt.tex
CEP2020
Includes two tex files - CEP_cybersecurity.tex and CEP_accreditationPostShadbolt.tex. Both follow the CEP template. They share a .bib file

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[CEP '20]{CEP '20: ACM Computing Education Practice}{January 9, 2020}{Durham, UK}
\acmBooktitle{CEP '20: Proceedings of the 3rd Conference on Computing Education Practice,
  June 9, 2020, Durham, UK}
\acmPrice{15.00}
\acmISBN{978-1-4503-9999-9/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Accreditation: Post Shadbolt}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

%\begin{comment}

\author{Tom Crick}
\affiliation{%
  \institution{Swansea University}
  \city{Swansea}
  \country{UK}
}
\email{thomas.crick@swansea.ac.uk}

\author{James H. Davenport}
\affiliation{%
  \institution{ University of Bath}
  \city{Bath}
  \country{UK}
}
\email{j.h.davenport@bath.ac.uk}

\author{Alastair irons}
\affiliation{%
  \institution{ Sunderland University}
  \city{Sunderland}
  \country{UK}
}
\email{alastair.irons@sunderland.ac.uk}

\author{Tom Prickett}
\affiliation{%
  \institution{ Northumbria University}
  \city{Newcastle upon Tyne}
  \country{UK}
}
\email{tom.prickett@northumbria.ac.uk}
%\end{comment}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Crick, Davenport,  Irons, and Prickett.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003456.10003457.10003527.10003529</concept_id>
<concept_desc>Social and professional topics~Accreditation</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Social and professional topics~Accreditation}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Accreditation}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section {What is it?}
A short description of the practice you're presenting

\section {Why are you doing it?}	What happened before? What is it changing / replacing / improving? What gap is it filling?

\section {Where does it fit?	}A short description of your teaching context. You may, for instance, include a description of intake, class size, curriculum sequence; anything that's necessary for others to understand your situation. How do things work at your institution?

\section {Does it work?}	How do you know? Give some evidence of effectiveness in context.
\section {Who else has done this?}	Where did you get the idea from? (If from published reports, please include references). How did you find out about it? Was it easy/hard to adopt? What did you change?
\section {What will you do next?}	Will you vary this, or develop it further?
\section{Why are you telling us this?}	What is interesting, or useful, about this to someone else?

\section{Acknowledgments}
The authors wish to thank Sally Pearce, Academic Accreditation Manager at BCS, The Chartered Institute for IT for supplying the summary information related to accreditation of UK degree programmes. Many people, accreditors and accredited, have contributed to acreditation pratice in the UK (and elsewhere), and spreading good practice.  All authors' institutions are members of the Institute of Coding, an initiative funded by the Office for Students (England) and the Higher Education Funding Council for Wales.
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
%\bibliography{sample-base}
\bibliography{CEP2020}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2019-09-29 CEP20/CEP_cybersecurity.tex
CEP2020
Includes two tex files - CEP_cybersecurity.tex and CEP_accreditationPostShadbolt.tex. Both follow the CEP template. They share a .bib file

--- 2019-09-29 CEP20/CEP_accreditationpostshadbolt.tex
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Accreditation: Post Shadbolt}
+\title{How well do we teach Cybersecurity}
 
@@
 <concept>
+<concept_id>10002978</concept_id>
+<concept_desc>Security and privacy</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
 <concept_id>10003456.10003457.10003527.10003529</concept_id>
@@
 
+\ccsdesc[500]{Security and privacy}
 \ccsdesc[500]{Social and professional topics~Accreditation}

### 2019-10-04 CEP20/CEP_accreditationpostshadbolt.tex
Draft of response to Shadbolt

@@
 \acmISBN{978-1-4503-9999-9/18/06}
+
+\usepackage{dirtytalk}
 
@@
 
-\author{Alastair irons}
+\author{Alastair Irons}
 \affiliation{%
@@
 \begin{abstract}
-  A clear and well-documented \LaTeX\ document is presented as an
-  article formatted for publication by ACM in a conference proceedings
-  or journal publication. Based on the ``acmart'' document class, this
-  article presents and explains many of the common variations, as well
-  as many of the formatting elements an author may use in the
-  preparation of the documentation of their work.
+The promotion of quality via Accreditation by Professional, Statutory and Regulatory Bodies (PSRBs) is very much a feature of higher education in the united kingdom (UK).
+In this changing  environment, there is a need for accreditation regimes to evolve in order to maximise the value they provide to Higher Education Institutions(HEIs).
+This paper provides a summary in the context of one professional body (BCS, The Chartered Institute For IT) of
+\begin{enumerate}
+\item what has happened in response to the recommendations of the Shadbolt review \cite{shadbolt2016shadbolt}
+\item ongoing enhancement projects and the related future conversations that will help steer further enhancements
+\end{enumerate}
 \end{abstract}
@@
 \section {What is it?}
-A short description of the practice you're presenting
+%A short description of the practice you're presenting
 
-\section {Why are you doing it?}	What happened before? What is it changing / replacing / improving? What gap is it filling?
+The Shadbolt Review \cite{shadbolt2016shadbolt} investigated high unemployment rates graduates of computer science and related degrees in the United Kingdom and the role of degree accreditation in promoting employability.The Shadbolt review noted that the rates of graduate unemployment were declining and varied considerably between geographic location and type of HEI.
+The discipline was already acting proactively to address the issue.
+Related professional bodies have enhanced their accreditation processes in response these challenges.
 
-\section {Where does it fit?	}A short description of your teaching context.
+BCS has made significant adjustments to their accreditation processes since 2015.
+These were partly in response to the recommendations of Shadbolt but also to other changes in the sector.
+These have been communicated to the assessor community and to those responsible for leading the development of HEI accreditation applications.
+However, there are many academics who are less intimately involved in accreditation, this paper serves to communicate these and future enhancements to them and the wider computer science education community.
+
+\section {Why are you doing it?}	
+%What happened before? What is it changing / replacing / improving? What gap is it filling?
+
+The Shadbolt review \cite[p.~8]{shadbolt2016shadbolt} makes two main recommendations for the enhancement of accreditation regimes.
+\begin{center} 
+	\begin{textit} 
+	"Recommendation 9 - Academic Accreditation of Degree Courses.
+
+		BCS, IET and Tech Partnership should ensure that existing systems of degree course
+	accreditation are flexible, agile, and enable HE providers to respond to changing
+	demand and emerging technological trends and developments. Accreditation of courses
+	should be focussed on outputs. Accrediting bodies should work to increase awareness
+	and value of accreditation so that it is valued by HE providers, students and employers,
+	and consider how their role can provide a forum for engagement between HE and
+	employers..
+
+	Recommendation 10 - Engaging industry in accreditation.
+
+	Employers, through employer groups, such as Tech Partnership, should engage more
+	consistently with HE providers and BCS \& IET to ensure accreditation is effective and
+	reflects current industry demand.
+	"
+	\end{textit}
+\end{center}
+
+Criticisms of the accreditation of degree programmes is not new, there is a history of claiming the processes are unnecessarily bureaucratic and constrain  innovation \cite{Harvey2004}.
+Equally its value has been highlighted particularly in the context of a potentially globally mobile workforce, and the dangers of accreditation streams being revenues streams in their own right rather than for the benefit of a discipline or wider society \cite{Knight_2015}.
+
+Hence, the challenge set for BCS and other accreditation providers that operate in the Computer Science discipline was broadly to maintain internationally recognised standards; reduce the perceived bureaucracy involved; respond to emerging technology trends and developments; promote enhancement and innovation; increase awareness and value of accreditation; and engage industry.
+
+Defining quality and what is required for quality improvement has been discussed by a plethora of authors.
+One early author, Joseph Juran indicated as part of his Quality Control Handbook that a trilogy was required, namely quality planning, quality control and quality improvement \cite{juran1962quality}.
+This serves to indicate that quality improvement efforts should be informed by the data available and are never complete.
+As such there is a need to continue to enhance accreditation processes, as there are all quality review and improvement processes.
+
+BCS is addressing these issues in the following ways.
+
+\subsection{The value of accreditation}
+What is the value of an accredited degree? This is a question that the BCS is in the process of exploring in much depth with a variety of stakeholders. The following are aspects of the value.
+
+\subsubsection{Independent peer review}
+Peer review is commonly used in quality regimes in higher education and in periodic review processes.
+In most cases the HEI being reviewed chooses the peer.
+As part of a BCS (or other professional body) a HEI has less choice regarding who completes a review and hence such a review is arguably more truly independent.
+BCS review panels contain a minimum of two experienced assessors who have strong awareness of the discipline norms across the UK sector.
+
+\subsubsection{Internationally agreed standards}
+The BCS accredits to internationally agreed standards, which evidence the global parity of the degree programmes accredited.
+This assists in the global mobility for the graduates of accredited programmes.
+
+\subsubsection{Ensuring curricula relevance}
+With the agreement of the sector (normally reached through the Council of Professors and Heads of Computing (CPHC)), it is possible to agree and promote curricula change.
+The most recent example of this is the inclusion of cybersecurity in all accredited degrees \cite{Cricketal2019}.
+BCS is in the process of mandating the inclusion of security in all accredited degree programmes to a standard agreed between professional bodies, industry and government.
+This is a work in progress \cite{Irons2016}.
+BCS has been requiring coverage since 2015 \cite[p.~17--18]{BCS2018a} with the result that all accredited universities must be compliant by 2020 (due to the five-year cycle).
+In the United States of America, the Association of Computing Machinery (ACM) has equally had cybersecurity (IAS: ``Information Assurance and Security'') in the curriculum since 2013~\cite{ACM2013a}, but it is not the accrediting body.
+The Accreditation Board for Engineering and Technology (ABET) is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''.
+This will mean that  all accredited universities should be compliant by 2025 (due to their six-year cycle).
+
+It is interesting to note this years Royal Institution Christmas lecture considers ethical questions related to computing and mathematics \cite{RoyalInstitution2019}.
+This is a curricula element that has been mandated by BCS accreditation in the form of legal, social, ethical and professional issues for a number of years \cite{Brooke2018}.
+
+\subsubsection{Disseminating good practice}
+BCS accreditation panels have been identifying aspects of good practice for wider dissemination for a number of years.
+By commending good practices and promoting these to the wider community.
+Since 2016, the process has been refined to more clearly signpost commendations.
+BCS is currently in the process of refining the promotion of good practice.
+
+One short term initiative, from autumn 2019, visited HEIs will be offered the opportunity to showcase an aspect of their provision or practice.
+The intention is to generate further good practice examples that a visiting panel can potentially commend.
+
+\subsubsection{Industry relevance}
+All BCS accreditation panels include an industrial assessor whose role is to ensure that programmes appropriately meet the needs of industry.
+Evidence that a visited HEIs' mechanisms for engaging with industry are supporting graduates to evidence they are achieving the expected exit standards are sought from a variety of sources including the curricula studied; the assessments and examinations sat; and the engagement mechanisms themselves.
+
+\subsubsection{Maintaining standards}
+BCS refuses accreditation for programmes that do not meet the expected standards or do not comply with the curricula expectations of the accreditation sought.
+
+\subsection{Driving improvement}
+A number of data sources have been employed to help drive the improvement efforts.
+BCS Secretariat continually evaluate via the use of opinion surveys and informal conversations.
+Views are gathered related to briefing sessions, pre-visit communications and visits.
+This information is fed into the ongoing reviews.
+These issues are discussed, where required at an appropriate committee meeting and used to identify any enhancement projects.
+Following a visit BCS panels engage in peer review.
+The feedback gained from assessor peer review is explored and opportunities for enhancement are discussed on an annual basis at a committee meeting.
+
+\subsection{Reducing bureaucracy and enhancing practices}
+A review of BCS Accreditation practice has been taking place since 2015 with the intention of adopting an enhancement oriented agenda; reducing the amount of bespoke documentation that is required; using technology to assist the process, whilst enforcing the international standards.
+
+Working through the process from a visited HEIs prospective the review results in the following changes.
+
+\subsubsection{HEI briefing}
+For a number of years all visited HEIs have been encouraged to attend a briefing to explore the BCS requirements with a focus upon any changes since the last visit and common challenges that present themselves as part of visits.
+Attendance was understandably mixed, partly due to travel.
+Since 2017 all Briefings have taken place by video conference.
+Feedback upon this approach has been very positive.
+
+\subsubsection{The application itself}
+From a paper and memory stick / CD submission before 2015, BCS has moved to a fully electronic submission.
+BCS is flexible in how the information is provided, a range of secure cloud based file sharing systems have been employed, submission via the use of Virtual Learning Environment has been completed and a minority of institutions have opted to create a website related to the submission.
+Most of the application consists of an evidence base which HEIs will already have.
+This is supplemented by a summary of the provision and each programme in which the department are welcome to reference existing resources.
+HEIs are still required to provide a mapping of where the BCS requirements are taught and assessed within each programme.
+This is required to evidence the where accreditation specific requirements are met in a particular programme.
+
+\subsubsection{Problematic areas}
+It is now normal practice for a BCS panel to communicate likely areas of discussion to a visited HEI prior the visit.
+There are some occasions when this is not possible, for example if a HEIs application was late or if other issues emerge as part of the visit.
+This has resulted in discussions tending to become more collegiate.
+
+\subsection{Promoting the value of accreditation}
+The value proposition discussed in this document is a definition of the value of accreditation from those who are intimately involved.
+BCS is in the process of initiating a conversation to explore the value proposition of accreditation from the point of view of other stakeholders.
+
+\section {Where does it fit?	}
+%A short description of your teaching context.
 You may, for instance, include a description of intake, class size, curriculum sequence; anything that's necessary for others to understand your situation.
@@
 
-\section {Does it work?}	How do you know? Give some evidence of effectiveness in context.
-\section {Who else has done this?}	Where did you get the idea from?
-(If from published reports, please include references).
-How did you find out about it?
-Was it easy/hard to adopt?
-What did you change?
-\section {What will you do next?}	Will you vary this, or develop it further?
-\section{Why are you telling us this?}	What is interesting, or useful, about this to someone else?
+In the United Kingdom the most common form of accreditation in the Computer Science and related discipline areas is by professional bodies.
+The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) are the main bodies involved.
+The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord \cite{Washington2019} and Seoul Accord \cite{Seoul2019}.
+These memoranda support the internationalising of the curriculum and promote consistency and parity in Computer Science education globally.
+These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
+This is reflected in IETs and BCSs long history of expecting coverage environmental factors within the programmes they accredit and BCSs expectation of significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}.
+
+As part of accreditation IET and BCS broadly check two things \cite{BCS2018a, IET2019}.
+
+\begin{enumerate}
+    \item Is the quality of experience provided by a HEI appropriate for accreditation?
+    A number of sources of information are used to determine this and a number of metrics are explored, including graduate employability.
+This is supplemented by other evidence of the quality of the provision, for example external examiners reports, the most recent subject review, annual review information, evidence of employer involvement, linkage with research.
+Together this evidences that a HEI is supporting their students in an appropriate manner.
+Accreditation will be refused for HEIs that do not offer appropriate support (including for example poor employability of graduates from a programme).
+    \item Are the exit standards of a programme consistent with the learning outcomes expected for the accreditation sought?
+The exit standards conform with one of the international memorandum (Washington or Seoul Accord or both).
+\end{enumerate}
+
+\section {Does it work?}	H
+%how do you know? Give some evidence of effectiveness in context.
+
+HEIs choose which programmes are submitted or not in an accreditation application. As such only a sample of provision may be considered.
+
+However a proxy for success is the anonymous feedback BCS obtains post visit. Some key aspects of this feedback are shown in the following table.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+%\begin{table}[]
+\begin{table}[h!]
+  \caption{Selected results from accreditation visit feedback survey}
+  \label{table:1}
+%\resizebox{\textwidth}{!}{%
+%\begin{tabular}{lllll}
+\begin{tabular}{ | p{5cm}|p{.75cm}|p{.75cm} |p{.75cm} |}
+\hline
+ & 2017 & 2018 & 2019   \\ \hline
+Responses & 12 & 14 & 16   \\
+\hline
+Attended pre-visit briefing & 6 & 14 & 16   \\
+\hline
+The visit felt to be worthwhile: & & &  \\
+Strongly Agree &92\% & 93\% &88\% \\
+Agree &0\% & 7\% &0\% \\
+Tend to Agree &0\% & 0\% &6\% \\
+Strongly Disagree &8\% & 0\% &0\% \\
+No reply &0\% & 0\% &6\% \\
+\hline
+Overall how satisfied  would you say you were with the visit? & & &  \\
+Very Satisfied &83\% & 86\% &94\% \\
+Quite Satisfied &8\% & 14\% &6\% \\
+Not Satisfied at all &8\% & 0\% &0\% \\
+\hline
+
+\end{tabular}%
+%}
+\end{table}
+
+Overall the results are positive, indicating the process is generally value and shows the increased attendance of the virtual briefing over the physical one.
+
+\section {Who else has done this?}	
+%Where did you get the idea from? (If from published reports, please include references). How did you find out about it? Was it easy/hard to adopt? What did you change?
+
+In addition IET and BCS, accreditation in the broad computing area in the UK is also being performed by a few different agencies.
+
+\subsubsection{Not for Profit Organisations}
+Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements \cite{TP2019}.
+Tech Partnership Degrees have a specialist scope, endorsing programmes in the area of IT Management for Business and Software Engineering for Business.
+Tech Partnership Degrees currently endorse 14 IT Management for Business Programmes and 5 Software Engineering for Business Programmes.
+As such Tech Partnership Degrees currently have limited impact upon more general Computer Science education.
+There is little published regarding the effectiveness or otherwise of this approach.
+
+The Institute of Coding (IoC) is a not for profit organisation that intends to enhance how Digital Skills are developed in Higher Education in the UK \cite{Davenportetal2019a}.
+Like the Tech Partnership focus is upon job market requirements.
+A micro-credentialing approach appears to being taken.
+It still is in the early stages of adoption.
+The BCS is actively working with IoC and intends to collaborate with the IoC in any initiatives of mutual benefit.
+
+\subsubsection{National Cyber Security Centre (NCSC)}
+The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
+The NCSC publishes and accredits to a number of cybersecurity standards \cite{NCSC2018a}.
+These standard are linked to the Association of Computer Machinary (ACM) recommendations for curricula.
+To date the major focus of NSSC accreditation has been upon Masters degrees specializing in cybersecurity.
+More recently the NCSC has also began accrediting integrative masters programmes, undergraduate degrees in cybersecurity and Computer Science degrees with a significant cybersecurity focus \cite{NCSC2018b}.
+
+Currently accredited are 15 cybersecurity MSc programmes with a further 11 provisional accredited, 3 integrated masters cybersecurity programmes and 1 cybersecurity Degree Programme with a further 2 provisionally accredited.
+Hence the extent of accreditation is currently reasonably limited in terms of reach to Computer Science programmes.
+This appears to be a positive initiative that will hopefully further develop over time.
+Again little has yet been published regarding the effectiveness of this initiative.
+
+\subsubsection{The wider Computer Science education community}
+The Computer Science discipline has responded actively to the challenges presented by Shadbolt with CPHC operating four working groups \cite{cphc_2016},  promotion of further engagement between industry and academia by initiatives such as the Royal Academy of Engineering Visiting Industrial Professor Scheme \cite{royal}
+
+\section {What will you do next?}	
+%Will you vary this, or develop it further?
+
+BCS reviews its accreditation processes on a regular basis. The current enhancement projects that are either in progress or planed for the future are as follows.
+
+\subsection{Internationalization}
+BCS accreditation primarily takes place in the UK, however there are a small but growing number of institutions outside the UK that BCS accredits or who are seeking accreditation from BCS.
+Under the terms of the Washington and Seoul Accords, BCS does not engage in accreditation in the jurisdictions of Accord signatories without first consulting with the related local professional body.
+On occasions BCS guidelines employ English / Welsh higher education terms without indicating that local equivalents are equally acceptable.
+The next set of guidelines will address this short coming.
+
+\subsection {Engineering Council and Accreditation of Higher Education Programmes (AHEP) Version 4}
+In the British jurisdiction of the Washington accord, the approach adopted is unusual in that the Engineering Council extends the license to accredit to a large number of professional bodies \cite{EC2019}.
+This is not the practice adopted in other jurisdictions, for example in the United States of America ABET is the sole body.
+A recent audit by the Washington Accord has highlighted divergence in practice in some areas with respect to Chartered Engineer Accreditation.
+The Engineering Council is in the process of seeking further consistency.
+This has lead to the new rules with respect of Compensation and Condonement for Eng accreditation \cite{EC2018}.
+Further changes are embedded in AHEP 4.
+The BCS requirements have been updated to reflect the expectations of Compensation and Condonement.
+Once AHEP 4 is finalised further work will be required to address these updates.
+Among other proposals, AHEP 4 intends to extend Chartered Engineering accreditation to include diversity and widening participation data (of staff and students) as part of the metrics that assess the quality of provision.
+The Computer Science discipline could clearly do better in this regard, so this should be a positive inclusion.
+
+\subsection{Support for HEIs seeking accreditation}
+As part of its operation BCS continually reviews and monitors the support process it employs for HEIs seeking accreditation and hence will continue to enhance processes in this area.
+
+\subsection{Agility of processes and changing external environment}
+There are considerable external pressures placed upon HEIs, by for example the Research Excellent Framework (REF) and the Teaching Excellence Framework(TEF), which could reduce the priority placed upon accreditation as a mechanism for supporting enhancement.
+Care is taken in visit reports to ensure positive aspects of provision are emphasised.
+This is in part intend to provide evidence that could be used to support future TEF related submissions.
+Discussions regarding the agility of processes are incorporated into two annual committee meetings to identify short, medium and long term projects.
+
+\subsection {Further role for employers in accreditation}
+As indicated previously an industrial assessor is a critical part of every visit panel.
+There are also interesting developments being lead by the IoC.
+Specifically, the IoC is proposing micro-credentialing as an extension to accreditation.
+This could potentially augment the current recognised pathways to professional accreditation by providing a more fine grained alternatives that could be useful to some employers or employees who wish to evidence their achievements in an accredited manner.
+The details of this proposal is not finalised, however the BCS continues to explore possibilities with the IoC with a view to supporting any initiative of mutual benefit.
+One of the challenges in this work is the lack of an agreed standard (in the way there is for the Seoul and Washington Accords) for micro-credentialing.
+
+\subsection {Exploration of the value of accreditation to all stakeholders}
+As discussed earlier,  BCS AAC is in the process of initiating a conversation to explore the value proposition of accreditation from other stakeholders points of view.
+
+\subsection {Improve communications}
+There is a strong and vibrant community of academics and industrialists who actively participate in accreditation.
+However, there are many involved with whom communication could benefit from being more active.
+A number of initiatives are in process to enhance these communications.
+For HEIs there are ongoing discussion within BCS regarding establishing a periodic bulletin which could be used to share updates of accreditation practice / procedure, highlight HEIs having their first visit, sharing good practice examples, promote assessor recruitment and so on.
+Assessor Peer Review has indicated more regular communication with assessors could be of benefit.
+Virtual link up sessions for assessors are proposed.
+The BCS are in the process of contacting assessors to determine likely interest and thoughts related to relevant topics for such link ups.
+
+\subsection {Mechanisms for sharing good practice}
+Whilst BCS has been identifying aspects of good practice as part of accreditation visits for a few years.
+They mechanism (publishing on the BCS website) does not appear to have been very effective.
+There are conversations ongoing regarding how this can be improved.
+
+\section{Why are you telling us this?}	
+%What is interesting, or useful, about this to someone else?
+
+One of the recommendations of Shadbolt is the value of accreditation should be more clearly communicated to stakeholders. This paper is part of a set of initiatives to achieve precisely that.
+
+Much has changed in BCS accreditation in the last few years.
+For many academics their only experience of BCS accreditation is the quinquennial BCS accreditation visit.
+Accreditation by the BCS is performed by panels of BCS assessors.
+An assessor may be an industrialist or an academic.
+The size of a panel varies depending upon the number of programmes a visited HEI is putting forward for consideration.
+A panel will always include an industrial assessor and two or more academic assessors.
+At most HEIs will have two academic assessors.
+There is a large pool of academic assessors and as such not all assessors will complete an accreditation visit every year.
+Not all HEIs have an academic assessor.
+Hence a significant set of the Computer Science education community are not informed regarding the evolution of accreditation practice.
+
+As such the intention of the paper is share to the wider community the developments made, the future enhancement aspirations and the forthcoming opportunities to engage in further conversations.
 

### 2019-10-07 CEP20/CEP_accreditationpostshadbolt.tex
JHD edits

@@
 \begin{abstract}
-The promotion of quality via Accreditation by Professional, Statutory and Regulatory Bodies (PSRBs) is very much a feature of higher education in the united kingdom (UK).
-In this changing  environment, there is a need for accreditation regimes to evolve in order to maximise the value they provide to Higher Education Institutions(HEIs).
+The promotion of quality via Accreditation by Professional, Statutory and Regulatory Bodies (PSRBs) is very much a feature of higher education in the United Kingdom (UK).
+In this changing  environment, there is a need for accreditation regimes to evolve in order to maximise the value they provide to Higher Education Institutions (HEIs).
 This paper provides a summary in the context of one professional body (BCS, The Chartered Institute For IT) of
@@
 
-The Shadbolt Review \cite{shadbolt2016shadbolt} investigated high unemployment rates graduates of computer science and related degrees in the United Kingdom and the role of degree accreditation in promoting employability.The Shadbolt review noted that the rates of graduate unemployment were declining and varied considerably between geographic location and type of HEI.
+The Shadbolt Review \cite{shadbolt2016shadbolt} investigated the relatively high unemployment rates graduates of computer science and related degrees in the United Kingdom and the role of degree accreditation in promoting employability.
+The Shadbolt review noted that the rates of graduate unemployment were declining and varied considerably between geographic location and type of HEI.
 The discipline was already acting proactively to address the issue.
@@
 BCS has made significant adjustments to their accreditation processes since 2015.
-These were partly in response to the recommendations of Shadbolt but also to other changes in the sector.
+These were partly in response to the recommendations of Shadbolt but also to other changes in the sector and the discipline.
 These have been communicated to the assessor community and to those responsible for leading the development of HEI accreditation applications.
@@
 
-Criticisms of the accreditation of degree programmes is not new, there is a history of claiming the processes are unnecessarily bureaucratic and constrain  innovation \cite{Harvey2004}.
-Equally its value has been highlighted particularly in the context of a potentially globally mobile workforce, and the dangers of accreditation streams being revenues streams in their own right rather than for the benefit of a discipline or wider society \cite{Knight_2015}.
+Criticisms of the accreditation of degree programmes is not new, there is a history of claiming the processes are unnecessarily bureaucratic and constrain  innovation \cite{Harvey2004},  and there are dangers of accreditation streams being revenues streams in their own right rather than for the benefit of a discipline or wider society \cite{Knight_2015}.
+Equally its value has been highlighted particularly in the context of a potentially globally mobile workforce.
 
@@
 In most cases the HEI being reviewed chooses the peer.
-As part of a BCS (or other professional body) a HEI has less choice regarding who completes a review and hence such a review is arguably more truly independent.
+As part of a BCS (or other professional body) accreditation a HEI has less choice regarding who completes a review and hence such a review is arguably more truly independent.
 BCS review panels contain a minimum of two experienced assessors who have strong awareness of the discipline norms across the UK sector.
@@
 The most recent example of this is the inclusion of cybersecurity in all accredited degrees \cite{Cricketal2019}.
-BCS is in the process of mandating the inclusion of security in all accredited degree programmes to a standard agreed between professional bodies, industry and government.
+%BCS is in the process of mandating %% JHD strengthens text
+BCS mandates the inclusion of security in all accredited degree programmes to a standard agreed between professional bodies, industry and government.
 This is a work in progress \cite{Irons2016}.
@@
 In the United Kingdom the most common form of accreditation in the Computer Science and related discipline areas is by professional bodies.
-The BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) are the main bodies involved.
-The accreditation provided by these institutes are underpinned by international Initiatives such as the Washington Accord \cite{Washington2019} and Seoul Accord \cite{Seoul2019}.
+BCS, The Chartered Institute for IT (BCS) and the Institution of Engineering and Technology (IET) are the main bodies involved.
+The accreditation provided by these %institutes
+are underpinned by international Initiatives such as the Washington Accord \cite[for CEng]{Washington2019} and Seoul Accord \cite[for CITP]{Seoul2019}.
 These memoranda support the internationalising of the curriculum and promote consistency and parity in Computer Science education globally.
-These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006}, \cite{Mutereko2017}.
+These professional bodies are also registered charities and hence have responsibilities for public good which extends beyond short term job market needs \cite{Stensaker2006,Mutereko2017}.
 This is reflected in IETs and BCSs long history of expecting coverage environmental factors within the programmes they accredit and BCSs expectation of significant coverage of Legal, Ethical, Social and Professional Issues \cite{Brooke2018}.
@@
 
-\section {Does it work?}	H
+\section {Does it work?}
 %how do you know? Give some evidence of effectiveness in context.
@@
 
-\subsubsection{Not for Profit Organisations}
+\subsection{Not for Profit Organisations}
 Tech Partnership Degrees is a Not For Profit Organisation that provides endorsements to Higher Education programmes with specific curricula elements aimed at job market requirements \cite{TP2019}.
@@
 
-\subsubsection{National Cyber Security Centre (NCSC)}
+\subsection{National Cyber Security Centre (NCSC)}
 The NCSC is a UK Government organisation tasked with enhancing the cybersecurity of the UK.
@@
 
-\subsubsection{The wider Computer Science education community}
+\subsection{The wider Computer Science education community}
 The Computer Science discipline has responded actively to the challenges presented by Shadbolt with CPHC operating four working groups \cite{cphc_2016},  promotion of further engagement between industry and academia by initiatives such as the Royal Academy of Engineering Visiting Industrial Professor Scheme \cite{royal}
@@
 On occasions BCS guidelines employ English / Welsh higher education terms without indicating that local equivalents are equally acceptable.
-The next set of guidelines will address this short coming.
+The next set of guidelines will address this shortcoming.
 
 \subsection {Engineering Council and Accreditation of Higher Education Programmes (AHEP) Version 4}
-In the British jurisdiction of the Washington accord, the approach adopted is unusual in that the Engineering Council extends the license to accredit to a large number of professional bodies \cite{EC2019}.
+In the British jurisdiction of the Washington accord, the approach adopted is unusual in that the Engineering Council extends the license to accredit \emph{Chartered Engineer} to a large number of professional bodies \cite{EC2019}.
 This is not the practice adopted in other jurisdictions, for example in the United States of America ABET is the sole body.
@@
 There is a large pool of academic assessors and as such not all assessors will complete an accreditation visit every year.
-Not all HEIs have an academic assessor.
+Conversely, not all HEIs have an academic assessor.
 Hence a significant set of the Computer Science education community are not informed regarding the evolution of accreditation practice.

### 2019-10-08 CEP20/CEP_cybersecurity.tex
Early draft

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{How well do we teach Cybersecurity}
+\title{The challenges of teaching cybersecurity as part of mainstream computer science and related disciplines?}
 
@@
 \renewcommand{\shortauthors}{Crick, Davenport,  Irons, and Prickett.}
-
 %%
@@
 \begin{abstract}
-  A clear and well-documented \LaTeX\ document is presented as an
-  article formatted for publication by ACM in a conference proceedings
-  or journal publication. Based on the ``acmart'' document class, this
-  article presents and explains many of the common variations, as well
-  as many of the formatting elements an author may use in the
-  preparation of the documentation of their work.
+  "Every Computer Science Degree Should Require a Course in Cybersecurity" **TO DO add reference**
+  argued a recent article in the Harvard Business review.
+Computer Science and related degrees in the united Kingdom are evolving to meet this expectation.
+The BCS, The Charterted Institute for IT has been mandating this for a number of years \cite{Cricketal2019}.
+This may have encouraged these developments, although the discipline has been embedding cybersecurity before it was mandated by the BCS.
+Delivering cybersecurity effectively presents a number of challenges related to research and pedagogy.
+This article explores the progress to date, challenges for the future and highlights areas of potential  enhancement.
+
 \end{abstract}
@@
 %%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm
 %% Please copy and paste the code instead of the example below.
@@
 
-%%
+% To do add cybersecurity to this
+
 %% Keywords. The author(s) should pick words that accurately describe
@@
 \section {What is it?}
-A short description of the practice you're presenting
+%A short description of the practice you're presenting
+
+This paper explores the challenges related to research and pedagogy of delivering the this cybersecurity content.
 
 \section {Why are you doing it?}	What happened before? What is it changing / replacing / improving? What gap is it filling?
+
+Cybersecurity is becoming increasing pivotal to the operation of organisations of all sizes and all organisations are increasingly expected to make reasonable adjustments to protect their activities
+
+\begin{quote}
+	``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''\\
+	\hfill Robert Hannigan~\cite{Hannigan2019a}, former Director of GCHQ
+\end{quote}
+
+This focus on cybersecurity, includes calls for formal education -- school-level as well as tertiary -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a,NCSC2017}.
+This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}
+
+Computer Science and related disciplines are evolving to meet these demands.
+However doing so is not without challenges.
+This paper explores the progress to date, challenges for the future and highlights a number of enhancement activities for the discipline.
 
@@
 \section {Does it work?}	How do you know? Give some evidence of effectiveness in context.
+
+The UK situation appears relatively advanced compared to other jurisdictions.
+61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping~\cite[Table 1]{Ruiz2019a}.
+As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.
+
+In many jurisdictions, Professional, Statutory and Regulatory Bodies (PSRBs) have responded to these expectations by adjusting their accreditation requirements.
+In the UK, BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015, with the result that all accredited universities should be compliant by 2020 (due to the five-year cycle).
+More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.
+
+It has been reported that this initiative is showing good progress \cite{Cricketal2019} with the majority university's already meeting these expectations and a further number evolving their provision so that they do so.
+
+Collective challenges still persist related to underpinning resources, staffing and expected pedagogies.
+
 \section {Who else has done this?}	Where did you get the idea from?
@@
 What did you change?
+
+In jurisdictions other than the UK, PSRB's have also adjusted their expectations to enhance the cybersecurity provision.
+For example in the United States of America (USA), the Association of Computing Machinery (ACM) has equally had cybersecurity (IAS: ``Information Assurance and Security'') in the curriculum since 2013~\cite{ACM2013a}, but it is not the accrediting body.
+The Accreditation Board for Engineering and Technology (ABET) is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''.
+This should mean that  all accredited universities should be compliant by 2025 (due to their six-year cycle)
+
+The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs):
+
+\begin{quote}
+	``{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
+			given the manner in which the topics are pervasive throughout other KAs.}}''
+\end{quote}
+
+It proposes nine ``core'' hours and 63.5 distributed across the other KAs. Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:
+
+\begin{quote}
+	``{\emph{Universities suffer shortcomings, as well.
+Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity.
+There is a big catch, however.
+Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}''
+\end{quote}
+
+However, whilst the accreditation regimes in the UK are mandating inclusion of cybersecurity there remain challenges in terms of the underpinning resources, the available Faculty resources and pedagogies required to effectively deliver the cybersecurity content.
+
 \section {What will you do next?}	Will you vary this, or develop it further?
